### 2013-03-12 inst/examples/salmonport.Rmd
Add main manuscript files in inst/examples

% Portfolio prioritization of salmon conservation
% Sean Anderson

# Abstract
First part is copied from what I wrote before (introduction and questions). Second part outlines the simulation. I haven't yet worked very much with the portfolios themselves, but I've set up the structure of the simulation. I outline how the simulation is currently working and show examples of some of the plotting functions I have to look at simulation and portfolio optimization.

# Introduction
[Based partly on Jon's notes.] Managing risk is fundamental to the conservation of an endangered species. When an endangered species exists as a metapopulation, we can manage risk at two levels: at the population level or at the metapopulation level. Typically we treat sources of risk at the metapopulation level as external and uncontrollable and so we manage risk by altering anthropogenic disturbances such as fishing or hunting on a population level as well as improving connectivity of populations.

The management of financial portfolios provides another way of considering risk. Economists consider the risk and performance of a financial portfolio as depending on the weighting of individual assets that make up the portfolio. Modern Portfolio Theory (MPT) proposes that there is a set of portfolios that achieves maximum expected return for a level of expected risk or minimum expected risk for a level of expected return [@Markowitz1952; @Markowitz1959]. This optimal set contains portfolios that range along a continuum of risk-tolerance; economists refer to this set as the efficient frontier.

Similarly, expected growth rate and variance at a metapopulation level will be a function of the properties of the populations. A portfolio approach to managing risk for a metapopulation therefore considers how conservation might affect the ``weight'' of each population in the metapopulation ``portfolio''.

Bring in salmon: we can manage risk at the population level or at the ESU level. ESU-level conservation is mandated but we generally lack tools to quantitatively manage risk at this level.

\begin{table}[h!]
\centering
\small
\caption{Components of salmon metapopulation portfolios}
\begin{tabular}{p{3.6cm}p{7.5cm}}
\toprule
Component & Definition for the salmon portfolio\\
\midrule
Assets & Viable salmonid populations? Stream-level populations? Depends on the 
system\\
Portfolio & The salmon metapopulation; possibly an ESU\\
Portfolio managers & Salmon managers\\
Investors & Salmon managers, conservation agency, or salmon fishers (each with 
their own goals and mandates)\\
Asset weights & Carrying capacity, or catches, or salmon spawner (or return) 
abundance\\
Asset returns & Annual rate of change of salmon returns or spawners; possibly 
annual rate of change of fisheries catches\\
Asset risk & Could be variance of asset returns; could be a coherent asymmetric 
risk metric such as CVaR\\
\bottomrule
\end{tabular}
\label{tab:port-components}
\end{table}

## Portfolio prioritization of salmon conservation
Here, we conduct a post-hoc assessment of the set of optimal salmon portfolios given various return-risk definitions and objectives. We ask: What does a portfolio approach to management tell us about salmonid conservation priorities? Does it produce any rules of thumb about what optimal portfolio management of salmonid metapopulations would entail?

## Questions
1. What are the attributes of the populations that form the optimal set? For example, are populations with high response diversity prioritized?

2. How much lower is stream-level and ESU-level extinction risk for the optimal portfolios than the less optimal portfolios? Extinction probability could be estimated through long-term simulation or through an asymmetric risk metric like conditional value at risk (CVaR). Given I am including straying, this will most likely be a quasi-extinction-risk metric.

3. How do our answers to other questions change depending on what we measure on the risk-return axes? For example, we could trade off the mean rate of change of salmon returns vs.\ variance of the rate of change. This would be most similar to a financial portfolio optimization. We could also trade off the salmon returns (or catches) themselves. Further, we could work with variance or a downside risk metric like CVaR.

4. How do constraints (e.g.\ viable salmonid population (VSP)-like constraints) affect the selected portfolios in risk-return space and in terms of common metrics such as population-level and ESU-level extinction risk?

5. How do the properties of the optimal portfolio set change based on different kinds of environmental driver patterns? For example, magnitude of driver variance, cyclical dynamics like the PDO, black swans or unexpected surprises, increases in temperature and variability through time (climate change).

Simulation
==========
See Figure \ref{fig:sim-diagram} for an illustration of the simulation 
structure.

\begin{figure}[htbp]
  \centering
    \includegraphics[height=7in]{simulation_diagram.pdf}
 \caption{Flow chart of the salmon-metapopulation simulation and portfolio optimization. There are $n$ salmon populations and $t$ generations. Blue text indicates values that are generated before the simulation progresses through time. Orange text indicates steps in which calculations are performed. Black text indicates values that are calculated. Grey text indicates parameters that can be set. Green text indicates the looping structure of the simulation.}
  \label{fig:sim-diagram}
\end{figure}

## Spawner-return relationship
I am using a Ricker curve,

\begin{equation}
R_{ti} = S_{ti}e^{a_{ti}(1-S_{ti}/b_i) + w_{ti}},
\end{equation}

\noindent
where $t$ represents a generation time, $i$ represents a population, $R$ is the number of returns, $S$ is the number of spawners, $a$ is the productivity parameter (which can vary with the environmental signal), and $b$ is the density-dependent term (which is used as the asset weights in the portfolios). The term $w_{ti}$ represents first-order autocorrelated error (AR1). Formally, $w_{ti} = w_{ti-1} \rho + v_{ti}$, where $v_{ti}$ represents independent and normally distributed error with mean 0 and standard deviation of $\sigma_v$. The parameter $\rho$ represents the correlation between residuals from subsequent years.

## Escapement targets
I am calculating spawning abundance MSY $S_{MSY}$ each generation using the equation $S_{MSY} = b(0.5-0.07a)$ from @Hilborn1992 p272, Table 7.2. I am using the precisely known $a$ and $b$ values each time and assume perfect implementation of the escapement goal (no outcome uncertainty) and no upper limit on how much the fishery can harvest. We may want to add a bit of noise here, say by assuming a constant value for $a$ or adding some implementation uncertainty. We need some level of harvesting to keep the spawning population away from carrying capacity. At carrying capacity, the productivity parameter $a$ has minimal effect on population dynamics, making it a poor way to integrate response diversity.

## Environmental signal
I have programmed five different kinds of environmental signals. I think it's easiest to think of these as temperature signals, but they could represent more general environmental anomalies, such as the PDO. The five kinds of environmental signals are described in Table \ref{tab:env-types} and illustrated in Figure \ref{fig:env-ts}. Later, I may include black-swan-like events, although the regime shifts could be seen this way if the changes are dramatic enough. So far all figures use a sine wave. The problem with using a randomly generated ARMA time series is that you get much greater differences in the portfolios from 
run to run.

\begin{table}[h!]
  \centering
  \small
  \caption{Environmental signals, parameters, and descriptions.}
\begin{tabular}{p{3.2cm}p{3.25cm}p{6.8cm}}
\toprule
Signal type & Parameters & Description \\
\midrule
Sine wave & $\overline{T} + A \cdot \sin(\omega t + \phi)$& $\overline{T}$ is 
the mean environmental value, $A$ is the amplitude, $\omega$ is the frequency in 
radians, and $\phi$ is the phase\\
ARMA & AR1, MA1, $\sigma_T$ & First-order autoregressive (AR1) and moving 
average (MA1) parameters and standard deviation; mean is assumed to be 0\\
Regime shifts & $T_{b0},T_{b1}, \ldots, T_{bn}$; $T_{v0},T_{v1}, \ldots, T_{vn}$ 
& Breakpoints ($T_{b}$) and values ($T_v$) for regimes $1$ through $n$\\
Linear change & $T_{\mathrm{min}}$, $T_{\mathrm{max}}$ & Minimum and maximum 
values\\
Constant & $T_{\mathrm{constant}}$& Constant value\\
\bottomrule
\end{tabular}
\label{tab:env-types}
\end{table}

```{r env-ts, fig.height=5, fig.width=4, fig.cap="Example temperature time series\\label{fig:env-ts}"}
source("plot_env_ts.R")
col_pal <- rev(gg_color_hue(10))
```

## Straying
[Insert a bit here about my research on straying and it's influence on metapopulation dynamics for salmon]. Essentially, straying should be enough to prevent local extinctions and ensure all possible habitat is occupied, but probably not strong enough to noticeably affect population dynamics or synchrony [@Schtickzelle2007].

I've implemented straying as in @Cooper1999. I generate a matrix that represents the fraction of straying between any two populations. I have set up the metapopulation (thus far) as in in a very simple scenario: the populations are arranged in a line and those that are nearer to each other are more likely to stray between each other [insert caveats]. Two parameters control the straying: the fraction of fish that stray from their natal stream in any given generation $f$ and the rate at which this straying between streams decays with distance $m$. I calculated the number of salmon straying from stream $i$ to stream $j$ as,

\begin{equation}
  \mathrm{strays}_{tij} = \frac{S_{ti} f e^{-m \lvert i-j \rvert }}{\sum\limits_{k \neq j}^{} e^{-m \lvert k-j \rvert }},
\end{equation}

where the denominator is a normalization constant to ensure that no salmon are gained or lost. The subscript $k$ represents a stream ID. [Not sure this is written correctly --- I'll think about it --- but it should be coded correctly.] There is no straying from and to the same stream. See Figure \ref{fig:stray-matrix} for an example straying matrix.

```{r stray-matrix, fig.height=3.7, fig.width=3.9, fig.cap="Example straying matrix. Darker blue colours indicate a greater rate of straying.\\label{fig:stray-matrix}"}
source("plot_straying_matrix.R")
```

## Diversity of response to the environmental conditions
[insert research I've done here on what can generate diversity of response to the environment across populations and how it is typically incorporated] This is a critical part to the paper.

I am incorporating diversity of response to environmental conditions, say temperature, via changes in the Ricker $a$ parameter. I am using thermal tolerance curves, based loosely on @Eliason2011. They identify unique thermal tolerance curves in terms of cardiorespiratory performance. I am using approximately the same temperature ranges and variability of those curves, but applying them to productivity. I show the curves I'm currently using in Figure \ref{fig:thermal-curves}. We could vary the distribution of the optimum temperatures, the maximum $a$ values, and the width of the curves if we wanted, although I think the current curves are a simple and straightforward base scenario and we may not need to do anything more complicated.

I calculate $a_{ti}$, the realized Ricker productivity parameters after accounting for thermal tolerance, as

\begin{equation}
  a_{ti} = a_i^* + \gamma_i (T_t - T_i^*)^2,
\end{equation}

where $a_i^*$ is the maximum possible productivity parameter for population $i$, $T_t$ is the temperature at time step $t$, $T_i^*$ is the optimal temperature for population $i$, and $\gamma_i$ controls the width of the thermal tolerance curve for each population. See Figure \ref{fig:thermal-curves} for the default thermal tolerance curves for ten populations.

Another possibility would be to draw the $a^*$, $\gamma$, and $T^*$ randomly. The advantage to setting it up as I have (with pre-specified thermal tolerance curves) is that it will hopefully make the post-hoc analysis of what affects the efficient frontier weights more straightforward.

```{r thermal-curves, cache=FALSE, fig.height=3.9, fig.width=5.4, fig.cap="Thermal tolerance curves for ten populations. Here, and throughout this document, I have used the same rainbow of colours to identify the salmon populations. The populations are arranged from blue (more productive with cooler temperatures) to red (more productive with warmer temperatures\\label{fig:thermal-curves}"}
source("plot_thermal_curves.R")
```

Portfolio optimization
======================
Details to follow. Quick version: I Monte Carlo the $b$ parameters (with a specified minimum value) and measure the resulting mean and variance of the rate of growth of the metapopulation salmon returns. I then identify the set of $b$ parameters (i.e.\ the set of weights) that results in the efficient frontier. We could later choose to work with risk metrics like conditional value at risk (CVaR), which measure the probability of a ``bad'' event specifically, not the symmetrical variance.

To find the efficient portfolios I divide all portfolios into bins by mean rate of change and find the portfolios with the minimum variance within each bin. I discard all portfolios with a mean rate of change below the minimum variance portfolio. These are ``undesirable'' portfolios.

Comparing portfolios
====================
This is a big area to think about. How do portfolios differ along the frontier? How do the frontier portfolios differ from non-frontier portfolios? What is the best way to visually compare them? What aspects to compare? Are there summary statistics we could use to compare them?

Some sample simulations
=======================
The following are options and their default values:

\begin{verbatim}
  n_t = 120, # number of years
  n_pop = 10, # number of populations
  stray_decay_rate = 0.3, # rate that straying decays with distance
  stray_fraction = 0.01, # fraction of fish that stray from natal streams
  b = rep(1000, n_pop), # Ricker density-dependent parameter
  spawners_0 = round(b), # spawners at start
  sigma_v = 0.1, # stock-recruit residual SD
  v_rho = 0.4, # stock-recruit residual AR1 correlation
  a_width_param = rep(0.02, n_pop), # width of thermal curves by pop
  optim_temp = seq(13, 19, length.out = n_pop), # optimal temperatures by pop
  max_a = rep(1.4, n_pop), # maximum Ricker a values by pop at optimum temp
  env_type = c("sine", "arma", "regime", "linear", "constant"),
  env_params = list(amplitude = 2.0, ang_frequency = 0.2, phase = 0, mean_value = 16),
  use_cache = FALSE, # regenerate stochastic values? 
  add_straying = TRUE # include or ignore straying
\end{verbatim}

First, let's run a couple example simulations:

```{r run-base-example1, fig.height=6.5, fig.width=5, fig.cap="Example simulation with all parameters at default values. I'm only showing a sample of 40 generations (after the burn-in period). From top to bottom: salmon returns each generation (catch plus escapement); Ricker $a$ parameter; the environmental signal (could think of it as temperature or a composite environmental signal like the PDO); number of salmon straying by population; number of straying salmon being received by population; fisheries catch based on escapement MSY calculation; salmon escapement; residuals from the spawner-return curve."}
set.seed(2)
eg1 <- meta_sim(b = rep(1000, 10), n_pop = 10)
plot_sim_ts(eg1, years_to_show = 40)
```

```{r plot-base-corrs, fig.width=4, fig.height=4, fig.cap="A plot comparing the log(returns) between each population. The population colours and numbers match in all figures. Note how populations 1 and 10 have asynchronous returns whereas populations with more similar thermal-tolerance curves (say populations 9 and 10) have more synchronous dynamics. Populations with thermal tolerance curves in the middle (e.g.\ population 6) are less correlated with other populations. Their population dynamics end up primarily driven by demographic stochasticity and less so by temperature-induced systematic changes in productivity."}
plot_correlation_between_returns(eg1, pal = col_pal)
```

```{r plot-base-rickers, fig.width=8, fig.height=5, cache=FALSE, fig.cap="We can also look at samples of the Ricker curves by population (population numbers are in the upper left corners). Here, I've sampled 40 of the $a$ parameters through time for each population. The shaded rectangles to indicate the range of spawners observed in the simulation. The colour of the line indicates the temperature anomaly in that year with blue indicating cooler years and red indicating warmer years. In this example all the populations have the same $b$ parameters. Notice how the productivity is more variable for populations nearer the limits of their thermal tolerance (e.g.\ population 1 vs.\ population 5). Also notice how population 1 is more productive in cool years whereas population 10 is more productive in warm years."}
plot_rickers(eg1, pal = col_pal)
```

```{r plot-various-options-ts-3pops, fig.width=4, fig.height=6.3, fig.cap="Example simulations with different parameter values. I am showing three populations here to make the plots easier to interpret.\\label{fig:sim-param-ts-3}"}
source("example_returns_3.R")
```

```{r plot-various-options-ts, fig.width=4, fig.height=6.3, fig.cap="Example simulations with different parameter values. This version has ten populations.\\label{fig:sim-param-ts-10}"}
source("example_returns_10.R")
```

\clearpage

An example of portfolio optimization
====================================
Now, let's run an example where we find the $b$ values that form the efficient frontier of metapopulation portfolios.

```{r set-a-seed, echo=FALSE}
set.seed(3)
```

```{r monte-carlo-portfolios}
weights_matrix <- create_asset_weights(n_pop = 10, n_sims = 100, 
  weight_lower_limit = 0.01)
mc_ports <- monte_carlo_portfolios(weights_matrix = weights_matrix,
  n_sims = 100, mean_b = 1000)
```

```{r eff-frontier-eg, fig.width=6.5, fig.height=3.5, fig.cap="Efficient frontier of metapopulation portfolios (red dots). Each dot represents a different set of weights of the Ricker $b$ parameters. Metapopulation growth rate is the first difference of the log of total metapopulation abundance. The colours in the right panel correspond to the other figures."}
ef_dat <- plot_efficient_portfolios(port_vals = mc_ports$port_vals, pal = 
  col_pal, weights_matrix = weights_matrix)
```

```{r plot-eff-ports, fig.height=6.5, fig.width=5, fig.cap="Simulation output from the minimum-variance metapopulation."}
min_var_sim <- mc_ports$sims_out[[ef_dat$min_var_port_id]]
plot_sim_ts(min_var_sim, pal = col_pal)
```

```{r fig.width=8, fig.height=5, fig.cap="Example Ricker curves for the minimum-variance metapopulation."}
plot_rickers(min_var_sim, pal = col_pal)
```

```{r plot-portfolio-timeseries, fig.width=6.5, fig.height=3.5, fig.cap="Time series for a random sample of portfolios (grey lines) and the minimum-variance portfolio (white lines). The top panel shows rate of change or metapopulation abundance at each time step and the bottom panel shows the metapopulation abundance."}
source("plot_portfolio_ts.R")
```

\clearpage

References
==========

### 2013-03-12 inst/examples/salmonport.tex
Add main manuscript files in inst/examples

--- 2013-03-12 inst/examples/salmonport.Rmd
@@
-% Portfolio prioritization of salmon conservation
-% Sean Anderson
+%\documentclass[]{article}
+\documentclass[12pt]{article}
 
-# Abstract
-First part is copied from what I wrote before (introduction and questions).
-Second part outlines the simulation.
-I haven't yet worked very much with the portfolios themselves, but I've set up the structure of the simulation.
-I outline how the simulation is currently working and show examples of some of the plotting functions I have to look at simulation and portfolio optimization.
+% my modifications:
+\usepackage{fontspec}
+\usepackage{booktabs}
+\defaultfontfeatures{Ligatures=TeX}
+\setsansfont{Calibri}
+\setmonofont{Inconsolata}
+\setmainfont{Linux Libertine}
+\DeclareTextCommandDefault{\nobreakspace}{\leavevmode\nobreak\ }
+\widowpenalty=1000
+\clubpenalty=1000
+\usepackage{setspace}
+\onehalfspacing
 
-# Introduction
-[Based partly on Jon's notes.] Managing risk is fundamental to the conservation of an endangered species.
-When an endangered species exists as a metapopulation, we can manage risk at two levels: at the population level or at the metapopulation level.
-Typically we treat sources of risk at the metapopulation level as external and uncontrollable and so we manage risk by altering anthropogenic disturbances such as fishing or hunting on a population level as well as improving connectivity of populations.
+% A bit bigger than one inch margins:
+%\topmargin -1.43cm
+%\oddsidemargin 0.4cm
+%\evensidemargin 0.4cm
+%\textwidth 14.5cm
+\textheight 22.5cm
 
-The management of financial portfolios provides another way of considering risk.
-Economists consider the risk and performance of a financial portfolio as depending on the weighting of individual assets that make up the portfolio.
-Modern Portfolio Theory (MPT) proposes that there is a set of portfolios that achieves maximum expected return for a level of expected risk or minimum expected risk for a level of expected return [@Markowitz1952; @Markowitz1959].
-This optimal set contains portfolios that range along a continuum of risk-tolerance; economists refer to this set as the efficient frontier.
+%\usepackage[left=1.5in,top=1.5in,right=1.5in,bottom=1.5in,nohead,nofoot]{geometry}
+\usepackage{geometry}
+\geometry{letterpaper}
 
-Similarly, expected growth rate and variance at a metapopulation level will be a function of the properties of the populations.
-A portfolio approach to managing risk for a metapopulation therefore considers how conservation might affect the ``weight'' of each population in the metapopulation ``portfolio''.
+\usepackage[round]{natbib}
+\bibliographystyle{cjfas}
+\bibpunct{(}{)}{;}{a}{}{;}
 
-Bring in salmon: we can manage risk at the population level or at the ESU level. ESU-level conservation is mandated but we generally lack tools to quantitatively manage risk at this level.
+%\usepackage[T1]{fontenc}
+%\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+% use upquote if available, for straight quotes in verbatim environments
+%\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+%\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  %\usepackage[utf8]{inputenc}
+%%\else % if luatex or xelatex
+  %\usepackage{fontspec}
+  %\ifxetex
+    %\usepackage{xltxtra,xunicode}
+  %\fi
+  %\defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
+  %\newcommand{\euro}{€}
+%%%%%\fi
+% use microtype if available
+%\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
+%%%\usepackage{natbib}
+%\bibliographystyle{plainnat}
+%%%%%\usepackage{graphicx}
+% We will generate all images so they have a width \maxwidth. This means
+% that they will get their normal width if they fit onto the page, but
+% are scaled down if they would overflow the margins.
+%\makeatletter
+%\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
+%\else\Gin@nat@width\fi}
+%\makeatother
+%\let\Oldincludegraphics\includegraphics
+%\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
+%\urlstyle{same}  % don't use monospace font for urls
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+%
+\title{Portfolio prioritization of salmon conservation}
+\author{Sean Anderson}
+\date{}
+
+\begin{document}
+\maketitle
+
+\section{Abstract}
+
+First part is copied from what I wrote before (introduction and
+questions). Second part outlines the simulation. I haven't yet worked
+very much with the portfolios themselves, but I've set up the structure
+of the simulation. I outline how the simulation is currently working and
+show examples of some of the plotting functions I have to look at
+simulation and portfolio optimization.
+
+\section{Introduction}
+
+{[}Based partly on Jon's notes.{]} Managing risk is fundamental to the
+conservation of an endangered species. When an endangered species exists
+as a metapopulation, we can manage risk at two levels: at the population
+level or at the metapopulation level. Typically we treat sources of risk
+at the metapopulation level as external and uncontrollable and so we
+manage risk by altering anthropogenic disturbances such as fishing or
+hunting on a population level as well as improving connectivity of
+populations.
+
+The management of financial portfolios provides another way of
+considering risk. Economists consider the risk and performance of a
+financial portfolio as depending on the weighting of individual assets
+that make up the portfolio. Modern Portfolio Theory (MPT) proposes that
+there is a set of portfolios that achieves maximum expected return for a
+level of expected risk or minimum expected risk for a level of expected
+return \citep{Markowitz1952, Markowitz1959}. This optimal set contains
+portfolios that range along a continuum of risk-tolerance; economists
+refer to this set as the efficient frontier.
+
+Similarly, expected growth rate and variance at a metapopulation level
+will be a function of the properties of the populations. A portfolio
+approach to managing risk for a metapopulation therefore considers how
+conservation might affect the
+\texttt{weight'' of each population in the metapopulation}portfolio''.
+
+Bring in salmon: we can manage risk at the population level or at the
+ESU level. ESU-level conservation is mandated but we generally lack
+tools to quantitatively manage risk at this level.
 
@@
 
-## Portfolio prioritization of salmon conservation
-Here, we conduct a post-hoc assessment of the set of optimal salmon portfolios given various return-risk definitions and objectives.
-We ask: What does a portfolio approach to management tell us about salmonid conservation priorities?
-Does it produce any rules of thumb about what optimal portfolio management of salmonid metapopulations would entail?
+\subsection{Portfolio prioritization of salmon conservation}
 
-## Questions
-1. What are the attributes of the populations that form the optimal set? For example, are populations with high response diversity prioritized?
+Here, we conduct a post-hoc assessment of the set of optimal salmon
+portfolios given various return-risk definitions and objectives. We ask:
+What does a portfolio approach to management tell us about salmonid
+conservation priorities? Does it produce any rules of thumb about what
+optimal portfolio management of salmonid metapopulations would entail?
 
-2.
-How much lower is stream-level and ESU-level extinction risk for the optimal portfolios than the less optimal portfolios?
-Extinction probability could be estimated through long-term simulation or through an asymmetric risk metric like conditional value at risk (CVaR).
-Given I am including straying, this will most likely be a quasi-extinction-risk metric.
+\subsection{Questions}
 
-3.
-How do our answers to other questions change depending on what we measure on the risk-return axes?
-For example, we could trade off the mean rate of change of salmon returns vs.\ variance of the rate of change.
-This would be most similar to a financial portfolio optimization.
-We could also trade off the salmon returns (or catches) themselves.
-Further, we could work with variance or a downside risk metric like CVaR.
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  What are the attributes of the populations that form the optimal set?
+  For example, are populations with high response diversity prioritized?
+\item
+  How much lower is stream-level and ESU-level extinction risk for the
+  optimal portfolios than the less optimal portfolios? Extinction
+  probability could be estimated through long-term simulation or through
+  an asymmetric risk metric like conditional value at risk (CVaR). Given
+  I am including straying, this will most likely be a
+  quasi-extinction-risk metric.
+\item
+  How do our answers to other questions change depending on what we
+  measure on the risk-return axes? For example, we could trade off the
+  mean rate of change of salmon returns vs.~variance of the rate of
+  change. This would be most similar to a financial portfolio
+  optimization. We could also trade off the salmon returns (or catches)
+  themselves. Further, we could work with variance or a downside risk
+  metric like CVaR.
+\item
+  How do constraints (e.g.~viable salmonid population (VSP)-like
+  constraints) affect the selected portfolios in risk-return space and
+  in terms of common metrics such as population-level and ESU-level
+  extinction risk?
+\item
+  How do the properties of the optimal portfolio set change based on
+  different kinds of environmental driver patterns? For example,
+  magnitude of driver variance, cyclical dynamics like the PDO, black
+  swans or unexpected surprises, increases in temperature and
+  variability through time (climate change).
+\end{enumerate}
 
-4.
-How do constraints (e.g.\ viable salmonid population (VSP)-like constraints) affect the selected portfolios in risk-return space and in terms of common metrics such as population-level and ESU-level extinction risk?
+\section{Simulation}
 
-5.
-How do the properties of the optimal portfolio set change based on different kinds of environmental driver patterns?
-For example, magnitude of driver variance, cyclical dynamics like the PDO, black swans or unexpected surprises, increases in temperature and variability through time (climate change).
-
-Simulation
-==========
 See Figure \ref{fig:sim-diagram} for an illustration of the simulation
@@
 
-## Spawner-return relationship
+\subsection{Spawner-return relationship}
+
 I am using a Ricker curve,
@@
 \noindent
-where $t$ represents a generation time, $i$ represents a population, $R$ is the number of returns, $S$ is the number of spawners, $a$ is the productivity parameter (which can vary with the environmental signal), and $b$ is the density-dependent term (which is used as the asset weights in the portfolios).
-The term $w_{ti}$ represents first-order autocorrelated error (AR1).
-Formally, $w_{ti} = w_{ti-1} \rho + v_{ti}$, where $v_{ti}$ represents independent and normally distributed error with mean 0 and standard deviation of $\sigma_v$.
-The parameter $\rho$ represents the correlation between residuals from subsequent years.
+where $t$ represents a generation time, $i$ represents a population, $R$
+is the number of returns, $S$ is the number of spawners, $a$ is the
+productivity parameter (which can vary with the environmental signal),
+and $b$ is the density-dependent term (which is used as the asset
+weights in the portfolios). The term $w_{ti}$ represents first-order
+autocorrelated error (AR1). Formally, $w_{ti} = w_{ti-1} \rho + v_{ti}$,
+where $v_{ti}$ represents independent and normally distributed error
+with mean 0 and standard deviation of $\sigma_v$. The parameter $\rho$
+represents the correlation between residuals from subsequent years.
 
-## Escapement targets
-I am calculating spawning abundance MSY $S_{MSY}$ each generation using the equation $S_{MSY} = b(0.5-0.07a)$ from @Hilborn1992 p272, Table 7.2.
-I am using the precisely known $a$ and $b$ values each time and assume perfect implementation of the escapement goal (no outcome uncertainty) and no upper limit on how much the fishery can harvest.
-We may want to add a bit of noise here, say by assuming a constant value for $a$ or adding some implementation uncertainty.
-We need some level of harvesting to keep the spawning population away from carrying capacity.
-At carrying capacity, the productivity parameter $a$ has minimal effect on population dynamics, making it a poor way to integrate response diversity.
+\subsection{Escapement targets}
 
-## Environmental signal
-I have programmed five different kinds of environmental signals.
-I think it's easiest to think of these as temperature signals, but they could represent more general environmental anomalies, such as the PDO.
-The five kinds of environmental signals are described in Table \ref{tab:env-types} and illustrated in Figure \ref{fig:env-ts}.
-Later, I may include black-swan-like events, although the regime shifts could be seen this way if the changes are dramatic enough.
-So far all figures use a sine wave.
-The problem with using a randomly generated ARMA time series is that you get much greater differences in the portfolios from
+I am calculating spawning abundance MSY $S_{MSY}$ each generation using
+the equation $S_{MSY} = b(0.5-0.07a)$ from \citet{Hilborn1992} p272,
+Table 7.2. I am using the precisely known $a$ and $b$ values each time
+and assume perfect implementation of the escapement goal (no outcome
+uncertainty) and no upper limit on how much the fishery can harvest. We
+may want to add a bit of noise here, say by assuming a constant value
+for $a$ or adding some implementation uncertainty. We need some level of
+harvesting to keep the spawning population away from carrying capacity.
+At carrying capacity, the productivity parameter $a$ has minimal effect
+on population dynamics, making it a poor way to integrate response
+diversity.
+
+\subsection{Environmental signal}
+
+I have programmed five different kinds of environmental signals. I think
+it's easiest to think of these as temperature signals, but they could
+represent more general environmental anomalies, such as the PDO. The
+five kinds of environmental signals are described in Table
+\ref{tab:env-types} and illustrated in Figure \ref{fig:env-ts}. Later, I
+may include black-swan-like events, although the regime shifts could be
+seen this way if the changes are dramatic enough. So far all figures use
+a sine wave. The problem with using a randomly generated ARMA time
+series is that you get much greater differences in the portfolios from
 run to run.
@@
 
-```{r env-ts, fig.height=5, fig.width=4, fig.cap="Example temperature time series\\label{fig:env-ts}"}
-source("plot_env_ts.R")
-col_pal <- rev(gg_color_hue(10))
-```
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/env-ts.pdf}
+\caption{Example temperature time series\label{fig:env-ts}}
+\end{figure}
 
-## Straying
-[Insert a bit here about my research on straying and it's influence on metapopulation dynamics for salmon].
-Essentially, straying should be enough to prevent local extinctions and ensure all possible habitat is occupied, but probably not strong enough to noticeably affect population dynamics or synchrony [@Schtickzelle2007].
+\subsection{Straying}
 
-I've implemented straying as in @Cooper1999.
-I generate a matrix that represents the fraction of straying between any two populations.
-I have set up the metapopulation (thus far) as in in a very simple scenario: the populations are arranged in a line and those that are nearer to each other are more likely to stray between each other [insert caveats].
-Two parameters control the straying: the fraction of fish that stray from their natal stream in any given generation $f$ and the rate at which this straying between streams decays with distance $m$.
-I calculated the number of salmon straying from stream $i$ to stream $j$ as,
+{[}Insert a bit here about my research on straying and it's influence on
+metapopulation dynamics for salmon{]}. Essentially, straying should be
+enough to prevent local extinctions and ensure all possible habitat is
+occupied, but probably not strong enough to noticeably affect population
+dynamics or synchrony \citep{Schtickzelle2007}.
+
+I've implemented straying as in \citet{Cooper1999}. I generate a matrix
+that represents the fraction of straying between any two populations. I
+have set up the metapopulation (thus far) as in in a very simple
+scenario: the populations are arranged in a line and those that are
+nearer to each other are more likely to stray between each other
+{[}insert caveats{]}. Two parameters control the straying: the fraction
+of fish that stray from their natal stream in any given generation $f$
+and the rate at which this straying between streams decays with distance
+$m$. I calculated the number of salmon straying from stream $i$ to
+stream $j$ as,
 
@@
 
-where the denominator is a normalization constant to ensure that no salmon are gained or lost.
-The subscript $k$ represents a stream ID.
-[Not sure this is written correctly --- I'll think about it --- but it should be coded correctly.] There is no straying from and to the same stream.
-See Figure \ref{fig:stray-matrix} for an example straying matrix.
+where the denominator is a normalization constant to ensure that no
+salmon are gained or lost. The subscript $k$ represents a stream ID.
+{[}Not sure this is written correctly --- I'll think about it --- but it
+should be coded correctly.{]} There is no straying from and to the same
+stream. See Figure \ref{fig:stray-matrix} for an example straying
+matrix.
 
-```{r stray-matrix, fig.height=3.7, fig.width=3.9, fig.cap="Example straying matrix. Darker blue colours indicate a greater rate of straying.\\label{fig:stray-matrix}"}
-source("plot_straying_matrix.R")
-```
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/stray-matrix.pdf}
+\caption{Example straying matrix. Darker blue colours indicate a greater
+rate of straying.\label{fig:stray-matrix}}
+\end{figure}
 
-## Diversity of response to the environmental conditions
-[insert research I've done here on what can generate diversity of response to the environment across populations and how it is typically incorporated] This is a critical part to the paper.
+\subsection{Diversity of response to the environmental conditions}
 
-I am incorporating diversity of response to environmental conditions, say temperature, via changes in the Ricker $a$ parameter.
-I am using thermal tolerance curves, based loosely on @Eliason2011.
-They identify unique thermal tolerance curves in terms of cardiorespiratory performance.
-I am using approximately the same temperature ranges and variability of those curves, but applying them to productivity.
-I show the curves I'm currently using in Figure \ref{fig:thermal-curves}.
-We could vary the distribution of the optimum temperatures, the maximum $a$ values, and the width of the curves if we wanted, although I think the current curves are a simple and straightforward base scenario and we may not need to do anything more complicated.
+{[}insert research I've done here on what can generate diversity of
+response to the environment across populations and how it is typically
+incorporated{]} This is a critical part to the paper.
 
-I calculate $a_{ti}$, the realized Ricker productivity parameters after accounting for thermal tolerance, as
+I am incorporating diversity of response to environmental conditions,
+say temperature, via changes in the Ricker $a$ parameter. I am using
+thermal tolerance curves, based loosely on \citet{Eliason2011}. They
+identify unique thermal tolerance curves in terms of cardiorespiratory
+performance. I am using approximately the same temperature ranges and
+variability of those curves, but applying them to productivity. I show
+the curves I'm currently using in Figure \ref{fig:thermal-curves}. We
+could vary the distribution of the optimum temperatures, the maximum $a$
+values, and the width of the curves if we wanted, although I think the
+current curves are a simple and straightforward base scenario and we may
+not need to do anything more complicated.
+
+I calculate $a_{ti}$, the realized Ricker productivity parameters after
+accounting for thermal tolerance, as
 
@@
 
-where $a_i^*$ is the maximum possible productivity parameter for population $i$, $T_t$ is the temperature at time step $t$, $T_i^*$ is the optimal temperature for population $i$, and $\gamma_i$ controls the width of the thermal tolerance curve for each population.
-See Figure \ref{fig:thermal-curves} for the default thermal tolerance curves for ten populations.
+where $a_i^*$ is the maximum possible productivity parameter for
+population $i$, $T_t$ is the temperature at time step $t$, $T_i^*$ is
+the optimal temperature for population $i$, and $\gamma_i$ controls the
+width of the thermal tolerance curve for each population. See Figure
+\ref{fig:thermal-curves} for the default thermal tolerance curves for
+ten populations.
 
-Another possibility would be to draw the $a^*$, $\gamma$, and $T^*$ randomly.
-The advantage to setting it up as I have (with pre-specified thermal tolerance curves) is that it will hopefully make the post-hoc analysis of what affects the efficient frontier weights more straightforward.
+Another possibility would be to draw the $a^*$, $\gamma$, and $T^*$
+randomly. The advantage to setting it up as I have (with pre-specified
+thermal tolerance curves) is that it will hopefully make the post-hoc
+analysis of what affects the efficient frontier weights more
+straightforward.
 
-```{r thermal-curves, cache=FALSE, fig.height=3.9, fig.width=5.4, fig.cap="Thermal tolerance curves for ten populations.
-Here, and throughout this document, I have used the same rainbow of colours to identify the salmon populations.
-The populations are arranged from blue (more productive with cooler temperatures) to red (more productive with warmer temperatures\\label{fig:thermal-curves}"}
-source("plot_thermal_curves.R")
-```
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/thermal-curves.pdf}
+\caption{Thermal tolerance curves for ten populations. Here, and
+throughout this document, I have used the same rainbow of colours to
+identify the salmon populations. The populations are arranged from blue
+(more productive with cooler temperatures) to red (more productive with
+warmer temperatures\label{fig:thermal-curves}}
+\end{figure}
 
-Portfolio optimization
-======================
-Details to follow.
-Quick version: I Monte Carlo the $b$ parameters (with a specified minimum value) and measure the resulting mean and variance of the rate of growth of the metapopulation salmon returns.
-I then identify the set of $b$ parameters (i.e.\ the set of weights) that results in the efficient frontier.
-We could later choose to work with risk metrics like conditional value at risk (CVaR), which measure the probability of a ``bad'' event specifically, not the symmetrical variance.
+\section{Portfolio optimization}
 
-To find the efficient portfolios I divide all portfolios into bins by mean rate of change and find the portfolios with the minimum variance within each bin.
-I discard all portfolios with a mean rate of change below the minimum variance portfolio.
-These are ``undesirable'' portfolios.
+Details to follow. Quick version: I Monte Carlo the $b$ parameters (with
+a specified minimum value) and measure the resulting mean and variance
+of the rate of growth of the metapopulation salmon returns. I then
+identify the set of $b$ parameters (i.e.~the set of weights) that
+results in the efficient frontier. We could later choose to work with
+risk metrics like conditional value at risk (CVaR), which measure the
+probability of a ``bad'' event specifically, not the symmetrical
+variance.
 
-Comparing portfolios
-====================
-This is a big area to think about.
-How do portfolios differ along the frontier?
-How do the frontier portfolios differ from non-frontier portfolios?
-What is the best way to visually compare them?
-What aspects to compare?
-Are there summary statistics we could use to compare them?
+To find the efficient portfolios I divide all portfolios into bins by
+mean rate of change and find the portfolios with the minimum variance
+within each bin. I discard all portfolios with a mean rate of change
+below the minimum variance portfolio. These are ``undesirable''
+portfolios.
 
-Some sample simulations
-=======================
+\section{Comparing portfolios}
+
+This is a big area to think about. How do portfolios differ along the
+frontier? How do the frontier portfolios differ from non-frontier
+portfolios? What is the best way to visually compare them? What aspects
+to compare? Are there summary statistics we could use to compare them?
+
+\section{Some sample simulations}
+
 The following are options and their default values:
@@
 
-```{r run-base-example1, fig.height=6.5, fig.width=5, fig.cap="Example simulation with all parameters at default values.
-I'm only showing a sample of 40 generations (after the burn-in period).
-From top to bottom: salmon returns each generation (catch plus escapement); Ricker $a$ parameter; the environmental signal (could think of it as temperature or a composite environmental signal like the PDO); number of salmon straying by population; number of straying salmon being received by population; fisheries catch based on escapement MSY calculation; salmon escapement; residuals from the spawner-return curve."}
-set.seed(2)
-eg1 <- meta_sim(b = rep(1000, 10), n_pop = 10)
-plot_sim_ts(eg1, years_to_show = 40)
-```
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/run-base-example1.pdf}
+\caption{Example simulation with all parameters at default values. I'm
+only showing a sample of 40 generations (after the burn-in period). From
+top to bottom: salmon returns each generation (catch plus escapement);
+Ricker $a$ parameter; the environmental signal (could think of it as
+temperature or a composite environmental signal like the PDO); number of
+salmon straying by population; number of straying salmon being received
+by population; fisheries catch based on escapement MSY calculation;
+salmon escapement; residuals from the spawner-return curve.}
+\end{figure}
 
-```{r plot-base-corrs, fig.width=4, fig.height=4, fig.cap="A plot comparing the log(returns) between each population.
-The population colours and numbers match in all figures.
-Note how populations 1 and 10 have asynchronous returns whereas populations with more similar thermal-tolerance curves (say populations 9 and 10) have more synchronous dynamics.
-Populations with thermal tolerance curves in the middle (e.g.\ population 6) are less correlated with other populations.
-Their population dynamics end up primarily driven by demographic stochasticity and less so by temperature-induced systematic changes in productivity."}
-plot_correlation_between_returns(eg1, pal = col_pal)
-```
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/plot-base-corrs.pdf}
+\caption{A plot comparing the log(returns) between each population. The
+population colours and numbers match in all figures. Note how
+populations 1 and 10 have asynchronous returns whereas populations with
+more similar thermal-tolerance curves (say populations 9 and 10) have
+more synchronous dynamics. Populations with thermal tolerance curves in
+the middle (e.g.~population 6) are less correlated with other
+populations. Their population dynamics end up primarily driven by
+demographic stochasticity and less so by temperature-induced systematic
+changes in productivity.}
+\end{figure}
 
-```{r plot-base-rickers, fig.width=8, fig.height=5, cache=FALSE, fig.cap="We can also look at samples of the Ricker curves by population (population numbers are in the upper left corners).
-Here, I've sampled 40 of the $a$ parameters through time for each population.
-The shaded rectangles to indicate the range of spawners observed in the simulation.
-The colour of the line indicates the temperature anomaly in that year with blue indicating cooler years and red indicating warmer years.
-In this example all the populations have the same $b$ parameters.
-Notice how the productivity is more variable for populations nearer the limits of their thermal tolerance (e.g.\ population 1 vs.\ population 5).
-Also notice how population 1 is more productive in cool years whereas population 10 is more productive in warm years."}
-plot_rickers(eg1, pal = col_pal)
-```
+\begin{verbatim}
+Error: could not find function "ricker_ar1"
+\end{verbatim}
 
-```{r plot-various-options-ts-3pops, fig.width=4, fig.height=6.3, fig.cap="Example simulations with different parameter values.
-I am showing three populations here to make the plots easier to interpret.\\label{fig:sim-param-ts-3}"}
-source("example_returns_3.R")
-```
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/plot-various-options-ts-3pops.pdf}
+\caption{Example simulations with different parameter values. I am
+showing three populations here to make the plots easier to
+interpret.\label{fig:sim-param-ts-3}}
+\end{figure}
 
-```{r plot-various-options-ts, fig.width=4, fig.height=6.3, fig.cap="Example simulations with different parameter values. This version has ten populations.\\label{fig:sim-param-ts-10}"}
-source("example_returns_10.R")
-```
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/plot-various-options-ts.pdf}
+\caption{Example simulations with different parameter values. This
+version has ten populations.\label{fig:sim-param-ts-10}}
+\end{figure}
 
@@
 
-An example of portfolio optimization
-====================================
-Now, let's run an example where we find the $b$ values that form the efficient frontier of metapopulation portfolios.
+\section{An example of portfolio optimization}
 
-```{r set-a-seed, echo=FALSE}
-set.seed(3)
-```
+Now, let's run an example where we find the $b$ values that form the
+efficient frontier of metapopulation portfolios.
 
-```{r monte-carlo-portfolios}
-weights_matrix <- create_asset_weights(n_pop = 10, n_sims = 100, 
-  weight_lower_limit = 0.01)
-mc_ports <- monte_carlo_portfolios(weights_matrix = weights_matrix,
-  n_sims = 100, mean_b = 1000)
-```
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/eff-frontier-eg.pdf}
+\caption{Efficient frontier of metapopulation portfolios (red dots).
+Each dot represents a different set of weights of the Ricker $b$
+parameters. Metapopulation growth rate is the first difference of the
+log of total metapopulation abundance. The colours in the right panel
+correspond to the other figures.}
+\end{figure}
 
-```{r eff-frontier-eg, fig.width=6.5, fig.height=3.5, fig.cap="Efficient frontier of metapopulation portfolios (red dots).
-Each dot represents a different set of weights of the Ricker $b$ parameters.
-Metapopulation growth rate is the first difference of the log of total metapopulation abundance.
-The colours in the right panel correspond to the other figures."}
-ef_dat <- plot_efficient_portfolios(port_vals = mc_ports$port_vals, pal = 
-  col_pal, weights_matrix = weights_matrix)
-```
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/plot-eff-ports.pdf}
+\caption{Simulation output from the minimum-variance metapopulation.}
+\end{figure}
 
-```{r plot-eff-ports, fig.height=6.5, fig.width=5, fig.cap="Simulation output from the minimum-variance metapopulation."}
-min_var_sim <- mc_ports$sims_out[[ef_dat$min_var_port_id]]
-plot_sim_ts(min_var_sim, pal = col_pal)
-```
+\begin{verbatim}
+Error: could not find function "ricker_ar1"
+\end{verbatim}
 
-```{r fig.width=8, fig.height=5, fig.cap="Example Ricker curves for the minimum-variance metapopulation."}
-plot_rickers(min_var_sim, pal = col_pal)
-```
-
-```{r plot-portfolio-timeseries, fig.width=6.5, fig.height=3.5, fig.cap="Time series for a random sample of portfolios (grey lines) and the minimum-variance portfolio (white lines).
-The top panel shows rate of change or metapopulation abundance at each time step and the bottom panel shows the metapopulation abundance."}
-source("plot_portfolio_ts.R")
-```
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/plot-portfolio-timeseries.pdf}
+\caption{Time series for a random sample of portfolios (grey lines) and
+the minimum-variance portfolio (white lines). The top panel shows rate
+of change or metapopulation abundance at each time step and the bottom
+panel shows the metapopulation abundance.}
+\end{figure}
 
@@
 
-References
-==========
+\renewcommand\refname{References}
+\bibliography{jshort,salmonport}
+
+\end{document}

### 2013-03-12 inst/examples/example_returns_10.R

@@ -1,0 +1,40 @@
+#' 
+
+set.seed(2)
+eg1 <- meta_sim(sigma_v = 0.1)
+set.seed(2)
+eg2 <- meta_sim(env_params = list(amplitude = .000001, ang_frequency = 0.2, 
+    phase = 0, mean_value = 16), sigma_v = 0.1)
+set.seed(2)
+eg3 <- meta_sim(stray_fraction = 0.00001, sigma_v = 0.1)
+set.seed(2)
+eg10 <- meta_sim(env_params = list(amplitude = 5.0001, ang_frequency = 0.2, 
+    phase = 0, mean_value = 16), sigma_v = 0.1)
+set.seed(2)
+eg4 <- meta_sim(sigma_v = 0.000001)
+set.seed(2)
+eg5 <- meta_sim(sigma_v = 0.15)
+set.seed(2)
+eg7 <- meta_sim(max_a = rep(0.9, 10),sigma_v = 0.1)
+set.seed(2)
+eg8 <- meta_sim(max_a = rep(1.8, 10), sigma_v = 0.1)
+set.seed(2)
+eg9 <- meta_sim(stray_fraction = 0.1, sigma_v = 0.1)
+
+source("plot_sim_ts_simple.R")
+par(mfrow = c(9, 1), mar = c(0,0,0,0), oma = c(4, 4, 1, 1), cex = 0.7, las = 1, xpd = FALSE)
+plot_sim_ts_simple(eg1, pal = col_pal, text = "Base case")
+plot_sim_ts_simple(eg2, pal = col_pal, text =  "No temperature fluctuations")
+plot_sim_ts_simple(eg10, pal = col_pal, text =  "Larger temperature fluctuations")
+plot_sim_ts_simple(eg3, pal = col_pal, text = "No straying")
+plot_sim_ts_simple(eg9, pal = col_pal, text = "More straying")
+plot_sim_ts_simple(eg4, pal = col_pal, text = "No spawner-return variance")
+plot_sim_ts_simple(eg5, pal = col_pal, text = "Larger spawner-return variance")
+#plot_sim_ts_simple(eg6, pal = col_pal, text = "No response diversity")
+plot_sim_ts_simple(eg7, pal = col_pal, text = "Lower productivity")
+plot_sim_ts_simple(eg8, pal = col_pal, text = "Higher productivity")
+axis(1, col = "grey50") 
[…]

### 2013-03-12 inst/examples/example_returns_3.R

@@ -1,0 +1,43 @@
+#' 
+
+col_pal <- rev(gg_color_hue(3))
+n_pop_temp <- 3
+set.seed(2)
+eg1 <- meta_sim(sigma_v = 0.1, n_pop = n_pop_temp)
+set.seed(2)
+eg2 <- meta_sim(env_params = list(amplitude = .000000001, ang_frequency = 0.2, 
+phase = 0, mean_value = 16), sigma_v = 0.1, n_pop = n_pop_temp)
+set.seed(2)
+eg3 <- meta_sim(stray_fraction = 0.00001, sigma_v = 0.1, n_pop = n_pop_temp)
+set.seed(2)
+eg10 <- meta_sim(env_params = list(amplitude = 5.0001, ang_frequency = 0.2, 
+phase = 0, mean_value = 16), sigma_v = 0.1, n_pop = n_pop_temp)
+set.seed(2)
+eg4 <- meta_sim(sigma_v = 0.000001, n_pop = n_pop_temp)
+set.seed(2)
+eg5 <- meta_sim(sigma_v = 0.15, n_pop = n_pop_temp)
+set.seed(2)
+eg7 <- meta_sim(max_a = rep(0.9, 10), sigma_v = 0.1, n_pop = n_pop_temp)
+set.seed(2)
+eg8 <- meta_sim(max_a = rep(1.8, 10), sigma_v = 0.1, n_pop = n_pop_temp)
+set.seed(2)
+eg9 <- meta_sim(stray_fraction = 0.1, sigma_v = 0.1, n_pop = n_pop_temp)
+
+source("plot_sim_ts_simple.R")
+par(mfrow = c(9, 1), mar = c(0,0,0,0), oma = c(4, 4, 1, 1), cex = 0.7, las = 1, xpd = FALSE)
+plot_sim_ts_simple(eg1, pal = col_pal, text = "Base case")
+plot_sim_ts_simple(eg2, pal = col_pal, text =  "No temperature fluctuations")
+plot_sim_ts_simple(eg10, pal = col_pal, text =  "Larger temperature fluctuations")
+plot_sim_ts_simple(eg3, pal = col_pal, text = "No straying")
+plot_sim_ts_simple(eg9, pal = col_pal, text = "More straying")
+plot_sim_ts_simple(eg4, pal = col_pal, text = "No spawner-return variance")
+plot_sim_ts_simple(eg5, pal = col_pal, text = "Larger spawner-return variance")
[…]

### 2013-03-12 inst/examples/generate_env_ts.R

@@ -1,0 +1,56 @@
+#' Create an environmental time series.
+#'
+#' @param n_t Length of time series.
+#' @param type Type of time series to produce.
+#' @param sine_params Parameters controlling sine wave time series.
+#' @param arma_params Parameters controlling ARMA time series.
+#' @param regime_params Parameters controlling regime-shift time series.
+#' @param linear_params Parameters controlling warming or cooling time series.
+#' @param constant_params Parameter controlling constant time series.
+#' @export
+#' @examples
+#' types <- c("sine", "arma", "regime", "linear", "constant")
+#' x <- list()
+#' for(i in 1:5) x[[i]] <- generate_env_ts(n_t = 100, type = types[i])
+#' op <- par(mfrow = c(5, 1), mar = c(3,3,1,0), cex = 0.7) 
+#' for(i in 1:5) plot(x[[i]], type = "o", main = types[i])
+#' par(op)
+
+generate_env_ts <- function(
+  n_t,
+  type = c("sine", "arma", "regime", "linear", "constant"),
+  sine_params = list(amplitude = 1, ang_frequency = 0.2, phase = 0, mean_value = 0),
+  arma_params = list(sigma_env = 0.30, ar = 0.6, ma = 0),
+  regime_params = list(break_pts = c(25, 75), break_vals = c(-1, 0, 1)),
+  linear_params = list(min_value = -1, max_value = 1),
+  constant_params = list(value = 0)
+  ) {
+  type <- type[1]
+  env_ts <- switch(type, 
+    arma = as.numeric(with(arma_params, arima.sim(model = list(ar = ar, ma = ma),
+        n = n_t, sd = sigma_env))), 
+    sine = {
+      x <- seq_len(n_t)
+      y <- with(sine_params, amplitude * sin(ang_frequency * x + phase) + mean_value) 
+      y
+    },
+    regime = {
+      regime_params$break_pts <- c(1, regime_params$break_pts, n_t)
+      y <- vector(length = n_t, mode = "numeric")
+      for(i in 2:length(regime_params$break_pts))
+      {
+        y[regime_params$break_pts[i-1]:regime_params$break_pts[i]] <-
+          regime_params$break_vals[i-1]
+      }
+      y
+    },
+    linear = {
+      with(linear_params, seq(min_value, max_value, length.out = n_t))
+    },
+    constant = {
+      rep(constant_params$value, n_t)
+    }
+    )
+  return(env_ts)
+}
+

### 2013-03-12 inst/examples/generate_straying_matrix.R

@@ -1,0 +1,30 @@
+# generate a matrix of straying proportions
+
+generate_straying_matrix <- function(n_pop, stray_fraction, stray_decay_rate){
+  
+  stray_mat <- matrix(ncol = n_pop, nrow = n_pop)
+  for(i in 1:n_pop) {
+    for(j in 1:n_pop) {
+      # Cooper and Mangel 1999 eq (2):
+      stray_mat[i,j] <- stray_fraction * 
+      exp(-stray_decay_rate * abs(i - j))
+    }
+  }
+  #browser()
+
+  diag(stray_mat) <- 0 # no self-straying
+  normalization <- rowSums(stray_mat) # so no salmon or lost or gained
+#   stray_mat_star <- matrix(ncol = ncol(stray_mat), nrow = nrow(stray_mat))
+#   for(i in 1:nrow(stray_mat)) {
+#     stray_mat_star[i, ] <- stray_mat[i, ] / normalization[i]
+#   }
+
+  #image(stray_mat_star)
+  #image((stray_mat / normalization) * stray_fraction)
+
+  stray_mat <- (stray_mat / normalization) * stray_fraction # to match the desired overall straying rate by stream
+  return(stray_mat)
+}
+
+#generate_straying_matrix(4, 0.01, 0.3)
+#image(stray_mat)

### 2013-03-12 inst/examples/manipulate_test.R

@@ -1,0 +1,5 @@
+#man_sim_run <- function(sigma_v2){
+#  eg <- meta_sim(b = rep(1000, 10), n_pop = 10, sigma_v = sigma_v2)
+#  plot_sim_ts(eg, pal = col_pal, years_to_show = 40)
+#}
+#manipulate(man_sim_run(sigma_v2 = 0.0011), sigma_v2 = slider(0.001, 0.5))

### 2013-03-12 inst/examples/plot_env_ts.R

@@ -1,0 +1,17 @@
+# plot example temperature time series
+
+types <- c("sine", "arma", "regime", "linear", "constant")
+x <- list()
+for(i in 1:5) x[[i]] <- generate_env_ts(n_t = 100, type = types[i], arma_params = list(sigma_env = 0.30, ar = 0.6, ma = 0.2))
+op <- par(mfrow = c(5, 1), mar = c(0,0,0,0), cex = 0.8, oma = c(3.1, 3.2, .5, 5), xpd = NA, mgp = c(2.5, 0.7, 0), tck = -0.06, las = 1) 
+for(i in 1:5) {
+  plot(x[[i]], type = "l",  axes = FALSE, ylim = c(-1.2, 1.12), xlab = "", ylab = "")
+  mtext(types[i], adj = 0.05, cex = 0.8, line = -1.5)
+  axis(2, at = c(-1, 0, 1), col = "grey50")
+  box(col = "grey50")
+}
+axis(1, col = "grey50")
+mtext("Generation", side = 1, line = 2.2, outer = TRUE)
+mtext("Anomaly", side = 2, line = 2.2, outer = TRUE, las = 0)
+par(op)
+

### 2013-03-12 inst/examples/plot_portfolio_ts.R

@@ -1,0 +1,44 @@
+
+par(mfrow = c(2, 1), cex = 0.8, mar = c(0, 4, 0, 1), oma = c(4, 0,1,0))
+burn <- 1:30
+port.x <- rowSums(min_var_sim$A[-burn, ])
+ret.x <- diff(log(port.x))
+plot(ret.x, type = "n", xaxt = "n", ylim = c(min(ret.x)*1.30, max(ret.x)*1.3))
+
+#for(i in ef_dat$ef_port_ids){
+  #this.port <- mc_ports$sims_out[[i]]
+  #port.x <- rowSums(this.port$A[-burn, ])
+  #ret.x <- diff(log(port.x))
+  #lines(1:length(ret.x), ret.x, type = "l", col = "#44444440")
+#}
+
+
+# all!
+for(i in sample(1:length(mc_ports$sims_out), 80)){
+  this.port <- mc_ports$sims_out[[i]]
+  port.x <- rowSums(this.port$A[-burn, ])
+  ret.x <- diff(log(port.x))
+  lines(1:length(ret.x), ret.x, type = "l", col = "#44444440")
+}
+port.x <- rowSums(min_var_sim$A[-burn, ])
+ret.x <- diff(log(port.x))
+#lines(1:length(ret.x), ret.x, lwd = 1.7, col = "#4F69FF")
+lines(1:length(ret.x), ret.x, lwd = 2.5, col = "white")
+
+# abundance?
+port.x <- rowSums(min_var_sim$A[-burn, ])
+plot(port.x, type = "n")
+for(i in sample(1:length(mc_ports$sims_out), 70)){
+  this.port <- mc_ports$sims_out[[i]]
+  port.x <- this.port$A[-burn, ]
+  port.x <- rowSums(this.port$A[-burn, ])
+  lines(1:length(port.x), port.x, type = "l", col = "#44444440")
+}
+
+port.x <- rowSums(min_var_sim$A[-burn, ])
+#ret.x <- diff(log(port.x))
+#lines(1:length(port.x), port.x, lwd = 1.7, col = "#4F69FF")
+lines(1:length(port.x), port.x, lwd = 2.5, col = "white")
+
+# pull out EF and plot the portfolios
+

### 2013-03-12 inst/examples/plot_sim_ts_simple.R

@@ -1,0 +1,26 @@
+# plot time series from the simulation
+# just the annual rate of growth and returns in this one
+
+plot_sim_ts_simple <- function(x, pal, burn = 1:30, text = "Returns") {
+# x is output from a simulation run; it's a list
+# pal is the colour palette
+
+to_show <- (max(burn)):(max(burn)+30) # years to show in time series example plots
+
+annotate <- function(text) {
+  mtext(text, side = 3, line = -1.5, cex = 0.7, adj = 0.05)
+}
+my.axis <- function(side) axis(side, col = "grey50", at = c(800, 1200))
+
+#par(mfrow = c(8, 1), mar = c(0,0,0,0), oma = c(4, 4, 1, 1), cex = 0.7, las = 1, xpd = FALSE)
+# returns:
+
+matplot(x$A[to_show, ], type = "l", col = pal, lty = 1, ylab = "Returns", xlab = "Time", xaxt = "n", axes = FALSE, xaxs = "i", ylim = c(580, 1450))
+annotate(text)
+box(col = "grey50")
+my.axis(2)
+
+#my.axis(1)
+
+}
+

### 2013-03-12 inst/examples/plot_straying_matrix.R

@@ -1,0 +1,34 @@
+# plot an example straying matrix
+
+library(colorspace)
+M <- generate_straying_matrix(10, 0.01, 0.3)
+library(reshape2)
+M <- melt(M)
+shades <- rev(sequential_hcl(20))
+cuts <- seq(min(M$value), max(M$value), length.out = 20)
+M$colour <- shades[findInterval(M$value, cuts)]
+
+#image(generate_straying_matrix(10, 0.01, 0.3), col = rev(sequential_hcl(11)))
+
+plot_straying_matrix <- function(x, pal) {
+
+par(mfrow = c(10, 10), cex = 0.7, mar = c(0,0,0,0), oma = c(4, 4, .5, .5))
+for(i in 1:10) {
+  for(j in 1:10) {
+    #if(j < i){ 
+      plot(1, 1, xlim = c(0, 1), ylim = c(0, 1), yaxs = "i", xaxs = "i", axes = FALSE, pch = 20, col = "grey20", type = "n")
+    with(x[x$Var1 == i & x$Var2 == j, ], rect(0, 0, 1, 1, col = colour, border = NA))
+    box(col = "grey50")
+    #}else{ 
+      #plot(1,1, type = "n", xlab = "", ylab = "", axes = FALSE)
+    #}
+    if(i == 10) mtext(j, side = 1, line = 1, cex = 0.8, col = pal[j])
+    if(j == 1) mtext(i, side = 2, las = 1, line = 1, cex = 0.8, col = pal[i])
+  }
+}
+    mtext("Receiving population", side = 1, line = 2.5, cex = 0.8, outer = TRUE)
+    mtext("Source population", side = 2, line = 2.5, cex = 0.8, outer = TRUE, las = 0)
+}
+
+plot_straying_matrix(M, pal = col_pal)
+

### 2013-03-12 inst/examples/plot_thermal_curves.R

@@ -1,0 +1,11 @@
+# plot thermal curves across populations as an example
+
+par(cex = 0.8, mgp = c(2.4, 0.65, 0), tck = -0.02)
+optim_temps <- seq(13, 19, length.out = 10)
+x <- seq(3, 29, length.out = 200)
+plot(1, 1, xlim = c(4, 28), ylim = c(-0.01, 1.4), ylab = "Ricker productivity parameter (a)", xlab = expression(Temperature~(degree*C)), type = "n", yaxs = "i", las = 1)
+for(i in 1:10) {
+  lines(x, thermal_curve_a(x, optim_temp = optim_temps[i], max_a = 1.3, width_param = 0.02), ylab = "a", xlab = "Temperature", type = "l", col = col_pal[i], lwd = 1.5)
+}
+
+

### 2013-03-12 inst/examples/run_multiple_mpt.R

@@ -1,0 +1,90 @@
+#' Run multiple optimizations
+
+# source files:
+rm(list = ls())
+setwd("~/Dropbox/ms/salmon_portfolios/")
+metafolio_files <-  list.files(path = "metafolio/R/")
+for(i in 1:length(metafolio_files)){
+  source(paste("metafolio/R/", metafolio_files[i], sep = ""));rm(i)}
+col_pal <- rev(gg_color_hue(10))
+
+# run it:
+get_min_var_weights <- function(n_sims = 500, n_pop = 10){
+  #browser()
+
+  weights_matrix <-  create_asset_weights(n_pop = n_pop, n_sims = n_sims, 
+    weight_lower_limit = 0.01)
+
+  mc_ports <- monte_carlo_portfolios(weights_matrix = weights_matrix,
+    n_sims = n_sims, mean_b = 1000)
+
+  ef_dat <- plot_efficient_portfolios(port_vals = mc_ports$port_vals, 
+    weights_matrix = weights_matrix, pal = col_pal, plot = FALSE)
+
+  min_var_sim <- mc_ports$sims_out[[ef_dat$min_var_port_id]]
+  weights <- min_var_sim$b
+
+  save(mc_ports, file = paste("mc_ports_", iter, ".rda", sep = ""))
+
+  # run an equal weighted portfolio:
+  b_vals_matrix <- rep(1000, n_pop)
+  port_vals = matrix(ncol = 2, nrow = 1)
+  b_vals <- b_vals_matrix
+  eq_sim <- meta_sim(b = b_vals, use_cache = TRUE, n_pop = n_pop)
+
+  save(eq_sim, file = paste("eq_port_", iter, ".rda", sep = ""))
+  burn <- 1:30
+  port.x <- rowSums(eq_sim$A[-burn, ])
+  ret.x <- diff(log(port.x))
+  var.x <- var(ret.x)
+  mean.x <- mean(ret.x)
+  eq_weight_cv <- var.x
+
+  port.minv <- rowSums(min_var_sim$A[-burn, ])
+  ret.minv <- diff(log(port.minv))
+  var.minv <- var(ret.minv)
+  mean.minv <- mean(ret.minv)
+  minv_cv <- var.minv
+
+
+## find out variance (and mean?) of equal-weighted portfolios and min-var ports
+## and record those values
+## how much better?
+
+  iter <<- iter + 1
[…]

### 2013-03-12 inst/examples/sub_grid.sh

@@ -1,0 +1,7 @@
+#!/bin/sh
+#PBS -l walltime=00:20:00
+#PBS -M sean_anderson@sfu.ca
+#PBS -l mem=300mb
+
+R CMD BATCH --no-save /home/anderson/Dropbox/ms/salmon_portfolios/run_multiple_mpt.R
+

### 2013-03-19 R/plot_sim_ts.R
Add estimated b and a to plot

@@ -5,7 +5,7 @@
 #' @param burn The number of years to discard as burnin a the beginning of the time 
 #' series.
 
-plot_sim_ts <- function(x, pal = rev(gg_color_hue(x$n_pop)), years_to_show = 30, burn = 1:30) {
+plot_sim_ts <- function(x, pal = rev(gg_color_hue(x$n_pop)), years_to_show = 30, burn = 1:50) {
 # x is output from a simulation run; it's a list
 # pal is the colour palette
 
@@ -18,7 +18,7 @@ annotate <- function(text) {
 }
 my.axis <- function(side) axis(side, col = "grey50")
 
-par(mfrow = c(8, 1), mar = c(0,0,0,0), oma = c(4, 4, 1, 1), cex = 0.7, las = 1, xpd = FALSE)
+par(mfrow = c(10, 1), mar = c(0,0,0,0), oma = c(4, 4, 1, 1), cex = 0.7, las = 1, xpd = FALSE)
 # returns:
 matplot(x$A[to_show, ], type = "l", col = pal, lty = 1, ylab = "Returns", xlab = "Time", xaxt = "n", axes = FALSE, xaxs = "i")
 annotate("Returns")
@@ -65,8 +65,19 @@ abline(h = 0, lty = 2, lwd = 1.5, col = "grey50")
 my.axis(2)
 annotate("Spawner-return curve residuals")
 
+matplot(x$Est_a[to_show, ], type = "l", col = pal, lty = 1, ylab = "Estimated a", xlab = "", axes = FALSE, xaxs = "i")
+box(col = "grey50")
+annotate("Estimated a")
+my.axis(2)
+
+matplot(x$Est_b[to_show, ], type = "l", col = pal, lty = 1, ylab = "Estimated b", xlab = "", axes = FALSE, xaxs = "i")
+box(col = "grey50")
+annotate("Estimated b")
+my.axis(2)
+
 my.axis(1)
 mtext("Generation", side = 1, outer = TRUE, line = 2.0, cex = 0.8)
+#browser()
 
 }
 

### 2013-03-19 inst/examples/salmonport.Rmd
Updating files before major work after Jon, Andy, Trevor meetings

@@
 
-# Abstract
+# Abstract #
 First part is copied from what I wrote before (introduction and questions).
@@
 
-# Introduction
-[Based partly on Jon's notes.] Managing risk is fundamental to the conservation of an endangered species.
+# Introduction #
+Managing risk is fundamental to the conservation of an endangered species.
 When an endangered species exists as a metapopulation, we can manage risk at two levels: at the population level or at the metapopulation level.
-Typically we treat sources of risk at the metapopulation level as external and uncontrollable and so we manage risk by altering anthropogenic disturbances such as fishing or hunting on a population level as well as improving connectivity of populations.
+Typically we treat sources of risk at the metapopulation level as external and uncontrollable and so we manage risk by altering fishing or hunting on a population level as well as improving the connectivity of populations.
 
 The management of financial portfolios provides another way of considering risk.
-Economists consider the risk and performance of a financial portfolio as depending on the weighting of individual assets that make up the portfolio.
-Modern Portfolio Theory (MPT) proposes that there is a set of portfolios that achieves maximum expected return for a level of expected risk or minimum expected risk for a level of expected return [@Markowitz1952; @Markowitz1959].
+Economists consider the risk and performance of a financial portfolio as a function of the weighting of individual assets that make up the portfolio.
+Modern Portfolio Theory (MPT) proposes that there is a set of portfolios that maximizes expected return for a level of expected risk or minimizes expected risk for a level of expected return [@Markowitz1952; @Markowitz1959].
 This optimal set contains portfolios that range along a continuum of risk-tolerance; economists refer to this set as the efficient frontier.
 
-Similarly, expected growth rate and variance at a metapopulation level will be a function of the properties of the populations.
-A portfolio approach to managing risk for a metapopulation therefore considers how conservation might affect the ``weight'' of each population in the metapopulation ``portfolio''.
+Similarly, expected growth rate and variance at a metapopulation level is a function of the variance, covariance, and size of the individual populations.
+A portfolio approach to managing risk for a metapopulation therefore considers how conservation might affect the "weight" of each population in a metapopulation "portfolio".
 
@@
 
-## Portfolio prioritization of salmon conservation
-Here, we conduct a post-hoc assessment of the set of optimal salmon portfolios given various return-risk definitions and objectives.
+## Portfolio prioritization of salmon conservation ##
+The basic goal is to conduct a post-hoc assessment of the set of optimal salmon portfolios given various return-risk definitions and objectives.
 We ask: What does a portfolio approach to management tell us about salmonid conservation priorities?
-Does it produce any rules of thumb about what optimal portfolio management of salmonid metapopulations would entail?
+Can we produce any rules of thumb about what optimal portfolio management of salmonid metapopulations would entail?
 
-## Questions
-1. What are the attributes of the populations that form the optimal set? For example, are populations with high response diversity prioritized?
+## Questions ##
+1. What are the attributes of the populations that form the optimal set? For example, does the optimal set tend to focus on populations with as diverse responses to the environment as possible?
 
@@
 How do our answers to other questions change depending on what we measure on the risk-return axes?
-For example, we could trade off the mean rate of change of salmon returns vs.\ variance of the rate of change.
+For example, we could trade off the mean rate of change of salmon returns vs.
+variance of the rate of change.
 This would be most similar to a financial portfolio optimization.
@@
 4.
-How do constraints (e.g.\ viable salmonid population (VSP)-like constraints) affect the selected portfolios in risk-return space and in terms of common metrics such as population-level and ESU-level extinction risk?
+How do constraints (e.g.
+viable salmonid population (VSP)-like constraints) affect the selected portfolios in risk-return space and in terms of common metrics such as population-level and ESU-level extinction risk?
 
@@
 
-Simulation
-==========
+# Simulation #
 See Figure \ref{fig:sim-diagram} for an illustration of the simulation 
@@
   \centering
-    \includegraphics[height=7in]{simulation_diagram.pdf}
+    \includegraphics[height=6.5in]{simulation_diagram.pdf}
  \caption{Flow chart of the salmon-metapopulation simulation and portfolio optimization.
@@
 
-## Spawner-return relationship
+## Spawner-return relationship ##
 I am using a Ricker curve,
@@
 Formally, $w_{ti} = w_{ti-1} \rho + v_{ti}$, where $v_{ti}$ represents independent and normally distributed error with mean 0 and standard deviation of $\sigma_v$.
-The parameter $\rho$ represents the correlation between residuals from subsequent years.
+The parameter $\rho$ represents the correlation between residuals from subsequent generations.
 
-## Escapement targets
+## Escapement targets ##
 I am calculating spawning abundance MSY $S_{MSY}$ each generation using the equation $S_{MSY} = b(0.5-0.07a)$ from @Hilborn1992 p272, Table 7.2.
@@
 
-## Environmental signal
+## Environmental signal ##
 I have programmed five different kinds of environmental signals.
-I think it's easiest to think of these as temperature signals, but they could represent more general environmental anomalies, such as the PDO.
-The five kinds of environmental signals are described in Table \ref{tab:env-types} and illustrated in Figure \ref{fig:env-ts}.
+I think it's easiest to think of these as temperature, but they could represent more general environmental anomalies, such as the PDO.
+In Table \ref{tab:env-types}, I describe the five kinds of environmental signals I have included.
+Examples of these time series are illustrated in Figure \ref{fig:env-ts}.
 Later, I may include black-swan-like events, although the regime shifts could be seen this way if the changes are dramatic enough.
-So far all figures use a sine wave.
-The problem with using a randomly generated ARMA time series is that you get much greater differences in the portfolios from
-run to run.
+So far, all simulation results in the rest of this document use a sine wave.
+The problem with using a randomly generated ARMA time series is that you get much greater differences in the results from
+run to run. This is fine in the long run, but hard to detect patterns initially.
 
@@
 average (MA1) parameters and standard deviation; mean is assumed to be 0\\
-Regime shifts & $T_{b0},T_{b1}, \ldots, T_{bn}$; $T_{v0},T_{v1}, \ldots, T_{vn}$ 
-& Breakpoints ($T_{b}$) and values ($T_v$) for regimes $1$ through $n$\\
+Regime shifts & $T_{b0},T_{b1}, \ldots, T_{br-1}$; $T_{v0},T_{v1}, \ldots, T_{vr}$ 
+& Breakpoints ($T_{b}$) and values ($T_v$) for regimes $1$ through $r$\\
 Linear change & $T_{\mathrm{min}}$, $T_{\mathrm{max}}$ & Minimum and maximum 
@@
 
-```{r env-ts, fig.height=5, fig.width=4, fig.cap="Example temperature time series\\label{fig:env-ts}"}
+```{r env-ts, fig.height=4, fig.width=4, cache = FALSE, fig.cap="Example temperature time series\\label{fig:env-ts}"}
+# library(metafolio)
+# or for rapid development:
+metafolio_files <-  list.files(path = "../../R/")
+for(i in 1:length(metafolio_files)){
+  source(paste("../..//R/", metafolio_files[i], sep = ""));rm(i)}
+
 source("plot_env_ts.R")
@@
 
-## Straying
+## Straying ##
 [Insert a bit here about my research on straying and it's influence on metapopulation dynamics for salmon].
@@
 I generate a matrix that represents the fraction of straying between any two populations.
-I have set up the metapopulation (thus far) as in in a very simple scenario: the populations are arranged in a line and those that are nearer to each other are more likely to stray between each other [insert caveats].
-Two parameters control the straying: the fraction of fish that stray from their natal stream in any given generation $f$ and the rate at which this straying between streams decays with distance $m$.
-I calculated the number of salmon straying from stream $i$ to stream $j$ as,
+I have set up the metapopulation (thus far) in a very simple scenario: the populations are arranged in a line and those that are nearer to each other are more likely to stray between each other [insert caveats].
+Two parameters control the straying: the fraction of fish $f$ that stray from their natal stream in any given generation and the rate at which this straying between streams decays with distance $m$.
+I calculated the number of salmon straying from stream $j$ to stream $i$ as,
 
 \begin{equation}
-  \mathrm{strays}_{tij} = \frac{S_{ti} f e^{-m \lvert i-j \rvert }}{\sum\limits_{k \neq j}^{} e^{-m \lvert k-j \rvert }},
+  \mathrm{strays}_{tij} = f R_{tj} \frac{e^{-m \lvert i-j \rvert }}{\displaystyle\sum\limits_{\substack{k = 1 \\
+  k \neq j}}^{n} e^{-m \lvert k-j \rvert }},
 \end{equation}
 
-where the denominator is a normalization constant to ensure that no salmon are gained or lost.
-The subscript $k$ represents a stream ID.
-[Not sure this is written correctly --- I'll think about it --- but it should be coded correctly.] There is no straying from and to the same stream.
+\noindent
+where $R_{tj}$ is the number of returning salmon at generation $t$ whose natal stream was stream $j$.
+The subscript $k$ represents a stream ID and $n$ the number of populations.
+The denominator is a normalizing constant to ensure the desired fraction of fish stray.
 See Figure \ref{fig:stray-matrix} for an example straying matrix.
@@
 
-## Diversity of response to the environmental conditions
+## Diversity of response to the environmental conditions ##
 [insert research I've done here on what can generate diversity of response to the environment across populations and how it is typically incorporated] This is a critical part to the paper.
@@
 I calculate $a_{ti}$, the realized Ricker productivity parameters after accounting for thermal tolerance, as
-
 \begin{equation}
@@
 \end{equation}
-
+\noindent
 where $a_i^*$ is the maximum possible productivity parameter for population $i$, $T_t$ is the temperature at time step $t$, $T_i^*$ is the optimal temperature for population $i$, and $\gamma_i$ controls the width of the thermal tolerance curve for each population.
@@
 
-```{r thermal-curves, cache=FALSE, fig.height=3.9, fig.width=5.4, fig.cap="Thermal tolerance curves for ten populations.
+```{r thermal-curves, fig.height=3.7, cache = FALSE, fig.width=5.0, fig.cap="Thermal tolerance curves for ten populations.
 Here, and throughout this document, I have used the same rainbow of colours to identify the salmon populations.
-The populations are arranged from blue (more productive with cooler temperatures) to red (more productive with warmer temperatures\\label{fig:thermal-curves}"}
+The populations are arranged from blue (more productive with cooler temperatures) to red (more productive with warmer temperatures).\\label{fig:thermal-curves}"}
 source("plot_thermal_curves.R")
@@
 
-Portfolio optimization
-======================
+# Portfolio optimization #
 Details to follow.
 Quick version: I Monte Carlo the $b$ parameters (with a specified minimum value) and measure the resulting mean and variance of the rate of growth of the metapopulation salmon returns.
-I then identify the set of $b$ parameters (i.e.\ the set of weights) that results in the efficient frontier.
+I then identify the set of $b$ parameters (i.e.
+the set of weights) that results in the efficient frontier.
 We could later choose to work with risk metrics like conditional value at risk (CVaR), which measure the probability of a ``bad'' event specifically, not the symmetrical variance.
@@
 
-Comparing portfolios
-====================
+# Comparing portfolios #
 This is a big area to think about.
@@
 
-Some sample simulations
-=======================
+# Some sample simulations #
 The following are options and their default values:
 
-\begin{verbatim}
-  n_t = 120, # number of years
+    n_t = 120, # number of generations
     n_pop = 10, # number of populations
@@
     add_straying = TRUE # include or ignore straying
-\end{verbatim}
 
@@
 plot_sim_ts(eg1, years_to_show = 40)
+
+o <- meta_sim(env_params = list(amplitude = .000000001, ang_frequency = 0.2, phase = 0, mean_value = 16), sigma_v = 0.01, n_pop = 10, stray_fraction = 0.00001, use_cache = FALSE)
+o <- meta_sim(env_params = list(amplitude = .000000001, ang_frequency = 0.2, phase = 0, mean_value = 16), sigma_v = 0.1, n_pop = 10, stray_fraction = 0.00001, use_cache = FALSE)
+#o<- meta_sim(b = rep(1000, 10), n_pop = 10)
+plot_sim_ts(o, years_to_show = 60)
+image(o$Est_b)
 ```
@@
 Note how populations 1 and 10 have asynchronous returns whereas populations with more similar thermal-tolerance curves (say populations 9 and 10) have more synchronous dynamics.
-Populations with thermal tolerance curves in the middle (e.g.\ population 6) are less correlated with other populations.
+Populations with thermal tolerance curves in the middle (e.g.
+population 6) are less correlated with other populations.
 Their population dynamics end up primarily driven by demographic stochasticity and less so by temperature-induced systematic changes in productivity."}
@@
 
-```{r plot-base-rickers, fig.width=8, fig.height=5, cache=FALSE, fig.cap="We can also look at samples of the Ricker curves by population (population numbers are in the upper left corners).
+```{r plot-base-rickers, fig.width=6, fig.height=3.75, cache=FALSE, fig.cap="We can also look at samples of the Ricker curves by population (population numbers are in the upper left corners).
 Here, I've sampled 40 of the $a$ parameters through time for each population.
@@
 In this example all the populations have the same $b$ parameters.
-Notice how the productivity is more variable for populations nearer the limits of their thermal tolerance (e.g.\ population 1 vs.\ population 5).
+Notice how the productivity is more variable for populations nearer the limits of their thermal tolerance (e.g.
+population 1 vs.
+population 5).
 Also notice how population 1 is more productive in cool years whereas population 10 is more productive in warm years."}
@@
 
-An example of portfolio optimization
-====================================
+# An example of portfolio optimization #
 Now, let's run an example where we find the $b$ values that form the efficient frontier of metapopulation portfolios.
@@
 ```{r set-a-seed, echo=FALSE}
-set.seed(3)
+set.seed(1)
 ```
@@
 weights_matrix <- create_asset_weights(n_pop = 10, n_sims = 100, 
-  weight_lower_limit = 0.01)
+  weight_lower_limit = 0.001)
 mc_ports <- monte_carlo_portfolios(weights_matrix = weights_matrix,
@@
 
-```{r fig.width=8, fig.height=5, fig.cap="Example Ricker curves for the minimum-variance metapopulation."}
+```{r fig.width=6, fig.height=3.75, fig.cap="Example Ricker curves for the minimum-variance metapopulation."}
 plot_rickers(min_var_sim, pal = col_pal)
@@
 
-```{r plot-portfolio-timeseries, fig.width=6.5, fig.height=3.5, fig.cap="Time series for a random sample of portfolios (grey lines) and the minimum-variance portfolio (white lines).
+```{r plot-portfolio-timeseries, fig.width=6, fig.height=3.5, fig.cap="Time series for a random sample of portfolios (grey lines) and the minimum-variance portfolio (white lines).
 The top panel shows rate of change or metapopulation abundance at each time step and the bottom panel shows the metapopulation abundance."}
@@
 
+```{r plot-weights-samples, fig.width=6, fig.height=6, fig.cap="A sample of 30 minimum-variance portfolios given the same set of starting parameters as for the example illustrated above.
+The dashed grey line indicates the weights if all populations had an equal carrying capacity.
+As it stands, the weights look fairly random to me."}
+source("look_at_cluster_dat.R")
+```
+
+```{r plot-weights-median, fig.width=4.2, fig.height=3.2, cache = FALSE, fig.cap="Median (solid line) and 25th and 75th quantiles of the weights across 40 minimum-variance portfolios.
+Just another way of looking at it."}
+source("min_var_ports_dist.R")
+```
+
 \clearpage
 
-References
-==========
+# Notes #
+- b is "unfished equilibrium stock size"
+
+```{r}
+```
+
+# References #

### 2013-03-19 inst/examples/salmonport.tex
Updating files before major work after Jon, Andy, Trevor meetings

@@
 \usepackage{booktabs}
+\usepackage{mathptmx}
 \defaultfontfeatures{Ligatures=TeX}
@@
 \setmonofont{Inconsolata}
-\setmainfont{Linux Libertine}
+%\setmainfont{Linux Libertine}
+%\setmainfont{Times New Roman}
+\setmainfont{Calibri}
+%\setmainfont{Helvetica}
 \DeclareTextCommandDefault{\nobreakspace}{\leavevmode\nobreak\ }
@@
 \onehalfspacing
+%\doublespacing
 
-% A bit bigger than one inch margins:
-%\topmargin -1.43cm
-%\oddsidemargin 0.4cm
-%\evensidemargin 0.4cm
-%\textwidth 14.5cm
-\textheight 22.5cm
+\usepackage{titlesec}
+\titleformat*{\section}{\Large\bfseries\sffamily}
+\titleformat*{\subsection}{\large\sc\sffamily}
+\titleformat*{\subsubsection}{\itshape\sffamily}
 
-%\usepackage[left=1.5in,top=1.5in,right=1.5in,bottom=1.5in,nohead,nofoot]{geometry}
 \usepackage{geometry}
@@
 %\usepackage{lmodern}
+
+% end customization except for commenting out lines
 \usepackage{amssymb,amsmath}
@@
 %\urlstyle{same}  % don't use monospace font for urls
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
+%\setlength{\parindent}{0pt}
+%\setlength{\parskip}{6pt plus 2pt minus 1pt}
 \setlength{\emergencystretch}{3em}  % prevent overfull lines
@@
 
-{[}Based partly on Jon's notes.{]} Managing risk is fundamental to the
-conservation of an endangered species. When an endangered species exists
-as a metapopulation, we can manage risk at two levels: at the population
-level or at the metapopulation level. Typically we treat sources of risk
-at the metapopulation level as external and uncontrollable and so we
-manage risk by altering anthropogenic disturbances such as fishing or
-hunting on a population level as well as improving connectivity of
-populations.
+Managing risk is fundamental to the conservation of an endangered
+species. When an endangered species exists as a metapopulation, we can
+manage risk at two levels: at the population level or at the
+metapopulation level. Typically we treat sources of risk at the
+metapopulation level as external and uncontrollable and so we manage
+risk by altering fishing or hunting on a population level as well as
+improving the connectivity of populations.
 
@@
 considering risk. Economists consider the risk and performance of a
-financial portfolio as depending on the weighting of individual assets
+financial portfolio as a function of the weighting of individual assets
 that make up the portfolio. Modern Portfolio Theory (MPT) proposes that
-there is a set of portfolios that achieves maximum expected return for a
-level of expected risk or minimum expected risk for a level of expected
+there is a set of portfolios that maximizes expected return for a level
+of expected risk or minimizes expected risk for a level of expected
 return \citep{Markowitz1952, Markowitz1959}. This optimal set contains
@@
 Similarly, expected growth rate and variance at a metapopulation level
-will be a function of the properties of the populations. A portfolio
-approach to managing risk for a metapopulation therefore considers how
-conservation might affect the
-\texttt{weight'' of each population in the metapopulation}portfolio''.
+is a function of the variance, covariance, and size of the individual
+populations. A portfolio approach to managing risk for a metapopulation
+therefore considers how conservation might affect the ``weight'' of each
+population in a metapopulation ``portfolio''.
 
@@
 
-Here, we conduct a post-hoc assessment of the set of optimal salmon
-portfolios given various return-risk definitions and objectives. We ask:
-What does a portfolio approach to management tell us about salmonid
-conservation priorities? Does it produce any rules of thumb about what
-optimal portfolio management of salmonid metapopulations would entail?
+The basic goal is to conduct a post-hoc assessment of the set of optimal
+salmon portfolios given various return-risk definitions and objectives.
+We ask: What does a portfolio approach to management tell us about
+salmonid conservation priorities? Can we produce any rules of thumb
+about what optimal portfolio management of salmonid metapopulations
+would entail?
 
@@
   What are the attributes of the populations that form the optimal set?
-  For example, are populations with high response diversity prioritized?
+  For example, does the optimal set tend to focus on populations with as
+  diverse responses to the environment as possible?
 \item
@@
 with mean 0 and standard deviation of $\sigma_v$. The parameter $\rho$
-represents the correlation between residuals from subsequent years.
+represents the correlation between residuals from subsequent
+generations.
 
@@
 I have programmed five different kinds of environmental signals. I think
-it's easiest to think of these as temperature signals, but they could
-represent more general environmental anomalies, such as the PDO. The
-five kinds of environmental signals are described in Table
-\ref{tab:env-types} and illustrated in Figure \ref{fig:env-ts}. Later, I
-may include black-swan-like events, although the regime shifts could be
-seen this way if the changes are dramatic enough. So far all figures use
+it's easiest to think of these as temperature, but they could represent
+more general environmental anomalies, such as the PDO. In Table
+\ref{tab:env-types}, I describe the five kinds of environmental signals
+I have included. Examples of these time series are illustrated in Figure
+\ref{fig:env-ts}. Later, I may include black-swan-like events, although
+the regime shifts could be seen this way if the changes are dramatic
+enough. So far, all simulation results in the rest of this document use
 a sine wave. The problem with using a randomly generated ARMA time
-series is that you get much greater differences in the portfolios from
-run to run.
+series is that you get much greater differences in the results from run
+to run. This is fine in the long run, but hard to detect patterns
+initially.
 
@@
 that represents the fraction of straying between any two populations. I
-have set up the metapopulation (thus far) as in in a very simple
-scenario: the populations are arranged in a line and those that are
-nearer to each other are more likely to stray between each other
-{[}insert caveats{]}. Two parameters control the straying: the fraction
-of fish that stray from their natal stream in any given generation $f$
-and the rate at which this straying between streams decays with distance
-$m$. I calculated the number of salmon straying from stream $i$ to
-stream $j$ as,
+have set up the metapopulation (thus far) in a very simple scenario: the
+populations are arranged in a line and those that are nearer to each
+other are more likely to stray between each other {[}insert caveats{]}.
+Two parameters control the straying: the fraction of fish $f$ that stray
+from their natal stream in any given generation and the rate at which
+this straying between streams decays with distance $m$. I calculated the
+number of salmon straying from stream $j$ to stream $i$ as,
 
 \begin{equation}
-  \mathrm{strays}_{tij} = \frac{S_{ti} f e^{-m \lvert i-j \rvert }}{\sum\limits_{k \neq j}^{} e^{-m \lvert k-j \rvert }},
+  \mathrm{strays}_{tij} = f R_{tj} \frac{e^{-m \lvert i-j \rvert }}{\displaystyle\sum\limits_{\substack{k = 1 \\
+  k \neq j}}^{n} e^{-m \lvert k-j \rvert }},
 \end{equation}
 
-where the denominator is a normalization constant to ensure that no
-salmon are gained or lost. The subscript $k$ represents a stream ID.
-{[}Not sure this is written correctly --- I'll think about it --- but it
-should be coded correctly.{]} There is no straying from and to the same
-stream. See Figure \ref{fig:stray-matrix} for an example straying
-matrix.
+\noindent
+where $R_{tj}$ is the number of returning salmon at generation $t$ whose
+natal stream was stream $j$. The subscript $k$ represents a stream ID
+and $n$ the number of populations. The denominator is a normalizing
+constant to ensure the desired fraction of fish stray. See Figure
+\ref{fig:stray-matrix} for an example straying matrix.
 
@@
 
+\noindent
 where $a_i^*$ is the maximum possible productivity parameter for
@@
 (more productive with cooler temperatures) to red (more productive with
-warmer temperatures\label{fig:thermal-curves}}
+warmer temperatures).\label{fig:thermal-curves}}
 \end{figure}
@@
 \begin{verbatim}
-  n_t = 120, # number of years
+n_t = 120, # number of generations
 n_pop = 10, # number of populations
@@
 
-\begin{verbatim}
-Error: could not find function "ricker_ar1"
-\end{verbatim}
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/plot-base-rickers.pdf}
+\caption{We can also look at samples of the Ricker curves by population
+(population numbers are in the upper left corners). Here, I've sampled
+40 of the $a$ parameters through time for each population. The shaded
+rectangles to indicate the range of spawners observed in the simulation.
+The colour of the line indicates the temperature anomaly in that year
+with blue indicating cooler years and red indicating warmer years. In
+this example all the populations have the same $b$ parameters. Notice
+how the productivity is more variable for populations nearer the limits
+of their thermal tolerance (e.g.~population 1 vs.~population 5). Also
+notice how population 1 is more productive in cool years whereas
+population 10 is more productive in warm years.}
+\end{figure}
 
@@
 
-\begin{verbatim}
-Error: could not find function "ricker_ar1"
-\end{verbatim}
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/unnamed-chunk-1.pdf}
+\caption{Example Ricker curves for the minimum-variance metapopulation.}
+\end{figure}
 
@@
 
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/plot-weights-samples.pdf}
+\caption{A sample of 30 minimum-variance portfolios given the same set
+of starting parameters as for the example illustrated above. The dashed
+grey line indicates the weights if all populations had an equal carrying
+capacity. As it stands, the weights look fairly random to me.}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/plot-weights-median.pdf}
+\caption{Median (solid line) and 25th and 75th quantiles of the weights
+across 40 minimum-variance portfolios. Just another way of looking at
+it.}
+\end{figure}
+
 \clearpage

### 2013-03-19 inst/examples/plot_portfolio_ts.R

@@ -1,9 +1,14 @@
 
-par(mfrow = c(2, 1), cex = 0.8, mar = c(0, 4, 0, 1), oma = c(4, 0,1,0))
+par(mfrow = c(2, 1), cex = 0.7, mar = c(0, 5, 0, 1), oma = c(4, 0,1,0), tck = -0.04, mgp = c(2, 0.7, 0), xpd = NA)
 burn <- 1:30
 port.x <- rowSums(min_var_sim$A[-burn, ])
 ret.x <- diff(log(port.x))
-plot(ret.x, type = "n", xaxt = "n", ylim = c(min(ret.x)*1.30, max(ret.x)*1.3))
+
+require(plyr)
+A_range <- ldply(mc_ports$sims_out, function(x) range(rowSums(x$A[-burn, ])))
+ret_range <- ldply(mc_ports$sims_out, function(x) range(diff(log(rowSums(x$A[-burn, ])))))
+
+plot(ret.x, type = "n", xaxt = "n", ylim = range(ret_range), ylab = "Generation rate of change of\nmetapopulation abundance", xlab = "")
 
 #for(i in ef_dat$ef_port_ids){
   #this.port <- mc_ports$sims_out[[i]]
@@ -12,9 +17,8 @@ plot(ret.x, type = "n", xaxt = "n", ylim = c(min(ret.x)*1.30, max(ret.x)*1.3))
   #lines(1:length(ret.x), ret.x, type = "l", col = "#44444440")
 #}
 
-
 # all!
-for(i in sample(1:length(mc_ports$sims_out), 80)){
+for(i in sample(1:length(mc_ports$sims_out), 100)){
   this.port <- mc_ports$sims_out[[i]]
   port.x <- rowSums(this.port$A[-burn, ])
   ret.x <- diff(log(port.x))
@@ -27,8 +31,9 @@ lines(1:length(ret.x), ret.x, lwd = 2.5, col = "white")
 
 # abundance?
 port.x <- rowSums(min_var_sim$A[-burn, ])
-plot(port.x, type = "n")
-for(i in sample(1:length(mc_ports$sims_out), 70)){
+plot(port.x, type = "n", ylim = range(A_range), ylab = "Metapopulation abundance", xlab = "Generation")
+
+for(i in sample(1:length(mc_ports$sims_out), 100)){
   this.port <- mc_ports$sims_out[[i]]
   port.x <- this.port$A[-burn, ]
   port.x <- rowSums(this.port$A[-burn, ])

### 2013-03-19 inst/examples/plot_sim_ts_simple.R

@@ -1,11 +1,11 @@
 # plot time series from the simulation
 # just the annual rate of growth and returns in this one
 
-plot_sim_ts_simple <- function(x, pal, burn = 1:30, text = "Returns") {
+plot_sim_ts_simple <- function(x, pal, burn = 1:50, text = "Returns") {
 # x is output from a simulation run; it's a list
 # pal is the colour palette
 
-to_show <- (max(burn)):(max(burn)+30) # years to show in time series example plots
+to_show <- (max(burn)):(max(burn)+50) # years to show in time series example plots
 
 annotate <- function(text) {
   mtext(text, side = 3, line = -1.5, cex = 0.7, adj = 0.05)
@@ -15,7 +15,7 @@ my.axis <- function(side) axis(side, col = "grey50", at = c(800, 1200))
 #par(mfrow = c(8, 1), mar = c(0,0,0,0), oma = c(4, 4, 1, 1), cex = 0.7, las = 1, xpd = FALSE)
 # returns:
 
-matplot(x$A[to_show, ], type = "l", col = pal, lty = 1, ylab = "Returns", xlab = "Time", xaxt = "n", axes = FALSE, xaxs = "i", ylim = c(580, 1450))
+matplot(x$A[to_show, ], type = "l", col = pal, lty = 1, ylab = "Returns", xlab = "Time", xaxt = "n", axes = FALSE, xaxs = "i", ylim = c(300, 1550))
 annotate(text)
 box(col = "grey50")
 my.axis(2)

### 2013-03-19 inst/examples/plot_thermal_curves.R

@@ -1,11 +1,22 @@
 # plot thermal curves across populations as an example
 
 par(cex = 0.8, mgp = c(2.4, 0.65, 0), tck = -0.02)
+#optim_temps <- seq(13, 19, length.out = 10)
+#x <- seq(3, 29, length.out = 200)
+#plot(1, 1, xlim = c(4, 28), ylim = c(-0.01, 1.4), ylab = "Ricker productivity parameter (a)", xlab = expression(Temperature~(degree*C)), type = "n", yaxs = "i", las = 1)
+#for(i in 1:10) {
+  #lines(x, thermal_curve_a(x, optim_temp = optim_temps[i], max_a = 1.3, width_param = 0.02), ylab = "a", xlab = "Temperature", type = "l", col = col_pal[i], lwd = 1.5)
+#}
+
+par(cex = 0.8, mgp = c(2.4, 0.65, 0), tck = -0.02)
 optim_temps <- seq(13, 19, length.out = 10)
+widths <- c(seq(0.06, 0.02, length.out = 5), rev(seq(0.06, 0.02, length.out = 5)))
+heights <- c(seq(1.8, 1.1, length.out = 5), rev(seq(1.8, 1.1, length.out = 5)))
+
 x <- seq(3, 29, length.out = 200)
-plot(1, 1, xlim = c(4, 28), ylim = c(-0.01, 1.4), ylab = "Ricker productivity parameter (a)", xlab = expression(Temperature~(degree*C)), type = "n", yaxs = "i", las = 1)
+plot(1, 1, xlim = c(4, 28), ylim = c(-0.01, 1.9), ylab = "Ricker productivity parameter (a)", xlab = expression(Temperature~(degree*C)), type = "n", yaxs = "i", las = 1)
 for(i in 1:10) {
-  lines(x, thermal_curve_a(x, optim_temp = optim_temps[i], max_a = 1.3, width_param = 0.02), ylab = "a", xlab = "Temperature", type = "l", col = col_pal[i], lwd = 1.5)
+  lines(x, thermal_curve_a(x, optim_temp = optim_temps[i], max_a = heights[i], width_param = widths[i]), ylab = "a", xlab = "Temperature", type = "l", col = col_pal[i], lwd = 1.5)
 }
 
 

### 2013-03-19 inst/examples/run_multiple_mpt.R

@@ -2,18 +2,18 @@
 
 # source files:
 rm(list = ls())
-setwd("~/Dropbox/ms/salmon_portfolios/")
-metafolio_files <-  list.files(path = "metafolio/R/")
+setwd("~/Dropbox/ms/metafolio/inst/examples/")
+metafolio_files <-  list.files(path = "../../R/")
 for(i in 1:length(metafolio_files)){
-  source(paste("metafolio/R/", metafolio_files[i], sep = ""));rm(i)}
+  source(paste("../..//R/", metafolio_files[i], sep = ""));rm(i)}
 col_pal <- rev(gg_color_hue(10))
 
 # run it:
-get_min_var_weights <- function(n_sims = 500, n_pop = 10){
+get_min_var_weights <- function(n_sims = 600, n_pop = 10){
   #browser()
 
   weights_matrix <-  create_asset_weights(n_pop = n_pop, n_sims = n_sims, 
-    weight_lower_limit = 0.01)
+    weight_lower_limit = 0.001)
 
   mc_ports <- monte_carlo_portfolios(weights_matrix = weights_matrix,
     n_sims = n_sims, mean_b = 1000)
@@ -24,7 +24,7 @@ get_min_var_weights <- function(n_sims = 500, n_pop = 10){
   min_var_sim <- mc_ports$sims_out[[ef_dat$min_var_port_id]]
   weights <- min_var_sim$b
 
-  save(mc_ports, file = paste("mc_ports_", iter, ".rda", sep = ""))
+  #save(mc_ports, file = paste("mc_ports_", iter, ".rda", sep = ""))
 
   # run an equal weighted portfolio:
   b_vals_matrix <- rep(1000, n_pop)
@@ -56,7 +56,7 @@ get_min_var_weights <- function(n_sims = 500, n_pop = 10){
 }
 
 iter <<- 1
-min_var_weights <- matrix(ncol = 10, nrow = 30)
+min_var_weights <- matrix(ncol = 10, nrow = 40)
 eq_weight_cv <- vector()
 minv_cv <- vector()
 
@@ -67,24 +67,5 @@ for(i in 1:nrow(min_var_weights)) {
   minv_cv[i] <- out$minv_cv
 }
 
-save(min_var_weights, eq_weight_cv, minv_cv, file = "min_var_weights_10pops_30mpt.rda")
[…]

### 2013-03-19 inst/examples/sub_grid.sh

@@ -1,7 +1,7 @@
 #!/bin/sh
-#PBS -l walltime=00:20:00
+#PBS -l walltime=00:50:00
 #PBS -M sean_anderson@sfu.ca
-#PBS -l mem=300mb
+#PBS -l mem=400mb
 
-R CMD BATCH --no-save /home/anderson/Dropbox/ms/salmon_portfolios/run_multiple_mpt.R
+R CMD BATCH --no-save /home/anderson/Dropbox/ms/metafolio/inst/examples/run_multiple_mpt.R
 

### 2013-03-19 inst/examples/salmonport.Rmd
Add impl. error on fishing

@@
 o <- meta_sim(env_params = list(amplitude = .000000001, ang_frequency = 0.2, phase = 0, mean_value = 16), sigma_v = 0.01, n_pop = 10, stray_fraction = 0.00001, use_cache = FALSE)
-o <- meta_sim(env_params = list(amplitude = .000000001, ang_frequency = 0.2, phase = 0, mean_value = 16), sigma_v = 0.1, n_pop = 10, stray_fraction = 0.00001, use_cache = FALSE)
+o <- meta_sim(env_params = list(amplitude = 2.5, ang_frequency = 0.2, phase = 0, mean_value = 16), sigma_v = 0.00001, n_pop = 10, stray_fraction = 0.00001, use_cache = FALSE)
+o <- meta_sim(env_params = list(amplitude = 2, ang_frequency = 0.2, phase = 0, mean_value = 16), sigma_v = 0.2, n_pop = 10, stray_fraction = 0.03, use_cache = FALSE)
 #o<- meta_sim(b = rep(1000, 10), n_pop = 10)

### 2013-03-19 R/meta_sim.R

@@ -25,9 +25,11 @@
 #' @param assessment_window Number of generations to use when fitting
 #' SR relationship for escapement targets; must be bigger than
 #' start_assessment
+#' @param sigma_impl implementation sd for beta distribution
 #' @param use_cache Use the stochastically generated values (SR residuals and
 #' possibly environmental time series) from the previous run?
 #' @param add_straying Implement straying between (sub)populations?
+#' @param add_impl_error Add implementation error?
 meta_sim <- function(
   n_t = 120, # number of years
   n_pop = 10, # number of subpopulations
@@ -44,8 +46,10 @@ meta_sim <- function(
   env_params = list(amplitude = 2.5, ang_frequency = 0.2, phase = 0, mean_value = 16),
   start_assessment = 30, # generation to start estimating SR relationship for escapement targets
   assessment_window = 25, # number of generations to use when fitting SR relationship for escapement targets; must be bigger than start_assessment
+  sigma_impl = 0.05, # sd on beta implementation error
   use_cache = FALSE, # regenerate stochastic values? 
   add_straying = TRUE # include or ignore straying
+  add_impl_error = TRUE # include implementation error?
 ) {
   
 #   source("generate_env_responses.R")
@@ -148,15 +152,18 @@ meta_sim <- function(
       Est_b[i,j]<-rick$b
 
       }
-    #browser()
     escapement_goals <- ricker_escapement(Est_a[i,],Est_b[i,])
     }
     #escapement_goals <- ricker_escapement(A_params[i,],b) # b is also a vector
+    if(add_impl_error) {
+      escapement_goals_fraction <- escapement_goals / A[i,]
+      escapement_goals_fraction[escapement_goals_fraction > .95] <- .95 # avoid rbeta errors if fraction too big - would have got reduced below anyways
+      escapement_goals_fraction_w_error <-  impl_error(mu = escapement_goals_fraction, sigma_impl = sigma_impl)
+      escapement_goals <- escapement_goals_fraction_w_error * A[i,]
+    }
     F[i, ] <- A[i, ] - escapement_goals # catch to leave escapement behind
     negative_F <- which(F[i, ] < 0)
     F[i, negative_F] <- 0
-# temporarily hammer them:
-    #F[i, ] <- A[i,]*runif(1, 0.1, 0.9)
     E[i, ] <- A[i, ] - F[i, ] # escapement
   }
   return(list(A = A, F = F, E = E, Eps = epsilon_mat, A_params = A_params, 

### 2013-03-28 inst/examples/salmonport.Rmd
Substantial additions to salmonport.Rmd

--- 2013-03-19 inst/examples/salmonport.Rmd
@@
 The term $w_{ti}$ represents first-order autocorrelated error (AR1).
-Formally, $w_{ti} = w_{ti-1} \rho + v_{ti}$, where $v_{ti}$ represents independent and normally distributed error with mean 0 and standard deviation of $\sigma_v$.
-The parameter $\rho$ represents the correlation between residuals from subsequent generations.
+Formally, $w_{ti} = w_{ti-1} \rho_w + v_{ti}$, where $v_{ti}$ represents independent and normally distributed error with mean 0 and standard deviation of $\sigma_v$.
+The parameter $\rho_w$ represents the correlation between residuals from subsequent generations.
 
@@
 ```{r env-ts, fig.height=4, fig.width=4, cache = FALSE, fig.cap="Example temperature time series\\label{fig:env-ts}"}
-# library(metafolio)
-# or for rapid development:
+#library(metafolio)
+#or for rapid development:
 metafolio_files <-  list.files(path = "../../R/")
@@
 I have set up the metapopulation (thus far) in a very simple scenario: the populations are arranged in a line and those that are nearer to each other are more likely to stray between each other [insert caveats].
-Two parameters control the straying: the fraction of fish $f$ that stray from their natal stream in any given generation and the rate at which this straying between streams decays with distance $m$.
+Two parameters control the straying: the fraction of fish $f_{stray}$ that stray from their natal stream in any given generation and the rate at which this straying between streams decays with distance $m$.
 I calculated the number of salmon straying from stream $j$ to stream $i$ as,
@@
 \begin{equation}
-  \mathrm{strays}_{tij} = f R_{tj} \frac{e^{-m \lvert i-j \rvert }}{\displaystyle\sum\limits_{\substack{k = 1 \\
+  \mathrm{strays}_{tij} = f_{stray} R_{tj} \frac{e^{-m \lvert i-j \rvert }}{\displaystyle\sum\limits_{\substack{k = 1 \\
   k \neq j}}^{n} e^{-m \lvert k-j \rvert }},
@@
 Here, and throughout this document, I have used the same rainbow of colours to identify the salmon populations.
-The populations are arranged from blue (more productive with cooler temperatures) to red (more productive with warmer temperatures).\\label{fig:thermal-curves}"}
+The populations are arranged from blue (more productive with cooler temperatures) to red (more productive with warmer temperatures).
+The vertical lines indicate mean, minimum, and maximum values for one scenario --- you can ignore them for now.\\label{fig:thermal-curves}"}
 source("plot_thermal_curves.R")
@@
 
+# Summary of simulation parameters
+
+\begin{table}[h!]
+\centering
+\small
+\caption{Salmon metapopulation parameters with base case and alternate values.}
+\begin{tabular}{p{7.0cm}p{1.6cm}p{3.2cm}}
+\toprule
+Description & Parameter & Base [lower, upper] \\
+\midrule
+Stock-recruit residual standard deviation (on log scale) & $\sigma_v$ & 0.30 [0.05, 0.50] \\
+First order (AR1) serial correlation of stock-recruit residuals & $\rho_w$ & 0.40 [0, 0.80] \\
+Fraction of fish that stray from natal streams & $f_{stray}$ & 0.02 [0, 0.10] \\
+Exponential rate of decay of straying with distance & $m$ & 0.30 [0.05, 0.50] \\
+Standard deviation of beta distribution for implementation error & $\sigma_{impl}$ & 0.05 [0, 0.20] \\
+Frequency of assessment (years) & $f_{assess}$ & 20 [5, 50] \\
+\bottomrule
+\end{tabular}
+\label{tab:salm-pars}
+\end{table}
+
+\clearpage
+
 # Portfolio optimization #
+
 Details to follow.
@@
 # Some sample simulations #
-The following are options and their default values:
-
-    n_t = 120, # number of generations
-    n_pop = 10, # number of populations
-    stray_decay_rate = 0.3, # rate that straying decays with distance
-    stray_fraction = 0.01, # fraction of fish that stray from natal streams
-    b = rep(1000, n_pop), # Ricker density-dependent parameter
-    spawners_0 = round(b), # spawners at start
-    sigma_v = 0.1, # stock-recruit residual SD
-    v_rho = 0.4, # stock-recruit residual AR1 correlation
-    a_width_param = rep(0.02, n_pop), # width of thermal curves by pop
-    optim_temp = seq(13, 19, length.out = n_pop), # optimal temperatures by pop
-    max_a = rep(1.4, n_pop), # maximum Ricker a values by pop at optimum temp
-    env_type = c("sine", "arma", "regime", "linear", "constant"),
-    env_params = list(amplitude = 2.0, ang_frequency = 0.2, phase = 0, mean_value = 16),
-    use_cache = FALSE, # regenerate stochastic values? 
-    add_straying = TRUE # include or ignore straying
-
-First, let's run a couple example simulations:
 
@@
 I'm only showing a sample of 40 generations (after the burn-in period).
-From top to bottom: salmon returns each generation (catch plus escapement); Ricker $a$ parameter; the environmental signal (could think of it as temperature or a composite environmental signal like the PDO); number of salmon straying by population; number of straying salmon being received by population; fisheries catch based on escapement MSY calculation; salmon escapement; residuals from the spawner-return curve."}
+From top to bottom: salmon returns each generation (catch plus escapement); Ricker $a$ parameter; the environmental signal (could think of it as temperature or a composite environmental signal like the PDO); number of salmon straying by population; number of straying salmon being received by population; fisheries catch based on escapement MSY calculation; salmon escapement; residuals from the spawner-return curve; estimated Ricker $a$ parameter; estimated Ricker $b$ parameter."}
 set.seed(2)
@@
 
-o <- meta_sim(env_params = list(amplitude = .000000001, ang_frequency = 0.2, phase = 0, mean_value = 16), sigma_v = 0.01, n_pop = 10, stray_fraction = 0.00001, use_cache = FALSE)
-o <- meta_sim(env_params = list(amplitude = .000000001, ang_frequency = 0.2, phase = 0, mean_value = 16), sigma_v = 0.1, n_pop = 10, stray_fraction = 0.00001, use_cache = FALSE)
-#o<- meta_sim(b = rep(1000, 10), n_pop = 10)
-plot_sim_ts(o, years_to_show = 60)
-image(o$Est_b)
+#o <- meta_sim(env_params = list(amplitude = .000000001, ang_frequency = 0.2, phase = 0, mean_value = 16), sigma_v = 0.01, n_pop = 10, stray_fraction = 0.00001, use_cache = FALSE)
+#o <- meta_sim(env_params = list(amplitude = 2.5, ang_frequency = 0.2, phase = 0, mean_value = 16), sigma_v = 0.00001, n_pop = 10, stray_fraction = 0.00001, use_cache = FALSE)
+#o <- meta_sim(env_params = list(amplitude = 2.2, ang_frequency = 0.2, phase = 0, mean_value = 16), sigma_v = 0.1, n_pop = 10, stray_fraction = 0.01, use_cache = FALSE)
+##o<- meta_sim(b = rep(1000, 10), n_pop = 10)
+#plot_sim_ts(o, years_to_show = 40)
+#image(o$Est_b)
 ```
@@
 
-```{r plot-various-options-ts, fig.width=4, fig.height=6.3, fig.cap="Example simulations with different parameter values. This version has ten populations.\\label{fig:sim-param-ts-10}"}
-source("example_returns_10.R")
-```
-
 \clearpage
@@
 # An example of portfolio optimization #
+
 Now, let's run an example where we find the $b$ values that form the efficient frontier of metapopulation portfolios.
@@
 
+```{r plot-cons-plans-curves, cache=FALSE, fig.width=3.3, fig.height=5.0, fig.cap="Different ways of prioritizing conservation.
+The curves show the environmental-tolerance curves for the populations that are conserved.
+The colours are the same as earlier and identify the different populations.
+These are in the same order as they are referred to in the next two plots: full range of environmental response diversity, most stable only, one half, other half, most asynchronous.", cache=FALSE}
+#Notes #
+#-b is "unfished equilibrium stock size"
+source("plot_thermal_curves_scenarios.R")
+```
+
+```{r plot-cons-plans-mv, fig.width=6.5, fig.height=3.5, fig.cap="Conservation plans in mean-variance space.
+The dots show portfolios from individual simulation runs.
+The dark and light shaded areas show 25% and 75% quartiles.
+The left panel shows a scenario with a stationary environmental signal and strong short-term environmental noise.
+The right panel shows a scenario with a linearly-increasing environmental signal and minimal short-term environmental noise.
+These two scenarios bound the two environmental extremes: strong short-term noise with minimal long-term signal vs.
+a strong long-term signal with minimal short-term signal.
+Notice what happens to the two *half* conservation strategies.
+A more realistic scenario combines the two.
+The efficient frontier is shown for the displayed portfolios.
+This isn't necessarily the overall efficient frontier.", cache=FALSE}
+#Notes #
+#-b is "unfished equilibrium stock size"
+require(RColorBrewer)
+cols <- brewer.pal(5, "Dark2")[c(1,2,4,5,3)] 
+source("cons_plans.R")
+```
+
+```{r plot-cons-plans-ts, fig.width=6, fig.height=3, fig.cap="Conservation plans in terms of metapopulation abundance.
+One realization with a periodic (sine wave) environmental signal is shown as an example.
+The light-grey lines in the background show 100 randomly-sampled portfolios with the same stock-recruit residuals and the same environmental signal.
+Notice how much less variable the scenarios with balanced response diversity are.
+The two scenarios with unbalanced response diversity tend to seesaw with the environment."}
+source("cons_plans_ts.R")
+```
+
 \clearpage
 
-# Notes #
-- b is "unfished equilibrium stock size"
+# Take home messages
+1. If the dominant environmental signal is a long-term shift, and you keep around one half of the response diversity, then you get different portfolio means but the same variances.
+2. If the dominant environmental signal is short-term, and you keep around one half of the response diversity, then you get different portfolio variances but the same means.
+3. Combine those two (which best reflects reality) and you get both different means and variances.
+4. Whether you invest in just stable, just asynchronous, or a mix of both, all that really matters is that you have as many populations as possible and as balanced as possible response diversity.
+5. An even weighting across as many possible streams (assuming balanced response diversity) gets you pretty close to the minimum-variance portfolio. [not shown currently]
+6.
+With systematic long-term environmental change that dominates: if you can forecast the environment and know the responses of salmon populations from individual streams then you can increase your expected rate of metapopulation growth with minimal effect on variance by prioritizing the populations that will do well.
+Although if you get it wrong you will do worse than if you had hedged your bets and conserved response diversity.
+7. But, if there are also strong short-term environmental fluctuations (and there probably are), then there's always a benefit to portfolio variance by having some response diversity.
+8. Variable stocks aren't inherently undesirable to conserve, as long as they are balanced or averaged out.
+9. Take home message: "To keep every cog and wheel is the first precaution of intelligent tinkering." But, if you must prioritize, keep a selection of different cogs and
+ wheels --- especially a spatially diverse set of cogs and wheels. That gives you the best chance of an efficient ecological portfolio.
 
-```{r}
-```
+# Open questions and next steps
+1. How do smaller or larger values of the various parameters affect the results?
+2. Are the results sensitive to the height and width of the thermal tolerance curves? (I think within reason, no.)
+3. Are these results robust to, say, cyclical or regime shift style environmental noise?
+4. Should this be a generic "environmental signal" or temperature?
+5. Show what happens for a given scenario as you conserve more streams (it gets closer to the minimum-variance portfolio).
+6. Show an example of where these conservation plans place you in mean-variance space across all possible weights. (Done but not shown.)
+7. Bring these conservation plans back to established suggestions (e.g. VSP).
+8. How can we simplify the main figures?
+9. Is there a way to combine the main figures to make the result more understandable (e.g. little response-diversity curves as a legend)?
+10. Redo the Monte Carlo portfolios to show an extreme example where you can see the weights shifting. For example, a linearly increasing environmental signal with some generation-to-generation noise.
+11. Generations or years? Years seem simpler to talk about. Generations seem more widely applicable and accurate.
+
+\clearpage
 

### 2013-03-28 inst/examples/bugaboo.R
Add new and modifed paper-specific R files

@@ -1,0 +1,3 @@
+#' 
+
+

### 2013-03-28 inst/examples/draw_contour.R

@@ -1,0 +1,78 @@
+draw.contour<-function(a,alpha=0.95,plot.dens=FALSE, line.width=2, line.type=1, limits=NULL, density.res=300,spline.smooth=-1,...){
+  ##a is a list or matrix of x and y coordinates (e.g., a=list("x"=rnorm(100),"y"=rnorm(100)))
+  ## if a is a list or dataframe, the components must be labeled "x" and "y"
+  ## if a is a matrix, the first column is assumed to be x, the second y
+  ##alpha is the contour level desired
+  ##if plot.dens==TRUE, then the joint density of x and y are plotted,
+  ##   otherwise the contour is added to the current plot.
+  ##density.res controls the resolution of the density plot
+
+  ##A key assumption of this function is that very little probability mass lies outside the limits of
+  ## the x and y values in "a". This is likely reasonable if the number of observations in a is large.
+  
+  require(MASS)
+  require(ks)
+  if(length(line.width)!=length(alpha)){
+    line.width <- rep(line.width[1],length(alpha))
+  }
+
+  if(length(line.type)!=length(alpha)){
+    line.type <- rep(line.type[1],length(alpha))
+  }
+  
+  if(is.matrix(a)){
+    a=list("x"=a[,1],"y"=a[,2])
+  }
+  ##generate approximate density values
+  if(is.null(limits)){
+    limits=c(range(a$x),range(a$y))
+  }
+  f1<-kde2d(a$x,a$y,n=density.res,lims=limits)
+
+  ##plot empirical density
+  if(plot.dens) image(f1,...)
+  
+  if(is.null(dev.list())){
+    ##ensure that there is a window in which to draw the contour
+    plot(a,type="n",xlim=limits[1:2],ylim=limits[3:4],...)
+  }
+    
+  ##estimate critical contour value
+  ## assume that density outside of plot is very small
+
+
+  zdens <- rev(sort(f1$z))
+  Czdens <- cumsum(zdens)
+  Czdens <- (Czdens/Czdens[length(zdens)])
+  crit.val <- zdens[max(which(Czdens<=.95))]    
+  lines(b.full[[1]])
+    b.full=contourLines(f1,levels=crit.val)
+
+  for(cont.level in 1:length(alpha)){
+    ##This loop allows for multiple contour levels
+    crit.val <- zdens[max(which(Czdens<=alpha[cont.level]))]
+   
+    ##determine coordinates of critical contour
[…]

### 2013-03-28 inst/examples/get_quantile_contour.R

@@ -1,0 +1,13 @@
+#' 
+
+get_quantile_contour <- function(x, alpha = 0.8) {
+  zdens <- rev(sort(x$z))
+  Czdens <- cumsum(zdens)
+  Czdens <- (Czdens/Czdens[length(zdens)])
+  crit.val <- zdens[max(which(Czdens<=alpha))]    
+  b.full=contourLines(x,levels=crit.val)
+  list(x = b.full[[1]]$x, y = b.full[[1]]$y)
+}
+
+#plot(get_quantile_contour(f1))
+#lines(get_quantile_contour(f1, alpha = 0.75))

### 2013-03-28 inst/examples/look_at_cluster_dat.R

@@ -1,0 +1,34 @@
+#' 
+
+
+load("cluster_dat/min_var_weights_10pops_40mpt.rda")
+
+ #matplot(t(log(min_var_weights)), type = "l")
+ #matplot(t(min_var_weights/10000*100), type = "l", ylab = "Percentage of total habitat")
+ #abline(h = 1/10*100, lwd = 2, col = "grey50")
+
+## plotting
+ #matplot(t(min_var_weights/10000*100), type = "l", ylab = "Percentage of total habitat", col = "#00000050", lty = 1, lwd = 1.5)
+ #abline(h = 1/10*100, lwd = 2, col = "grey50")
+#par(mfrow = c(6, 2), cex = 0.7, mar = c(0,0,0,0));for(i in 1:12) acf(min_var_weights[i, ])
+#ar1 <- apply(min_var_weights, 1, function(x) {
+  #m <-arima(x, order = c(1, 0,0))
+  #as.numeric(coef(m)[1])
+    #})
+#hist(ar1)
+##plot(density(ar1))
+
+par(mfrow = c(6, 5), mar = c(0,0,0,0), oma = c(4, 4, .5, .5), cex = 0.7);for(i in 1:30) {
+  plot(min_var_weights[i,]/100, type = "l", axes = FALSE, ylim = c(1, 20))
+  box(col = "grey50")
+  abline(h = 10, col = "grey70", lty = 2)
+  if(i %in% seq(1, 26, 5)) axis(2, col = "grey50", las = 1)
+  if(i %in% seq(26, 30)) axis(1, col = "grey50", las = 1)
+}
+mtext("Carrying capacity weight (percentage)", side = 2, outer = TRUE, line = 2.5)
+mtext("Population", side = 1, outer = TRUE, line = 2.5)
+## what is the variance of the even weighted portfolio compared to the minimum 
+## variance portfolio?
+#plot(jitter(rep(1, length(minv_cv))), minv_cv/eq_weight_cv, xlim = c(0.8, 1.2))
+#hist(minv_cv/eq_weight_cv)
+

### 2013-03-28 inst/examples/min_var_ports_dist.R

@@ -1,0 +1,17 @@
+#'  distribution of portfolio weights across runs:
+
+load("cluster_dat/min_var_weights_10pops_40mpt.rda")
+
+library(plyr)
+
+p <- adply(min_var_weights, 2, function(x) data.frame(l = quantile(x/100, 0.25), m = median(x/100), u = quantile(x/100, 0.75)))
+p$X1 <- as.numeric(as.character(p$X1))
+names(p)[1] <- "pop"
+
+library(ggplot2)
+par(cex = 0.8)
+with(p, plot(pop, m, type = "l", lwd = 1.5, ylim = c(min(l), max(u)), xlab = "Population", ylab = "Weight", las = 1))
+with(p, polygon(c(pop, rev(pop)), c(l, rev(u)), border = NA, col = "#00000030"))
+abline(h = 10, lty = 2)
+
+

### 2013-03-28 inst/examples/plot_cons_plans.R

@@ -1,0 +1,58 @@
+#' 
+
+add_dens_polygon <- function(x, y, col, lwd = 1, alpha = 0.7, add_pts = FALSE) {
+  source("get_quantile_contour.R")
+  require(MASS)
+  k <- get_quantile_contour(kde2d(x,y), alpha = 0.75)
+  polygon(k$x, k$y, border = col, col = paste(col, "30", sep = ""), lwd = lwd)
+  k <- get_quantile_contour(kde2d(x,y), alpha = 0.25)
+  polygon(k$x, k$y, border = col, col = paste(col, "70", sep = ""), lwd = lwd)
+  if(add_pts) points(x, y, pch = 20, col = paste(col, "60", sep = ""))
+}
+
+plot_cons_plans <- function(plans_mv, plans_name, xlim = NULL, ylim = NULL, add_pts = TRUE) {
+
+  if(is.null(xlim)) {
+lims <- ldply(plans_mv, function(x) data.frame(x.max = max(x$v), x.min = min(x$v), y.max = max(x$m), y.min = min(x$m)))
+  xlim = c(min(lims$x.min), max(lims$x.max)*0.8)
+ ylim = c(min(lims$y.min), max(lims$y.max))
+  }
+
+#par(cex = 0.8, mar = c(4,5,0,0), oma = c(.5,.5,.5,.5), mgp = c(3, .5, 0), tck = -0.01, las = 1)
+plot(1, 1, xlim = xlim, ylim = ylim, type = "n", xlab = "Variance of generation-to-generation\nrate of change", ylab = "Mean of generation-to-generation\nrate of change", axes = FALSE)
+box(col = "grey50")
+axis(1, col=  "grey50")
+axis(2, col=  "grey50")
+
+#require(RColorBrewer)
+#cols <- brewer.pal(5, "Dark2")[c(3, 2, 5, 4, 1)]
+#cols <- brewer.pal(5, "Dark2")[c(1,2,4,5,3)] 
+#cols <- gg_color_hue(5)[c(3, 2, 1, 4, 5)]
+#cols <- c(cols, cols[3])
+#for(i in 1:length(plans_mv)) {
+for(i in w_show) {
+  add_dens_polygon(plans_mv[[i]]$v, plans_mv[[i]]$m, col = cols[i], add_pts = add_pts)
+}
+legend("topright", legend = plans_name[w_show], fill = paste(cols[w_show], "80", sep = ""), bty = "n", cex = 0.9)
+
[…]

### 2013-03-28 inst/examples/plot_env_ts.R

@@ -2,7 +2,7 @@
 
 types <- c("sine", "arma", "regime", "linear", "constant")
 x <- list()
-for(i in 1:5) x[[i]] <- generate_env_ts(n_t = 100, type = types[i], arma_params = list(sigma_env = 0.30, ar = 0.6, ma = 0.2))
+for(i in 1:5) x[[i]] <- generate_env_ts(n_t = 100, type = types[i])
 op <- par(mfrow = c(5, 1), mar = c(0,0,0,0), cex = 0.8, oma = c(3.1, 3.2, .5, 5), xpd = NA, mgp = c(2.5, 0.7, 0), tck = -0.06, las = 1) 
 for(i in 1:5) {
   plot(x[[i]], type = "l",  axes = FALSE, ylim = c(-1.2, 1.12), xlab = "", ylab = "")

### 2013-03-28 inst/examples/plot_sim_ts_simple.R

@@ -1,11 +1,11 @@
 # plot time series from the simulation
 # just the annual rate of growth and returns in this one
 
-plot_sim_ts_simple <- function(x, pal, burn = 1:50, text = "Returns") {
+plot_sim_ts_simple <- function(x, pal, burn = 1:60, text = "Returns") {
 # x is output from a simulation run; it's a list
 # pal is the colour palette
 
-to_show <- (max(burn)):(max(burn)+50) # years to show in time series example plots
+to_show <- (max(burn)):(max(burn)+40) # years to show in time series example plots
 
 annotate <- function(text) {
   mtext(text, side = 3, line = -1.5, cex = 0.7, adj = 0.05)
@@ -15,7 +15,7 @@ my.axis <- function(side) axis(side, col = "grey50", at = c(800, 1200))
 #par(mfrow = c(8, 1), mar = c(0,0,0,0), oma = c(4, 4, 1, 1), cex = 0.7, las = 1, xpd = FALSE)
 # returns:
 
-matplot(x$A[to_show, ], type = "l", col = pal, lty = 1, ylab = "Returns", xlab = "Time", xaxt = "n", axes = FALSE, xaxs = "i", ylim = c(300, 1550))
+matplot(x$A[to_show, ], type = "l", col = pal, lty = 1, ylab = "Returns", xlab = "Time", xaxt = "n", axes = FALSE, xaxs = "i", ylim = c(300, 2450))
 annotate(text)
 box(col = "grey50")
 my.axis(2)

### 2013-03-28 inst/examples/salmonport.tex
Update all remaining files

@@
 \setmonofont{Inconsolata}
-%\setmainfont{Linux Libertine}
+\setmainfont{Linux Libertine}
 %\setmainfont{Times New Roman}
-\setmainfont{Calibri}
+%\setmainfont{Calibri}
 %\setmainfont{Helvetica}
@@
 weights in the portfolios). The term $w_{ti}$ represents first-order
-autocorrelated error (AR1). Formally, $w_{ti} = w_{ti-1} \rho + v_{ti}$,
-where $v_{ti}$ represents independent and normally distributed error
-with mean 0 and standard deviation of $\sigma_v$. The parameter $\rho$
-represents the correlation between residuals from subsequent
-generations.
+autocorrelated error (AR1). Formally,
+$w_{ti} = w_{ti-1} \rho_w + v_{ti}$, where $v_{ti}$ represents
+independent and normally distributed error with mean 0 and standard
+deviation of $\sigma_v$. The parameter $\rho_w$ represents the
+correlation between residuals from subsequent generations.
 
@@
 other are more likely to stray between each other {[}insert caveats{]}.
-Two parameters control the straying: the fraction of fish $f$ that stray
-from their natal stream in any given generation and the rate at which
-this straying between streams decays with distance $m$. I calculated the
-number of salmon straying from stream $j$ to stream $i$ as,
+Two parameters control the straying: the fraction of fish $f_{stray}$
+that stray from their natal stream in any given generation and the rate
+at which this straying between streams decays with distance $m$. I
+calculated the number of salmon straying from stream $j$ to stream $i$
+as,
 
 \begin{equation}
-  \mathrm{strays}_{tij} = f R_{tj} \frac{e^{-m \lvert i-j \rvert }}{\displaystyle\sum\limits_{\substack{k = 1 \\
+  \mathrm{strays}_{tij} = f_{stray} R_{tj} \frac{e^{-m \lvert i-j \rvert }}{\displaystyle\sum\limits_{\substack{k = 1 \\
   k \neq j}}^{n} e^{-m \lvert k-j \rvert }},
@@
 (more productive with cooler temperatures) to red (more productive with
-warmer temperatures).\label{fig:thermal-curves}}
+warmer temperatures). The vertical lines indicate mean, minimum, and
+maximum values for one scenario --- you can ignore them for
+now.\label{fig:thermal-curves}}
 \end{figure}
+
+\section{Summary of simulation parameters}
+
+\begin{table}[h!]
+\centering
+\small
+\caption{Salmon metapopulation parameters with base case and alternate values.}
+\begin{tabular}{p{7.0cm}p{1.6cm}p{3.2cm}}
+\toprule
+Description & Parameter & Base [lower, upper] \\
+\midrule
+Stock-recruit residual standard deviation (on log scale) & $\sigma_v$ & 0.30 [0.05, 0.50] \\
+First order (AR1) serial correlation of stock-recruit residuals & $\rho_w$ & 0.40 [0, 0.80] \\
+Fraction of fish that stray from natal streams & $f_{stray}$ & 0.02 [0, 0.10] \\
+Exponential rate of decay of straying with distance & $m$ & 0.30 [0.05, 0.50] \\
+Standard deviation of beta distribution for implementation error & $\sigma_{impl}$ & 0.05 [0, 0.20] \\
+Frequency of assessment (years) & $f_{assess}$ & 20 [5, 50] \\
+\bottomrule
+\end{tabular}
+\label{tab:salm-pars}
+\end{table}
+
+\clearpage
 
@@
 
-The following are options and their default values:
-
-\begin{verbatim}
-n_t = 120, # number of generations
-n_pop = 10, # number of populations
-stray_decay_rate = 0.3, # rate that straying decays with distance
-stray_fraction = 0.01, # fraction of fish that stray from natal streams
-b = rep(1000, n_pop), # Ricker density-dependent parameter
-spawners_0 = round(b), # spawners at start
-sigma_v = 0.1, # stock-recruit residual SD
-v_rho = 0.4, # stock-recruit residual AR1 correlation
-a_width_param = rep(0.02, n_pop), # width of thermal curves by pop
-optim_temp = seq(13, 19, length.out = n_pop), # optimal temperatures by pop
-max_a = rep(1.4, n_pop), # maximum Ricker a values by pop at optimum temp
-env_type = c("sine", "arma", "regime", "linear", "constant"),
-env_params = list(amplitude = 2.0, ang_frequency = 0.2, phase = 0, mean_value = 16),
-use_cache = FALSE, # regenerate stochastic values? 
-add_straying = TRUE # include or ignore straying
-\end{verbatim}
-
-First, let's run a couple example simulations:
-
 \begin{figure}[htbp]
@@
 by population; fisheries catch based on escapement MSY calculation;
-salmon escapement; residuals from the spawner-return curve.}
+salmon escapement; residuals from the spawner-return curve; estimated
+Ricker $a$ parameter; estimated Ricker $b$ parameter.}
 \end{figure}
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{figure/plot-various-options-ts.pdf}
-\caption{Example simulations with different parameter values. This
-version has ten populations.\label{fig:sim-param-ts-10}}
-\end{figure}
-
 \clearpage
@@
 efficient frontier of metapopulation portfolios.
+
+\begin{verbatim}
+Error: ignoring SIGPIPE signal
+\end{verbatim}
 
@@
 
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/plot-cons-plans-curves.pdf}
+\caption{Different ways of prioritizing conservation. The curves show
+the environmental-tolerance curves for the populations that are
+conserved. The colours are the same as earlier and identify the
+different populations. These are in the same order as they are referred
+to in the next two plots: full range of environmental response
+diversity, most stable only, one half, other half, most asynchronous.}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/plot-cons-plans-mv.pdf}
+\caption{Conservation plans in mean-variance space. The dots show
+portfolios from individual simulation runs. The dark and light shaded
+areas show 25\% and 75\% quartiles. The left panel shows a scenario with
+a stationary environmental signal and strong short-term environmental
+noise. The right panel shows a scenario with a linearly-increasing
+environmental signal and minimal short-term environmental noise. These
+two scenarios bound the two environmental extremes: strong short-term
+noise with minimal long-term signal vs.~a strong long-term signal with
+minimal short-term signal. Notice what happens to the two \emph{half}
+conservation strategies. A more realistic scenario combines the two. The
+efficient frontier is shown for the displayed portfolios. This isn't
+necessarily the overall efficient frontier.}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/plot-cons-plans-ts.pdf}
+\caption{Conservation plans in terms of metapopulation abundance. One
+realization with a periodic (sine wave) environmental signal is shown as
+an example. The light-grey lines in the background show 100
+randomly-sampled portfolios with the same stock-recruit residuals and
+the same environmental signal. Notice how much less variable the
+scenarios with balanced response diversity are. The two scenarios with
+unbalanced response diversity tend to seesaw with the environment.}
+\end{figure}
+
+\clearpage
+
+\section{Take home messages}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  If the dominant environmental signal is a long-term shift, and you
+  keep around one half of the response diversity, then you get different
+  portfolio means but the same variances.
+\item
+  If the dominant environmental signal is short-term, and you keep
+  around one half of the response diversity, then you get different
+  portfolio variances but the same means.
+\item
+  Combine those two (which best reflects reality) and you get both
+  different means and variances.
+\item
+  Whether you invest in just stable, just asynchronous, or a mix of
+  both, all that really matters is that you have as many populations as
+  possible and as balanced as possible response diversity.
+\item
+  An even weighting across as many possible streams (assuming balanced
+  response diversity) gets you pretty close to the minimum-variance
+  portfolio. {[}not shown currently{]}
+\item
+  With systematic long-term environmental change that dominates: if you
+  can forecast the environment and know the responses of salmon
+  populations from individual streams then you can increase your
+  expected rate of metapopulation growth with minimal effect on variance
+  by prioritizing the populations that will do well. Although if you get
+  it wrong you will do worse than if you had hedged your bets and
+  conserved response diversity.
+\item
+  But, if there are also strong short-term environmental fluctuations
+  (and there probably are), then there's always a benefit to portfolio
+  variance by having some response diversity.
+\item
+  Variable stocks aren't inherently undesirable to conserve, as long as
+  they are balanced or averaged out.
+\item
+  Take home message: ``To keep every cog and wheel is the first
+  precaution of intelligent tinkering.'' But, if you must prioritize,
+  keep a selection of different cogs and wheels --- especially a
+  spatially diverse set of cogs and wheels. That gives you the best
+  chance of an efficient ecological portfolio.
+\end{enumerate}
+
+\section{Open questions and next steps}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  How do smaller or larger values of the various parameters affect the
+  results?
+\item
+  Are the results sensitive to the height and width of the thermal
+  tolerance curves? (I think within reason, no.)
+\item
+  Are these results robust to, say, cyclical or regime shift style
+  environmental noise?
+\item
+  Should this be a generic ``environmental signal'' or temperature?
+\item
+  Show what happens for a given scenario as you conserve more streams
+  (it gets closer to the minimum-variance portfolio).
+\item
+  Show an example of where these conservation plans place you in
+  mean-variance space across all possible weights. (Done but not shown.)
+\item
+  Bring these conservation plans back to established suggestions
+  (e.g.~VSP).
+\item
+  How can we simplify the main figures?
+\item
+  Is there a way to combine the main figures to make the result more
+  understandable (e.g.~little response-diversity curves as a legend)?
+\item
+  Redo the Monte Carlo portfolios to show an extreme example where you
+  can see the weights shifting. For example, a linearly increasing
+  environmental signal with some generation-to-generation noise.
+\item
+  Generations or years? Years seem simpler to talk about. Generations
+  seem more widely applicable and accurate.
+\end{enumerate}
+
 \clearpage

### 2013-03-28 R/generate_env_ts.R

@@ -18,19 +18,19 @@
 generate_env_ts <- function(
   n_t,
   type = c("sine", "arma", "regime", "linear", "constant"),
-  sine_params = list(amplitude = 1, ang_frequency = 0.2, phase = 0, mean_value = 0),
-  arma_params = list(sigma_env = 0.30, ar = 0.6, ma = 0),
+  sine_params = list(amplitude = 1, ang_frequency = 0.2, phase = 0, mean_value = 0, slope = 0, sigma_env = 0.10),
+  arma_params = list(mean_value = 0, sigma_env = 0.50, ar = 0.4, ma = 0),
   regime_params = list(break_pts = c(25, 75), break_vals = c(-1, 0, 1)),
-  linear_params = list(min_value = -1, max_value = 1),
+  linear_params = list(min_value = -1, max_value = 1, sigma_env = 0.10),
   constant_params = list(value = 0)
   ) {
   type <- type[1]
   env_ts <- switch(type, 
-    arma = as.numeric(with(arma_params, arima.sim(model = list(ar = ar, ma = ma),
+    arma = as.numeric(with(arma_params, mean_value + arima.sim(model = list(ar = ar, ma = ma),
         n = n_t, sd = sigma_env))), 
     sine = {
       x <- seq_len(n_t)
-      y <- with(sine_params, amplitude * sin(ang_frequency * x + phase) + mean_value) 
+      y <- with(sine_params, amplitude * sin(ang_frequency * x + phase) + mean_value + x * slope + rnorm(n_t, mean = 0, sd = sigma_env))
       y
     },
     regime = {
@@ -44,7 +44,7 @@ generate_env_ts <- function(
       y
     },
     linear = {
-      with(linear_params, seq(min_value, max_value, length.out = n_t))
+      with(linear_params, seq(min_value, max_value, length.out = n_t) + rnorm(n_t, mean = 0, sd = sigma_env))
     },
     constant = {
       rep(constant_params$value, n_t)

### 2013-03-28 R/meta_sim.R

@@ -26,6 +26,8 @@
 #' SR relationship for escapement targets; must be bigger than
 #' start_assessment
 #' @param sigma_impl implementation sd for beta distribution
+#' @param assess_freq How many generations before re-assessing a and b
+#' parameters.
 #' @param use_cache Use the stochastically generated values (SR residuals and
 #' possibly environmental time series) from the previous run?
 #' @param add_straying Implement straying between (sub)populations?
@@ -34,21 +36,22 @@ meta_sim <- function(
   n_t = 120, # number of years
   n_pop = 10, # number of subpopulations
   stray_decay_rate = 0.3, # rate that straying decays with distance
-  stray_fraction = 0.01, # fraction of fish that stray from natal streams
+  stray_fraction = 0.000001, # fraction of fish that stray from natal streams
   b = rep(1000, n_pop), # Ricker density-dependent parameter
   spawners_0 = round(b), # spawners at start
-  sigma_v = 0.1, # stock-recruit residual SD
+  sigma_v = 0.3, # stock-recruit residual SD
   v_rho = 0.4, # stock-recruit residual AR1 correlation
-  a_width_param = c(seq(0.06, 0.02, length.out = n_pop/2), rev(seq(0.06, 0.02, length.out = n_pop/2))), # width of thermal curves by pop
+  a_width_param = c(seq(0.05, 0.02, length.out = n_pop/2), rev(seq(0.05, 0.02, length.out = n_pop/2))), # width of thermal curves by pop
   optim_temp = seq(13, 19, length.out = n_pop), # optimal temperatures by pop
-  max_a = c(seq(1.8, 1.1, length.out = 5), rev(seq(1.8, 1.1, length.out = 5))), # maximum Ricker a values by pop at optimum temp
+  max_a = c(seq(2.8, 2.2, length.out = 5), rev(seq(2.8, 2.2, length.out = 5))), # maximum Ricker a values by pop at optimum temp
   env_type = c("sine", "arma", "regime", "linear", "constant"),
-  env_params = list(amplitude = 2.5, ang_frequency = 0.2, phase = 0, mean_value = 16),
+  env_params = list(amplitude = 3.2, ang_frequency = 0.2, phase = runif(1, -pi, pi), mean_value = 15, slope = 0, sigma_env = 0.30),
   start_assessment = 30, # generation to start estimating SR relationship for escapement targets
[…]

### 2013-03-28 R/monte_carlo_portfolios.R

@@ -10,6 +10,7 @@ monte_carlo_portfolios <- function(weights_matrix, n_sims = 500, mean_b = 1000, 
 
   sims_out <- list() # to store simulation output
 
+
   for(k in 1:n_sims) {
     b_vals <- b_vals_matrix[k, ]
     if(k == 1)
@@ -24,3 +25,35 @@ monte_carlo_portfolios <- function(weights_matrix, n_sims = 500, mean_b = 1000, 
   }
 return(list(port_vals = port_vals, n_sims = 500, sims_out = sims_out))
 }
+
+monte_carlo_portfolios_foreach <- function(weights_matrix, n_sims = 500, mean_b = 1000, burn = 1:30, ...) {
+
+  b_vals_matrix <- weights_matrix * mean_b * ncol(weights_matrix)
+  n_pop <- ncol(weights_matrix)
+
+  port_vals = matrix(ncol = 2, nrow = n_sims)
+
+  sims_out <- list() # to store simulation output
+
+#require(foreach)
+
+# just set cache:
+junk <- meta_sim(b = b_vals_matrix[1,], use_cache = FALSE, n_pop = n_pop, ...)
+  
+sims_out <- foreach(k = 1:n_sims) %dopar% {
+    b_vals <- b_vals_matrix[k, ]
+    meta_sim(b = b_vals, use_cache = TRUE, n_pop = n_pop, ...)
+    #port.x <- rowSums(sims_out[[k]]$A[-burn, ])
+    #ret.x <- diff(log(port.x))
+    #var.x <- var(ret.x)
+    #mean.x <- mean(ret.x)
+    #port_vals[k, ] <- c(mean.x, var.x)
+  }
+
+#return(list(port_vals = port_vals, n_sims = 500, sims_out = sims_out))
+sims_out
+}                                                                                          
+
+  
+
+

### 2013-03-28 R/ricker_v_t.R

@@ -8,7 +8,11 @@
 #' x <- seq(1, 100)
 #' v_t <- as.numeric(arima.sim(n = 100, model = list(ar = 0.3), sd = 0.1, mean = 0))
 #' plot(x, ricker_v_t(spawners = x, a = 1.1, b = 60, v_t))
-ricker_v_t <- function(spawners, a, b, v_t) {
-  spawners * exp(a * (1 - spawners / b) + v_t)
+#ricker_v_t <- function(spawners, a, b, v_t) {
+  #spawners * exp(a * (1 - spawners / b) + v_t)
+#}
+
+ricker_v_t <- function(spawners, a, b, d, v_t) {
+  spawners^d * exp(a * (1 - spawners^d / b) + v_t)
 }
 

### 2013-04-09 inst/examples/salmonport.Rmd
at copy sent to committee before 2013-04-08

@@
 % Sean Anderson
+
+*Other title ideas:*  
+- Portfolio optimization reveals...  
+- something more general with animals instead of just salmon?  
+- Portfolio theory reveals conservation strategies that...  
+- Intelligent tinkering with ecological portfolios: patterns of environmental response diversity drive salmon conservation priorities  
+- Keeping every cog and wheel...
 
@@
 #or for rapid development:
-metafolio_files <-  list.files(path = "../../R/")
-for(i in 1:length(metafolio_files)){
-  source(paste("../..//R/", metafolio_files[i], sep = ""));rm(i)}
-
+library(devtools)
+load_all("../../") # assuming you're 2 folders deeper than the package
 source("plot_env_ts.R")
@@
 Fraction of fish that stray from natal streams & $f_{stray}$ & 0.02 [0, 0.10] \\
-Exponential rate of decay of straying with distance & $m$ & 0.30 [0.05, 0.50] \\
+Exponential rate of decay of straying with distance & $m$ & 0.3 [0.05, 0.5] \\
 Standard deviation of beta distribution for implementation error & $\sigma_{impl}$ & 0.05 [0, 0.20] \\
@@
 
-```{r fig.width=6, fig.height=3.75, fig.cap="Example Ricker curves for the minimum-variance metapopulation."}
+```{r fig.width=6, fig.height=3.75, fig.cap="Example Ricker curves for the minimum-variance metapopulation. (Ignore right now)"}
 plot_rickers(min_var_sim, pal = col_pal)
@@
 ```{r plot-weights-median, fig.width=4.2, fig.height=3.2, cache = FALSE, fig.cap="Median (solid line) and 25th and 75th quantiles of the weights across 40 minimum-variance portfolios.
-Just another way of looking at it."}
+Just another way of looking at it.
+(Nothing to see here right now.)"}
 source("min_var_ports_dist.R")
@@
 Notice what happens to the two *half* conservation strategies.
-A more realistic scenario combines the two.
+A more realistic scenario combines the two kinds of environmental noise.
 The efficient frontier is shown for the displayed portfolios.
@@
 
-# Take home messages
+# Some thoughts on the results
 1. If the dominant environmental signal is a long-term shift, and you keep around one half of the response diversity, then you get different portfolio means but the same variances.
@@
 
+\includepdf{salmonport-mindmap.pdf}
+
 # References #

### 2013-04-09 inst/examples/salmonport.tex
copy sent to committee before 2013-04-08
updating whole project

@@
 %\doublespacing
+\usepackage{pdflscape}
+\usepackage{pdfpages}
 
@@
 \maketitle
+
+\emph{Other title ideas:}\\- Portfolio optimization reveals\ldots{}\\-
+something more general with animals instead of just salmon?\\- Portfolio
+theory reveals conservation strategies that\ldots{}\\- Intelligent
+tinkering with ecological portfolios: patterns of environmental response
+diversity drive salmon conservation priorities\\- Keeping every cog and
+wheel\ldots{}
 
@@
 \includegraphics{figure/unnamed-chunk-1.pdf}
-\caption{Example Ricker curves for the minimum-variance metapopulation.}
+\caption{Example Ricker curves for the minimum-variance metapopulation.
+(Ignore right now)}
 \end{figure}
@@
 across 40 minimum-variance portfolios. Just another way of looking at
-it.}
+it. (Nothing to see here right now.)}
 \end{figure}
@@
 minimal short-term signal. Notice what happens to the two \emph{half}
-conservation strategies. A more realistic scenario combines the two. The
-efficient frontier is shown for the displayed portfolios. This isn't
-necessarily the overall efficient frontier.}
+conservation strategies. A more realistic scenario combines the two
+kinds of environmental noise. The efficient frontier is shown for the
+displayed portfolios. This isn't necessarily the overall efficient
+frontier.}
 \end{figure}
@@
 
-\section{Take home messages}
+\section{Some thoughts on the results}
 
@@
 
+\includepdf{salmonport-mindmap.pdf}
+
 \renewcommand\refname{References}

### 2013-04-09 inst/examples/salmportms.Rmd
copy sent to committee before 2013-04-08
updating whole project

% Portfolio prioritization of salmon conservation
% Sean C. Anderson and others
%

*Authors:*  
  Jonathan W. Moore  
  Trevor A. Branch  
  Andrew B. Cooper  
  Nicholas K. Dulvy  
  Maybe we should bring Michelle McClure back in

*Other title ideas:*  
- Portfolio optimization reveals...  
- something more general with animals instead of just salmon?  
- Portfolio theory reveals conservation strategies that...  
- Intelligent tinkering with ecological portfolios: patterns of environmental response diversity drive salmon conservation priorities  
- Keeping every cog and wheel...

  see Table A1 in "Viable Salmonid Populations and the Recovery of Evolutionarily Significant Units":
  straying rates from the literature for Pacific salmon. range from 0 to 90... usually around 4 or 5

# Introduction #

managing risk fundamental to conservation of metapopulations

portfolios provide another way of managing risk - explain mean-variance optimization and how this translates to ecology
[@Figge2004]
[@Hoekstra2012] 
[@Ando2011,@Ando2012] 
[@Markowitz1952,@Markowitz1959]

we use salmon as an example, describe the components of the example, provide references on metapopulation existence
[@Schindler2010]

research questions, and briefly our approach - Monte Carlo simulation modelling, as simple a model as possible while retaining important realistic aspects of salmon metapopulations

# Methods #

Data - if we're basing this at all on reality... probably not

## Simulation model components

general overview

## Salmon metapopulation dynamics ##

- spawner-recruit relationship paragraph
- straying paragraph

## Environmental dynamics ##

## Environmental response diversity ##
Scenarios? More variable are also more productive?...

## Portfolio optimization ##

## Conservation rules of thumb ##

# Results #

## Base case ##

## Effect of X on ... ##

## Effect of Y on ... ##

# Discussion #

# Acknowledgements #

# References #

# Tables #

1. Simulation parameters, symbols, base and alternate values

# Figures #

1. Flow chart of Monte Carlo procedure
2. Climate scenarios
3. Climate-productivity curves (response-diversity scenarios)
4. Time series components of an example simulation
5. Effect on returns of ramping up or down each main parameter
6. Efficient frontier given various climate scenarios
7. Rules of thumb (in the spirit of VSP guidelines) in risk-reward space
8. Potentially a plot that shows how some summary metric of the efficient frontier or rules of thumb for high and low scenarios of important simulation parameters

# Supplementary code #

The `metafolio` `R` package and documentation.

# Supplementary figures #

1. Straying matrix
2. Efficient frontiers and/or rules-of-thumb plots varying each major simulation variable up and down

### 2013-04-09 R/fit_ricker.R

@@ -1,0 +1,15 @@
+#' Fit a Ricker curve
+
+#S <- seq(100, 1000, length.out = 25)
+#R <- sapply(S, function(x) ricker_ar1(spawners = x, a = 1.9, b = 900, sigma_sq_v = 0.01)$recruits)
+#plot(S, log(R/S))
+
+fit_ricker <- function(S, R) {
+  m <- lm(log(R/S) ~ S)
+  m.coef <- coef(m)
+  a <- as.numeric(m.coef[1])
+  b <- -as.numeric(a/m.coef[2])
+  return(list(a = a, b = b))
+}
+
+

### 2013-04-09 R/get_port_vals.R

@@ -1,0 +1,10 @@
+#' 
+
+get_port_vals <- function(x, burn = 1:30) {
+    port.x <- rowSums(x$A[-burn, ])
+    ret.x <- diff(log(port.x))
+    var.x <- var(ret.x)
+    mean.x <- mean(ret.x)
+    data.frame(m = mean.x, v = var.x)
+}
+

### 2013-04-09 R/impl_error.R

@@ -1,0 +1,24 @@
+#' Add implementation error
+
+# From Pestes et al. 2008:
+
+#portion harvested between 0 and 1:
+#Hact,yr = beta(βh, γh), (1)
+#where values of βh and γh were the beta distribution’s lo- cation and shape parameters, with βh = Htar,yr for a given year and γh = 0.1 (the latter determined by numerical ex- periment to approximate the observed implementation uncertainty).
+
+impl_error <- function(mu, sigma_impl) {
+  var <- sigma_impl ^ 2
+
+est_beta_params <- function(mu, var) {
+  alpha <- ((1 - mu) / var - 1 / mu) * mu ^ 2
+  beta <- alpha * (1 / mu - 1)
+  return(list(alpha = alpha, beta = beta))
+}
+
+bp <- est_beta_params(mu, var)
+#print(bp)
+#hist(rbeta(10000, bp$alpha, bp$beta))
+#rbeta(1, bp$alpha, bp$beta)
+mapply(rbeta, n = 1, shape1 = bp$alpha, shape2 = bp$beta)
+}
+

### 2013-04-09 inst/examples/bug.sh

@@ -1,0 +1,7 @@
+#!/bin/sh
+#PBS -l walltime=00:60:00
+#PBS -M sean_anderson@sfu.ca
+#PBS -l mem=800mb
+
+R CMD BATCH --no-save /home/anderson/Dropbox/ms/metafolio/inst/examples/bugaboo.R
+

### 2013-04-09 inst/examples/extract_bib.rb

@@ -1,0 +1,10 @@
+#!/usr/bin/env ruby
+# from https://gist.github.com/dsanson/1383132
+ 
+bibfile = "~/Dropbox/tex/ref2.bib"
+ 
+matches = ARGF.read.scan(/@(.*?)[\.,:;\] ]/)
+reg = "\\(" + matches.join("\\)\\|\\(") + "\\)"
+ 
+system 'bibtool --preserve.key.case=on -X \'' + reg + '\' ' + bibfile
+

### 2013-04-09 inst/examples/cons_plans_n.R
fix description, correct data loaded

@@ -1,8 +1,6 @@
-#' A wider variety of conservation plans; also try and better mimic
-#' what VSP guidelines do and do not suggest
+#' show difference of increasing the number of streams you maintain
 
-# first show difference of increasing the number of streams you
-# maintain:
+USE_CACHE <- TRUE
 
 require(plyr)
 num_pops <- c(2, 4, 8, 16, 32)
@@ -16,7 +14,7 @@ plans_name <- num_pops
 w_show <- seq_len(length(num_pops))
 
 if(USE_CACHE) {
-  load("plans_mv.rda")
+  load("plans_mv_n.rda")
 }else{
   plans_mv_arma_n <- list()
   plans_mv_linear_n <- list()

### 2013-04-09 inst/examples/salmonport.Rmd
add effect of different numbers of subpopulations on MV plot

@@
 
+```{r plot-cons-plans-n-mv, fig.width=6.5, fig.height=3.5, fig.cap="Conservation plans in mean-variance space.
+*This version shows conservation plans with balanced response diversity but different numbers of populations.** The dots show portfolios from individual simulation runs.
+The dark and light shaded areas show 25% and 75% quartiles.
+The left panel shows a scenario with a stationary environmental signal and strong short-term environmental noise.
+The right panel shows a scenario with a linearly-increasing environmental signal and minimal short-term environmental noise.
+These two scenarios bound the two environmental extremes: strong short-term noise with minimal long-term signal vs.
+a strong long-term signal with minimal short-term signal.
+Notice what happens to the two *half* conservation strategies.
+A more realistic scenario combines the two kinds of environmental noise.
+The efficient frontier is shown for the displayed portfolios.
+This isn't necessarily the overall efficient frontier.", cache=FALSE}
+source("cons_plans_n.R")
+```
+
 \clearpage

### 2013-04-10 inst/examples/salmonport.Rmd
modificiations right before 2013-04-10 meeting

--- 2013-04-09 inst/examples/salmonport.Rmd
@@
 - Keeping every cog and wheel...  
+- A yin for every yang...
 
@@
 library(devtools)
-load_all("../../") # assuming you're 2 folders deeper than the package
+metafolio_files <-  list.files(path = "../../R/")
+for(i in 1:length(metafolio_files)){
+   source(paste("../..//R/", metafolio_files[i], sep = ""));rm(i)}
+#load_all("../../") # Not working with knitr?
 source("plot_env_ts.R")
@@
 The colours are the same as earlier and identify the different populations.
-These are in the same order as they are referred to in the next two plots: full range of environmental response diversity, most stable only, one half, other half, most asynchronous.", cache=FALSE}
+These are in the same order as they are referred to in the next two plots: full range of environmental response diversity, most stable only, one half, other half, most asynchronous."}
 #Notes #
 #-b is "unfished equilibrium stock size"
+col_pal <- rev(gg_color_hue(10))
 source("plot_thermal_curves_scenarios.R")
@@
 
-```{r plot-cons-plans-mv, fig.width=6.5, fig.height=3.5, fig.cap="Conservation plans in mean-variance space.
+\begin{figure}[htbp]
+  \centering
+    \includegraphics[height=5.0in]{thermal-curve-scenarios.pdf}
+ \caption{Different ways of prioritizing conservation.
+The curves show the environmental-tolerance curves for the populations that are conserved.
+The colours are the same as earlier and identify the different populations.
+These are in the same order as they are referred to in the next two plots: full range of environmental response diversity, most stable only, one half, other half, most asynchronous.}
+  \label{fig:sim-diagram}
+\end{figure}
+
+```{r plot-cons-plans-mv, fig.width=6.2, fig.height=3.3, fig.cap="Conservation plans in mean-variance space.
 The dots show portfolios from individual simulation runs.
@@
 The efficient frontier is shown for the displayed portfolios.
-This isn't necessarily the overall efficient frontier.", cache=FALSE}
+This isn't necessarily the overall efficient frontier.\\label{fig:cons-plans-mv}", cache=FALSE}
 #Notes #
@@
 source("cons_plans_ts.R")
+```
+
+```{r plot-cons-plans-n-mv, fig.width=6.2, fig.height=3.3, fig.cap="Conservation plans in mean-variance space.
+**This version shows conservation plans with balanced response diversity but different numbers of populations.** Notice how the left panel (strong short-term environmental variability) and the right panel (strong long-term environmental shift) look approximately the same.
+This illustrates (in this contrived example) how the number of populations decreases metapopulation variance through statistical averaging of non-systematic sources of stochasticity.
+The perfectly balanced environmental tolerances counteract the environmental fluctuations.
+In the real world, you wouldn't know the exact environmental tolerances, and so conserving more populations might increase your chances of retaining populations with a diversity of environmental tolerances.
+The orange blob in this figure matches the ``Full range of responses'' blob in Fig.\\ \\ref{fig:cons-plans-mv}.
+The dots show portfolios from individual simulation runs.
+The dark and light shaded areas show 25% and 75% quartiles.", cache=FALSE}
+source("cons_plans_n.R")
 ```

### 2013-04-10 inst/examples/salmonport.tex
modificiations right before 2013-04-10 meeting

@@
 diversity drive salmon conservation priorities\\- Keeping every cog and
-wheel\ldots{}
+wheel\ldots{}\\- A yin for every yang\ldots{}
 
@@
   \centering
-\includegraphics{figure/plot-cons-plans-curves.pdf}
-\caption{Different ways of prioritizing conservation. The curves show
-the environmental-tolerance curves for the populations that are
-conserved. The colours are the same as earlier and identify the
-different populations. These are in the same order as they are referred
-to in the next two plots: full range of environmental response
-diversity, most stable only, one half, other half, most asynchronous.}
+    \includegraphics[height=5.0in]{thermal-curve-scenarios.pdf}
+ \caption{Different ways of prioritizing conservation.
+The curves show the environmental-tolerance curves for the populations that are conserved.
+The colours are the same as earlier and identify the different populations.
+These are in the same order as they are referred to in the next two plots: full range of environmental response diversity, most stable only, one half, other half, most asynchronous.}
+  \label{fig:sim-diagram}
 \end{figure}
@@
 displayed portfolios. This isn't necessarily the overall efficient
-frontier.}
+frontier.\label{fig:cons-plans-mv}}
 \end{figure}
@@
 unbalanced response diversity tend to seesaw with the environment.}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figure/plot-cons-plans-n-mv.pdf}
+\caption{Conservation plans in mean-variance space. \textbf{This version
+shows conservation plans with balanced response diversity but different
+numbers of populations.} Notice how the left panel (strong short-term
+environmental variability) and the right panel (strong long-term
+environmental shift) look approximately the same. This illustrates (in
+this contrived example) how the number of populations decreases
+metapopulation variance through statistical averaging of non-systematic
+sources of stochasticity. The perfectly balanced environmental
+tolerances counteract the environmental fluctuations. In the real world,
+you wouldn't know the exact environmental tolerances, and so conserving
+more populations might increase your chances of retaining populations
+with a diversity of environmental tolerances. The orange blob in this
+figure matches the ``Full range of responses'' blob in
+Fig.~\ref{fig:cons-plans-mv}. The dots show portfolios from individual
+simulation runs. The dark and light shaded areas show 25\% and 75\%
+quartiles.}
 \end{figure}

### 2013-04-10 inst/examples/plot_thermal_curves_scenarios.R

@@ -1,6 +1,7 @@
 #' plot thermal curves for different conservation scenarios
 
-par(cex = 0.6, mgp = c(2.4, 0.65, 0), tck = -0.02, mfrow = c(5,1), mar = c(0,5,0,.5), oma = c(4, 0, .5, 0))
+pdf("thermal-curve-scenarios.pdf", width = 3.3, height = 5.0)
+par(cex = 0.6, mgp = c(2.4, 0.55, 0), tck = -0.03, mfrow = c(5,1), mar = c(0,5,0,.5), oma = c(4, 0, .5, 0))
 
 optim_temps <- seq(13, 19, length.out = 10)
 
@@ -30,3 +31,4 @@ mtext(expression(Temperature~(degree*C)), side = 1, outer = FALSE, line = 2.5, l
 #abline(v = c(16-3.5, c = 16+3.5), lty = 3)
 
 
+dev.off()

### 2013-05-01 inst/examples/salmonport.Rmd
Bankruptcy commit

@@
 ```{r env-ts, fig.height=4, fig.width=4, cache = FALSE, fig.cap="Example temperature time series\\label{fig:env-ts}"}
-#library(metafolio)
-#or for rapid development:
+
 library(devtools)
+
+RELOAD_R_PACKAGE <- FALSE
+
+if(RELOAD_R_PACKAGE) {
+  document("../../")
+  build("../../")
+  if("metafolio" %in% loaded_packages()) 
+    detach("package:metafolio", unload = TRUE)
+  install.packages("~/Dropbox/ms/metafolio_0.1.tar.gz", type = "source", repos = NULL, INSTALL_opts = "--byte-compile")
+  library(metafolio)
+}
+
 metafolio_files <-  list.files(path = "../../R/")
@@
    source(paste("../..//R/", metafolio_files[i], sep = ""));rm(i)}
-#load_all("../../") # Not working with knitr?
+
+#load_all("../../")
+
 source("plot_env_ts.R")

### 2013-05-01 R/create_asset_weights.R

@@ -1,4 +1,5 @@
-# create asset weights matrix for portfolios
+#' create asset weights matrix for portfolios
+#' @export
 create_asset_weights <- function(
   n_pop, # number of subpops
   n_sims, # number of simulations

### 2013-05-01 R/generate_env_ts.R

@@ -7,6 +7,7 @@
 #' @param regime_params Parameters controlling regime-shift time series.
 #' @param linear_params Parameters controlling warming or cooling time series.
 #' @param constant_params Parameter controlling constant time series.
+#' @export
 #' @examples
 #' types <- c("sine", "arma", "regime", "linear", "constant")
 #' x <- list()

### 2013-05-01 R/get_port_vals.R

@@ -1,4 +1,7 @@
+#' Get portfolio mean and variance values from a meta_sim() object 
 #' 
+#' @param x A list object as returned from \code{meta_sim}
+#' @param burn Number of years to throw out at burn in
 
 get_port_vals <- function(x, burn = 1:30) {
     port.x <- rowSums(x$A[-burn, ])

### 2013-05-01 R/gg_color_hue.R

@@ -6,6 +6,7 @@
 #' @param l Luminance in the range [0,100]
 #' @param c Chroma of the colour. 
 #' @details See the \code{hcl} function for details.
+#' @export
 #' @examples
 #' gg_color_hue(10)
 #' 

### 2013-05-01 R/meta_sim.R

@@ -152,8 +152,8 @@ meta_sim <- function(
           #if(j == 1) plot(spawners, log(recruits/spawners), main = i)
           #if(j == 1) print(rick)
           # bounds for sanity:
-          if(rick$a > 3) rick$a <- 3
-          if(rick$a < 0.05) rick$a <- 0.05
+          if(rick$a > 4) rick$a <- 4
+          if(rick$a < 0.02) rick$a <- 0.02
           if(rick$b > Est_b[i - 1, j] * 1.5) rick$b <- Est_b[i - 1, j] * 1.5 # at most increase b by 50% 
           if(rick$b < Est_b[i - 1, j] * 0.5) rick$b <- Est_b[i - 1, j] * 0.5 # at most decrease b by 50% 
           Est_a[i,j]<-rick$a

### 2013-05-01 R/monte_carlo_portfolios.R

@@ -1,5 +1,6 @@
 #' Monte Carlo the weights into portfolios and record the simulation
 # output and portfolio metrics (mean and variance for now)
+#' @export
 
 monte_carlo_portfolios <- function(weights_matrix, n_sims = 500, mean_b = 1000, burn = 1:30, ...) {
 

### 2013-05-01 R/plot_correlation_between_returns.R

@@ -3,6 +3,7 @@
 #' @param x A list output object from \code{meta_sim}.
 #' @param burn Number of years to discard at start as burnin.
 #' @param pal Colours to plot with for stocks.
+#' @export
 
 plot_correlation_between_returns <- function(x, burn = 1:30, pal = gg_color_hue(x$n_pop)) {
   # x is output from a simulation run

### 2013-05-01 R/plot_efficient_portfolios.R

@@ -1,4 +1,5 @@
 #' Basic efficient portfolio plots and return efficient portfolio data.
+#' @export
 
 plot_efficient_portfolios <- function(port_vals, weights_matrix, pal, plot = TRUE) {
 

### 2013-05-01 R/plot_rickers.R

@@ -6,6 +6,7 @@
 #' @param add_y_axes_pops Panels to add y axes on
 #' @param add_x_axes_pops Panels to add x axes on
 #' @param burn Number of initial years to throw out as burnin
+#' @export
 #' 
 plot_rickers <- function(
   x,  # sim output list

### 2013-05-01 R/plot_sim_ts.R

@@ -4,6 +4,7 @@
 #' @param years_to_show How many years to plot after the burn in period.
 #' @param burn The number of years to discard as burnin a the beginning of the time 
 #' series.
+#' @export
 
 plot_sim_ts <- function(x, pal = rev(gg_color_hue(x$n_pop)), years_to_show = 30, burn = 1:50) {
 # x is output from a simulation run; it's a list

### 2013-05-01 R/ricker_ar1.R

@@ -1,4 +1,5 @@
 # Ricker model with AR1 error structure
+#' @export
 
 ricker_ar1 <- function(spawners, a, b, sigma_sq_v = 0, v_t_minus1 = 0, rho = 0) {
   v_t <- rnorm(1, mean = 0, sd = sqrt(sigma_sq_v)) # process noise

### 2013-05-01 R/ricker_escapement.R

@@ -5,6 +5,7 @@
 #' @param a Ricker productivity parameter.
 #' @param b Ricker density-dependent parameter.
 #' @references Hilborn, R.W. and Walters, C. 1992. Quantitative fisheries stock assessment: Choice, dynamics, and uncertainty. Chapman and Hall, London.
+#' @export
 #' @examples
 #' ricker_escapement(1.1, 1000)
 ricker_escapement <- function(a, b) {

### 2013-05-01 R/ricker_v_t.R

@@ -4,6 +4,7 @@
 #' @param a Ricker productivity paramater. Recruits are e^a at the origin.
 #' @param b Ricker density dependent parameter.
 #' @param v_t Residual on the curve. Will be exponentiated.
+#' @export
 #' @examples
 #' x <- seq(1, 100)
 #' v_t <- as.numeric(arima.sim(n = 100, model = list(ar = 0.3), sd = 0.1, mean = 0))

### 2013-05-01 R/thermal_curve_a.R

@@ -11,6 +11,7 @@
 #' @param width_param A parameter to control the width of the
 #' parabola. Smaller numbers make wider parabolas.
 #' @return A productivity parameter given the location on a thermal tolerance curve.
+#' @export
 #' @examples
 #' x <- seq(5, 30, length.out = 200)
 #' plot(x, thermal_curve_a(x), ylab = "a", xlab = "Temperature", type = "l")

### 2013-05-01 inst/examples/optim.R

@@ -1,0 +1,30 @@
+#' 
+
+
+
+f2 <- function(b = rep(1000, 9)) {
+  if(sum(b) > 10000){
+    r<- Inf
+  }else{
+    b <- c(b, 10000-sum(b))
+    p <- meta_sim(b = b,
+      n_pop = length(b),
+      env_type = "linear",
+      env_params = list(min_value = 10, max_value = 20, sigma_env = 0.2),
+      use_cache = TRUE,
+      start_assessment = 20,
+      assessment_window = 50,
+      sigma_impl = 1e-9
+      )
+    mv <- get_port_vals(p)
+    r <- -log(as.numeric(mv[2]))
+  }
+  return(r)
+}
+
+x<-f2()
+
+library(DEoptim)
+
+j <- DEoptim(f2, lower = rep(10, 9), upper = rep(9999, 9), control = list(itermax = 20))
+plot(j)

### 2013-05-02 inst/examples/cons_plans_n.R
Remove linear-trend plot (by commenting out)

@@ -9,38 +9,54 @@ w <- list()
 for(i in 1:length(num_pops)) {
   w[[i]] <- matrix(rep(1, num_pops[i]) * (tot_abund / num_pops[i]), nrow = 1)
 }
-
 plans_name <- num_pops
-w_show <- seq_len(length(num_pops))
+
+#### in this version, the pops are wiped out; total abundance changes
+w <- list()
+
+for(i in 1:4) {
+  w[[i]] <- matrix(rep(625, 16), nrow = 1)
+}
+  
+w[[1]][-c(1, 16)] <- 5 # 2 pops
+w[[2]][-c(1,2,15,16)] <- 5 # 4 pops
+w[[3]][-c(1:4,13:16)] <- 5 # 8 pops
+plans_name <- num_pops[1:4]
+####
+
+w_show <- seq_len(length(w))
 
 if(USE_CACHE) {
   load("plans_mv_n.rda")
 }else{
   plans_mv_arma_n <- list()
-  plans_mv_linear_n <- list()
-  for(i in 1:length(num_pops)) {
-    #plans_mv_arma_n[[i]] <- get_conserv_plans_mv(
-      #weights = w[[i]], 
-      #env_type = "arma",
-      #env_params = list(mean_value = 16, ar = 0.4, sigma_env = 1.4, ma = 0)
-      #)
-    plans_mv_linear_n[[i]] <- get_conserv_plans_mv(
+  #plans_mv_linear_n <- list()
+  for(i in 1:length(w)) {
+    plans_mv_arma_n[[i]] <- get_conserv_plans_mv(
       weights = w[[i]], 
-      env_type = "linear",
-      env_params = list(min_value = 10, max_value = 20, sigma_env = 0.2))
-    print(paste("Completed", i, "of", length(num_pops), "conservation plans to evaluate"))
+      reps = 30,
+      env_type = "arma",
+      env_params = list(mean_value = 16, ar = 0.4, sigma_env = 1.4, ma = 0)
+      )
+    #plans_mv_linear_n[[i]] <- get_conserv_plans_mv(
+      #weights = w[[i]], 
+      #reps = 30,
+      #env_type = "linear",
+      #env_params = list(min_value = 10, max_value = 20, sigma_env = 0.2))
+    print(paste("Completed", i, "of", length(w), "conservation plans to evaluate"))
   }
 
[…]

### 2013-05-11 inst/examples/salmonport.tex
Update md and tex after makefile run

@@
 
-\emph{Other title ideas:}\\- Portfolio optimization reveals\ldots{}\\-
-something more general with animals instead of just salmon?\\- Portfolio
-theory reveals conservation strategies that\ldots{}\\- Intelligent
+\emph{Other title ideas:} - Portfolio optimization reveals\ldots{} -
+something more general with animals instead of just salmon? - Portfolio
+theory reveals conservation strategies that\ldots{} - Intelligent
 tinkering with ecological portfolios: patterns of environmental response
-diversity drive salmon conservation priorities\\- Keeping every cog and
-wheel\ldots{}\\- A yin for every yang\ldots{}
+diversity drive salmon conservation priorities - Keeping every cog and
+wheel\ldots{} - A yin for every yang\ldots{}
 

### 2013-05-20 inst/ms/ms.tex
Add main ms and SOM files and folder

\emph{Other title ideas:}

\begin{itemize}
\item
  Portfolio optimization reveals\ldots{}
\item
  something more general with animals instead of just salmon?
\item
  Portfolio theory reveals conservation strategies that\ldots{}
\item
  Intelligent tinkering with ecological portfolios: patterns of
  environmental response diversity drive salmon conservation priorities
\item
  Keeping every cog and wheel\ldots{}
\end{itemize}

\section{Abstract}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Managing risk is fundamental to the conservation of an endangered
  species. When an endangered species exists as a metapopulation, we can
  manage risk at two levels: at the population level or at the
  metapopulation level. Whereas risk is typically managed at the
  population level, a portfolio approach to managing risk might consider
  how conservation affects the ``weight'' of each population in a
  metapopulation ``portfolio''.
\item
  Here, we ask how a portfolio approach to managing risk can inform the
  spatial conservation of metapopulations in a changing world. To answer
  this, we develop a salmon metapopulation simulation in which
  population-specific productivity is driven by spatially-distributed
  environmental tolerance and patterns of short- and long-term
  environmental change. We then implement different spatial conservation
  ``rules of thumb'' that control the population-specific carrying
  capacities and evaluate the salmon portfolios along risk and return
  axes, similarly to how financial portfolios are assessed.
\item
  Our results show, first, that maintaining populations with a variety
  of environmental tolerances gives the best chance at an efficient
  ecological portfolio --- minimizing metapopulation variance while
  maximizing metapopulation growth rate. This finding emphasizes the
  risk of allowing large spatial blocks of habitat destruction, say
  through the development of dams. Second, we show that focusing on
  well-performing stocks now at the detriment of others is at best
  equivalent to a risky but efficient portfolio and is more likely a
  risky and inefficient portfolio --- it neither minimizes
  metapopulation variance nor maximizes growth rate compared to other
  strategies. Third, we show that maintaining more populations reduces
  metapopulation risk for the same spatial conservation strategy. Given
  a lack of knowledge of how populations respond to the environment, the
  most risk-averse approach is to conserve as many populations as
  possible.
\item
  Our findings highlight three key points: (1) the conservation priority
  of maintaining biocomplexity and therefore environmental response
  diversity,
\end{enumerate}

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\setcounter{enumi}{1}
\itemsep1pt\parskip0pt\parsep0pt
\item
  the research priority of identifying differences in environmental
  tolerance given predicted environmental changes, and (3) the utility
  of considering risk for groups of fish stocks --- especially given
  environmental, biological, and implementation uncertainty --- through
  the lens of portfolio theory.
\end{enumerate}

\section{Introduction}

Managing risk fundamental to conservation of metapopulations

Portfolios provide another way of managing risk - explain mean-variance
optimization and how this translates to ecology \citep{Figge2004}
\citep{Hoekstra2012} \citep[\citet{Ando2012}]{Ando2011}
\citep[\citet{Markowitz1959}]{Markowitz1952}

We use salmon as an example, describe the components of the example,
provide references on metapopulation existence \citep{Schindler2010}

Research questions, and briefly our approach - Monte Carlo simulation
modelling, as simple a model as possible while retaining important
realistic aspects of salmon metapopulations

\section{Methods}

\subsection{Simulation model components}

general overview

See Figure \ref{f:sim-flow} for an illustration of the simulation
structure.

\subsection{Salmon metapopulation dynamics}

We are using a Ricker curve,

\begin{equation}
R_{ti} = S_{ti}e^{a_{ti}(1-S_{ti}/b_i) + w_{ti}},
\end{equation}

\noindent
where $t$ represents a generation time, $i$ represents a population, $R$
is the number of returns, $S$ is the number of spawners, $a$ is the
productivity parameter (which can vary with the environmental signal),
and $b$ is the density-dependent term (which is used as the asset
weights in the portfolios). The term $w_{ti}$ represents first-order
autocorrelated error (AR1). Formally,
$w_{ti} = w_{ti-1} \rho_w + v_{ti}$, where $v_{ti}$ represents
independent and normally distributed error with mean 0 and standard
deviation of $\sigma_v$. The parameter $\rho_w$ represents the
correlation between residuals from subsequent generations.

@cooper1999

We implemented straying as in \citet{Cooper1999}. We generate a matrix
that represents the fraction of straying between any two populations. We
have set up the metapopulation (thus far) in a very simple scenario: the
populations are arranged in a line and those that are nearer to each
other are more likely to stray between each other {[}insert caveats{]}.
Two parameters control the straying: the fraction of fish $f_{stray}$
that stray from their natal stream in any given generation and the rate
at which this straying between streams decays with distance $m$. We
calculated the number of salmon straying from stream $j$ to stream $i$
as,

\begin{equation}
  \mathrm{strays}_{tij} = f_{stray} R_{tj}
    \frac{e^{-m \lvert i-j \rvert }}
      {\displaystyle\sum\limits_{\substack{k = 1 \\
    k \neq j}}^{n} e^{-m \lvert k-j \rvert }},
  \label{eq:stray}
\end{equation}

\noindent
where $R_{tj}$ is the number of returning salmon at generation $t$ whose
natal stream was stream $j$. The subscript $k$ represents a stream ID
and $n$ the number of populations. The denominator is a normalizing
constant to ensure the desired fraction of fish stray. See Figure
\ref{f:stray} for an example straying matrix.

\subsection{Environmental dynamics}

\subsection{Environmental response diversity}

Scenarios? More variable are also more productive?\ldots{}

\subsection{Portfolio optimization}

\subsection{Conservation rules of thumb}

\section{Results}

Fig. \ref{f:eg-n-linear}

Fig. \ref{f:eg-n-arma}

Fig. \ref{f:eg-sp-arma}

Fig. \ref{f:eg-sp-linear}

\subsection{Base case}

\subsection{Effect of X on \ldots{}}

\subsection{Effect of Y on \ldots{}}

\section{Discussion}

\section{Acknowledgements}

\section{References}

\bibliographystyle{apalike}

\bibliography{jshort,ms}

\clearpage

\section{Tables}

\begin{table}[h!]
\centering
\small
\caption{Components of salmon metapopulation portfolios}
\begin{tabular}{p{3.6cm}p{7.5cm}}
\toprule
Component & Definition for the salmon portfolio\\
\midrule
Assets             & Stream-level salmon populations; possibly a Viable
                     Salmonid Population\\
Portfolio          & The salmon metapopulation; possibly an Evolutionarily
                     Significant Unit\\
Portfolio managers & Salmon managers\\
Investors          & Salmon managers, conservation agency, or salmon fishers\\
Asset weights      & Carrying capacity (specifically the $b$ parameter in
                     a Ricker model)\\
Asset returns      & Rate of change of generation-to-generation salmon
                     metapopulation abundance\\
Asset risk         & Variance of generation-to-generation salmon metapopulation
                     abundance\\
\bottomrule
\end{tabular}
\label{tab:port-components}
\end{table}

\clearpage

\begin{table}[h!]
\centering
\small
\caption{Salmon metapopulation parameters with base case and alternate values.}
\begin{tabular}{p{7.0cm}p{1.6cm}p{3.2cm}}
\toprule
Description                                                      & Parameter       & Base [lower, upper] \\
\midrule
Stock-recruit residual standard deviation (on log scale)         & $\sigma_v$      & 0.30 [0.05, 0.50] \\
First order (AR1) serial correlation of stock-recruit residuals  & $\rho_w$        & 0.40 [0, 0.80] \\
Fraction of fish that stray from natal streams                   & $f_{stray}$     & 0.02 [0, 0.10] \\
Exponential rate of decay of straying with distance              & $m$             & 0.3 [0.05, 0.5] \\
Standard deviation of beta distribution for implementation error & $\sigma_{impl}$ & 0.05 [0, 0.20] \\
Frequency of assessment (years)                                  & $f_{assess}$    & 20 [5, 50] \\
\bottomrule
\end{tabular}
\label{tab:salm-pars}
\end{table}

\clearpage

\section{Figures}

`

\begin{figure}[htbp]
\centering
\includegraphics[height=5.5in]{../examples/simulation-diagram2.pdf}
\caption{Flow chart of the salmon-metapopulation simulation. There are $n$ 
salmon populations and $t$ generations. Blue text indicates values that are 
generated before the simulation progresses through time. Red text indicates 
steps in which calculations are performed. Black text indicates values that are 
calculated. Grey text indicates parameters that can be set. Green text 
indicates the looping structure of the simulation.}
\label{f:sim-flow}
\end{figure}

`

\begin{figure}[htbp]
\centering
\includegraphics[width=4.0in]{../examples/spatial-arma-sim.pdf}
\caption{The components of an example metapopulation simulation.  We show, from 
top to bottom, the environmental signal, the resulting productivity parameter 
(Ricker $a$), the salmon returns, fisheries catch, salmon escapement, salmon 
straying from their natal streams, salmon joining from other streams, 
stock-recruit residuals on a log scale, and the estimated $a$ and $b$ 
parameters in the fitted Ricker curve. The colored lines indicate populations 
that thrive at low (cool colours) to high (warm colours) environmental values.}
\label{f:sp-eg}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=3.0in]{../examples/thermal-curve-scenarios.pdf}
\caption{Different ways of prioritizing response-diversity conservation. The 
coloured lines represent environmental-tolerance curves for conserved 
populations. The colours identify populations that thrive at low (cool colours) 
to high (warm colours).}
\label{f:curves}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=5.0in]{../examples/spatial-mv.pdf}
\caption{The importance of preserving environmental response diversity through 
spatial conservation strategies. The conservation strategies correspond to 
Fig.~\ref{f:curves} and represent conserving a range of responses (green), the 
most stable populations only (orange), or one type of environmental response 
(purple and pink).  In risk-return space we show environmental scenarios that 
are comprised primarily of (a) short-term and (b) long-term environmental 
fluctuations (see Fig.~X). The dots show simulated metapopulations and the 
contours show 25\% and 75\% quantiles across 500 simulations per strategy. We 
also show example metapopulation abundance time series for the (c, e) 
short-term and (d, f) long-term  environmental-fluctuation scenarios.}
\label{f:sp-mv}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=5.0in]{../examples/cons-plans-n.pdf}
\caption{The importance of preserving as many salmon populations as possible 
when we don't know how response diversity is distributed. In risk-return space 
we show environmental scenarios that are comprised primarily of (a) short-term 
and (b) long-term environmental fluctuations (see Fig.~X). We show 
metapopulations in which 2 (red), 4 (orange), 8 (yellow), or 16 (green) 
populations of random response diversity are conserved. The dots show simulated 
metapopulations and the contours show 25\% and 75\% quantiles across 500 
simulations per strategy. We also show example metapopulation (c) 
rate-of-change and (d) abundance time series for a the short-term 
environmental-fluctuation scenario.}
\label{f:n-mv}
\end{figure}

### 2013-05-20 inst/ms/salmonfolios.tex
Add main ms and SOM files and folder

\documentclass[12pt]{article}
\usepackage{geometry}
\geometry{letterpaper}
\usepackage{amssymb,amsmath,amsfonts}
\usepackage{graphicx}

\usepackage{fontspec}
%\usepackage{pdfsync}
\defaultfontfeatures{Ligatures=TeX}
\setsansfont{Calibri}
\setmonofont{Inconsolata}
\setmainfont{Linux Libertine}
%\setmainfont{Times New Roman}
\usepackage[sectionbib]{chapterbib}

\usepackage{titlesec}
\titleformat*{\section}{\Large\bfseries\sffamily}
\titleformat*{\subsection}{\large\sc\sffamily}
\titleformat*{\subsubsection}{\itshape\sffamily}
\DeclareTextCommandDefault{\nobreakspace}{\leavevmode\nobreak\ }
\usepackage{setspace}
\usepackage{booktabs}
\bibliographystyle{apalike}
\setcounter{secnumdepth}{-2}

\usepackage[round]{natbib}
\bibliographystyle{apalike}
\bibpunct{(}{)}{;}{a}{}{;}

\title{Portfolio prioritization of salmon conservation }
\author{Sean C. Anderson$^{1*}$ \and Jonathan W. Moore$^1$ \and Trevor A. Branch
  \and Andrew B. Cooper$^2$ \and Nicholas K.  Dulvy$^1$}
\date{}

\begin{document}
%\onehalfspacing
\doublespacing
\maketitle

\noindent
  \textsuperscript{1}Department of Biological Sciences, Simon Fraser
  University, Burnaby BC, V5A 1S6, Canada

\noindent
	\textsuperscript{2}School of Resource and Environmental Management, Simon
  Fraser University, Burnaby, BC, V5A 1S6, Canada

\noindent
	\textsuperscript{*}Corresponding author: sean\_anderson@sfu.ca

\clearpage

\include{ms}

% redefine the command that creates the Table number.
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{table}{0} % reset counter

% redefine the command that creates the Figure number.
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\figurename}{Fig.}
\setcounter{figure}{0}  % reset counter

\include{som}

\end{document}

### 2013-05-20 inst/ms/som.tex
Add main ms and SOM files and folder

\section{Supporting online material}

\subsection{Supplementary code}

The \texttt{metafolio} \texttt{R} package and documentation.

\subsection{Supplementary figures}

\begin{figure}[htbp]
\centering
\includegraphics[width=4.0in]{../examples/figure/plot-various-options-ts-3pops.pdf}
\caption{The impact of increasing or decreasing various parameter values in the simulations. The different lines represent different salmon populations.}
\label{f:eg-sens}
\end{figure}

\clearpage

\begin{figure}[htbp]
\centering
\includegraphics[width=4.0in]{../examples/figure/stray-matrix.pdf}
\caption{ An example straying matrix. The rows and columns represent different populations (indicated by population number). Dark blue indicates a high rate of straying and light blue indicates a low rate of straying.  The straying matrix is generated through equation \ref{eq:stray}.}
\label{f:stray}
\end{figure}

\clearpage

\begin{figure}[htbp]
\centering
\includegraphics[width=4.5in]{../examples/thermal-curves.pdf}
\caption{The full range of environmental tolerance curves shown for 10 populations. The vertical dotted lines indicate the general range of environmental fluctuations in the main simulations, and the vertical dashed line indicates the mean environmental value in the main simulations.}
\label{f:all-curves}
\end{figure}

\clearpage

\begin{figure}[htbp]
\centering
\includegraphics[width=4.5in]{../examples/spatial-arma-sim.pdf}
\caption{Spatial and short-term environmental fluctuations}
\label{f:eg-sp-arma}
\end{figure}

\clearpage

\begin{figure}[htbp]
\centering
\includegraphics[width=4.5in]{../examples/spatial-linear-sim.pdf}
\caption{Spatial and long-term environmental fluctuations}
\label{f:eg-sp-linear}
\end{figure}

\clearpage

\begin{figure}[htbp]
\centering
\includegraphics[width=4.5in]{../examples/n-arma-sim.pdf}
\caption{Number and short-term environmental fluctuations}
\label{f:eg-n-arma}
\end{figure}

\clearpage

\begin{figure}[htbp]
\centering
\includegraphics[width=4.5in]{../examples/n-linear-sim.pdf}
\caption{Number and long-term environmental change}
\label{f:eg-n-linear}
\end{figure}

\clearpage

A placeholder: \citep{Schindler2010}

\bibliographystyle{apalike}

\bibliography{jshort,som}

### 2013-05-28 inst/ms/ms.tex
A variety of ms updates; split into multiple files

@@
   of maintaining biocomplexity and therefore environmental response
-  diversity,
-\end{enumerate}
-
-\begin{enumerate}
-\def\labelenumi{(\arabic{enumi})}
-\setcounter{enumi}{1}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  the research priority of identifying differences in environmental
-  tolerance given predicted environmental changes, and (3) the utility
-  of considering risk for groups of fish stocks --- especially given
-  environmental, biological, and implementation uncertainty --- through
-  the lens of portfolio theory.
+  diversity, (2) the research priority of identifying differences in
+  environmental tolerance given predicted environmental changes, and (3)
+  the utility of considering risk for groups of fish stocks ---
+  especially given environmental, biological, and implementation
+  uncertainty --- through the lens of portfolio theory.
 \end{enumerate}
@@
 
-Managing risk fundamental to conservation of metapopulations
+Managing risk is fundamental to the conservation of an endangered
+species. When an endangered species exists as a metapopulation, we can
+manage risk at two levels: at the population level or at the
+metapopulation level. Typically we treat sources of risk at the
+metapopulation level as external and uncontrollable and so we manage
+risk by altering fishing or hunting on a population level as well as
+improving the connectivity of populations.
+
+The management of financial portfolios provides another way of
+considering risk. Economists consider the risk and performance of a
+financial portfolio as a function of the weighting of individual assets
+that make up the portfolio. Modern Portfolio Theory (MPT) proposes that
+there is a set of portfolios that maximizes expected return for a level
+of expected risk or minimizes expected risk for a level of expected
+return \citep{Markowitz1952, Markowitz1959}. This optimal set contains
+portfolios that range along a continuum of risk-tolerance; economists
+refer to this set as the efficient frontier.
 
 Portfolios provide another way of managing risk - explain mean-variance
-optimization and how this translates to ecology \citep{Figge2004}
-\citep{Hoekstra2012} \citep[\citet{Ando2012}]{Ando2011}
-\citep[\citet{Markowitz1959}]{Markowitz1952}
+optimization and how this translates to ecology
+\citep{Figge2004}\citep{Hoekstra2012}
+\citep[\citet{Ando2012}]{Ando2011}\citep[\citet{Markowitz1959}]{Markowitz1952}
 
@@
 
+Here, we ask how a portfolio approach to managing risk can inform the
+spatial conservation of metapopulations in a changing world. To answer
+this, we develop a salmon metapopulation simulation in which
+population-specific productivity is driven by spatially-distributed
+environmental tolerance and patterns of short- and long-term
+environmental change. We then implement different conservation ``rules
+of thumb'' that control the population-specific carrying capacities and
+evaluate the salmon portfolios along risk and return axes, similarly to
+how financial portfolios are assessed. We show\ldots{} (briefly)
+
 \section{Methods}
 
-\subsection{Simulation model components}
+\subsection{Overview}
 
-general overview
+We developed a salmon metapopulation simulation model that includes a
+stock-recruit relationship, demographic stochasticity, straying between
+populations, varying responses to the environment, escapement target
+setting, and implementation uncertainty. Under two kinds of
+environmental regimes we tested different conservation rules of thumb
+and evaluated these plans in risk-return space similar to how financial
+managers evaluate financial portfolios. See Figure \ref{f:sim-flow} for
+an illustration of the overall simulation structure.
 
-See Figure \ref{f:sim-flow} for an illustration of the simulation
-structure.
+\subsection{Defining the ecological portfolio}
+
+In our ecological portfolio, we defined assets as stream-level
+populations and the portfolio as a salmon metapopulation (Table
+\ref{t:port}). The portfolio investors were the stakeholders in the
+fishery and metapopulation performance. For example, we could consider
+fisheries managers, conservation agencies, or first Nations groups as
+investors. We defined asset value as abundance (returns) of stream
+populations and value of the portfolio as the overall metapopulation
+abundance. In this scenario, the equivalent to financial rate of return
+is the generation-to-generation rate of change of metapopulation
+abundance. We defined the financial asset investment weights as the
+carrying capacity of the stream populations, specifically the unfinished
+equilibrium stock size.
 
@@
 
-We are using a Ricker curve,
+The salmon metapopulation dynamics in our simulation were governed by a
+spawner-return relationship with stochastic noise and straying between
+populations.
+
+\subsubsection{Spawner-return relationship}
+
+We defined the spawner-return relationship with a Ricker model
+\citep{Ricker1954},
 
@@
 
-\noindent
-where $t$ represents a generation time, $i$ represents a population, $R$
-is the number of returns, $S$ is the number of spawners, $a$ is the
-productivity parameter (which can vary with the environmental signal),
+\noindent where $t$ represents a generation time, $i$ represents a
+population, $R$ is the number of returns, $S$ is the number of spawners,
+$a$ is the productivity parameter (which can vary with the environment),
 and $b$ is the density-dependent term (which is used as the asset
 weights in the portfolios). The term $w_{ti}$ represents first-order
-autocorrelated error (AR1). Formally,
-$w_{ti} = w_{ti-1} \rho_w + v_{ti}$, where $v_{ti}$ represents
-independent and normally distributed error with mean 0 and standard
-deviation of $\sigma_v$. The parameter $\rho_w$ represents the
-correlation between residuals from subsequent generations.
+autocorrelated error. Formally, $w_{ti} = w_{ti-1} \rho_w + v_{ti}$,
+where $v_{ti}$ represents independent and normally distributed error
+with mean 0 and standard deviation of $\sigma_v$. The parameter $\rho_w$
+represents the correlation between residuals from subsequent
+generations.
 
-@cooper1999
+\subsubsection{Straying}
 
-We implemented straying as in \citet{Cooper1999}. We generate a matrix
-that represents the fraction of straying between any two populations. We
-have set up the metapopulation (thus far) in a very simple scenario: the
-populations are arranged in a line and those that are nearer to each
-other are more likely to stray between each other {[}insert caveats{]}.
-Two parameters control the straying: the fraction of fish $f_{stray}$
-that stray from their natal stream in any given generation and the rate
-at which this straying between streams decays with distance $m$. We
-calculated the number of salmon straying from stream $j$ to stream $i$
-as,
+We implemented straying as in \citet{Cooper1999}. We set up the
+metapopulation in a simple scenario: the populations were arranged in a
+line and those that were nearer to each other were more likely to stray
+between each other. Two parameters controlled the straying: the fraction
+of fish $f_{\mathrm{stray}}$ that stray from their natal stream in any
+generation and the rate $m$ at which this straying between streams
+decays with distance. We calculated the number of salmon straying from
+stream $j$ to stream $i$ as,
 
 \begin{equation}
-  \mathrm{strays}_{tij} = f_{stray} R_{tj}
+  \mathrm{strays}_{tij} = f_{\mathrm{stray}} R_{tj}
     \frac{e^{-m \lvert i-j \rvert }}
@@
 
+\subsection{Fishing}
+
+Our simulation used a simple set of rules to establish escapement
+targets and implement harvesting. Every 5(TODO CAN VARY) years our
+simulation fitted a spawner-return function and target harvest rate
+$H_{\mathrm{tar}}$ was set based on \citet{Hilborn1992} as
+
+\begin{equation}
+  H_{\mathrm{tar}} = \frac{A}{b (0.5 - 0.07a)}
+  \label{eq:esc}
+\end{equation}
+
+\noindent
+where $A$ represents the return abundance and $a$ and $b$ represent the
+Ricker model parameters. We included implementation uncertainty in the
+actual harvest rate $H_{\mathrm{act}}$ as
+
+\begin{equation}
+  H_{\mathrm{act}} = \mathrm{beta}(\alpha_h, \beta_h)
+\end{equation}
+
+\noindent
+where $\alpha_h$ and $\beta_h$ are the location and shape parameters in
+a beta distribution. They can be calculated from the desired mean
+$\mu_h$ and standard deviation $\sigma_h$ as: TODO ADD REFERENCES
+
+\begin{align}
+  \alpha_h &= \frac{\mu_h^2(1 - \mu_h)}{\sigma_h - 1/\mu_h}\\
+   \beta_h &= \frac{\alpha}{\mu_h - 1}.
+\end{align}
+
+\noindent
+Further, to establish a range of spawner-return values and to mimic the
+start of an open-access fishery, for the first 20 (TODO variable) years
+we drew the fraction of fish harvested randomly from a uniform
+distribution between 0.1 and 0.9. TODO CONSIDER BETTER RANGE HERE
+
 \subsection{Environmental dynamics}
 
-\subsection{Environmental response diversity}
+We can generally break environmental dynamics down into short-term and
+long-term changes. We evaluated the contribution of the environment to
+metapopulation performance under these two components separately.
 
-Scenarios? More variable are also more productive?\ldots{}
-
-\subsection{Portfolio optimization}
+short-term forcing: stationary AR1; long-term forcing: linear increase
 
 \subsection{Conservation rules of thumb}
+
+We evaluated two sets of conservation rules of thumb: (1) spatial
+response diversity rules of thumb in an idealistic situation where you
+can detect response diversity, and (2) a more realistic scenario where
+we know little about response diversity and we're left with a choice of
+how many populations to conserve. We then considered how optimal these
+strategies are across all possible conservation-prioritization
+strategies.
+
+\subsubsection{Spatial conservation strategies}
+
+We evaluated four spatial conservation rules of thumb (Figure
+\ref{f:curves}). In all scenarios, four populations are conserved and
+the unfished equilibrium biomass of the remaining populations is reduced
+to near elimination (five salmon). These reduced populations can
+therefore still receive straying salmon but are unlikely to re-build on
+their own to any substantial volume of salmon.
+
+The four scenarios are:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  Conserve an even sampling of response diversity. In other words,
+  conserve alternating subpopulations.
+\item
+  Conserve the most stable populations only.
+\item
+  Conserve one half of the metapopulation.
+\item
+  Conserve the other half of the metapopulation.
+\end{enumerate}
+
+\subsubsection{How many populations to conserve?}
+
+In reality we rarely know precise levels of response diversity. We
+therefore additionally considered a case where the conservation is
+randomly assigned with respect to response diversity but where different
+numbers of streams can be conserved. We considered conserving from two
+to 16 streams. Similarly to the spatial strategies, we reduced the
+carrying capacity of the remaining streams to the nominal level of five
+salmon.
+
+\subsubsection{How optimal are the conservation strategies?}
+
+To test how optimal the strategies are, we generated metapopulation
+portfolios in which the investment weights (Ricker $b$ parameters) were
+randomly assigned. This Monte-Carlo strategy illustrates a large sample
+of possible portfolios in risk-return space. We then placed the
+conservation strategies listed above into this space to determine how
+close to the efficient frontier they lie.
 
@@
 \midrule
-Assets             & Stream-level salmon populations; possibly a Viable
-                     Salmonid Population\\
-Portfolio          & The salmon metapopulation; possibly an Evolutionarily
-                     Significant Unit\\
+Assets             & Stream-level salmon populations; possibly a Viable Salmonid Population\\
+Portfolio          & The salmon metapopulation; possibly an Evolutionarily Significant Unit\\
 Portfolio managers & Salmon managers\\
 Investors          & Salmon managers, conservation agency, or salmon fishers\\
-Asset weights      & Carrying capacity (specifically the $b$ parameter in
-                     a Ricker model)\\
-Asset returns      & Rate of change of generation-to-generation salmon
-                     metapopulation abundance\\
-Asset risk         & Variance of generation-to-generation salmon metapopulation
-                     abundance\\
+Asset weights      & Carrying capacity (specifically the $b$ parameters in a Ricker model)\\
+Asset returns      & Rate of change of generation-to-generation salmon metapopulation abundance\\
+Asset risk         & Variance of generation-to-generation salmon metapopulation abundance\\
 \bottomrule
 \end{tabular}
-\label{tab:port-components}
+\label{t:port}
 \end{table}
@@
 First order (AR1) serial correlation of stock-recruit residuals  & $\rho_w$              & 0.40 [0, 0.80] \\
-Fraction of fish that stray from natal streams                   & $f_{stray}$     & 0.02 [0, 0.10] \\
+Fraction of fish that stray from natal streams                   & $f_{\mathrm{stray}}$  & 0.02 [0, 0.10] \\
 Exponential rate of decay of straying with distance              & $m$                   & 0.3 [0.05, 0.5] \\
-Standard deviation of beta distribution for implementation error & $\sigma_{impl}$ & 0.05 [0, 0.20] \\
-Frequency of assessment (years)                                  & $f_{assess}$    & 20 [5, 50] \\
+Standard deviation of beta distribution for implementation error & $\sigma_{h}$          & 0.05 [0, 0.20] \\
+Frequency of assessment (years)                                  & $f_{\mathrm{assess}}$ & 20 [5, 50] \\
 \bottomrule
@@
 
-`
+\clearpage
 
@@
 
+\clearpage
+
 \begin{figure}[htbp]
@@
 \end{figure}
+
+\clearpage
 
@@
 simulations per strategy. We also show example metapopulation (c) 
-rate-of-change and (d) abundance time series for a the short-term 
+rate-of-change and (d) abundance time series for the short-term 
 environmental-fluctuation scenario.}
@@
 \end{figure}
+
+\clearpage
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=5.0in]{../examples/quasi-extinct.pdf}
+\caption{First draft of a quasi-extinction figure. The y-axis shows the 
+cumulative proportion of subpopulations that have had abundance dip below a 
+threshold of 20\% of unfished equilibrium abundance. We could show a (binomial) 
+measure of uncertainty around these if we want to. Some interesting points: the 
+cumulative proportion of ``quasi-extinction'' is higher for both ``one half'' 
+scenarios. If it's set up correctly, then this means that there's a benefit to 
+the subpopulations to investing in a more ``optimal'' portfolio. The }
+\label{f:n-mv}
+\end{figure}

### 2013-06-05 inst/ms/ms.tex
Make a variety of ms edits

@@
 of expected risk or minimizes expected risk for a level of expected
-return \citep{Markowitz1952, Markowitz1959}. This optimal set contains
+return \citep{markowitz1952, markowitz1959}. This optimal set contains
 portfolios that range along a continuum of risk-tolerance; economists
@@
 optimization and how this translates to ecology
-\citep{Figge2004}\citep{Hoekstra2012}
-\citep[\citet{Ando2012}]{Ando2011}\citep[\citet{Markowitz1959}]{Markowitz1952}
+\citep{figge2004}\citep{hoekstra2012}
+\citep[\citet{ando2012}]{ando2011}\citep[\citet{markowitz1959}]{markowitz1952}
 
 We use salmon as an example, describe the components of the example,
-provide references on metapopulation existence \citep{Schindler2010}
+provide references on metapopulation existence \citep{schindler2010}
 
@@
 
-\subsection{Overview}
-
 We developed a salmon metapopulation simulation model that includes a
@@
 managers evaluate financial portfolios. See Figure \ref{f:sim-flow} for
-an illustration of the overall simulation structure.
+an illustration of the overall simulation structure. We provide a
+package \texttt{metafolio} for the statistical software \texttt{R}
+\citep{r2013} as an appendix, which can be used to carry out the
+simulations and analyses we describe in this paper.
 
@@
 
-In our ecological portfolio, we defined assets as stream-level
-populations and the portfolio as a salmon metapopulation (Table
-\ref{t:port}). The portfolio investors were the stakeholders in the
-fishery and metapopulation performance. For example, we could consider
-fisheries managers, conservation agencies, or first Nations groups as
-investors. We defined asset value as abundance (returns) of stream
-populations and value of the portfolio as the overall metapopulation
-abundance. In this scenario, the equivalent to financial rate of return
-is the generation-to-generation rate of change of metapopulation
-abundance. We defined the financial asset investment weights as the
-carrying capacity of the stream populations, specifically the unfinished
-equilibrium stock size.
+In our ecological portfolios, we defined assets as stream-level
+populations and the portfolios as salmon metapopulations (Table
+\ref{t:port}). We use the terms \textit{stream} and \textit{populations}
+interchangeably to represent the portfolio assets. The portfolio
+investors were the stakeholders in the fishery and metapopulation
+performance. For example, we could consider fisheries managers,
+conservation agencies, or First Nations groups as investors. We defined
+asset value as the abundance of returning salmon in each stream and
+value of the portfolio as the overall metapopulation abundance. In this
+scenario, the equivalent to financial rate of return is the
+generation-to-generation rate of change of metapopulation abundance. We
+defined the financial asset investment weights as the carrying capacity
+of the stream populations --- specifically the unfinished equilibrium
+stock size --- since maintaining or restoring habitat requires money,
+time, and resources.
 
@@
 The salmon metapopulation dynamics in our simulation were governed by a
-spawner-return relationship with stochastic noise and straying between
-populations.
+spawner-return relationship with stochastic noise and by straying
+between populations.
 
@@
 We defined the spawner-return relationship with a Ricker model
-\citep{Ricker1954},
+\citep{ricker1954},
 
 \begin{equation}
-R_{ti} = S_{ti}e^{a_{ti}(1-S_{ti}/b_i) + w_{ti}},
+R_{ti} = S_{ti}e^{a_{ti}(1-S_{ti}/b_i) + w_{ti}}
 \end{equation}
@@
 
+We manipulated the capacity and productivity parameters $b$ and $a$ as
+part of the portfolio simulation. The capacity parameters $b_i$ were
+controlled by the investment weights in the populations. For example, a
+large investment in a stream was represented by a larger unfished
+equilibrium stock size $b$ for stream $i$. The productivity parameters
+$a_{ti}$ were controlled by the interaction between an environmental
+signal and the stream-level population environmental tolerance curves.
+The environmental tolerance curves were generated as parabolas according
+to
+
+\begin{equation}
+  a_{ti} =
+  \begin{cases}
+    W_i (e_t - e_i^{\mathrm{opt}})^2 + a_i^{\mathrm{max}},
+      & \text{if } a_{ti} > 0\\
+      0, & \text{if } a_{ti} \leq 0
+  \end{cases}
+\end{equation}
+
+\noindent where $W_i$ controls the width of the curve for population
+$i$, $e_t$ represents the environmental value at generation $t$,
+$e_i^{\mathrm{opt}}$ represents the optimal environmental value for
+population $i$, and $a_i^{\mathrm{max}}$ represents the maximum possible
+$a$ value for population $i$. See Figure \ref{f:all-curves} for example
+environmental tolerance curves.
+
 \subsubsection{Straying}
 
-We implemented straying as in \citet{Cooper1999}. We set up the
-metapopulation in a simple scenario: the populations were arranged in a
+We implemented straying as in \citet{cooper1999}. We set up the
+metapopulation in a simple scenario: we arranged the populations in a
 line and those that were nearer to each other were more likely to stray
@@
 decays with distance. We calculated the number of salmon straying from
-stream $j$ to stream $i$ as,
+stream $j$ to stream $i$ as
 
@@
       {\displaystyle\sum\limits_{\substack{k = 1 \\
-    k \neq j}}^{n} e^{-m \lvert k-j \rvert }},
+    k \neq j}}^{n} e^{-m \lvert k-j \rvert }}
   \label{eq:stray}
@@
 
-\noindent
-where $R_{tj}$ is the number of returning salmon at generation $t$ whose
-natal stream was stream $j$. The subscript $k$ represents a stream ID
-and $n$ the number of populations. The denominator is a normalizing
-constant to ensure the desired fraction of fish stray. See Figure
-\ref{f:stray} for an example straying matrix.
+\noindent where $R_{tj}$ is the number of returning salmon at generation
+$t$ whose natal stream was stream $j$. The subscript $k$ represents a
+stream ID and $n$ the number of populations. The denominator is a
+normalizing constant to ensure the desired fraction of fish stray. See
+Figure \ref{f:stray} for an example straying matrix.
 
@@
 Our simulation used a simple set of rules to establish escapement
-targets and implement harvesting. Every 5(TODO CAN VARY) years our
-simulation fitted a spawner-return function and target harvest rate
-$H_{\mathrm{tar}}$ was set based on \citet{Hilborn1992} as
+targets and harvest the fish. Every $f_\mathrm{assess}$ years (base case
+of five years) our simulation fitted a spawner-return function and
+target harvest rate $H_{\mathrm{tar}}$ was set based on
+\citet{hilborn1992} as
 
@@
 
-\noindent
-where $A$ represents the return abundance and $a$ and $b$ represent the
-Ricker model parameters. We included implementation uncertainty in the
-actual harvest rate $H_{\mathrm{act}}$ as
+\noindent where $A$ represents the return abundance and $a$ and $b$
+represent the Ricker model parameters. We included implementation
+uncertainty in the actual harvest rate $H_{\mathrm{act}}$ as
 
@@
 
-\noindent
-where $\alpha_h$ and $\beta_h$ are the location and shape parameters in
-a beta distribution. They can be calculated from the desired mean
-$\mu_h$ and standard deviation $\sigma_h$ as: TODO ADD REFERENCES
+\noindent where $\alpha_h$ and $\beta_h$ are the location and shape
+parameters in a beta distribution. They can be calculated from the
+desired mean $\mu_h$ and standard deviation $\sigma_h$ as
+\citep[p.~97]{morgan1990}
 
 \begin{align}
-  \alpha_h &= \frac{\mu_h^2(1 - \mu_h)}{\sigma_h - 1/\mu_h}\\
-   \beta_h &= \frac{\alpha}{\mu_h - 1}.
+  \alpha_h &= \mu_h^2
+                \left(
+                \frac{1 - \mu_h}{\sigma_h^2} - \frac{1}{\mu_h}
+                \right)\\
+   \beta_h &= \alpha \left({\frac{1}{\mu_h} - 1}\right).
 \end{align}
 
-\noindent
-Further, to establish a range of spawner-return values and to mimic the
-start of an open-access fishery, for the first 20 (TODO variable) years
-we drew the fraction of fish harvested randomly from a uniform
-distribution between 0.1 and 0.9. TODO CONSIDER BETTER RANGE HERE
+\noindent Further, to establish a range of spawner-return values and to
+mimic the start of an open-access fishery, for the first 20 (TODO
+variable) years we drew the fraction of fish harvested randomly from a
+uniform distribution between 0.1 and 0.9. TODO CONSIDER BETTER RANGE
+HERE
 
@@
 
-We can generally break environmental dynamics down into short-term and
-long-term changes. We evaluated the contribution of the environment to
-metapopulation performance under these two components separately.
+We evaluated portfolio performance under short- and long-term
+environmental dynamics. We represented short-term dynamics as a
+stationary first-order autoregressive process, AR(1), with correlation
+$\rho_e$
 
-short-term forcing: stationary AR1; long-term forcing: linear increase
+\begin{equation}
+  e_t = e_{t-1} \rho_e + d_t, d_t \sim \mathrm{N}(0, \sigma_d)
+\end{equation}
+
+\noindent where $e_t$ represents the environmental value in generation
+$t$ and $d$ represents normally distributed deviations of mean 0 and
+standard deviation $\sigma_d$. We represented long-term environmental
+dynamics as a linear shift in the environmental value through time
+
+\begin{equation}
+  e_t = \alpha_e + \beta_e t
+\end{equation}
+
+\noindent where $\alpha_e$ represents the intercept and $\beta_e$ the
+slope. To maintain a balanced response, we centered the trend so that
+midway through the simulation (after any burn-in period) the
+environmental value was at the mean environmental tolerance.
 
@@
 We evaluated two sets of conservation rules of thumb: (1) spatial
-response diversity rules of thumb in an idealistic situation where you
+response diversity rules of thumb in an idealistic scenario where you
 can detect response diversity, and (2) a more realistic scenario where
@@
 how many populations to conserve. We then considered how optimal these
-strategies are across all possible conservation-prioritization
+strategies were across all possible conservation-prioritization
 strategies.
@@
 We evaluated four spatial conservation rules of thumb (Figure
-\ref{f:curves}). In all scenarios, four populations are conserved and
-the unfished equilibrium biomass of the remaining populations is reduced
-to near elimination (five salmon). These reduced populations can
-therefore still receive straying salmon but are unlikely to re-build on
-their own to any substantial volume of salmon.
-
-The four scenarios are:
+\ref{f:curves}). In all spatial scenarios, we conserved four populations
+and set the unfished equilibrium biomass of the remaining populations to
+near elimination (five salmon). These reduced populations could
+therefore still receive straying salmon but were unlikely to re-build on
+their own to a substantial abundance. The four scenarios were:
 
@@
 \def\labelenumi{\arabic{enumi}.}
+\itemsep1pt\parskip0pt\parsep0pt
 \item
@@
 In reality we rarely know precise levels of response diversity. We
-therefore additionally considered a case where the conservation is
+therefore additionally considered a case where the conservation was
 randomly assigned with respect to response diversity but where different
@@
 
-To test how optimal the strategies are, we generated metapopulation
+To test how optimal the strategies were, we generated metapopulation
 portfolios in which the investment weights (Ricker $b$ parameters) were
-randomly assigned. This Monte-Carlo strategy illustrates a large sample
-of possible portfolios in risk-return space. We then placed the
+randomly assigned such that the total available ``investment'' remained
+constant. This Monte-Carlo strategy illustrates a large sample of
+possible portfolios in risk-return space. We then placed the
 conservation strategies listed above into this space to determine how
-close to the efficient frontier they lie.
+close to the efficient frontier they lie. {[}Figure not shown yet{]}
 
@@
 
-Fig. \ref{f:eg-n-linear}
+Figure \ref{f:curves} Figure \ref{f:sp-mv} Figure \ref{f:n-mv}
 
-Fig. \ref{f:eg-n-arma}
-
-Fig. \ref{f:eg-sp-arma}
-
-Fig. \ref{f:eg-sp-linear}
-
-\subsection{Base case}
-
-\subsection{Effect of X on \ldots{}}
-
-\subsection{Effect of Y on \ldots{}}
+Figure \ref{f:eg-n-linear} Figure \ref{f:eg-n-arma} Figure
+\ref{f:eg-sp-arma} Figure \ref{f:eg-sp-linear}
 
@@
 \section{Acknowledgements}
-
-\section{References}
 
@@
 Stock-recruit residual standard deviation (on log scale)         & $\sigma_v$            & 0.30 [0.05, 0.50] \\
-First order (AR1) serial correlation of stock-recruit residuals  & $\rho_w$              & 0.40 [0, 0.80] \\
+First order AR(1) serial correlation of stock-recruit residuals  & $\rho_w$              & 0.40 [0, 0.80] \\
 Fraction of fish that stray from natal streams                   & $f_{\mathrm{stray}}$  & 0.02 [0, 0.10] \\
@@
 Standard deviation of beta distribution for implementation error & $\sigma_{h}$          & 0.05 [0, 0.20] \\
-Frequency of assessment (years)                                  & $f_{\mathrm{assess}}$ & 20 [5, 50] \\
+Frequency of assessment (years)                                  & $f_{\mathrm{assess}}$ & 5 [1, 20] \\
+Environmental tolerance parameters ... & & \\
 \bottomrule
 \end{tabular}
-\label{tab:salm-pars}
+\label{t:pars}
 \end{table}
@@
 \includegraphics[height=5.5in]{../examples/simulation-diagram2.pdf}
-\caption{Flow chart of the salmon-metapopulation simulation. There are $n$ 
-salmon populations and $t$ generations. Blue text indicates values that are 
-generated before the simulation progresses through time. Red text indicates 
-steps in which calculations are performed. Black text indicates values that are 
-calculated. Grey text indicates parameters that can be set. Green text 
-indicates the looping structure of the simulation.}
+\caption{Flow chart of the salmon-metapopulation simulation.
+There are $n$ salmon populations and $t$ generations.
+Blue text indicates values that are generated before the simulation progresses through time.
+Red text indicates steps in which calculations are performed through time.
+Black text indicates values that are calculated.
+Grey text indicates parameters that can be set.
+Green text indicates the looping structure of the simulation.}
 \label{f:sim-flow}
@@
 \includegraphics[width=4.0in]{../examples/spatial-arma-sim.pdf}
-\caption{The components of an example metapopulation simulation.  We show, from 
-top to bottom, the environmental signal, the resulting productivity parameter 
-(Ricker $a$), the salmon returns, fisheries catch, salmon escapement, salmon 
-straying from their natal streams, salmon joining from other streams, 
-stock-recruit residuals on a log scale, and the estimated $a$ and $b$ 
-parameters in the fitted Ricker curve. The colored lines indicate populations 
-that thrive at low (cool colours) to high (warm colours) environmental values.}
+\caption{The components of an example metapopulation simulation.
+We show, from top to bottom, the environmental signal, the resulting productivity parameter (Ricker $a$), the salmon returns, fisheries catch, salmon escapement, salmon straying from their natal streams, salmon joining from other streams, stock-recruit residuals on a log scale, and the estimated $a$ and $b$ parameters in the fitted Ricker curve.
+The colored lines indicate populations that thrive at low (cool colours) to high (warm colours) environmental values.}
 \label{f:sp-eg}
@@
 \includegraphics[width=3.0in]{../examples/thermal-curve-scenarios.pdf}
-\caption{Different ways of prioritizing response-diversity conservation. The 
-coloured lines represent environmental-tolerance curves for conserved 
-populations. The colours identify populations that thrive at low (cool colours) 
-to high (warm colours).}
+\caption{Different ways of prioritizing response-diversity conservation.
+Panel (a) shows the thermal tolerance cures for ten possible populations and panels (b--e) show different ways of prioritizing four of those populations.
+The curves describe how productivity varies with the environment for a given population.
+Some populations thrive at low environmental values (cool colours) and some at high (warm colours) values.
+Some are tolerant to a wider range of environmental conditions (yellow-to-green colours) but with a lower maximum productivity.
+The total possible productivity across environmental values (the area under the curves) is the same.}
 \label{f:curves}
@@
 \includegraphics[width=5.0in]{../examples/spatial-mv.pdf}
-\caption{The importance of preserving environmental response diversity through 
-spatial conservation strategies. The conservation strategies correspond to 
-Fig.~\ref{f:curves} and represent conserving a range of responses (green), the 
-most stable populations only (orange), or one type of environmental response 
-(purple and pink).  In risk-return space we show environmental scenarios that 
-are comprised primarily of (a) short-term and (b) long-term environmental 
-fluctuations (see Fig.~X). The dots show simulated metapopulations and the 
-contours show 25\% and 75\% quantiles across 500 simulations per strategy. We 
-also show example metapopulation abundance time series for the (c, e) 
-short-term and (d, f) long-term  environmental-fluctuation scenarios.}
+\caption{The importance of preserving environmental response diversity through spatial conservation strategies.
+The conservation strategies correspond to Fig.~\ref{f:curves} and represent conserving a range of responses (green), the most stable populations only (orange), or one type of environmental response (purple and pink).
+In risk-return space we show environmental scenarios that are comprised primarily of (a) short-term and (b) long-term environmental fluctuations (see Fig.~X).
+The dots show simulated metapopulations and the contours show 25\% and 75\% quantiles across 500 simulations per strategy.
+We also show example metapopulation abundance time series for the (c, e) short-term and (d, f) long-term  environmental-fluctuation scenarios.}
 \label{f:sp-mv}
@@
 \includegraphics[width=5.0in]{../examples/cons-plans-n.pdf}
-\caption{The importance of preserving as many salmon populations as possible 
-when we don't know how response diversity is distributed. In risk-return space 
-we show environmental scenarios that are comprised primarily of (a) short-term 
-and (b) long-term environmental fluctuations (see Fig.~X). We show 
-metapopulations in which 2 (red), 4 (orange), 8 (yellow), or 16 (green) 
-populations of random response diversity are conserved. The dots show simulated 
-metapopulations and the contours show 25\% and 75\% quantiles across 500 
-simulations per strategy. We also show example metapopulation (c) 
-rate-of-change and (d) abundance time series for the short-term 
-environmental-fluctuation scenario.}
+\caption{The importance of preserving as many salmon populations as possible when we don't know how response diversity is distributed.
+In risk-return space we show environmental scenarios that are comprised primarily of (a) short-term and (b) long-term environmental fluctuations (see Fig.~X).
+We show metapopulations in which 2 (red), 4 (orange), 8 (yellow), or 16 (green) populations of random response diversity are conserved.
+The dots show simulated metapopulations and the contours show 25\% and 75\% quantiles across 500 simulations per strategy.
+We also show example metapopulation (c) rate-of-change and (d) abundance time series for the short-term environmental-fluctuation scenario.}
 \label{f:n-mv}
@@
 \includegraphics[width=5.0in]{../examples/quasi-extinct.pdf}
-\caption{First draft of a quasi-extinction figure. The y-axis shows the 
-cumulative proportion of subpopulations that have had abundance dip below a 
-threshold of 20\% of unfished equilibrium abundance. We could show a (binomial) 
+\caption{First draft of a quasi-extinction figure I THINK THIS WILL BE 
+SCRAPPED OR CHANGED. The y-axis shows the cumulative proportion of 
+subpopulations that have had abundance dip below a threshold of 20\% 
+of unfished equilibrium abundance. We could show a (binomial) 
 measure of uncertainty around these if we want to. Some interesting points: the 
@@
 the subpopulations to investing in a more ``optimal'' portfolio. The }
-\label{f:n-mv}
+\label{f:ext}
 \end{figure}

### 2013-06-06 R/add_inset_env.r
Move two functions to package

@@ -1,0 +1,11 @@
+#' @export
+add_inset_env <- function(env, x = 0.12, y = -0.013, size = c(1, .5), ...) {
+  TeachingDemos::subplot(plot(seq(0.1, 0.15, length.out =
+        length(env)), env, axes = FALSE, ann = FALSE, type =
+      "l", col = "grey10"), x = x, y = y, size = size, ...)
+  par(xpd = NA)
+  text(x, y, "Environment", adj = c(.5, 3.0), col = "black", cex = 1)
+  par(xpd = FALSE)
+}
+
+

### 2013-06-06 R/plot_sp_A_ts.r

@@ -1,0 +1,34 @@
+#' @export
+plot_sp_A_ts <- function(X, ylim, x_axis = TRUE, y_axis = TRUE, rate = FALSE, lwd = 1.7, y_axis_ticks = NULL, start_new_plots = 1, labels = NULL, burn = 30, ...) {
+  #A_range <- ldply(X, function(x) range(rowSums(x$A[-burn, ])))
+  burn <- 1:burn
+  for(i in 1:4){
+    if(i %in% start_new_plots) {
+      plot(1,1,ylim = ylim, xlim = c(1, 70), type = "n",
+        xlab = "", ylab = "", xaxt = "n", axes = FALSE, yaxs = "i", ...)
+      if(!is.null(labels)) {
+        mtext(labels[i], side = 3, line = -1.2, cex = 0.8, adj = 0.02)
+      }
+    if(y_axis)  {
+      if(is.null(y_axis_ticks)) {
+        axis(2, col=  "grey50", at = pretty(axTicks(2), n = 2))
+      } else {
+        axis(2, col=  "grey50", at = y_axis_ticks)
+      }
+    }
+    }
+
+    x <- X[[i]]$A[-burn, ]
+    port.x <- rowSums(x)
+    if(!rate) {
+    lines(1:length(port.x), port.x, col = cols[i], lwd = lwd, lty = 1)
+    } else {
+    lines(2:length(port.x), diff(log(port.x)), col = cols[i], lwd = lwd, lty = 1)
+    }
+    box(col = "grey50")
+  }
+  if(x_axis)
+    axis(1, col=  "grey50")
+}
+
+

### 2013-06-06 inst/ms/ms.tex
A variety of ms updates; especially adding outlines

@@
   Portfolio optimization reveals\ldots{}
-\item
-  something more general with animals instead of just salmon?
 \item
@@
 
-\section{Abstract}
+\section{Long and rambling abstract that's a bit old now}
 
@@
 
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
   Managing risk is fundamental to the conservation of an endangered
-species. When an endangered species exists as a metapopulation, we can
-manage risk at two levels: at the population level or at the
-metapopulation level. Typically we treat sources of risk at the
-metapopulation level as external and uncontrollable and so we manage
-risk by altering fishing or hunting on a population level as well as
-improving the connectivity of populations.
+  species.
 
+  \begin{itemize}
+  \itemsep1pt\parskip0pt\parsep0pt
+  \item
+    risk is usually assessed on a population by population basis
+  \item
+    for metapopulations, we can consider risk for components or risk for
+    the whole
+  \end{itemize}
+\item
   The management of financial portfolios provides another way of
-considering risk. Economists consider the risk and performance of a
-financial portfolio as a function of the weighting of individual assets
-that make up the portfolio. Modern Portfolio Theory (MPT) proposes that
-there is a set of portfolios that maximizes expected return for a level
-of expected risk or minimizes expected risk for a level of expected
-return \citep{markowitz1952, markowitz1959}. This optimal set contains
-portfolios that range along a continuum of risk-tolerance; economists
-refer to this set as the efficient frontier.
+  considering risk for metapopulations.
 
-Portfolios provide another way of managing risk - explain mean-variance
-optimization and how this translates to ecology
-\citep{figge2004}\citep{hoekstra2012}
-\citep[\citet{ando2012}]{ando2011}\citep[\citet{markowitz1959}]{markowitz1952}
+  \begin{itemize}
+  \itemsep1pt\parskip0pt\parsep0pt
+  \item
+    performance of weighted assets
+  \item
+    efficient frontier concept
+  \item
+    the parallels with ecological metapopulations
+  \end{itemize}
+\item
+  Pacific salmon are an ideal model to consider under the framework of
+  portfolio prioritization because of their metapopulation structure,
+  their societal importance and threatened status, and the need to
+  prioritize conservation efforts.
 
-We use salmon as an example, describe the components of the example,
-provide references on metapopulation existence \citep{schindler2010}
+  \begin{itemize}
+  \itemsep1pt\parskip0pt\parsep0pt
+  \item
+    portfolio-like structure
 
-Research questions, and briefly our approach - Monte Carlo simulation
-modelling, as simple a model as possible while retaining important
-realistic aspects of salmon metapopulations
+    \begin{itemize}
+    \itemsep1pt\parskip0pt\parsep0pt
+    \item
+      documented metapopulation structure
+    \item
+      predators and fisheries often integrate across stocks
+    \end{itemize}
+  \item
+    Highly threatened; highly valued; face many threats
 
-Here, we ask how a portfolio approach to managing risk can inform the
-spatial conservation of metapopulations in a changing world. To answer
-this, we develop a salmon metapopulation simulation in which
-population-specific productivity is driven by spatially-distributed
-environmental tolerance and patterns of short- and long-term
-environmental change. We then implement different conservation ``rules
-of thumb'' that control the population-specific carrying capacities and
-evaluate the salmon portfolios along risk and return axes, similarly to
-how financial portfolios are assessed. We show\ldots{} (briefly)
+    \begin{itemize}
+    \itemsep1pt\parskip0pt\parsep0pt
+    \item
+      damns
+    \item
+      roads
+    \item
+      other habitat degradation
+    \item
+      logging
+    \item
+      fishing
+    \item
+      changing climate
+    \end{itemize}
+  \item
+    There's a need for prioritization
+
+    \begin{itemize}
+    \itemsep1pt\parskip0pt\parsep0pt
+    \item
+      because of the highly threatened nature
+    \item
+      conflicting goals for resource use
+    \item
+      limited resources
+    \item
+      Existing prioritization schemes
+    \end{itemize}
+  \end{itemize}
+\item
+  Portfolio stabilization through systematic asynchrony in salmon
+  population dynamics can be driven by stocks responding uniquely to the
+  same environment (`response diversity') or by physical features of the
+  environment causing stocks to experience different environmental
+  forces (`environmental filter').
+
+  \begin{itemize}
+  \itemsep1pt\parskip0pt\parsep0pt
+  \item
+    different responses to the environment: response diversity /
+    biocomplexity
+
+    \begin{itemize}
+    \itemsep1pt\parskip0pt\parsep0pt
+    \item
+      genetic variation by stock
+    \item
+      can be driven through historical conditions
+    \item
+      we focus on this here
+    \end{itemize}
+  \item
+    experiencing different environment: environmental filter concept
+    (Moran like)
+
+    \begin{itemize}
+    \itemsep1pt\parskip0pt\parsep0pt
+    \item
+      spatially spread out, geographical and hydrological features
+      filter the environment to create a variety of forces
+    \item
+      Typically a combination of both
+    \end{itemize}
+  \item
+    Conversely, forces acting against asynchrony
+
+    \begin{itemize}
+    \itemsep1pt\parskip0pt\parsep0pt
+    \item
+      inverse: loss of population diversity
+    \item
+      inverse: moran
+    \item
+      but also: dispersal
+    \item
+      and: trophic interactions with synchronized species
+    \end{itemize}
+  \end{itemize}
+\end{itemize}
+
+Here, we ask how a portfolio approach to management can inform the
+conservation of metapopulations in a changing world. We ask two primary
+questions: (1) What does portfolio optimization tell us about different
+spatial approaches to prioritizing metapopulation conservation assuming
+that response diversity is spatially distributed? (2) In a scenario
+where we don't know the population-level response diversity, what does
+portfolio optimization tell us about the number of populations
+conserved? To answer these questions, we develop a salmon metapopulation
+simulation in which population-specific productivity is driven by
+spatially-distributed environmental tolerance and patterns of short- and
+long-term climatic change. We then implement different conservation
+``rules of thumb'' that control the population-level carrying
+capacities. We evaluate the salmon portfolios along risk and return
+axes, similarly to how financial portfolios are assessed. Our findings
+illustrate that: (1) conserving response diversity buffers
+metapopulation risk given short-term climate forcing and rate of return
+given long-term climate forcing, (2) conserving more subpopulations
+buffers risk regardless of the climate trend, and (3) considering
+metapopulations through the lens of portfolio theory provides a useful
+additional dimension through which we can evaluate conservation
+strategies.
 
@@
 managers evaluate financial portfolios. See Figure \ref{f:sim-flow} for
-an illustration of the overall simulation structure. We provide a
-package \texttt{metafolio} for the statistical software \texttt{R}
-\citep{r2013} as an appendix, which can be used to carry out the
-simulations and analyses we describe in this paper.
+an illustration of the overall simulation structure. See Table
+\ref{t:pars} for a list of the input parameters to our simulation and
+their default values. We provide a package \texttt{metafolio} for the
+statistical software \texttt{R} \citep{r2013} as an appendix, to carry
+out the simulations and analyses we describe in this paper.
 
@@
 \ref{t:port}). We use the terms \textit{stream} and \textit{populations}
-interchangeably to represent the portfolio assets. The portfolio
-investors were the stakeholders in the fishery and metapopulation
-performance. For example, we could consider fisheries managers,
-conservation agencies, or First Nations groups as investors. We defined
-asset value as the abundance of returning salmon in each stream and
-value of the portfolio as the overall metapopulation abundance. In this
-scenario, the equivalent to financial rate of return is the
-generation-to-generation rate of change of metapopulation abundance. We
-defined the financial asset investment weights as the carrying capacity
-of the stream populations --- specifically the unfinished equilibrium
-stock size --- since maintaining or restoring habitat requires money,
-time, and resources.
+interchangeably to represent the portfolio assets. We defined the
+portfolio investors as the stakeholders in the fishery and
+metapopulation performance. For example, we could consider fisheries
+managers, conservation agencies, or First Nations groups as investors.
+We defined asset value as the abundance of returning salmon in each
+stream and value of the portfolio as the overall metapopulation
+abundance. In this scenario, the equivalent to financial rate of return
+is the generation-to-generation rate of change of metapopulation
+abundance. We defined the financial asset investment weights as the
+capacity of the stream populations --- specifically the unfinished
+equilibrium stock size --- since maintaining or restoring habitat
+requires money, time, and resources.
 
@@
 The salmon metapopulation dynamics in our simulation were governed by a
-spawner-return relationship with stochastic noise and by straying
-between populations.
+spawner-return relationship with stochastic demographic noise and by
+straying between populations.
 
@@
 $a_{ti}$ were controlled by the interaction between an environmental
-signal and the stream-level population environmental tolerance curves.
-The environmental tolerance curves were generated as parabolas according
-to
+signal and the stream-level population environmental-tolerance curves.
+
+We generated the environmental-tolerance parabolas according to
 
@@
 population $i$, and $a_i^{\mathrm{max}}$ represents the maximum possible
-$a$ value for population $i$. See Figure \ref{f:all-curves} for example
-environmental tolerance curves.
+$a$ value for population $i$. We set the $W_i$ parameters (Table
+\ref{t:pars}) and calculated the $a_i^{\mathrm{max}}$ parameters so that
+the area under each curve $A_i$ was equal. See Figure \ref{f:curves}a
+for example environmental-tolerance curves.
 
@@
 Our simulation used a simple set of rules to establish escapement
-targets and harvest the fish. Every $f_\mathrm{assess}$ years (base case
+targets and harvest the fish. Every $f_\mathrm{assess}$ years (default
 of five years) our simulation fitted a spawner-return function and
@@
 \noindent Further, to establish a range of spawner-return values and to
-mimic the start of an open-access fishery, for the first 20 (TODO
-variable) years we drew the fraction of fish harvested randomly from a
-uniform distribution between 0.1 and 0.9. TODO CONSIDER BETTER RANGE
-HERE
+mimic the start of an open-access fishery, for the first 30 years we
+drew the fraction of fish harvested randomly from a uniform distribution
+between 0.1 and 0.9. We discarded these initial 30 years as a burn-in
+period throughout our analyses.
 
@@
 We evaluated two sets of conservation rules of thumb: (1) spatial
-response diversity rules of thumb in an idealistic scenario where you
-can detect response diversity, and (2) a more realistic scenario where
-we know little about response diversity and we're left with a choice of
-how many populations to conserve. We then considered how optimal these
-strategies were across all possible conservation-prioritization
-strategies.
+response diversity conservation strategies in an idealistic scenario
+where you can detect response diversity, and (2) a more realistic
+scenario where we know little about response diversity and we're left
+with a choice of how many populations to conserve.
 
@@
 We evaluated four spatial conservation rules of thumb (Figure
-\ref{f:curves}). In all spatial scenarios, we conserved four populations
-and set the unfished equilibrium biomass of the remaining populations to
-near elimination (five salmon). These reduced populations could
-therefore still receive straying salmon but were unlikely to re-build on
+\ref{f:curves}b--e). In all spatial scenarios, we conserved four
+populations and set the unfished equilibrium biomass of the remaining
+populations to near elimination (five salmon). These reduced populations
+could still receive straying salmon but were unlikely to rebuild on
 their own to a substantial abundance. The four scenarios were:
@@
 \item
-  Conserve an even sampling of response diversity. In other words,
-  conserve alternating subpopulations.
+  Conserve an even sampling of response diversity.
 \item
@@
 randomly assigned with respect to response diversity but where different
-numbers of streams can be conserved. We considered conserving from two
+numbers of streams could be conserved. We considered conserving from two
 to 16 streams. Similarly to the spatial strategies, we reduced the
-carrying capacity of the remaining streams to the nominal level of five
-salmon.
-
-\subsubsection{How optimal are the conservation strategies?}
-
-To test how optimal the strategies were, we generated metapopulation
-portfolios in which the investment weights (Ricker $b$ parameters) were
-randomly assigned such that the total available ``investment'' remained
-constant. This Monte-Carlo strategy illustrates a large sample of
-possible portfolios in risk-return space. We then placed the
-conservation strategies listed above into this space to determine how
-close to the efficient frontier they lie. {[}Figure not shown yet{]}
+capacity of the remaining streams to the nominal level of five salmon.
 
@@
 
-Figure \ref{f:curves} Figure \ref{f:sp-mv} Figure \ref{f:n-mv}
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  Spatial conservation
+
+  \begin{itemize}
+  \itemsep1pt\parskip0pt\parsep0pt
+  \item
+    Short-term environment
+
+    \begin{itemize}
+    \itemsep1pt\parskip0pt\parsep0pt
+    \item
+      Response diversity conservation affects risk properties of
+      portfolio
+    \item
+      Nearly the same portfolio whether you invest in the most stable or
+      the full range of responses
+    \item
+      Local demographic extinction may be different
+    \item
+      You can see this for an example metapopulation time series: c, e:
+      generation-to-generation fluctuations dampened by response
+      diversity - a ying for every yang
+    \end{itemize}
+  \item
+    Long-term environment
+
+    \begin{itemize}
+    \itemsep1pt\parskip0pt\parsep0pt
+    \item
+      Response diversity conservation affects return properties of
+      portfolio
+    \item
+      Depends whether you ``get it right'' in choosing the right
+      populations to conserve
+    \item
+      Notably, those that are most productive at first are no longer so
+      later
+    \item
+      You can see what's happening in the example time series: d, f: You
+      get it right or wrong in the long run. Response diversity means
+      that when one does poorly another is doing well, longterm rate of
+      return is stable
+    \end{itemize}
+  \end{itemize}
+\item
+  Number
+
+  \begin{itemize}
+  \itemsep1pt\parskip0pt\parsep0pt
+  \item
+    Short-term environment
+
+    \begin{itemize}
+    \itemsep1pt\parskip0pt\parsep0pt
+    \item
+      More is usually better for risk (give ratios)
+    \item
+      The unknown response diversity causes a risk spread
+    \item
+      Example time series: Fig 4c: Can see the stabilizing effect of
+      number on the rate of change
+    \item
+      Example time series Fig 4d: The way we've set it up, this also
+      affects the absolute abundance - a double whammy
+    \end{itemize}
+  \item
+    Long-term environment
+
+    \begin{itemize}
+    \itemsep1pt\parskip0pt\parsep0pt
+    \item
+      The unknown response diversity causes a spread in rate of return -
+      some cases do well, some do poorly depending on the selection of
+      traits
+    \item
+      More is also usually better in terms of risk
+    \item
+      More dampens demographic stochasticity
+    \item
+      More likely to end up with response diversity?
+    \end{itemize}
+  \end{itemize}
+\end{itemize}
+
+Figure \ref{f:sp-mv}
+
+Figure \ref{f:n-mv}
 
@@
 \small
-\caption{Components of salmon metapopulation portfolios}
+\caption{Components of salmon metapopulation portfolios KEEP THIS?}
 \begin{tabular}{p{3.6cm}p{7.5cm}}
@@
 \centering
-\small
-\caption{Salmon metapopulation parameters with base case and alternate values.}
-\begin{tabular}{p{7.0cm}p{1.6cm}p{3.2cm}}
+\footnotesize
+\caption{Input parameters to the salmon metapopulation simulation with default values.}
+\begin{tabular}{p{7.7cm}p{1.4cm}p{3.6cm}}
 \toprule
-Description                                                      & Parameter             & Base [lower, upper] \\
+Description                                                                                & Symbol                & Value \\
 \midrule
-Stock-recruit residual standard deviation (on log scale)         & $\sigma_v$            & 0.30 [0.05, 0.50] \\
-First order AR(1) serial correlation of stock-recruit residuals  & $\rho_w$              & 0.40 [0, 0.80] \\
-Fraction of fish that stray from natal streams                   & $f_{\mathrm{stray}}$  & 0.02 [0, 0.10] \\
-Exponential rate of decay of straying with distance              & $m$                   & 0.3 [0.05, 0.5] \\
-Standard deviation of beta distribution for implementation error & $\sigma_{h}$          & 0.05 [0, 0.20] \\
-Frequency of assessment (years)                                  & $f_{\mathrm{assess}}$ & 5 [1, 20] \\
-Environmental tolerance parameters ... & & \\
+
+\textit{Population dynamics parameters}                                                             &                       & \\
+Stock-recruit residual standard deviation (on log scale)                                   & $\sigma_v$            & 0.30  \\
+AR(1) serial correlation of stock-recruit residuals                                        & $\rho_w$              & 0.40  \\
+Fraction of fish that stray from natal streams                                             & $f_{\mathrm{stray}}$  & 0.02  \\
+Exponential rate of decay of straying with distance                                        & $m$                   & 0.3  \\
+
+\noalign{\vskip 3mm} 
+\textit{Environmental parameters}                                                                   &                       & \\
+Width of environmental-tolerance curves for populations $i$ 1 to $n$                       & $W_i$                 & seq(0.05, 0.02, 0.02 0.05, length = n pop)\\
+Optimum environmental value for populations $i$ 1 to $n$                                   & $e_i^{\mathrm{opt}}$  & seq(13, 19, length = n pop)\\
+Area under each environmental-tolerance curve in environmental units                       & $A$                   & 30\\
+AR(1) autocorrelation of the environmental signal for the short-term fluctuation scenarios & $\rho_e$              & 0.1 \\
+Standard deviation of environmental noise for the short-term fluctuation scenarios         & $\sigma_d$            & 2 \\
+
+\noalign{\vskip 3mm} 
+\textit{Fishery parameters}                                                                         &                       & \\
+Standard deviation of beta distribution for implementation error                           & $\sigma_{h}$          & 0.05  \\
+Frequency of assessment (years)                                                            & $f_{\mathrm{assess}}$ & 5  \\
 \bottomrule
@@
 \centering
-\includegraphics[width=4.0in]{../examples/spatial-arma-sim.pdf}
-\caption{The components of an example metapopulation simulation.
-We show, from top to bottom, the environmental signal, the resulting productivity parameter (Ricker $a$), the salmon returns, fisheries catch, salmon escapement, salmon straying from their natal streams, salmon joining from other streams, stock-recruit residuals on a log scale, and the estimated $a$ and $b$ parameters in the fitted Ricker curve.
-The colored lines indicate populations that thrive at low (cool colours) to high (warm colours) environmental values.}
-\label{f:sp-eg}
+\includegraphics[width=3.0in]{../examples/thermal-curve-scenarios.pdf}
+\caption{Different ways of prioritizing response-diversity conservation.
+Panel a shows the thermal tolerance cures for ten possible populations and panels b--e show different ways of prioritizing four of those populations.
+The curves describe how productivity varies with the environment for a given population.
+Some populations thrive at low environmental values (cool colours) and some at high (warm colours) values.
+Some are tolerant to a wider range of environmental conditions (yellow-to-green colours) but with a lower maximum productivity.
+The total possible productivity (the area under the curves) is the same for each population.}
+\label{f:curves}
 \end{figure}
@@
 \centering
-\includegraphics[width=3.0in]{../examples/thermal-curve-scenarios.pdf}
-\caption{Different ways of prioritizing response-diversity conservation.
-Panel (a) shows the thermal tolerance cures for ten possible populations and panels (b--e) show different ways of prioritizing four of those populations.
-The curves describe how productivity varies with the environment for a given population.
-Some populations thrive at low environmental values (cool colours) and some at high (warm colours) values.
-Some are tolerant to a wider range of environmental conditions (yellow-to-green colours) but with a lower maximum productivity.
-The total possible productivity across environmental values (the area under the curves) is the same.}
-\label{f:curves}
+\includegraphics[width=4.0in]{../examples/spatial-arma-sim.pdf}
+\caption{The components of an example metapopulation simulation.
+We show, from top to bottom, the environmental signal, the resulting productivity parameter (Ricker $a$), the salmon returns, fisheries catch, salmon escapement, salmon straying from their natal streams, salmon joining from other streams, stock-recruit residuals on a log scale, and the estimated $a$ and $b$ parameters in the fitted Ricker curve.
+The colored lines indicate populations that thrive at low (cool colours) to high (warm colours) environmental values.}
+\label{f:sp-eg}
 \end{figure}
@@
 \includegraphics[width=5.0in]{../examples/cons-plans-n.pdf}
-\caption{The importance of preserving as many salmon populations as possible when we don't know how response diversity is distributed.
+\caption{The importance of preserving as many subpopulations as possible when we don't know how response diversity is distributed.
 In risk-return space we show environmental scenarios that are comprised primarily of (a) short-term and (b) long-term environmental fluctuations (see Fig.~X).
@@
 \clearpage
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=5.0in]{../examples/quasi-extinct.pdf}
-\caption{First draft of a quasi-extinction figure I THINK THIS WILL BE 
-SCRAPPED OR CHANGED. The y-axis shows the cumulative proportion of 
-subpopulations that have had abundance dip below a threshold of 20\% 
-of unfished equilibrium abundance. We could show a (binomial) 
-measure of uncertainty around these if we want to. Some interesting points: the 
-cumulative proportion of ``quasi-extinction'' is higher for both ``one half'' 
-scenarios. If it's set up correctly, then this means that there's a benefit to 
-the subpopulations to investing in a more ``optimal'' portfolio. The }
-\label{f:ext}
-\end{figure}

### 2013-06-06 inst/ms/som.tex
A variety of ms updates; split into multiple files
Make a variety of ms edits
A variety of ms updates; especially adding outlines

@@
 \includegraphics[width=4.0in]{../examples/figure/plot-various-options-ts-3pops.pdf}
-\caption{The impact of increasing or decreasing various parameter values in the simulations. The different lines represent different salmon populations.}
+\caption{The impact of increasing or decreasing various parameter values in the simulations.
+The different lines represent different salmon populations.
+(NEED TO ADD PARAMETER VALUES AND EXPAND THIS SLIGHTLY)}
 \label{f:eg-sens}
@@
 \includegraphics[width=4.0in]{../examples/figure/stray-matrix.pdf}
-\caption{ An example straying matrix.
-The rows and columns represent different populations (indicated by population number).
-Dark blue indicates a high rate of straying and light blue indicates a low rate of straying.
-The straying matrix is generated through equation \ref{eq:stray}.}
+\caption{An example straying matrix. The rows and columns represent different 
+populations (indicated by population number). Dark blue indicates a high rate 
+of straying and light blue indicates a low rate of straying. The straying 
+matrix is generated through equation \ref{eq:stray}.}
 \label{f:stray}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=4.5in]{../examples/thermal-curves.pdf}
-\caption{The full range of environmental tolerance curves shown for 10 populations.
-The vertical dotted lines indicate the general range of environmental fluctuations in the main simulations, and the vertical dashed line indicates the mean environmental value in the main simulations.}
-\label{f:all-curves}
 \end{figure}
@@
 
-A placeholder: \citep{Schindler2010}
+A placeholder: \citep{schindler2010}
 

### 2013-06-06 inst/ms/extra-param-table.tex
Move extra parameter table rows to temporary file

Target harvest rate                                              & 
$H_\mathrm{tar}$      & \\
Actual harvest rate                                              & $H_\mathrm{act}$     & \\
Salmon return abundance                & $R$ & \\
Salmon spawning abundance                & $S$ & \\
Ricker capacity parameter & $b$ & \\
Ricker productivity parameter & $a$ & \\
Environmental value at generation $t$ & $e_t$ & \\

Independent and normally distributed environmental deviations with mean 0 & $d$ & \\

Location and shape parameters for the implementation uncertainty beta 
distribution & $\alpha_h$, $\beta_h$ & \\
Mean of implementation uncertainty beta distribution & $\mu_h$ & \\

\midrule
Indices & & \\
\midrule
Stream population ID & $i$ and $j$ & \\
Generation & $t$ & \\
Number of populations & $n$ & \\

### 2013-06-06 inst/examples/cons_plans2.R
Minor tweaks to conservation plan scripts

@@ -3,7 +3,7 @@
 # This version uses the new package functions
 
 set.seed(2)
-USE_CACHE <- FALSE
+USE_CACHE <- TRUE
 
 w_plans <- list() 
 w_plans[[1]] <- c(5, 1000, 5, 1000, 5, 5, 1000, 5, 1000, 5)
@@ -36,9 +36,9 @@ arma_env_params <- list(mean_value = 16, ar = 0.1, sigma_env = 2, ma = 0)
   #years_to_show = 100, burn = 0)
 
 pdf("spatial-arma-sim.pdf", width = 5, height = 7)
-plot_sim_ts(meta_sim(b = w[[1]][[1]], n_pop = 10, env_params =
-    arma_env_params, env_type = "arma", assess_freq = 5),
-  years_to_show = 70, burn = 30)
+eg_arma <- meta_sim(b = w[[1]][[1]], n_pop = 10, env_params = arma_env_params,
+  env_type = "arma", assess_freq = 5)
+plot_sim_ts(eg_arma, years_to_show = 70, burn = 30)
 dev.off()
 
 if(!USE_CACHE) {
@@ -60,9 +60,9 @@ linear_env_params <- list(min_value = 12, max_value = 20, sigma_env = 0.001,
   #years_to_show = 100, burn = 0)
 
 pdf("spatial-linear-sim.pdf", width = 5, height = 7)
-plot_sim_ts(meta_sim(b = w[[1]][[1]], n_pop = 10, env_params =
-    linear_env_params, env_type = "linear", assess_freq = 5),
-  years_to_show = 70, burn = 30)
+eg_linear <- meta_sim(b = w[[1]][[1]], n_pop = 10, env_params =
+    linear_env_params, env_type = "linear", assess_freq = 5)
+plot_sim_ts(eg_linear, years_to_show = 70, burn = 30)
 dev.off()
 
 if(!USE_CACHE) {
@@ -76,8 +76,6 @@ if(!USE_CACHE) {
 
 cols <- RColorBrewer::brewer.pal(5, "Dark2")
 
-#pdf("cons-plans-spatial.pdf", width = 7, height = 3.7)
-#par(mfrow = c(1, 2))
 pdf("spatial-mv.pdf", width = 6.5, height = 6.8)
 layout(rbind(
   c(1, 2),
@@ -95,12 +93,17 @@ ylim <- c(-0.017, 0.017)
[…]

### 2013-06-06 inst/examples/cons_plans_n_random_resp_div.R

@@ -7,7 +7,7 @@ set.seed(1)
 USE_CACHE <- FALSE
 
 # in this version, the pops are wiped out; total abundance changes
-n_trials <- 500 # number of trials at each n conservation plan
+n_trials <- 800 # number of trials at each n conservation plan
 num_pops <- c(2, 4, 8, 16) # n pops to conserve
 n_plans <- length(num_pops) # number of plans
 w <- list()
@@ -28,9 +28,10 @@ arma_env_params <- list(mean_value = 16, ar = 0.1, sigma_env = 2, ma = 0)
   #years_to_show = 100, burn = 0)
 
 pdf("n-arma-sim.pdf", width = 5, height = 7)
-plot_sim_ts(meta_sim(b = w[[1]][[2]], n_pop = 16, env_params =
-    arma_env_params, env_type = "arma", assess_freq = 5, max_a = thermal_integration(16)),
-  years_to_show = 70, burn = 30)
+eg_arma <- meta_sim(b = w[[1]][[2]], n_pop = 16, env_params =
+  arma_env_params, env_type = "arma", assess_freq = 5, max_a =
+  thermal_integration(16))
+plot_sim_ts(eg_arma, years_to_show = 70, burn = 30)
 dev.off()
 
 # TODO need to fix it here to get max_a integrated based on n_pops
@@ -53,9 +54,10 @@ linear_env_params <- list(min_value = 12, max_value = 20, sigma_env = 0.001,
   #years_to_show = 100, burn = 0)
 
 pdf("n-linear-sim.pdf", width = 5, height = 7)
-plot_sim_ts(meta_sim(b = w[[4]][[1]], n_pop = 16, env_params =
-    linear_env_params, env_type = "linear", assess_freq = 5, max_a = thermal_integration(16)),
-  years_to_show = 70, burn = 30)
+eg_linear <- meta_sim(b = w[[4]][[1]], n_pop = 16, env_params =
+  linear_env_params, env_type = "linear", assess_freq = 5, max_a =
+  thermal_integration(16))
+plot_sim_ts(eg_linear, years_to_show = 70, burn = 30)
 dev.off()
 
 if(!USE_CACHE) {
@@ -97,12 +99,19 @@ ylim <- c(-0.017, 0.017)
 par(las = 1, cex = 0.8, mar = c(0, 0, 0, 0), oma = c(4, 5.2, 1.5, .5), tck = -0.02, mgp = c(2, .6, 0)) 
[…]

### 2013-06-11 inst/ms/ms.tex
Develop the main text

@@
 \begin{itemize}
+\item
+  Prioritizing metapopulation conservation through the lens of portfolio
+  theory
 \item
@@
 \end{itemize}
+
+\textbf{Ecology Letters}
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  letters: 5000 words, 6 figures or tables
+\item
+  ideas and perspectives: 7500 words, 10 figures (needs 300 word
+  proposal)
+\end{itemize}
+
+\begin{quote}
+Ecology Letters is a forum for the very rapid publication of the most
+novel research in ecology, research that is not yet in the public
+domain. Manuscripts relating to the ecology of all taxa, in any biome
+and geographic area will be considered, and priority will be given to
+those papers exploring or testing clearly stated hypotheses. The journal
+publishes concise papers that merit urgent publication by virtue of
+their originality, general interest and their contribution to new
+developments in ecology. We discourage purely descriptive papers and
+those merely confirming or extending results of previous work.
+\end{quote}
+
+\textbf{Ecological Applications}
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  60 pages including everything
+\end{itemize}
+
+\begin{quote}
+Ecological Applications is concerned broadly with the applications of
+ecological science to environmental problems. It publishes papers that
+develop scientific principles to support environmental decision-making,
+as well as papers that discuss the application of ecological concepts to
+environmental issues, policy, and management. Papers may report on
+experimental tests, actual applications, scientific decision support
+techniques, economic analyses, social implications of environmental
+issues, or other relevant topics. Statistical or experimental methods
+papers that support research and applications are welcome. Papers
+submitted to Ecological Applications should be accessible to both
+scholars and practitioners.
+\end{quote}
 
@@
   \item
-    risk is usually assessed on a population by population basis
+    risk is usually assessed on a population-by-population basis
   \item
@@
 spatial approaches to prioritizing metapopulation conservation assuming
-that response diversity is spatially distributed? (2) In a scenario
-where we don't know the population-level response diversity, what does
-portfolio optimization tell us about the number of populations
-conserved? To answer these questions, we develop a salmon metapopulation
-simulation in which population-specific productivity is driven by
-spatially-distributed environmental tolerance and patterns of short- and
-long-term climatic change. We then implement different conservation
-``rules of thumb'' that control the population-level carrying
-capacities. We evaluate the salmon portfolios along risk and return
-axes, similarly to how financial portfolios are assessed. Our findings
-illustrate that: (1) conserving response diversity buffers
+that response diversity is spatially distributed? (2) If we don't know
+how response diversity is distribute, what does portfolio optimization
+tell us about how many populations we should conserve? To answer these
+questions, we develop a salmon metapopulation simulation in which
+population-specific productivity is driven by spatially-distributed
+environmental tolerance and patterns of short- and long-term climatic
+change. We then implement different conservation ``rules of thumb'' that
+control the population-level carrying capacities and evaluate the salmon
+portfolios along risk and return axes, as a portfolio manager would. Our
+findings illustrate that: (1) conserving response diversity buffers
 metapopulation risk given short-term climate forcing and rate of return
@@
 equilibrium stock size --- since maintaining or restoring habitat
-requires money, time, and resources.
+requires money, time, and resources. Investment in a population
+therefore represents investing in salmon habitat conservation or
+reconstruction.
 
@@
 The salmon metapopulation dynamics in our simulation were governed by a
-spawner-return relationship with stochastic demographic noise and by
+spawner-return relationship with demographic stochasticity and by
 straying between populations.
@@
 weights in the portfolios). The term $w_{ti}$ represents first-order
-autocorrelated error. Formally, $w_{ti} = w_{ti-1} \rho_w + v_{ti}$,
-where $v_{ti}$ represents independent and normally distributed error
-with mean 0 and standard deviation of $\sigma_v$. The parameter $\rho_w$
+autocorrelated error. Formally, $w_{ti} = w_{ti-1} \rho_w + r_{ti}$,
+where $r_{ti}$ represents independent and normally distributed error
+with mean 0 and standard deviation of $\sigma_r$. The parameter $\rho_w$
 represents the correlation between residuals from subsequent
@@
 \begin{equation}
-  e_t = \alpha_e + \beta_e t
+  e_t = \beta_e t - \overline{\beta_e t}
 \end{equation}
 
-\noindent where $\alpha_e$ represents the intercept and $\beta_e$ the
-slope. To maintain a balanced response, we centered the trend so that
-midway through the simulation (after any burn-in period) the
-environmental value was at the mean environmental tolerance.
+\noindent where $\beta_e$ represents the slope. To maintain a balanced
+response, we centered the trend by subtracting the mean
+$\overline{\beta_e t}$ so that midway through the simulation (after any
+burn-in period) the environmental value was at the mean environmental
+tolerance.
 
@@
 with a choice of how many populations to conserve.
-
-\subsubsection{Spatial conservation strategies}
 
@@
 
-\subsubsection{How many populations to conserve?}
-
 In reality we rarely know precise levels of response diversity. We
@@
 
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Spatial conservation
+\subsection{Which populations to conserve?}
 
-  \begin{itemize}
-  \itemsep1pt\parskip0pt\parsep0pt
-  \item
-    Short-term environment
+\subsubsection{Short-term environment}
 
-    \begin{itemize}
-    \itemsep1pt\parskip0pt\parsep0pt
-    \item
-      Response diversity conservation affects risk properties of
-      portfolio
-    \item
-      Nearly the same portfolio whether you invest in the most stable or
-      the full range of responses
-    \item
-      Local demographic extinction may be different
-    \item
-      You can see this for an example metapopulation time series: c, e:
-      generation-to-generation fluctuations dampened by response
-      diversity - a ying for every yang
-    \end{itemize}
-  \item
-    Long-term environment
+Given strong short-term environmental fluctuations, conserving response
+diversity buffers the risk properties of an ecological portfolio (Fig.
+4a). In our simulation, the median variance of generation-to-generation
+rate of change in abundance was X times lower given balanced response
+diversity (full range of responses or most stable only vs.~conserving
+one half or the other). In fact, even though by conserving the full
+range of responses, the portfolio was comprised of warm and
+cool-thriving populations that were more variable on their own, each was
+balanced by an opposing population. The portfolio risk was therefore
+comparable between the full range of responses and most stable only
+portfolios.
 
-    \begin{itemize}
-    \itemsep1pt\parskip0pt\parsep0pt
-    \item
-      Response diversity conservation affects return properties of
-      portfolio
-    \item
-      Depends whether you ``get it right'' in choosing the right
-      populations to conserve
-    \item
-      Notably, those that are most productive at first are no longer so
-      later
-    \item
-      You can see what's happening in the example time series: d, f: You
-      get it right or wrong in the long run. Response diversity means
-      that when one does poorly another is doing well, longterm rate of
-      return is stable
-    \end{itemize}
-  \end{itemize}
-\item
-  Number
+We can see the mechanism behind these portfolio properties by inspecting
+example population time series (Fig. 4c, d). If only the upper or lower
+half of response diversity is conserved, the portfolio tends to do well
+or poorly depending on the environmental conditions (Fig. 4d). This risk
+is buffered with balanced response diversity (Fig. 4c).
 
-  \begin{itemize}
-  \itemsep1pt\parskip0pt\parsep0pt
-  \item
-    Short-term environment
+\subsubsection{Long-term environment}
 
-    \begin{itemize}
-    \itemsep1pt\parskip0pt\parsep0pt
-    \item
-      More is usually better for risk (give ratios)
-    \item
-      The unknown response diversity causes a risk spread
-    \item
-      Example time series: Fig 4c: Can see the stabilizing effect of
-      number on the rate of change
-    \item
-      Example time series Fig 4d: The way we've set it up, this also
-      affects the absolute abundance - a double whammy
-    \end{itemize}
-  \item
-    Long-term environment
+Given long-term environmental change, the choice of which populations to
+conserve affects the return properties of an ecological portfolio (Fig.
+4b). By conserving balance response diversity, an ecological manager is
+hedging his or her bets on what will happen with the environment and how
+the populations will respond. The typical return for a balanced response
+diversity strategy was zero --- the metapopulation neither increased or
+decreased in abundance in the long run. By conserving only the upper or
+lower half of response diversity, a conservation manager is putting all
+his or her eggs in one basket --- the metapopulation might do really
+well through time or it might do really poorly. The example
+metapopulation abundance time series (Fig. 4d, f) illustrate this
+effect. By conserving response diversity, when one population is doing
+poorly, another is doing well and the metapopulation abundance remains
+stationary through time.
 
-    \begin{itemize}
-    \itemsep1pt\parskip0pt\parsep0pt
-    \item
-      The unknown response diversity causes a spread in rate of return -
-      some cases do well, some do poorly depending on the selection of
-      traits
-    \item
-      More is also usually better in terms of risk
-    \item
-      More dampens demographic stochasticity
-    \item
-      More likely to end up with response diversity?
-    \end{itemize}
-  \end{itemize}
-\end{itemize}
+Notably, in theses simulations, if a managers invested in the
+populations that were doing well at the beginning they would have had
+the lowest rate-of-return portfolio in the end (purple portfolios in
+Fig. 4b).
 
-Figure \ref{f:sp-mv}
+Spatial conservation strategies in the face of longterm environmental
+change hinge on whether you ``get it right'' --- whether you choose just
+the right populations to conserve.
 
-Figure \ref{f:n-mv}
+\subsection{How many populations to conserve?}
 
-Figure \ref{f:eg-n-linear} Figure \ref{f:eg-n-arma} Figure
-\ref{f:eg-sp-arma} Figure \ref{f:eg-sp-linear}
+Given a scenario where we don't know the distribution of
+population-level response diversity, portfolio optimization informs us
+about the risk buffering from maintaining multiple populations (Fig. 5).
+In short, investing in more populations buffers portfolio risk.
+
+\subsubsection{Short-term environment}
+
+Given short-term environmental noise, conserving more populations
+buffers portfolio risk while the random conservation of response
+diversity creates a spread of metapopulation risk for the same number of
+populations conserved (Fig. 5a). For example, a metapopulation with
+eight conserved populations is X times less risky than a metapopulation
+with only four. We can see this risk-buffering effect through example
+metapopulations in Fig. 5c and 5d. We note that the risk-return axes of
+portfolio optimization ignore the absolute-abundance dimension (Fig.
+5d). As one would expect, conserving fewer populations also results in
+lower-abundance metapopulations.
+
+\subsubsection{Long-term environment}
+
+Given long-term environmental noise, conserving more populations also
+buffers portfolio risk. However, in contrast to the short-term
+environmental noise scenario, the unknown response diversity creates a
+spread of possible metapopulation return for the same number of
+conserved populations (Fig. 5b). Here, the number of populations
+conserved buffers non-systematic (i.e.~not-environmentally-driven)
+stochasticity.
 
@@
 \textit{Population dynamics parameters}                              &                       & \\
-Stock-recruit residual standard deviation (on log scale)                                   & $\sigma_v$            & 0.30  \\
+Stock-recruit residual standard deviation (on log scale)             & $\sigma_r$            & 0.30  \\
 AR(1) serial correlation of stock-recruit residuals                  & $\rho_w$              & 0.40  \\
@@
 Area under each environmental-tolerance curve in environmental units & $A$                   & 30\\
-AR(1) autocorrelation of the environmental signal for the short-term fluctuation scenarios & $\rho_e$              & 0.1 \\
-Standard deviation of environmental noise for the short-term fluctuation scenarios         & $\sigma_d$            & 2 \\
+
+Standard deviation of short-term environmental fluctuations          & $\sigma_d$            & 2 \\
+AR(1) autocorrelation of short-term environmental fluctuations       & $\rho_e$              & 0.1 \\
+Slope of long-term environmental signal                              & $\beta_e$             & 0.114 \\
 

### 2013-06-12 inst/ms/ms.tex
Develop writing

@@
 \item
-  Prioritizing metapopulation conservation through the lens of portfolio
-  theory
+  Prioritizing metapopulation conservation through the lens of portfolio theory
 \item
@@
 \item
-  Intelligent tinkering with ecological portfolios: patterns of
-  environmental response diversity drive salmon conservation priorities
+  Intelligent tinkering with ecological portfolios: patterns of environmental response diversity drive salmon conservation priorities
 \item
@@
 \item
-  ideas and perspectives: 7500 words, 10 figures (needs 300 word
-  proposal)
+  ideas and perspectives: 7500 words, 10 figures (needs 300 word proposal)
 \end{itemize}
@@
 \begin{quote}
-Ecology Letters is a forum for the very rapid publication of the most
-novel research in ecology, research that is not yet in the public
-domain. Manuscripts relating to the ecology of all taxa, in any biome
-and geographic area will be considered, and priority will be given to
-those papers exploring or testing clearly stated hypotheses. The journal
-publishes concise papers that merit urgent publication by virtue of
-their originality, general interest and their contribution to new
-developments in ecology. We discourage purely descriptive papers and
-those merely confirming or extending results of previous work.
+Ecology Letters is a forum for the very rapid publication of the most novel research in ecology, research that is not yet in the public domain.
+Manuscripts relating to the ecology of all taxa, in any biome and geographic area will be considered, and priority will be given to those papers exploring or testing clearly stated hypotheses.
+The journal publishes concise papers that merit urgent publication by virtue of their originality, general interest and their contribution to new developments in ecology.
+We discourage purely descriptive papers and those merely confirming or extending results of previous work.
 \end{quote}
@@
 \begin{quote}
-Ecological Applications is concerned broadly with the applications of
-ecological science to environmental problems. It publishes papers that
-develop scientific principles to support environmental decision-making,
-as well as papers that discuss the application of ecological concepts to
-environmental issues, policy, and management. Papers may report on
-experimental tests, actual applications, scientific decision support
-techniques, economic analyses, social implications of environmental
-issues, or other relevant topics. Statistical or experimental methods
-papers that support research and applications are welcome. Papers
-submitted to Ecological Applications should be accessible to both
-scholars and practitioners.
+Ecological Applications is concerned broadly with the applications of ecological science to environmental problems.
+It publishes papers that develop scientific principles to support environmental decision-making, as well as papers that discuss the application of ecological concepts to environmental issues, policy, and management.
+Papers may report on experimental tests, actual applications, scientific decision support techniques, economic analyses, social implications of environmental issues, or other relevant topics.
+Statistical or experimental methods papers that support research and applications are welcome.
+Papers submitted to Ecological Applications should be accessible to both scholars and practitioners.
 \end{quote}
@@
 \item
-  Managing risk is fundamental to the conservation of an endangered
-  species. When an endangered species exists as a metapopulation, we can
-  manage risk at two levels: at the population level or at the
-  metapopulation level. Whereas risk is typically managed at the
-  population level, a portfolio approach to managing risk might consider
-  how conservation affects the ``weight'' of each population in a
-  metapopulation ``portfolio''.
+  Managing risk is fundamental to the conservation of an endangered species.
+When an endangered species exists as a metapopulation, we can manage risk at two levels: at the population level or at the metapopulation level.
+Whereas risk is typically managed at the population level, a portfolio approach to managing risk might consider how conservation affects the ``weight'' of each population in a metapopulation ``portfolio''.
 \item
-  Here, we ask how a portfolio approach to managing risk can inform the
-  spatial conservation of metapopulations in a changing world. To answer
-  this, we develop a salmon metapopulation simulation in which
-  population-specific productivity is driven by spatially-distributed
-  environmental tolerance and patterns of short- and long-term
-  environmental change. We then implement different spatial conservation
-  ``rules of thumb'' that control the population-specific carrying
-  capacities and evaluate the salmon portfolios along risk and return
-  axes, similarly to how financial portfolios are assessed.
+  Here, we ask how a portfolio approach to managing risk can inform the spatial conservation of metapopulations in a changing world.
+To answer this, we develop a salmon metapopulation simulation in which population-specific productivity is driven by spatially-distributed environmental tolerance and patterns of short- and long-term environmental change.
+We then implement different spatial conservation ``rules of thumb'' that control the population-specific carrying capacities and evaluate the salmon portfolios along risk and return axes, similarly to how financial portfolios are assessed.
 \item
-  Our results show, first, that maintaining populations with a variety
-  of environmental tolerances gives the best chance at an efficient
-  ecological portfolio --- minimizing metapopulation variance while
-  maximizing metapopulation growth rate. This finding emphasizes the
-  risk of allowing large spatial blocks of habitat destruction, say
-  through the development of dams. Second, we show that focusing on
-  well-performing stocks now at the detriment of others is at best
-  equivalent to a risky but efficient portfolio and is more likely a
-  risky and inefficient portfolio --- it neither minimizes
-  metapopulation variance nor maximizes growth rate compared to other
-  strategies. Third, we show that maintaining more populations reduces
-  metapopulation risk for the same spatial conservation strategy. Given
-  a lack of knowledge of how populations respond to the environment, the
-  most risk-averse approach is to conserve as many populations as
-  possible.
+  Our results show, first, that maintaining populations with a variety of environmental tolerances gives the best chance at an efficient ecological portfolio --- minimizing metapopulation variance while maximizing metapopulation growth rate.
+This finding emphasizes the risk of allowing large spatial blocks of habitat destruction, say through the development of dams.
+Second, we show that focusing on well-performing stocks now at the detriment of others is at best equivalent to a risky but efficient portfolio and is more likely a risky and inefficient portfolio --- it neither minimizes metapopulation variance nor maximizes growth rate compared to other strategies.
+Third, we show that maintaining more populations reduces metapopulation risk for the same spatial conservation strategy.
+Given a lack of knowledge of how populations respond to the environment, the most risk-averse approach is to conserve as many populations as possible.
 \item
-  Our findings highlight three key points: (1) the conservation priority
-  of maintaining biocomplexity and therefore environmental response
-  diversity, (2) the research priority of identifying differences in
-  environmental tolerance given predicted environmental changes, and (3)
-  the utility of considering risk for groups of fish stocks ---
-  especially given environmental, biological, and implementation
-  uncertainty --- through the lens of portfolio theory.
+  Our findings highlight three key points: (1) the conservation priority of maintaining biocomplexity and therefore environmental response diversity, (2) the research priority of identifying differences in environmental tolerance given predicted environmental changes, and (3) the utility of considering risk for groups of fish stocks --- especially given environmental, biological, and implementation uncertainty --- through the lens of portfolio theory.
 \end{enumerate}
@@
 \item
-  Managing risk is fundamental to the conservation of an endangered
-  species.
+  Managing risk is fundamental to the conservation of an endangered species.
 
@@
   \item
-    for metapopulations, we can consider risk for components or risk for
-    the whole
+    for metapopulations, we can consider risk for components or risk for the whole
   \end{itemize}
 \item
-  The management of financial portfolios provides another way of
-  considering risk for metapopulations.
+  The management of financial portfolios provides another way of considering risk for metapopulations.
 
@@
 \item
-  Pacific salmon are an ideal model to consider under the framework of
-  portfolio prioritization because of their metapopulation structure,
-  their societal importance and threatened status, and the need to
-  prioritize conservation efforts.
+  Pacific salmon are an ideal model to consider under the framework of portfolio prioritization because of their metapopulation structure, their societal importance and threatened status, and the need to prioritize conservation efforts.
 
@@
 \item
-  Portfolio stabilization through systematic asynchrony in salmon
-  population dynamics can be driven by stocks responding uniquely to the
-  same environment (`response diversity') or by physical features of the
-  environment causing stocks to experience different environmental
-  forces (`environmental filter').
+  Portfolio stabilization through systematic asynchrony in salmon population dynamics can be driven by stocks responding uniquely to the same environment (`response diversity') or by physical features of the environment causing stocks to experience different environmental forces (`environmental filter').
 
@@
   \item
-    different responses to the environment: response diversity /
-    biocomplexity
+    different responses to the environment: response diversity / biocomplexity
 
@@
   \item
-    experiencing different environment: environmental filter concept
-    (Moran like)
+    experiencing different environment: environmental filter concept (Moran like)
 
@@
     \item
-      spatially spread out, geographical and hydrological features
-      filter the environment to create a variety of forces
+      spatially spread out, geographical and hydrological features filter the environment to create a variety of forces
     \item
@@
 
-Here, we ask how a portfolio approach to management can inform the
-conservation of metapopulations in a changing world. We ask two primary
-questions: (1) What does portfolio optimization tell us about different
-spatial approaches to prioritizing metapopulation conservation assuming
-that response diversity is spatially distributed? (2) If we don't know
-how response diversity is distribute, what does portfolio optimization
-tell us about how many populations we should conserve? To answer these
-questions, we develop a salmon metapopulation simulation in which
-population-specific productivity is driven by spatially-distributed
-environmental tolerance and patterns of short- and long-term climatic
-change. We then implement different conservation ``rules of thumb'' that
-control the population-level carrying capacities and evaluate the salmon
-portfolios along risk and return axes, as a portfolio manager would. Our
-findings illustrate that: (1) conserving response diversity buffers
-metapopulation risk given short-term climate forcing and rate of return
-given long-term climate forcing, (2) conserving more subpopulations
-buffers risk regardless of the climate trend, and (3) considering
-metapopulations through the lens of portfolio theory provides a useful
-additional dimension through which we can evaluate conservation
-strategies.
+Here, we ask how a portfolio approach to management can inform the conservation of metapopulations in a changing world.
+We ask two primary questions: (1) What does portfolio optimization tell us about different spatial approaches to prioritizing metapopulation conservation assuming that response diversity is spatially distributed?
+(2) If we don't know how response diversity is distribute, what does portfolio optimization tell us about how many populations we should conserve?
+To answer these questions, we develop a salmon metapopulation simulation in which population-specific productivity is driven by spatially-distributed environmental tolerance and patterns of short- and long-term climatic change.
+We then implement different conservation ``rules of thumb'' that control the population-level carrying capacities and evaluate the salmon portfolios along risk and return axes, as a portfolio manager would.
+Our findings illustrate that: (1) conserving response diversity buffers metapopulation risk given short-term climate forcing and rate of return given long-term climate forcing, (2) conserving more subpopulations buffers risk regardless of the climate trend, and (3) considering metapopulations through the lens of portfolio theory provides a useful additional dimension through which we can evaluate conservation strategies.
 
@@
 
-We developed a salmon metapopulation simulation model that includes a
-stock-recruit relationship, demographic stochasticity, straying between
-populations, varying responses to the environment, escapement target
-setting, and implementation uncertainty. Under two kinds of
-environmental regimes we tested different conservation rules of thumb
-and evaluated these plans in risk-return space similar to how financial
-managers evaluate financial portfolios. See Figure \ref{f:sim-flow} for
-an illustration of the overall simulation structure. See Table
-\ref{t:pars} for a list of the input parameters to our simulation and
-their default values. We provide a package \texttt{metafolio} for the
-statistical software \texttt{R} \citep{r2013} as an appendix, to carry
-out the simulations and analyses we describe in this paper.
+We developed a salmon metapopulation simulation model that includes a stock-recruit relationship, demographic stochasticity, straying between populations, varying responses to the environment, escapement target setting, and implementation uncertainty.
+Under two kinds of environmental regimes we tested different conservation rules of thumb and evaluated these plans in risk-return space similar to how financial managers evaluate financial portfolios.
+See Figure \ref{f:sim-flow} for an illustration of the overall simulation structure.
+See Table \ref{t:pars} for a list of the input parameters to our simulation and their default values.
+We provide a package \texttt{metafolio} for the statistical software \texttt{R} \citep{r2013} as an appendix, to carry out the simulations and analyses we describe in this paper.
 
@@
 
-In our ecological portfolios, we defined assets as stream-level
-populations and the portfolios as salmon metapopulations (Table
-\ref{t:port}). We use the terms \textit{stream} and \textit{populations}
-interchangeably to represent the portfolio assets. We defined the
-portfolio investors as the stakeholders in the fishery and
-metapopulation performance. For example, we could consider fisheries
-managers, conservation agencies, or First Nations groups as investors.
-We defined asset value as the abundance of returning salmon in each
-stream and value of the portfolio as the overall metapopulation
-abundance. In this scenario, the equivalent to financial rate of return
-is the generation-to-generation rate of change of metapopulation
-abundance. We defined the financial asset investment weights as the
-capacity of the stream populations --- specifically the unfinished
-equilibrium stock size --- since maintaining or restoring habitat
-requires money, time, and resources. Investment in a population
-therefore represents investing in salmon habitat conservation or
-reconstruction.
+In our ecological portfolios, we defined assets as stream-level populations and the portfolios as salmon metapopulations (Table \ref{t:port}).
+We use the terms \emph{stream} and \emph{populations} interchangeably to represent the portfolio assets.
+We defined the portfolio investors as the stakeholders in the fishery and metapopulation performance.
+For example, we could consider fisheries managers, conservation agencies, or First Nations groups as investors.
+We defined asset value as the abundance of returning salmon in each stream and value of the portfolio as the overall metapopulation abundance.
+In this scenario, the equivalent to financial rate of return is the generation-to-generation rate of change of metapopulation abundance.
+We defined the financial asset investment weights as the capacity of the stream populations --- specifically the unfinished equilibrium stock size --- since maintaining or restoring habitat requires money, time, and resources.
+Investment in a population therefore represents investing in salmon habitat conservation or reconstruction.
 
@@
 
-The salmon metapopulation dynamics in our simulation were governed by a
-spawner-return relationship with demographic stochasticity and by
-straying between populations.
+The salmon metapopulation dynamics in our simulation were governed by a spawner-return relationship with demographic stochasticity and by straying between populations.
 
@@
 
-We defined the spawner-return relationship with a Ricker model
-\citep{ricker1954},
+We defined the spawner-return relationship with a Ricker model \citep{ricker1954},
 
@@
 
-\noindent where $t$ represents a generation time, $i$ represents a
-population, $R$ is the number of returns, $S$ is the number of spawners,
-$a$ is the productivity parameter (which can vary with the environment),
-and $b$ is the density-dependent term (which is used as the asset
-weights in the portfolios). The term $w_{ti}$ represents first-order
-autocorrelated error. Formally, $w_{ti} = w_{ti-1} \rho_w + r_{ti}$,
-where $r_{ti}$ represents independent and normally distributed error
-with mean 0 and standard deviation of $\sigma_r$. The parameter $\rho_w$
-represents the correlation between residuals from subsequent
-generations.
+\noindent where $t$ represents a generation time, $i$ represents a population, $R$ is the number of returns, $S$ is the number of spawners, $a$ is the productivity parameter (which can vary with the environment), and $b$ is the density-dependent term (which is used as the asset weights in the portfolios).
+The term $w_{ti}$ represents first-order autocorrelated error.
+Formally, $w_{ti} = w_{ti-1} \rho_w + r_{ti}$, where $r_{ti}$ represents independent and normally distributed error with mean 0 and standard deviation of $\sigma_r$.
+The parameter $\rho_w$ represents the correlation between residuals from subsequent generations.
 
-We manipulated the capacity and productivity parameters $b$ and $a$ as
-part of the portfolio simulation. The capacity parameters $b_i$ were
-controlled by the investment weights in the populations. For example, a
-large investment in a stream was represented by a larger unfished
-equilibrium stock size $b$ for stream $i$. The productivity parameters
-$a_{ti}$ were controlled by the interaction between an environmental
-signal and the stream-level population environmental-tolerance curves.
+We manipulated the capacity and productivity parameters $b$ and $a$ as part of the portfolio simulation.
+The capacity parameters $b_i$ were controlled by the investment weights in the populations.
+For example, a large investment in a stream was represented by a larger unfished equilibrium stock size $b$ for stream $i$.
+The productivity parameters $a_{ti}$ were controlled by the interaction between an environmental signal and the stream-level population environmental-tolerance curves.
 
@@
 
-\noindent where $W_i$ controls the width of the curve for population
-$i$, $e_t$ represents the environmental value at generation $t$,
-$e_i^{\mathrm{opt}}$ represents the optimal environmental value for
-population $i$, and $a_i^{\mathrm{max}}$ represents the maximum possible
-$a$ value for population $i$. We set the $W_i$ parameters (Table
-\ref{t:pars}) and calculated the $a_i^{\mathrm{max}}$ parameters so that
-the area under each curve $A_i$ was equal. See Figure \ref{f:curves}a
-for example environmental-tolerance curves.
+\noindent where $W_i$ controls the width of the curve for population $i$, $e_t$ represents the environmental value at generation $t$, $e_i^{\mathrm{opt}}$ represents the optimal environmental value for population $i$, and $a_i^{\mathrm{max}}$ represents the maximum possible $a$ value for population $i$.
+We set the $W_i$ parameters (Table \ref{t:pars}) and calculated the $a_i^{\mathrm{max}}$ parameters so that the area under each curve $A_i$ was equal.
+See Figure \ref{f:curves}a for example environmental-tolerance curves.
 
@@
 
-We implemented straying as in \citet{cooper1999}. We set up the
-metapopulation in a simple scenario: we arranged the populations in a
-line and those that were nearer to each other were more likely to stray
-between each other. Two parameters controlled the straying: the fraction
-of fish $f_{\mathrm{stray}}$ that stray from their natal stream in any
-generation and the rate $m$ at which this straying between streams
-decays with distance. We calculated the number of salmon straying from
-stream $j$ to stream $i$ as
+We implemented straying as in \citet{cooper1999}.
+We set up the metapopulation in a simple scenario: we arranged the populations in a line and those that were nearer to each other were more likely to stray between each other.
+Two parameters controlled the straying: the fraction of fish $f_{\mathrm{stray}}$ that stray from their natal stream in any generation and the rate $m$ at which this straying between streams decays with distance.
+We calculated the number of salmon straying from stream $j$ to stream $i$ as
 
@@
 
-\noindent where $R_{tj}$ is the number of returning salmon at generation
-$t$ whose natal stream was stream $j$. The subscript $k$ represents a
-stream ID and $n$ the number of populations. The denominator is a
-normalizing constant to ensure the desired fraction of fish stray. See
-Figure \ref{f:stray} for an example straying matrix.
+\noindent where $R_{tj}$ is the number of returning salmon at generation $t$ whose natal stream was stream $j$.
+The subscript $k$ represents a stream ID and $n$ the number of populations.
+The denominator is a normalizing constant to ensure the desired fraction of fish stray.
+See Figure \ref{f:stray} for an example straying matrix.
 
@@
 
-Our simulation used a simple set of rules to establish escapement
-targets and harvest the fish. Every $f_\mathrm{assess}$ years (default
-of five years) our simulation fitted a spawner-return function and
-target harvest rate $H_{\mathrm{tar}}$ was set based on
-\citet{hilborn1992} as
+Our simulation used a simple set of rules to establish escapement targets and harvest the fish.
+Every $f_\mathrm{assess}$ years (default of five years) our simulation fitted a spawner-return function and target harvest rate $H_{\mathrm{tar}}$ was set based on \citet{hilborn1992} as
 
@@
 
-\noindent where $A$ represents the return abundance and $a$ and $b$
-represent the Ricker model parameters. We included implementation
-uncertainty in the actual harvest rate $H_{\mathrm{act}}$ as
+\noindent where $A$ represents the return abundance and $a$ and $b$ represent the Ricker model parameters. We included implementation uncertainty in the actual harvest rate $H_{\mathrm{act}}$ as
 
@@
 
-\noindent where $\alpha_h$ and $\beta_h$ are the location and shape
-parameters in a beta distribution. They can be calculated from the
-desired mean $\mu_h$ and standard deviation $\sigma_h$ as
-\citep[p.~97]{morgan1990}
+\noindent where $\alpha_h$ and $\beta_h$ are the location and shape parameters in a beta distribution.
+They can be calculated from the desired mean $\mu_h$ and standard deviation $\sigma_h$ as \citep[p.~97]{morgan1990}
 
@@
 
-\noindent Further, to establish a range of spawner-return values and to
-mimic the start of an open-access fishery, for the first 30 years we
-drew the fraction of fish harvested randomly from a uniform distribution
-between 0.1 and 0.9. We discarded these initial 30 years as a burn-in
-period throughout our analyses.
+\noindent Further, to establish a range of spawner-return values and to mimic the start of an open-access fishery, for the first 30 years we drew the fraction of fish harvested randomly from a uniform distribution between 0.1 and 0.9.
+We discarded these initial 30 years as a burn-in period throughout our analyses.
 
@@
 
-We evaluated portfolio performance under short- and long-term
-environmental dynamics. We represented short-term dynamics as a
-stationary first-order autoregressive process, AR(1), with correlation
-$\rho_e$
+We evaluated portfolio performance under short- and long-term environmental dynamics.
+We represented short-term dynamics as a stationary first-order autoregressive process, AR(1), with correlation $\rho_e$
 
@@
 
-\noindent where $e_t$ represents the environmental value in generation
-$t$ and $d$ represents normally distributed deviations of mean 0 and
-standard deviation $\sigma_d$. We represented long-term environmental
-dynamics as a linear shift in the environmental value through time
+\noindent where $e_t$ represents the environmental value in generation $t$ and $d$ represents normally distributed deviations of mean 0 and standard deviation $\sigma_d$.
+We represented long-term environmental dynamics as a linear shift in the environmental value through time
 
@@
 
-\noindent where $\beta_e$ represents the slope. To maintain a balanced
-response, we centered the trend by subtracting the mean
-$\overline{\beta_e t}$ so that midway through the simulation (after any
-burn-in period) the environmental value was at the mean environmental
-tolerance.
+\noindent where $\beta_e$ represents the slope.
+To maintain a balanced response, we centered the trend by subtracting the mean $\overline{\beta_e t}$ so that midway through the simulation (after any burn-in period) the environmental value was at the mean environmental tolerance.
 
@@
 
-We evaluated two sets of conservation rules of thumb: (1) spatial
-response diversity conservation strategies in an idealistic scenario
-where you can detect response diversity, and (2) a more realistic
-scenario where we know little about response diversity and we're left
-with a choice of how many populations to conserve.
+We evaluated two sets of conservation rules of thumb: (1) spatial response diversity conservation strategies in an idealistic scenario where you can detect response diversity, and (2) a more realistic scenario where we know little about response diversity and we're left with a choice of how many populations to conserve.
 
-We evaluated four spatial conservation rules of thumb (Figure
-\ref{f:curves}b--e). In all spatial scenarios, we conserved four
-populations and set the unfished equilibrium biomass of the remaining
-populations to near elimination (five salmon). These reduced populations
-could still receive straying salmon but were unlikely to rebuild on
-their own to a substantial abundance. The four scenarios were:
+We evaluated four spatial conservation rules of thumb (Figure \ref{f:curves}b--e).
+In all spatial scenarios, we conserved four populations and set the unfished equilibrium biomass of the remaining populations to near elimination (five salmon).
+These reduced populations could still receive straying salmon but were unlikely to rebuild on their own to a substantial abundance.
+The four scenarios were:
 
@@
 
-In reality we rarely know precise levels of response diversity. We
-therefore additionally considered a case where the conservation was
-randomly assigned with respect to response diversity but where different
-numbers of streams could be conserved. We considered conserving from two
-to 16 streams. Similarly to the spatial strategies, we reduced the
-capacity of the remaining streams to the nominal level of five salmon.
+In reality we rarely know precise levels of response diversity.
+We therefore additionally considered a case where the conservation was randomly assigned with respect to response diversity but where different numbers of streams could be conserved.
+We considered conserving from two to 16 streams.
+Similarly to the spatial strategies, we reduced the capacity of the remaining streams to the nominal level of five salmon.
 
@@
 
-Given strong short-term environmental fluctuations, conserving response
-diversity buffers the risk properties of an ecological portfolio (Fig.
-4a). In our simulation, the median variance of generation-to-generation
-rate of change in abundance was X times lower given balanced response
-diversity (full range of responses or most stable only vs.~conserving
-one half or the other). In fact, even though by conserving the full
-range of responses, the portfolio was comprised of warm and
-cool-thriving populations that were more variable on their own, each was
-balanced by an opposing population. The portfolio risk was therefore
-comparable between the full range of responses and most stable only
-portfolios.
+Given strong short-term environmental fluctuations, conserving response diversity buffers the risk properties of an ecological portfolio (Fig.~\ref{f:sp-mv}a).
+In our simulation, the median variance of generation-to-generation rate of change in abundance was X times lower given balanced response diversity (full range of responses or most stable only vs.~conserving one half or the other).
+In fact, even though by conserving the full range of responses, the portfolio was comprised of warm and cool-thriving populations that were more variable on their own, each was balanced by an opposing population.
+The portfolio risk was therefore comparable between the full range of responses and most stable only portfolios.
 
-We can see the mechanism behind these portfolio properties by inspecting
-example population time series (Fig. 4c, d). If only the upper or lower
-half of response diversity is conserved, the portfolio tends to do well
-or poorly depending on the environmental conditions (Fig. 4d). This risk
-is buffered with balanced response diversity (Fig. 4c).
+We can see the mechanism behind these portfolio properties by inspecting example population time series (Fig.~\ref{f:sp-mv}c, d).
+If only the upper or lower half of response diversity is conserved, the portfolio tends to do well or poorly depending on the environmental conditions (Fig.~\ref{f:sp-mv}d).
+This risk is buffered with balanced response diversity (Fig.~\ref{f:sp-mv}c).
 
@@
 
-Given long-term environmental change, the choice of which populations to
-conserve affects the return properties of an ecological portfolio (Fig.
-4b). By conserving balance response diversity, an ecological manager is
-hedging his or her bets on what will happen with the environment and how
-the populations will respond. The typical return for a balanced response
-diversity strategy was zero --- the metapopulation neither increased or
-decreased in abundance in the long run. By conserving only the upper or
-lower half of response diversity, a conservation manager is putting all
-his or her eggs in one basket --- the metapopulation might do really
-well through time or it might do really poorly. The example
-metapopulation abundance time series (Fig. 4d, f) illustrate this
-effect. By conserving response diversity, when one population is doing
-poorly, another is doing well and the metapopulation abundance remains
-stationary through time.
+Given long-term environmental change, the choice of which populations to conserve affects the return properties of an ecological portfolio (Fig.~\ref{f:sp-mv}b).
+By conserving balance response diversity, an ecological manager is hedging his or her bets on what will happen with the environment and how the populations will respond.
+The typical return for a balanced response diversity strategy was zero --- the metapopulation neither increased or decreased in abundance in the long run.
+By conserving only the upper or lower half of response diversity, a conservation manager is putting all his or her eggs in one basket --- the metapopulation might do really well through time or it might do really poorly.
+The example metapopulation abundance time series (Fig.~\ref{f:sp-mv}d, f) illustrate this effect.
+By conserving response diversity, when one population is doing poorly, another is doing well and the metapopulation abundance remains stationary through time.
 
-Notably, in theses simulations, if a managers invested in the
-populations that were doing well at the beginning they would have had
-the lowest rate-of-return portfolio in the end (purple portfolios in
-Fig. 4b).
+Notably, in theses simulations, if a managers invested in the populations that were doing well at the beginning they would have had the lowest rate-of-return portfolio in the end (purple portfolios in Fig.~\ref{f:sp-mv}b).
 
-Spatial conservation strategies in the face of longterm environmental
-change hinge on whether you ``get it right'' --- whether you choose just
-the right populations to conserve.
+Spatial conservation strategies in the face of longterm environmental change hinge on whether you ``get it right'' --- whether you choose just the right populations to conserve.
 
@@
 
-Given a scenario where we don't know the distribution of
-population-level response diversity, portfolio optimization informs us
-about the risk buffering from maintaining multiple populations (Fig. 5).
+Given a scenario where we don't know the distribution of population-level response diversity, portfolio optimization informs us about the risk buffering from maintaining multiple populations (Fig.~\ref{f:n-mv}).
 In short, investing in more populations buffers portfolio risk.
@@
 
-Given short-term environmental noise, conserving more populations
-buffers portfolio risk while the random conservation of response
-diversity creates a spread of metapopulation risk for the same number of
-populations conserved (Fig. 5a). For example, a metapopulation with
-eight conserved populations is X times less risky than a metapopulation
-with only four. We can see this risk-buffering effect through example
-metapopulations in Fig. 5c and 5d. We note that the risk-return axes of
-portfolio optimization ignore the absolute-abundance dimension (Fig.
-5d). As one would expect, conserving fewer populations also results in
-lower-abundance metapopulations.
+Given short-term environmental noise, conserving more populations buffers portfolio risk while the random conservation of response diversity creates a spread of metapopulation risk for the same number of populations conserved (Fig.~\ref{f:n-mv}a).
+For example, a metapopulation with eight conserved populations is X times less risky than a metapopulation with only four.
+We can see this risk-buffering effect through example metapopulations in Fig.~\ref{f:n-mv}c and \ref{f:n-mv}d.
+We note that the risk-return axes of portfolio optimization ignore the absolute-abundance dimension (Fig.~\ref{f:n-mv}d).
+As one would expect, conserving fewer populations also results in lower-abundance metapopulations.
 
@@
 
-Given long-term environmental noise, conserving more populations also
-buffers portfolio risk. However, in contrast to the short-term
-environmental noise scenario, the unknown response diversity creates a
-spread of possible metapopulation return for the same number of
-conserved populations (Fig. 5b). Here, the number of populations
-conserved buffers non-systematic (i.e.~not-environmentally-driven)
-stochasticity.
+Given long-term environmental noise, conserving more populations also buffers portfolio risk.
+However, in contrast to the short-term environmental noise scenario, the unknown response diversity creates a spread of possible metapopulation return for the same number of conserved populations (Fig.~\ref{f:n-mv}b).
+Here, the number of populations conserved buffers non-systematic (i.e.~not-environmentally-driven) stochasticity.
 

### 2013-07-08 inst/ms/ms.tex
Major additions to text; especially introduction

@@
-\emph{Other title ideas:}
+\section{Abstract}
 
-\begin{itemize}
-\item
-  Prioritizing metapopulation conservation through the lens of portfolio theory
-\item
-  Portfolio optimization reveals\ldots{}
-\item
-  Portfolio theory reveals conservation strategies that\ldots{}
-\item
-  Intelligent tinkering with ecological portfolios: patterns of environmental response diversity drive salmon conservation priorities
-\item
-  Keeping every cog and wheel\ldots{}
-\end{itemize}
-
-\textbf{Ecology Letters}
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  letters: 5000 words, 6 figures or tables
-\item
-  ideas and perspectives: 7500 words, 10 figures (needs 300 word proposal)
-\end{itemize}
-
-\begin{quote}
-Ecology Letters is a forum for the very rapid publication of the most novel research in ecology, research that is not yet in the public domain.
-Manuscripts relating to the ecology of all taxa, in any biome and geographic area will be considered, and priority will be given to those papers exploring or testing clearly stated hypotheses.
-The journal publishes concise papers that merit urgent publication by virtue of their originality, general interest and their contribution to new developments in ecology.
-We discourage purely descriptive papers and those merely confirming or extending results of previous work.
-\end{quote}
-
-\textbf{Ecological Applications}
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  60 pages including everything
-\end{itemize}
-
-\begin{quote}
-Ecological Applications is concerned broadly with the applications of ecological science to environmental problems.
-It publishes papers that develop scientific principles to support environmental decision-making, as well as papers that discuss the application of ecological concepts to environmental issues, policy, and management.
-Papers may report on experimental tests, actual applications, scientific decision support techniques, economic analyses, social implications of environmental issues, or other relevant topics.
-Statistical or experimental methods papers that support research and applications are welcome.
-Papers submitted to Ecological Applications should be accessible to both scholars and practitioners.
-\end{quote}
-
-\section{Long and rambling abstract that's a bit old now}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  Managing risk is fundamental to the conservation of an endangered species.
-When an endangered species exists as a metapopulation, we can manage risk at two levels: at the population level or at the metapopulation level.
-Whereas risk is typically managed at the population level, a portfolio approach to managing risk might consider how conservation affects the ``weight'' of each population in a metapopulation ``portfolio''.
-\item
-  Here, we ask how a portfolio approach to managing risk can inform the spatial conservation of metapopulations in a changing world.
-To answer this, we develop a salmon metapopulation simulation in which population-specific productivity is driven by spatially-distributed environmental tolerance and patterns of short- and long-term environmental change.
-We then implement different spatial conservation ``rules of thumb'' that control the population-specific carrying capacities and evaluate the salmon portfolios along risk and return axes, similarly to how financial portfolios are assessed.
-\item
-  Our results show, first, that maintaining populations with a variety of environmental tolerances gives the best chance at an efficient ecological portfolio --- minimizing metapopulation variance while maximizing metapopulation growth rate.
-This finding emphasizes the risk of allowing large spatial blocks of habitat destruction, say through the development of dams.
-Second, we show that focusing on well-performing stocks now at the detriment of others is at best equivalent to a risky but efficient portfolio and is more likely a risky and inefficient portfolio --- it neither minimizes metapopulation variance nor maximizes growth rate compared to other strategies.
-Third, we show that maintaining more populations reduces metapopulation risk for the same spatial conservation strategy.
-Given a lack of knowledge of how populations respond to the environment, the most risk-averse approach is to conserve as many populations as possible.
-\item
-  Our findings highlight three key points: (1) the conservation priority of maintaining biocomplexity and therefore environmental response diversity, (2) the research priority of identifying differences in environmental tolerance given predicted environmental changes, and (3) the utility of considering risk for groups of fish stocks --- especially given environmental, biological, and implementation uncertainty --- through the lens of portfolio theory.
-\end{enumerate}
+{[}@convertino2013; @wade2013{]}
 
@@
 
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
 Managing risk is fundamental to the conservation of an endangered species.
+When an endangered species exists as a metapopulation, we can manage risk at two levels: at the population level or at the metapopulation level {[}REF{]}.
+Typically we treat sources of risk for metapopulations as exogenous and manage risk by altering fishing or hunting, implementing reserves, or improving connectivity for subpopulations {[}e.g.
+@akcakaya2007{]}.
 
-  \begin{itemize}
-  \itemsep1pt\parskip0pt\parsep0pt
-  \item
-    risk is usually assessed on a population-by-population basis
-  \item
-    for metapopulations, we can consider risk for components or risk for the whole
-  \end{itemize}
-\item
-  The management of financial portfolios provides another way of considering risk for metapopulations.
+The management of financial portfolios provides another way to consider risk.
+Economists consider the risk and performance of a financial portfolio based on the weighting of individual investments (called assets) that make up the portfolio.
+Modern Portfolio Theory (MPT) proposes that there is a set of portfolios that maximizes expected return for a level of expected risk or minimizes expected risk for a level of expected return \citep{markowitz1952, markowitz1959}.
+Similarly, expected growth rate and variance of a metapopulation is a function of the variance, covariance, and size of the individual populations.
+A portfolio approach to managing risk for a metapopulation might therefore consider how conservation actions affect the weight of each population in a metapopulation portfolio.
 
-  \begin{itemize}
-  \itemsep1pt\parskip0pt\parsep0pt
-  \item
-    performance of weighted assets
-  \item
-    efficient frontier concept
-  \item
-    the parallels with ecological metapopulations
-  \end{itemize}
-\item
-  Pacific salmon are an ideal model to consider under the framework of portfolio prioritization because of their metapopulation structure, their societal importance and threatened status, and the need to prioritize conservation efforts.
+Managing Pacific salmon under the uncertainty of climate change is an ideal scenario to consider through the lens of portfolio theory for three reasons.
+(1) Salmon often form metapopulations {[}REF{]} and we can consider the metapopulation in a river-catchment as a portfolio and the stream populations as assets.
+Continuing the analogy, predators and fisheries often integrate across multiple streams, acting as investors in the salmon portfolio.
+Fisheries managers and conservation agencies then act as portfolio managers through actions such choosing which salmon habitat to prioritize for protection or restoration.
+(2) Many Pacific salmon metapopulations are highly threatened {[}REFs{]} and will likely become more at risk as threats such as overfishing, dams, roads, logging, and particularly climate change intensify {[}REFs throughout{]}.
+(3) Salmon are highly valued by society, fishers, conservation groups, and indigenous people {[}REFs{]}.
+Although civil society allocates extensive resources to conserving salmonids (e.g.~XX in XX) the scale of the problem demands a prioritization of conservation efforts {[}REFs{]}.
 
-  \begin{itemize}
-  \itemsep1pt\parskip0pt\parsep0pt
-  \item
-    portfolio-like structure
-
-    \begin{itemize}
-    \itemsep1pt\parskip0pt\parsep0pt
-    \item
-      documented metapopulation structure
-    \item
-      predators and fisheries often integrate across stocks
-    \end{itemize}
-  \item
-    Highly threatened; highly valued; face many threats
-
-    \begin{itemize}
-    \itemsep1pt\parskip0pt\parsep0pt
-    \item
-      damns
-    \item
-      roads
-    \item
-      other habitat degradation
-    \item
-      logging
-    \item
-      fishing
-    \item
-      changing climate
-    \end{itemize}
-  \item
-    There's a need for prioritization
-
-    \begin{itemize}
-    \itemsep1pt\parskip0pt\parsep0pt
-    \item
-      because of the highly threatened nature
-    \item
-      conflicting goals for resource use
-    \item
-      limited resources
-    \item
-      Existing prioritization schemes
-    \end{itemize}
-  \end{itemize}
-\item
-  Portfolio stabilization through systematic asynchrony in salmon population dynamics can be driven by stocks responding uniquely to the same environment (`response diversity') or by physical features of the environment causing stocks to experience different environmental forces (`environmental filter').
-
-  \begin{itemize}
-  \itemsep1pt\parskip0pt\parsep0pt
-  \item
-    different responses to the environment: response diversity / biocomplexity
-
-    \begin{itemize}
-    \itemsep1pt\parskip0pt\parsep0pt
-    \item
-      genetic variation by stock
-    \item
-      can be driven through historical conditions
-    \item
-      we focus on this here
-    \end{itemize}
-  \item
-    experiencing different environment: environmental filter concept (Moran like)
-
-    \begin{itemize}
-    \itemsep1pt\parskip0pt\parsep0pt
-    \item
-      spatially spread out, geographical and hydrological features filter the environment to create a variety of forces
-    \item
-      Typically a combination of both
-    \end{itemize}
-  \item
-    Conversely, forces acting against asynchrony
-
-    \begin{itemize}
-    \itemsep1pt\parskip0pt\parsep0pt
-    \item
-      inverse: loss of population diversity
-    \item
-      inverse: moran
-    \item
-      but also: dispersal
-    \item
-      and: trophic interactions with synchronized species
-    \end{itemize}
-  \end{itemize}
-\end{itemize}
+A diversified financial portfolio reduces risk by exploiting the covariance between assets --- ideally the same market forces cause one stock to rise if another one falls --- and there are two key mechanisms that reduce metapopulation risk by creating asynchronous population dynamics.
+First, localized habitat features can filter the environment creating unique conditions within a spread out metapopulation {[}REFS{]} (i.e.~Moran effect REF).
+For example, landscape topology can affect stream flow causing different streams of salmon to experience unique conditions {[}REF{]}.
+Second, diversity of life-histories, behaviours, personalities, and genetics can cause populations to respond uniquely to the same environment (i.e.~response diversity REF and biocomplexity REF).
+These unique traits can be phenotypic or genetic and derived from short-term adaptations {[}REF{]} or long-term evolution {[}REF{]} to historical conditions.
+In this paper we focus on the latter, response-diversity mechanism.
 
 Here, we ask how a portfolio approach to management can inform the conservation of metapopulations in a changing world.
-We ask two primary questions: (1) What does portfolio optimization tell us about different spatial approaches to prioritizing metapopulation conservation assuming that response diversity is spatially distributed?
-(2) If we don't know how response diversity is distribute, what does portfolio optimization tell us about how many populations we should conserve?
-To answer these questions, we develop a salmon metapopulation simulation in which population-specific productivity is driven by spatially-distributed environmental tolerance and patterns of short- and long-term climatic change.
-We then implement different conservation ``rules of thumb'' that control the population-level carrying capacities and evaluate the salmon portfolios along risk and return axes, as a portfolio manager would.
-Our findings illustrate that: (1) conserving response diversity buffers metapopulation risk given short-term climate forcing and rate of return given long-term climate forcing, (2) conserving more subpopulations buffers risk regardless of the climate trend, and (3) considering metapopulations through the lens of portfolio theory provides a useful additional dimension through which we can evaluate conservation strategies.
+We ask two primary questions: (1) What does portfolio theory tell us about spatial approaches to prioritizing metapopulation conservation?
+(2) If we don't know how response-diversity is distributed, what does portfolio theory tell us about how many populations to conserve?
+To answer these questions, we develop a salmon metapopulation simulation in which spatially-distributed thermal tolerance and patterns of short- and long-term climatic change drive population-specific productivity.
+We then implement different conservation ``rules of thumb'' that control the population-level carrying capacities and evaluate the salmon portfolios along risk and return axes, as a financial portfolio manager might.
+We show that conserving response diversity buffers metapopulation risk given short-term climate forcing and metapopulation growth rate given long-term climate warming.
+We then show that conserving more subpopulations buffers risk regardless of response diversity or climate trend, and we conclude that considering metapopulations through portfolio theory provides a useful additional dimension to evaluate conservation strategies.
 
@@
 Under two kinds of environmental regimes we tested different conservation rules of thumb and evaluated these plans in risk-return space similar to how financial managers evaluate financial portfolios.
-See Figure \ref{f:sim-flow} for an illustration of the overall simulation structure.
-See Table \ref{t:pars} for a list of the input parameters to our simulation and their default values.
+We illustrate the overall simulation structure in Fig.~1.
 We provide a package \texttt{metafolio} for the statistical software \texttt{R} \citep{r2013} as an appendix, to carry out the simulations and analyses we describe in this paper.
@@
 
-In our ecological portfolios, we defined assets as stream-level populations and the portfolios as salmon metapopulations (Table \ref{t:port}).
+In our ecological portfolios, we defined assets as stream-level populations and the portfolios as salmon metapopulations.
 We use the terms \emph{stream} and \emph{populations} interchangeably to represent the portfolio assets.
 We defined the portfolio investors as the stakeholders in the fishery and metapopulation performance.
-For example, we could consider fisheries managers, conservation agencies, or First Nations groups as investors.
+CHANGE THIS For example, we could consider fisheries managers, conservation agencies, or First Nations groups as investors.
 We defined asset value as the abundance of returning salmon in each stream and value of the portfolio as the overall metapopulation abundance.
@@
 
-\begin{equation}
-R_{ti} = S_{ti}e^{a_{ti}(1-S_{ti}/b_i) + w_{ti}}
-\end{equation}
+\[R_{i(t)} = S_{i(t)}e^{a_{i(t)}(1-S_{i(t)}/b_i) + w_{i(t)}}\]
 
-\noindent where $t$ represents a generation time, $i$ represents a population, $R$ is the number of returns, $S$ is the number of spawners, $a$ is the productivity parameter (which can vary with the environment), and $b$ is the density-dependent term (which is used as the asset weights in the portfolios).
-The term $w_{ti}$ represents first-order autocorrelated error.
-Formally, $w_{ti} = w_{ti-1} \rho_w + r_{ti}$, where $r_{ti}$ represents independent and normally distributed error with mean 0 and standard deviation of $\sigma_r$.
+\noindent where $i$ represents a population, $t$ a generation time, $R$ the number of returns, $S$ the number of spawners, $a$ the productivity parameter (which can vary with the environment), and $b$ the density-dependent term (which is used as the asset weights in the portfolios).
+The term $w_{i(t)}$ represents first-order autocorrelated error.
+Formally, $w_{i(t)} = w_{ti-1} \rho_w + r_{i(t)}$, where $r_{i(t)}$ represents independent and normally distributed error with mean 0 and standard deviation of $\sigma_r$.
 The parameter $\rho_w$ represents the correlation between residuals from subsequent generations.
@@
 For example, a large investment in a stream was represented by a larger unfished equilibrium stock size $b$ for stream $i$.
-The productivity parameters $a_{ti}$ were controlled by the interaction between an environmental signal and the stream-level population environmental-tolerance curves.
+The productivity parameters $a_{i(t)}$ were controlled by the interaction between an environmental signal and the stream-level population environmental-tolerance curves.
 
@@
 
-\begin{equation}
-  a_{ti} =
+\[a_{i(t)} =
   \begin{cases}
     W_i (e_t - e_i^{\mathrm{opt}})^2 + a_i^{\mathrm{max}},
-      & \text{if } a_{ti} > 0\\
-      0, & \text{if } a_{ti} \leq 0
-  \end{cases}
-\end{equation}
+      & \text{if } a_{i(t)} > 0\\
+      0, & \text{if } a_{i(t)} \leq 0
+  \end{cases}\]
 
 \noindent where $W_i$ controls the width of the curve for population $i$, $e_t$ represents the environmental value at generation $t$, $e_i^{\mathrm{opt}}$ represents the optimal environmental value for population $i$, and $a_i^{\mathrm{max}}$ represents the maximum possible $a$ value for population $i$.
-We set the $W_i$ parameters (Table \ref{t:pars}) and calculated the $a_i^{\mathrm{max}}$ parameters so that the area under each curve $A_i$ was equal.
-See Figure \ref{f:curves}a for example environmental-tolerance curves.
+We set the $W_i$ parameters (Table S1) and calculated the $a_i^{\mathrm{max}}$ parameters so that the area under each curve $A_i$ was equal.
+See Figure 2a for example environmental-tolerance curves.
 
@@
 
-\begin{equation}
-  \mathrm{strays}_{tij} = f_{\mathrm{stray}} R_{tj}
+\[\mathrm{strays}_{tij} = f_{\mathrm{stray}} R_{tj}
     \frac{e^{-m \lvert i-j \rvert }}
-      {\displaystyle\sum\limits_{\substack{k = 1 \\
-    k \neq j}}^{n} e^{-m \lvert k-j \rvert }}
-  \label{eq:stray}
-\end{equation}
+      {\displaystyle\sum\limits_{
+        \substack{k = 1 \\ k \neq j}}^{n} 
+        e^{-m \lvert k-j \rvert }}\]
 
@@
 Our simulation used a simple set of rules to establish escapement targets and harvest the fish.
-Every $f_\mathrm{assess}$ years (default of five years) our simulation fitted a spawner-return function and target harvest rate $H_{\mathrm{tar}}$ was set based on \citet{hilborn1992} as
+Every five years our simulation fitted a spawner-return function and target harvest rate $H_{\mathrm{tar}}$ was set based on \citet{hilborn1992} as
 
-\begin{equation}
-  H_{\mathrm{tar}} = \frac{A}{b (0.5 - 0.07a)}
-  \label{eq:esc}
-\end{equation}
+\[H_{\mathrm{tar}} = \frac{A}{b (0.5 - 0.07a)}
+  \label{eq:esc}\]
 
@@
 
-\begin{equation}
-  H_{\mathrm{act}} = \mathrm{beta}(\alpha_h, \beta_h)
-\end{equation}
+\[H_{\mathrm{act}} = \mathrm{beta}(\alpha_h, \beta_h)\]
 
 \noindent where $\alpha_h$ and $\beta_h$ are the location and shape parameters in a beta distribution.
-They can be calculated from the desired mean $\mu_h$ and standard deviation $\sigma_h$ as \citep[p.~97]{morgan1990}
+They can be calculated from the desired mean $H_{\mathrm{tar}}$ and standard deviation $\sigma_h$ as \citep[p.~97]{morgan1990}
 
-\begin{align}
-  \alpha_h &= \mu_h^2
+\[\begin{aligned}
+  \alpha_h &= H_{\mathrm{tar}}^2
                 \left(
-                \frac{1 - \mu_h}{\sigma_h^2} - \frac{1}{\mu_h}
+                \frac{1 - H_{\mathrm{tar}}}{\sigma_h^2} - \frac{1}{H_{\mathrm{tar}}}
                 \right)\\
-   \beta_h &= \alpha \left({\frac{1}{\mu_h} - 1}\right).
-\end{align}
+   \beta_h &= \alpha \left({\frac{1}{H_{\mathrm{tar}}} - 1}\right).\end{aligned}\]
 
-\noindent Further, to establish a range of spawner-return values and to mimic the start of an open-access fishery, for the first 30 years we drew the fraction of fish harvested randomly from a uniform distribution between 0.1 and 0.9.
-We discarded these initial 30 years as a burn-in period throughout our analyses.
+Further, to establish a range of spawner-return values and to mimic the start of an open-access fishery, for the first 30 years we drew the fraction of fish harvested randomly from a uniform distribution between 0.1 and 0.9.
+We discarded these initial 30 years as a burn-in period.
 
@@
 
-\begin{equation}
-  e_t = e_{t-1} \rho_e + d_t, d_t \sim \mathrm{N}(0, \sigma_d)
-\end{equation}
+\[e_t = e_{t-1} \rho_e + d_t, d_t \sim \mathrm{N}(0, \sigma_d)\]
 
@@
 
-\begin{equation}
-  e_t = \beta_e t - \overline{\beta_e t}
-\end{equation}
+\[e_t = \beta_e t - \overline{\beta_e t}\]
 
@@
 To maintain a balanced response, we centered the trend by subtracting the mean $\overline{\beta_e t}$ so that midway through the simulation (after any burn-in period) the environmental value was at the mean environmental tolerance.
+
+Mention Fig.~3 and maybe put it before Fig.~2?
 
@@
 
-We evaluated four spatial conservation rules of thumb (Figure \ref{f:curves}b--e).
+We evaluated four spatial conservation rules of thumb (Fig.~2b--e).
 In all spatial scenarios, we conserved four populations and set the unfished equilibrium biomass of the remaining populations to near elimination (five salmon).
@@
 
-Given strong short-term environmental fluctuations, conserving response diversity buffers the risk properties of an ecological portfolio (Fig.~\ref{f:sp-mv}a).
+Given strong short-term environmental fluctuations, conserving response diversity buffers the risk properties of an ecological portfolio (Fig.~4a).
 In our simulation, the median variance of generation-to-generation rate of change in abundance was X times lower given balanced response diversity (full range of responses or most stable only vs.~conserving one half or the other).
@@
 
-We can see the mechanism behind these portfolio properties by inspecting example population time series (Fig.~\ref{f:sp-mv}c, d).
-If only the upper or lower half of response diversity is conserved, the portfolio tends to do well or poorly depending on the environmental conditions (Fig.~\ref{f:sp-mv}d).
-This risk is buffered with balanced response diversity (Fig.~\ref{f:sp-mv}c).
+We can see the mechanism behind these portfolio properties by inspecting example population time series (Fig.~4c, d).
+If only the upper or lower half of response diversity is conserved, the portfolio tends to do well or poorly depending on the environmental conditions (Fig.~4d).
+This risk is buffered with balanced response diversity (Fig.~4c).
 
@@
 
-Given long-term environmental change, the choice of which populations to conserve affects the return properties of an ecological portfolio (Fig.~\ref{f:sp-mv}b).
-By conserving balance response diversity, an ecological manager is hedging his or her bets on what will happen with the environment and how the populations will respond.
+Given long-term environmental change, the choice of which populations to conserve affects the return properties of an ecological portfolio (Fig.~4b).
+By conserving balanced response diversity, an ecological manager is hedging his or her bets on what will happen with the environment and how the populations will respond.
 The typical return for a balanced response diversity strategy was zero --- the metapopulation neither increased or decreased in abundance in the long run.
 By conserving only the upper or lower half of response diversity, a conservation manager is putting all his or her eggs in one basket --- the metapopulation might do really well through time or it might do really poorly.
-The example metapopulation abundance time series (Fig.~\ref{f:sp-mv}d, f) illustrate this effect.
+The example metapopulation abundance time series (Fig.~4d, f) illustrate this effect.
 By conserving response diversity, when one population is doing poorly, another is doing well and the metapopulation abundance remains stationary through time.
 
-Notably, in theses simulations, if a managers invested in the populations that were doing well at the beginning they would have had the lowest rate-of-return portfolio in the end (purple portfolios in Fig.~\ref{f:sp-mv}b).
+Notably, in these simulations, if a managers invested in the populations that were doing well at the beginning they would have had the lowest rate-of-return portfolio in the end (purple portfolios in Fig.~4b).
 
@@
 
-Given a scenario where we don't know the distribution of population-level response diversity, portfolio optimization informs us about the risk buffering from maintaining multiple populations (Fig.~\ref{f:n-mv}).
+Given a scenario where we don't know the distribution of population-level response diversity, portfolio optimization informs us about the risk buffering from maintaining multiple populations (Fig.~5).
 In short, investing in more populations buffers portfolio risk.
@@
 
-Given short-term environmental noise, conserving more populations buffers portfolio risk while the random conservation of response diversity creates a spread of metapopulation risk for the same number of populations conserved (Fig.~\ref{f:n-mv}a).
+Given short-term environmental noise, conserving more populations buffers portfolio risk while the random conservation of response diversity creates a spread of metapopulation risk for the same number of populations conserved (Fig.~5a).
 For example, a metapopulation with eight conserved populations is X times less risky than a metapopulation with only four.
-We can see this risk-buffering effect through example metapopulations in Fig.~\ref{f:n-mv}c and \ref{f:n-mv}d.
-We note that the risk-return axes of portfolio optimization ignore the absolute-abundance dimension (Fig.~\ref{f:n-mv}d).
+We can see this risk-buffering effect through example metapopulations in Fig.~5c and 5d.
+We note that the risk-return axes of portfolio optimization ignore the absolute-abundance dimension (Fig.~5d).
 As one would expect, conserving fewer populations also results in lower-abundance metapopulations.
@@
 Given long-term environmental noise, conserving more populations also buffers portfolio risk.
-However, in contrast to the short-term environmental noise scenario, the unknown response diversity creates a spread of possible metapopulation return for the same number of conserved populations (Fig.~\ref{f:n-mv}b).
+However, in contrast to the short-term environmental noise scenario, the unknown response diversity creates a spread of possible metapopulation return for the same number of conserved populations (Fig.~5b).
 Here, the number of populations conserved buffers non-systematic (i.e.~not-environmentally-driven) stochasticity.
@@
 
-\bibliographystyle{apalike}
+\bibliographystyle{ecologyletters2}
 
 \bibliography{jshort,ms}
-
-\clearpage
-
-\section{Tables}
-
-\begin{table}[h!]
-\centering
-\small
-\caption{Components of salmon metapopulation portfolios KEEP THIS?}
-\begin{tabular}{p{3.6cm}p{7.5cm}}
-\toprule
-Component          & Definition for the salmon portfolio\\
-\midrule
-Assets             & Stream-level salmon populations; possibly a Viable Salmonid Population\\
-Portfolio          & The salmon metapopulation; possibly an Evolutionarily Significant Unit\\
-Portfolio managers & Salmon managers\\
-Investors          & Salmon managers, conservation agency, or salmon fishers\\
-Asset weights      & Carrying capacity (specifically the $b$ parameters in a Ricker model)\\
-Asset returns      & Rate of change of generation-to-generation salmon metapopulation abundance\\
-Asset risk         & Variance of generation-to-generation salmon metapopulation abundance\\
-\bottomrule
-\end{tabular}
-\label{t:port}
-\end{table}
-
-\clearpage
-
-\begin{table}[h!]
-\centering
-\footnotesize
-\caption{Input parameters to the salmon metapopulation simulation with default values.}
-\begin{tabular}{p{7.7cm}p{1.4cm}p{3.6cm}}
-\toprule
-Description                                                          & Symbol                & Value \\
-\midrule
-
-\textit{Population dynamics parameters}                              &                       & \\
-Stock-recruit residual standard deviation (on log scale)             & $\sigma_r$            & 0.30  \\
-AR(1) serial correlation of stock-recruit residuals                  & $\rho_w$              & 0.40  \\
-Fraction of fish that stray from natal streams                       & $f_{\mathrm{stray}}$  & 0.02  \\
-Exponential rate of decay of straying with distance                  & $m$                   & 0.3  \\
-
-\noalign{\vskip 3mm}
-\textit{Environmental parameters}                                    &                       & \\
-Width of environmental-tolerance curves for populations $i$ 1 to $n$ & $W_i$                 & seq(0.05, 0.02, 0.02 0.05, length = n pop)\\
-Optimum environmental value for populations $i$ 1 to $n$             & $e_i^{\mathrm{opt}}$  & seq(13, 19, length = n pop)\\
-Area under each environmental-tolerance curve in environmental units & $A$                   & 30\\
-
-Standard deviation of short-term environmental fluctuations          & $\sigma_d$            & 2 \\
-AR(1) autocorrelation of short-term environmental fluctuations       & $\rho_e$              & 0.1 \\
-Slope of long-term environmental signal                              & $\beta_e$             & 0.114 \\
-
-\noalign{\vskip 3mm}
-\textit{Fishery parameters}                                          &                       & \\
-Standard deviation of beta distribution for implementation error     & $\sigma_{h}$          & 0.05  \\
-Frequency of assessment (years)                                      & $f_{\mathrm{assess}}$ & 5  \\
-\bottomrule
-\end{tabular}
-\label{t:pars}
-\end{table}
 

### 2013-07-08 inst/ms/salmonfolios.tex
A variety of ms updates; split into multiple files
Make a variety of ms edits
A variety of ms updates; especially adding outlines
Develop the main text
Major additions to text; especially introduction

@@
 \geometry{letterpaper}
-\usepackage{amssymb,amsmath,amsfonts}
+\usepackage{amssymb}
+\usepackage{amsmath}
+\usepackage{amsfonts}
 \usepackage{graphicx}
-
+\usepackage{lineno}
 \usepackage{fontspec}
-%\usepackage{pdfsync}
-\defaultfontfeatures{Ligatures=TeX}
-\setsansfont{Calibri}
-\setmonofont{Inconsolata}
-\setmainfont{Linux Libertine}
-%\setmainfont{Times New Roman}
-\usepackage[sectionbib]{chapterbib}
-
-\usepackage{titlesec}
-\titleformat*{\section}{\Large\bfseries\sffamily}
-\titleformat*{\subsection}{\large\sc\sffamily}
-\titleformat*{\subsubsection}{\itshape\sffamily}
-\DeclareTextCommandDefault{\nobreakspace}{\leavevmode\nobreak\ }
+\textheight 20.5cm
 \usepackage{setspace}
 \usepackage{booktabs}
-\bibliographystyle{apalike}
-\setcounter{secnumdepth}{-2}
-
+\setlength{\parskip}{3pt} % changes vertical space between paragraphs
 \usepackage[round]{natbib}
-\bibliographystyle{apalike}
 \bibpunct{(}{)}{;}{a}{}{;}
 
-\title{Portfolio prioritization of salmon conservation }
-\author{Sean C. Anderson$^{1*}$ \and Jonathan W. Moore$^1$ \and Trevor A. Branch
-  \and Andrew B. Cooper$^2$ \and Nicholas K.  Dulvy$^1$}
+\hyphenation{meta-pop-ulation meta-pop-ulations sub-pop-ulations sub-pop-ulation e-con-o-mist}
+
+\title{Portfolio conservation of metapopulations under climate change}
+
+\author{
+  Sean C. Anderson$^{1*}$, 
+  Jonathan W. Moore$^1$,
+  Trevor A. Branch?,\\
+  Nicholas K. Dulvy$^1$,
+  Michelle M. McClure?,
+  Andrew B. Cooper$^2$
+}
 \date{}
+
+%\\Climate 
+  %change and portfolio conservation of metapopulations\\Climate change and 
+  %portfolio conservation\\Portfolio prioritization of salmon 
+  %conservation\\Portfolio prioritization of metapopulation conservation
+
+%-   Prioritizing metapopulation conservation through the lens of
+    %portfolio theory
+
+%-   Portfolio optimization reveals...
+
+%-   Portfolio theory reveals conservation strategies that...
+
+%-   Intelligent tinkering with ecological portfolios: patterns of
+    %environmental response diversity drive salmon conservation
+    %priorities
+
+%-   Keeping every cog and wheel...
+
+%**Ecological Applications**
+
+%-   60 pages including everything
+
+%> Ecological Applications is concerned broadly with the applications of
+%> ecological science to environmental problems. It publishes papers that
+%> develop scientific principles to support environmental
+%> decision-making, as well as papers that discuss the application of
+%> ecological concepts to environmental issues, policy, and management.
+%> Papers may report on experimental tests, actual applications,
+%> scientific decision support techniques, economic analyses, social
+%> implications of environmental issues, or other relevant topics.
+%> Statistical or experimental methods papers that support research and
+%> applications are welcome. Papers submitted to Ecological Applications
+%> should be accessible to both scholars and practitioners.
 
@@
 %\onehalfspacing
-\doublespacing
+\linenumbers
+\modulolinenumbers[2]
+\begin{spacing}{1.3}
+%\doublespacing
 \maketitle
@@
 \noindent
-	\textsuperscript{*}Corresponding author: sean\_anderson@sfu.ca
+Statement of authorship: 
+%Contributions by authors should be listed on the title page and will be printed at the end of the manuscript.
+This statement should be appropriate to the study described in the manuscript and should clarify who designed the study, who performed the research, who provided new methods or materials, and who wrote the manuscript.
+We encourage concise statements such as “JW performed phylogenetic analyses, MH collected data, performed modeling work and analyzed output data, and PK performed the meta-analysis.
+MH wrote the first draft of the manuscript, and all authors contributed substantially to revisions.”
+
+\noindent
+Running title: 
+%(abbreviated form of title) of less than 45 characters including spaces
+
+\noindent
+Keywords:
+%up to 10
+
+\noindent
+Type of article: Letters
+
+\noindent
+Abstract:
+
+\noindent
+Main text:
+%(excluding abstract, acknowledgements, references, table and figure legends), and the number of words in each text box
+
+\noindent
+References:
+
+\noindent
+Number of figures:
+
+\noindent
+Number of tables:
+
+ %- the name and complete mailing address (including telephone and fax numbers and e-mail address) of the person to whom correspondence should be sent
+\noindent
+	\textsuperscript{*}Corresponding author: Sean C. Anderson; Department of Biological Sciences, Simon Fraser University, Burnaby BC, V5A 1S6; Phone: 1-778-782-3989; E-mail: sean\_anderson@sfu.ca
 
@@
 
+\end{spacing}
 \end{document}

### 2013-08-03 inst/ms/ms.tex
Make substantial additions to main text

@@
 \section{Abstract}
-
-{[}@convertino2013; @wade2013{]}
 
@@
 
-Managing risk is fundamental to the conservation of an endangered species.
-When an endangered species exists as a metapopulation, we can manage risk at two levels: at the population level or at the metapopulation level {[}REF{]}.
-Typically we treat sources of risk for metapopulations as exogenous and manage risk by altering fishing or hunting, implementing reserves, or improving connectivity for subpopulations {[}e.g.
-@akcakaya2007{]}.
+Managing risk is fundamental to the conservation of an endangered species \citep{burgman2005, iucn2009}.
+When an endangered species exists as a metapopulation, we can manage risk at two levels: at the population level or at the metapopulation level.
+Typically we treat sources of risk at the metapopulation level as exogenous and manage risk at the population level by hunting or fishing, implementing reserves, or improving connectivity \citep[e.g.][]{akcakaya2007}.
 
-The management of financial portfolios provides another way to consider risk.
+The management of financial portfolios provides another way of considering risk \citep{figge2004, koellner2006, ando2012}.
 Economists consider the risk and performance of a financial portfolio based on the weighting of individual investments (called assets) that make up the portfolio.
-Modern Portfolio Theory (MPT) proposes that there is a set of portfolios that maximizes expected return for a level of expected risk or minimizes expected risk for a level of expected return \citep{markowitz1952, markowitz1959}.
-Similarly, expected growth rate and variance of a metapopulation is a function of the variance, covariance, and size of the individual populations.
+Modern Portfolio Theory proposes that there is a set of portfolios that maximizes expected return for a level of expected risk or minimizes expected risk for a level of expected return \citep{markowitz1952, markowitz1959}.
+Similarly, expected growth rate and variance of a metapopulation is a function of the variance, covariance, and size of the individual populations \citep{moore2010}.
 A portfolio approach to managing risk for a metapopulation might therefore consider how conservation actions affect the weight of each population in a metapopulation portfolio.
@@
 Managing Pacific salmon under the uncertainty of climate change is an ideal scenario to consider through the lens of portfolio theory for three reasons.
-(1) Salmon often form metapopulations {[}REF{]} and we can consider the metapopulation in a river-catchment as a portfolio and the stream populations as assets.
-Continuing the analogy, predators and fisheries often integrate across multiple streams, acting as investors in the salmon portfolio.
+(1) Pacific salmon form metapopulations \citep{rieman2000, schtickzelle2007} and we can consider, for example, the metapopulation in a river-catchment as a portfolio and the stream populations as assets \citep{schindler2010, moore2010}.
+Continuing the analogy, predators and fisheries often integrate across multiple populations \citep{hilborn2003, schindler2008}, acting as investors in the salmon portfolio.
 Fisheries managers and conservation agencies then act as portfolio managers through actions such choosing which salmon habitat to prioritize for protection or restoration.
-(2) Many Pacific salmon metapopulations are highly threatened {[}REFs{]} and will likely become more at risk as threats such as overfishing, dams, roads, logging, and particularly climate change intensify {[}REFs throughout{]}.
-(3) Salmon are highly valued by society, fishers, conservation groups, and indigenous people {[}REFs{]}.
-Although civil society allocates extensive resources to conserving salmonids (e.g.~XX in XX) the scale of the problem demands a prioritization of conservation efforts {[}REFs{]}.
+(2) Many Pacific salmon metapopulations are highly threatened \citep{mcclure2003, gustafson2007, peterman2012} and will likely become more at risk as threats such as overfishing, dams, roads, logging, and particularly climate change intensify \citep[e.g.][]{lackey2003}.
+(3) Salmon are highly valued by society, fishers, conservation groups, and indigenous people \citep{nrc1996}.
+Although civil society allocates extensive resources to conserving salmonids the scale of the problem demands a prioritization of conservation efforts \citep{allendorf1997}.
 
-A diversified financial portfolio reduces risk by exploiting the covariance between assets --- ideally the same market forces cause one stock to rise if another one falls --- and there are two key mechanisms that reduce metapopulation risk by creating asynchronous population dynamics.
-First, localized habitat features can filter the environment creating unique conditions within a spread out metapopulation {[}REFS{]} (i.e.~Moran effect REF).
-For example, landscape topology can affect stream flow causing different streams of salmon to experience unique conditions {[}REF{]}.
-Second, diversity of life-histories, behaviours, personalities, and genetics can cause populations to respond uniquely to the same environment (i.e.~response diversity REF and biocomplexity REF).
-These unique traits can be phenotypic or genetic and derived from short-term adaptations {[}REF{]} or long-term evolution {[}REF{]} to historical conditions.
-In this paper we focus on the latter, response-diversity mechanism.
+A diversified financial portfolio reduces risk by exploiting the covariance between assets --- ideally one stocks rises while another one falls --- and there are two key mechanisms that generate asynchrony in metapopulation dynamics.
+First, localized habitat features can filter the environment generating unique conditions for subpopulations \citep{schindler2008} (i.e.~Moran effect).
+For salmon, this mechanism acts primarily during freshwater life-history stages and since many salmon species spend more time in the ocean \citep{quinn2005}, a second mechanism is also of key importance --- diversity of response to the environment \citetext{\citealp[i.e.~response diversity,][]{elmqvist2003}; \citealp[and biocomplexity][]{colwell1998}; \citealp{hilborn2003}}.
+This second mechanism can result from unique phenotypic or genetic traits \citep{crozier2008, kovach2012} and be derived from rapid contemporary adaption \citep{stockwell2003, fraser2011} or long-term evolution to historical conditions \citep{eliason2011}.
+In this paper we focus on this latter, response-diversity mechanism, and evaluate it in the context of thermal response diversity.
+
+Salmon are strongly affected by climate warming and yet show a remarkable diversity of tolerance to temperature \citep{beacham1989, crozier2006, battin2007, crozier2008}.
+In addition to climate warming posing perhaps the greatest threat to global biodiversity \citep{thomas2004}, warming poses a particular threat to riverine species whose ranges are largely confined to existing habitat \citep{thomas2010} and where shallower water can be affected more rapidly by thermal changes than larger water basins \citep{isaak2010} and by snow-melt timing and extreme hydrological events \citep{crozier2008}.
+These consequences, and in particular warmer water itself, can combine to limit salmon productivity through oxygen limitation \citep{portner2007} leading to decreased cardiac performance \citep{eliason2011} and having profound implications for behaviour \citep{goniea2006}, disease resistance (REF), energetic costs (REF Rand 1998 from Pess 2012?), and ultimately survival \citep{peterman1998, eliason2011}.
+With 29\% of Pacific Northwest salmon population having gone extinct in the last XX years and XX\% of remaining stocks classified as at risk by XX, these populations are at risk from any further decline in fitness \citep{crozier2008, mcclure2003}.
+Yet, adverse stream temperatures are already impeding recovery of some PNW salmon populations \citep{mccullough1999} and leading to earlier and more condensed migration of salmon \citep{kovach2013a}\ldots{} finish this.
 
 Here, we ask how a portfolio approach to management can inform the conservation of metapopulations in a changing world.
-We ask two primary questions: (1) What does portfolio theory tell us about spatial approaches to prioritizing metapopulation conservation?
-(2) If we don't know how response-diversity is distributed, what does portfolio theory tell us about how many populations to conserve?
+We ask two primary questions: (1) Given a plausible distribution of response diversity, how does portfolio theory inform spatial approaches to prioritizing metapopulation conservation?
+(2) If we don't know how response-diversity is distributed, what does portfolio theory tell us about the number of populations to conserve?
 To answer these questions, we develop a salmon metapopulation simulation in which spatially-distributed thermal tolerance and patterns of short- and long-term climatic change drive population-specific productivity.
-We then implement different conservation ``rules of thumb'' that control the population-level carrying capacities and evaluate the salmon portfolios along risk and return axes, as a financial portfolio manager might.
-We show that conserving response diversity buffers metapopulation risk given short-term climate forcing and metapopulation growth rate given long-term climate warming.
-We then show that conserving more subpopulations buffers risk regardless of response diversity or climate trend, and we conclude that considering metapopulations through portfolio theory provides a useful additional dimension to evaluate conservation strategies.
+We then implement conservation rules of thumb that control the population-level carrying capacities and evaluate the salmon portfolios along risk and return axes, as a financial portfolio manager might.
+We show that conserving response diversity buffers metapopulation risk given short-term climate forcing and ensures metapopulation persistence given long-term climate warming.
+We then show that conserving more subpopulations buffers risk regardless of response diversity or climate trend, and conclude that considering metapopulations through portfolio theory provides a useful additional dimension through which we can consider conservation strategies.
 
@@
 
-We developed a salmon metapopulation simulation model that includes a stock-recruit relationship, demographic stochasticity, straying between populations, varying responses to the environment, escapement target setting, and implementation uncertainty.
+We developed a 100-year salmon metapopulation simulation model that includes a spawner-return relationship, demographic stochasticity, straying between populations, varying responses to the environment, escapement target setting, and implementation uncertainty.
 Under two kinds of environmental regimes we tested different conservation rules of thumb and evaluated these plans in risk-return space similar to how financial managers evaluate financial portfolios.
 We illustrate the overall simulation structure in Fig.~1.
-We provide a package \texttt{metafolio} for the statistical software \texttt{R} \citep{r2013} as an appendix, to carry out the simulations and analyses we describe in this paper.
+We provide a package \texttt{metafolio} for the statistical software \texttt{R} \citep{r2013} as an appendix, to carry out the simulations and analyses described in this paper.
 
@@
 
-In our ecological portfolios, we defined assets as stream-level populations and the portfolios as salmon metapopulations.
+In our ecological portfolios, we defined assets as stream-level populations and portfolios as salmon metapopulations.
 We use the terms \emph{stream} and \emph{populations} interchangeably to represent the portfolio assets.
 We defined the portfolio investors as the stakeholders in the fishery and metapopulation performance.
-CHANGE THIS For example, we could consider fisheries managers, conservation agencies, or First Nations groups as investors.
-We defined asset value as the abundance of returning salmon in each stream and value of the portfolio as the overall metapopulation abundance.
-In this scenario, the equivalent to financial rate of return is the generation-to-generation rate of change of metapopulation abundance.
+For example, the investors could be conservation agencies, First Nations groups, or civil society as a whole.
+The fisheries management agency then becomes the portfolio manager.
+We defined the asset value as the abundance of returning salmon in each stream and the value of the portfolio as the overall metapopulation abundance.
+In this scenario, the equivalent to financial rate of return is the generation-to-generation metapopulation growth rate.
 We defined the financial asset investment weights as the capacity of the stream populations --- specifically the unfinished equilibrium stock size --- since maintaining or restoring habitat requires money, time, and resources.
-Investment in a population therefore represents investing in salmon habitat conservation or reconstruction.
+Investment in a population therefore represents investing in salmon habitat conservation or restoration.
 
@@
 
-We manipulated the capacity and productivity parameters $b$ and $a$ as part of the portfolio simulation.
+We manipulated the capacity and productivity parameters $b_i$ and $a_{i(t)}$ as part of the portfolio simulation.
 The capacity parameters $b_i$ were controlled by the investment weights in the populations.
 For example, a large investment in a stream was represented by a larger unfished equilibrium stock size $b$ for stream $i$.
-The productivity parameters $a_{i(t)}$ were controlled by the interaction between an environmental signal and the stream-level population environmental-tolerance curves.
+The productivity parameters $a_{i(t)}$ were controlled by the interaction between a temperature time series and the subpopulation thermal-tolerance curves.
 
-We generated the environmental-tolerance parabolas according to
+We generated the thermal-tolerance performance curves according to
 
@@
 
-\noindent where $W_i$ controls the width of the curve for population $i$, $e_t$ represents the environmental value at generation $t$, $e_i^{\mathrm{opt}}$ represents the optimal environmental value for population $i$, and $a_i^{\mathrm{max}}$ represents the maximum possible $a$ value for population $i$.
+\noindent where $W_i$ controls the width of the curve for population $i$, $e_t$ represents the environmental value at generation $t$, $e_i^{\mathrm{opt}}$ represents the optimal temperature for population $i$, and $a_i^{\mathrm{max}}$ represents the maximum possible $a$ value for population $i$.
 We set the $W_i$ parameters (Table S1) and calculated the $a_i^{\mathrm{max}}$ parameters so that the area under each curve $A_i$ was equal.
-See Figure 2a for example environmental-tolerance curves.
+We chose parameter values such that some populations were warm-tolerant, some were cold-tolerant, and some had a wider range of thermal-tolerance, but a lower maximum productivity (see Fig.~2a and Appendix S1 in Supplementary Information).
+Although we refer to a thermal-tolerance curve because temperature is a dominant driver of salmonid performance \citep{mccullough1999}, our model applies to any environmental tolerance (e.g.~tolerance to stream flow volume or changes in snow melt timing \citep{crozier2008}).
 
@@
 We implemented straying as in \citet{cooper1999}.
-We set up the metapopulation in a simple scenario: we arranged the populations in a line and those that were nearer to each other were more likely to stray between each other.
+We arranged the subpopulations in a line and salmon were more likely to stray to streams near their natal stream.
 Two parameters controlled the straying: the fraction of fish $f_{\mathrm{stray}}$ that stray from their natal stream in any generation and the rate $m$ at which this straying between streams decays with distance.
@@
 
-\[\mathrm{strays}_{tij} = f_{\mathrm{stray}} R_{tj}
+\[\mathrm{strays}_{ij(t)} = f_{\mathrm{stray}} R_{j(t)}
     \frac{e^{-m \lvert i-j \rvert }}
@@
 
-\noindent where $R_{tj}$ is the number of returning salmon at generation $t$ whose natal stream was stream $j$.
+\noindent where $R_{j(t)}$ is the number of returning salmon at generation $t$ whose natal stream was stream $j$.
 The subscript $k$ represents a stream ID and $n$ the number of populations.
@@
 
+Our simulation did not account for the homogenization of diversity due to straying.
+For example, all salmon in population A maintained the same thermal-tolerance curve regardless of how many salmon it received from population B.
+However, the low straying rates in our simulation rate should render this effect minimal except for populations at low abundance.
+In our model, the main implication is that, in reality, populations that reach a low abundance may reach demographic extinction, potentially suffer consequences such as inbreeding depression \citep{wang2002} and subsequently contribute little to response diversity.
+
 \subsection{Fishing}
@@
 Our simulation used a simple set of rules to establish escapement targets and harvest the fish.
-Every five years our simulation fitted a spawner-return function and target harvest rate $H_{\mathrm{tar}}$ was set based on \citet{hilborn1992} as
+Every five years our simulation fitted a spawner-return function and the target harvest rate $H_{\mathrm{tar}}$ was set based on \citet{hilborn1992} as
 
@@
 
+\noindent
 Further, to establish a range of spawner-return values and to mimic the start of an open-access fishery, for the first 30 years we drew the fraction of fish harvested randomly from a uniform distribution between 0.1 and 0.9.
@@
 
-We evaluated portfolio performance under short- and long-term environmental dynamics.
+Environmental dynamics typically have both short- and long-term fluctuations.
+We evaluated portfolio performance under these two components separately.
 We represented short-term dynamics as a stationary first-order autoregressive process, AR(1), with correlation $\rho_e$
@@
 
-Mention Fig.~3 and maybe put it before Fig.~2?
+Combining salmon population dynamics, fishing, and environmental dynamics, we illustrate the components of an example simulation in Fig.~3.
 
@@
 
-We evaluated two sets of conservation rules of thumb: (1) spatial response diversity conservation strategies in an idealistic scenario where you can detect response diversity, and (2) a more realistic scenario where we know little about response diversity and we're left with a choice of how many populations to conserve.
+We evaluated two sets of conservation rules of thumb: (1) strategies to conserve spatial response diversity, and (2) a scenario where we know little about response diversity and we're left with a choice of how many populations to conserve.
 
@@
 These reduced populations could still receive straying salmon but were unlikely to rebuild on their own to a substantial abundance.
-The four scenarios were:
+The four scenarios we considered were:
 
@@
 \item
-  Conserve the most stable populations only.
+  Conserve the middle section of the metapopulation (conserve stable populations).
 \item
-  Conserve one half of the metapopulation.
+  Conserve the lower half of the metapopulation (conserve cool-tolerant populations).
 \item
-  Conserve the other half of the metapopulation.
+  Conserve the upper half of the metapopulation (conserve warm-tolerant populations).
 \end{enumerate}
@@
 In reality we rarely know precise levels of response diversity.
-We therefore additionally considered a case where the conservation was randomly assigned with respect to response diversity but where different numbers of streams could be conserved.
+We therefore considered a case where the conservation was randomly assigned with respect to response diversity but where different numbers of streams could be conserved.
 We considered conserving from two to 16 streams.
@@
 
-\subsubsection{Short-term environment}
-
-Given strong short-term environmental fluctuations, conserving response diversity buffers the risk properties of an ecological portfolio (Fig.~4a).
-In our simulation, the median variance of generation-to-generation rate of change in abundance was X times lower given balanced response diversity (full range of responses or most stable only vs.~conserving one half or the other).
-In fact, even though by conserving the full range of responses, the portfolio was comprised of warm and cool-thriving populations that were more variable on their own, each was balanced by an opposing population.
-The portfolio risk was therefore comparable between the full range of responses and most stable only portfolios.
-
-We can see the mechanism behind these portfolio properties by inspecting example population time series (Fig.~4c, d).
-If only the upper or lower half of response diversity is conserved, the portfolio tends to do well or poorly depending on the environmental conditions (Fig.~4d).
+\emph{Given short-term environmental fluctuations}, conserving response diversity buffers the risk properties of an ecological portfolio (Fig.~4a).
+In our model, the median variance of metapopulation growth rate was X times lower given balanced response diversity (full range of responses or most stable only vs.~conserving one half or the other).
+This increase in stability occurred despite the portfolio being comprised of warm and cool-thriving populations that were more variable on their own than populations with wide thermal-tolerance curves.We can see the mechanism behind these portfolio properties by inspecting example population time series (Fig.~4c, d).If only the upper or lower half of response diversity is conserved, the portfolio tends to alternate between performing well and poorly depending on the environmental conditions resulting in a risker portfolio (Fig.~4e).
 This risk is buffered with balanced response diversity (Fig.~4c).
 
-\subsubsection{Long-term environment}
-
-Given long-term environmental change, the choice of which populations to conserve affects the return properties of an ecological portfolio (Fig.~4b).
-By conserving balanced response diversity, an ecological manager is hedging his or her bets on what will happen with the environment and how the populations will respond.
-The typical return for a balanced response diversity strategy was zero --- the metapopulation neither increased or decreased in abundance in the long run.
-By conserving only the upper or lower half of response diversity, a conservation manager is putting all his or her eggs in one basket --- the metapopulation might do really well through time or it might do really poorly.
-The example metapopulation abundance time series (Fig.~4d, f) illustrate this effect.
+\emph{Given long-term environmental change}, the choice of which populations to conserve affects the ``return'' (metapopulation growth rate) properties of an ecological portfolio (Fig.~4b).
+By conserving balanced response diversity, an ecological manager is hedging his or her bets on the environmental trend and how the populations will respond.The typical metapopulation growth rate for a balanced response diversity strategy was zero --- the metapopulation neither increased or decreased in abundance in the long run.The example metapopulation abundance time series (Fig.~4d, f) illustrate the mechanism.
 By conserving response diversity, when one population is doing poorly, another is doing well and the metapopulation abundance remains stationary through time.
-
-Notably, in these simulations, if a managers invested in the populations that were doing well at the beginning they would have had the lowest rate-of-return portfolio in the end (purple portfolios in Fig.~4b).
-
-Spatial conservation strategies in the face of longterm environmental change hinge on whether you ``get it right'' --- whether you choose just the right populations to conserve.
+If a manager had invested in the populations that were doing well at the beginning they would have had the lowest rate-of-return portfolio in the end (purple portfolios in Fig.~4f).
 
@@
 
-Given a scenario where we don't know the distribution of population-level response diversity, portfolio optimization informs us about the risk buffering from maintaining multiple populations (Fig.~5).
-In short, investing in more populations buffers portfolio risk.
+In a scenario where we don't know the distribution of population-level response diversity, portfolio optimization informs us that investing in more populations buffers portfolio risk regardless of environmental trend.
+(Fig.~5).
 
-\subsubsection{Short-term environment}
-
-Given short-term environmental noise, conserving more populations buffers portfolio risk while the random conservation of response diversity creates a spread of metapopulation risk for the same number of populations conserved (Fig.~5a).
+\emph{Given short-term environmental fluctuations}, conserving more populations buffers portfolio risk (Fig.~5a), and at the same time, the random conservation of response diversity creates a spread of metapopulation risk for the same number of populations conserved.
 For example, a metapopulation with eight conserved populations is X times less risky than a metapopulation with only four.
 We can see this risk-buffering effect through example metapopulations in Fig.~5c and 5d.
-We note that the risk-return axes of portfolio optimization ignore the absolute-abundance dimension (Fig.~5d).
-As one would expect, conserving fewer populations also results in lower-abundance metapopulations.
+We note that the risk-return axes of portfolio optimization ignore the absolute-abundance dimension (Fig.~5d) --- as one would expect, conserving fewer populations also results in lower-abundance metapopulations.
 
-\subsubsection{Long-term environment}
-
-Given long-term environmental noise, conserving more populations also buffers portfolio risk.
+\emph{Given long-term environmental change}, conserving more populations also buffers portfolio risk.
 However, in contrast to the short-term environmental noise scenario, the unknown response diversity creates a spread of possible metapopulation return for the same number of conserved populations (Fig.~5b).
-Here, the number of populations conserved buffers non-systematic (i.e.~not-environmentally-driven) stochasticity.
+Here, the number of populations conserved buffers non-systematic (primarily demographic) stochasticity.
 
 \section{Discussion}
+
+Our model demonstrates how population diversity can stabilize realistic metapopulations under climate change.
+We show how maintaining populations with a variety of thermal-tolerances reduces risk caused by short-term environmental stochasticity and optimizes chances for long-term persistence given climate change.
+Further, without knowing how response diversity is distributed, conserving more subpopulations results in less risky metapopulations regardless of environmental trend and response-diversity distribution.
+Combined, these two findings illustrate that focusing on well-performing stocks now at the detriment of others is at best equivalent to a risky but efficient portfolio, and at worst a poorly-performing risky portfolio.
+In this discussion, we begin by linking our model with real-world conservation issues for Pacific Northwest salmon and considering the implications of our results for salmon conservation and management.
+We then consider broader research priorities and extensions that our results suggest for metapopulation conservation.
+
+The primary lesson from our simulation model is the importance of response diversity in buffering risk and ensuring long-term persistence in the face of an uncertain world.
+First, our results emphasize the need to make response diversity a primary goal of conservation and recovery efforts \citep{moore2010} and suggest that precautionary management should not just focus on abundance but also be precautionary with respect to response diversity.
+Thus, our results highlight the need to preserve the processes that maintain response diversity, with habitat heterogeneity, which can lead to local adaptation \citep{taylor1991, fraser2011} and hence response diversity, being a key process.
+Second, our results can inform habitat prioritization decisions.
+If closer populations are more correlated with each other and conservation actions must be prioritized, then our model suggests we should focus on populations that aren't spatially contiguous, as long as sufficient straying is maintained, to maximize diversity of response to the environment.
+Third, where possible, avoid structures that artificially remove response diversity.
+For salmon, dams are a prominent example \citep{waples2007, pess2012}.
+Dams have a double impact whereby their introduction can selectively eliminate a large swath of habitat, perhaps analogous to our scenario X (Fig.~X), and then mitigation approaches such as hatcheries and small transportation can further reduce population diversity and likely response diversity (REFS).
+
+Our simulation model emphasizes the importance of thinking at the portfolio level --- considering aggregate risk and performance for ecological portfolios.
+In many ways our results are intuitive, however, we often don't consider aggregate risk despite the many benefits to doing so.
+Only by simultaneously considering aggregate risk and return can we optimally diversify our ecological systems \citep{ando2012}.
+To that end we can think of portfolio risk-return performance as another dimension of conservation.
+This dimension emphasizes the value of a broader scale of management; it emphasizes the value of integrated management planning (REF).
+Further, it emphasizes the need to not just focus on productive populations but also including less productive populations to lower the long-term aggregate risk for the system \citep{hilborn2003}.
+This perspective does not negate an understanding at a finer scale, say of individual stocks.
+Rather, as in finance, it requires an understanding at a finer scale but also an understanding of the covariance between stocks to predict the properties of the portfolio as a whole.
+
+TODO: Across the range of Pacific Northwest salmonids, United States and Canadian guidelines are increasingly well-structured to benefit from the findings of our simulation model.
+In the United States, Evolutionary Significant Units (ESUs) represent\ldots{} and Viable Salmonid Populations (VSPs) represent\ldots{} \citep{mceacheran2000, waples1991, moritz1994}.
+In Canada, the Wild Salmon Policy \citep{dfo2005wsp} structures salmon populations into Conservation Units (CUs).
+Our recommendations support a number of VSP and CU guidelines.
+For example\ldots{}
+
+Our model is unique in that it is the first to consider risk for the metapopulation as a whole while identifying the role of response diversity in buffering that risk.
+Our model follows from the work of \citet{cooper1999} who showed the importance of metapopulation habitat structure for salmonid conservation and \citet{moore2010} who linked salmon portfolio performance to asynchrony in population dynamics.
+Our model complements existing simulation-based salmon-habitat prioritization models that emphasize detailed assessment of individual stocks and we compare two of them here.
+One a frequently used simulation-habitat-prioritization model, the Shiraz model \citep{scheuerell2006}, focuses on detailed conditioning of the habitat-population-dynamics relationship for a single salmon population.
+While the Shiraz model can be applied to an entire watershed it combines the populations together as a single unit.
+In contrast to our model, the Shiraz model breaks the spawner-recruit model into multiple stages.
+A second prioritization scheme proposes a ranking of priority based on population viability analysis (PVA) and extinction consequences \citep{allendorf1997}.
+They suggests using PVA or surrogate measures of population viability and then considering the genetic consequences of conserving various populations.
+\citet{allendorf1997}, however, also focus on the assessment of individual stocks without considering covariance between them and therefore the performance of the salmon portfolio as a whole.
+Compared to these detailed population-specific models, we emphasize that our model is, at this stage, primarily conceptual, but could be conditioned on real systems or expanded to incorporate greater life-stage realism.
+Our intention was to develop a parsimonious model to explore the salient aspects of metapopulation-portfolio performance.
+
+There are number of specific extensions to our simulation model that could be useful.
+First our model could be extended beyond thermal response diversity.
+For the example of salmon this might include response diversity to inter-decadal climate signals such as the Pacific Decadal Oscillation \citep{mantua2002}, stream flow, or dissolved oxygen \citep{priede1988}.
+Second, aspects of our model were tailored to salmon but could be extended to metapopulations of other taxa, to ecological communities, and to meta-communities.
+Third, an extension of our model would consider the Moran or environmental-filter concept \citep{schindler2008} where populations further from each other are more likely to experience different environmental forces.
+Fourth, a model could consider the contribution of contemporary evolution \citep{stockwell2003}.
+This rapid adaptation to changes in the environment could strongly affect portfolio performance and emphasize the importance of maintaining genetic diversity.
+Finally, a fruitful extension would be to condition our model on a system of interest, for example a given river basin, and optimize that metapopulation portfolio across conservation and restoration options as part of a formal decision analysis.
+
+Our results suggest three broad research priorities.
+First our work highlights the importance of identifying response diversity \citep[e.g.][]{eliason2011}, spatial asynchrony \citep[e.g.][]{thorson2013} and in particular the management schemes that promote these processes.
+Our work highlights how these processes can contribute to reducing risk and increasing the likelihood of persistence of metapopulations under uncertainty.
+Second, our work suggests the importance of identifying what drives productivity in different scenarios.
+For example, what is the relative effect on productivity of population density, an environment-recruitment relationship, and environmental regime shifts \citep{vert-pre2013} and what management approaches are robust to our uncertainty about these contributions?
+Third, our work suggests research that looks for real-life case studies as in situ experiments testing our conclusions.
+The question remains, does managing metapopulations portfolios using modern portfolio theory result in better outcomes that are in line with the societal goals?
+
+In conclusion, our results demonstrate the utility of considering metapopulations in a portfolio context and the power of portfolio simulation to consider robust management strategies that minimize metapopulation risk and maximize growth under uncertain environmental conditions.
+Our findings show that response diversity buffers metapopulation risk given short-term environmental stochasticity and ensures persistence given long-term regime changes.
+Further, our findings show that, independent of response-diversity, conserving more populations buffers metapopulation risk.
+These results point to the critical importance of choosing management, conservation, and restoration actions that preserve and promote response diversity.
 
@@
 \centering
-\includegraphics[height=5.5in]{../examples/simulation-diagram2.pdf}
+\includegraphics[height=5.5in]{../examples/simulation-diagram3.pdf}
 \caption{Flow chart of the salmon-metapopulation simulation.
@@
 \includegraphics[width=4.0in]{../examples/spatial-arma-sim.pdf}
-\caption{The components of an example metapopulation simulation.
-We show, from top to bottom, the environmental signal, the resulting productivity parameter (Ricker $a$), the salmon returns, fisheries catch, salmon escapement, salmon straying from their natal streams, salmon joining from other streams, stock-recruit residuals on a log scale, and the estimated $a$ and $b$ parameters in the fitted Ricker curve.
+\caption{[TODO clean up a bit - especially estimated b y-axes] The components of an example metapopulation simulation.
+We show, from top to bottom, the environmental signal, the resulting productivity parameter (Ricker $a$), the salmon returns, fisheries catch, salmon escapement, salmon straying from their natal streams, salmon joining from other streams, spawner-return residuals on a log scale, and the estimated $a$ and $b$ parameters in the fitted Ricker curve.
 The colored lines indicate populations that thrive at low (cool colours) to high (warm colours) environmental values.}
@@
 \includegraphics[width=5.0in]{../examples/cons-plans-n.pdf}
-\caption{The importance of preserving as many subpopulations as possible when we don't know how response diversity is distributed.
+\caption{[TODO move environmental ts to label, add arrow to high ext.
+risk, add labels to lower panels pointing out result] The importance of preserving as many subpopulations as possible when we don't know how response diversity is distributed.
 In risk-return space we show environmental scenarios that are comprised primarily of (a) short-term and (b) long-term environmental fluctuations (see Fig.~X).

### 2013-08-03 inst/ms/salmonfolios.tex
Make substantial additions to main text

@@
 \usepackage{lineno}
-\usepackage{fontspec}
 \textheight 20.5cm
@@
 \usepackage{booktabs}
+\usepackage[sectionbib]{chapterbib}
 \setlength{\parskip}{3pt} % changes vertical space between paragraphs
@@
 
-\title{Portfolio conservation of metapopulations under climate change}
+\title{Portfolio conservation of metapopulations\\under climate change}
 
@@
   Michelle M. McClure?,
-  Andrew B. Cooper$^2$
+  Andrew B. Cooper$^2$\\(Order to be discussed)
 }
@@
 \begin{document}
-%\onehalfspacing
-\linenumbers
-\modulolinenumbers[2]
-\begin{spacing}{1.3}
+\begin{spacing}{1.1}
 %\doublespacing
@@
 \noindent
-Statement of authorship: 
+\textit{Statement of authorship}: SCA, JWM, NKD, and ABC designed the analyses.
+SCA wrote the simulation and analyzed the results.
+SCA and XX \ldots wrote the manuscript.
+All authors participated in conception, discussion of the results, and manuscript preparation.
 %Contributions by authors should be listed on the title page and will be printed at the end of the manuscript.
@@
 \noindent
-Running title: 
+\textit{Running title}:
+Portfolio conservation of metapopulations
 %(abbreviated form of title) of less than 45 characters including spaces
@@
 \noindent
-Keywords:
+\textit{Keywords}: 
+biocomplexity, 
+Modern Portfolio Theory, 
+portfolio effect, 
+prioritization, 
+resource management, 
+range contraction, 
+response diversity, 
+risk assessment, 
+salmon, 
+stochastic simulation 
+(maximum 10)
 %up to 10
@@
 \noindent
-Type of article: Letters
+\textit{Type of article}: Letters
 
 \noindent
-Abstract:
+\textit{Words in abstract}: (maximum 150)
 
 \noindent
-Main text:
+\textit{Words in main text}: (maximum 5000)
 %(excluding abstract, acknowledgements, references, table and figure legends), and the number of words in each text box
@@
 \noindent
-References:
+\textit{Number of references}:
 
 \noindent
-Number of figures:
+\textit{Number of figures}: 5
 
 \noindent
-Number of tables:
+\textit{Number of tables}: 1?
 
@@
 
+\end{spacing}
 \clearpage
 
+\linenumbers
+\modulolinenumbers[2]
+\begin{spacing}{1.4}
+%\onehalfspacing
 \include{ms}

### 2013-08-07 inst/ms/ms.tex
Add to ms; move ms to all Pandoc

@@
+\section{Portfolio conservation of metapopulations under climate change}
+
+\setlength{\parskip}{6pt} \setlength{\parindent}{0cm}
+
+Sean C.
+Anderson\textsuperscript{1*}, Jonathan W.
+Moore\textsuperscript{1}, Trevor A.
+Branch?, Nicholas K.
+Dulvy\textsuperscript{1}, Michelle M.
+McClure?, Andrew B.
+Cooper\textsuperscript{2} (Order to be discussed)
+
+\textsuperscript{1}Department of Biological Sciences, Simon Fraser University, Burnaby BC, V5A 1S6, Canada
+
+\textsuperscript{2}School of Resource and Environmental Management, Simon Fraser University, Burnaby, BC, V5A 1S6, Canada
+
+\emph{Statement of authorship}: SCA, JWM, NKD, and ABC designed the analyses.
+SCA wrote the simulation, analyzed the results, and wrote the first draft of the manuscript.
+All authors contributed substantially to revisions.
+
+\emph{Running title}: Portfolio conservation of metapopulations (maximum 45 characters)
+
+\emph{Keywords}: biocomplexity, Modern Portfolio Theory, portfolio effect, prioritization, resource management, range contraction, response diversity, risk assessment, salmon, stochastic simulation (maximum 10)
+
+\emph{Type of article}: Letters
+
+\emph{Words in abstract and main text}: (maximum 150, 5000)
+
+\emph{Number of references}: (maximum 50)
+
+\emph{Number of tables and figures}: 1, 5
+
+\textsuperscript{*}Corresponding author: Sean C. Anderson; Department of Biological Sciences, Simon Fraser University, Burnaby BC, V5A 1S6; Phone: 1-778-782-3989; E-mail: sean\_anderson@sfu.ca
+
+\setlength{\parskip}{2pt} \setlength{\parindent}{16pt}
+
 \section{Abstract}
+
+Managing risk is fundamental to the conservation of an endangered species.
+When an endangered species exists as a metapopulation, we typically manage risk at the subpopulation level.
+However, a portfolio approach might consider how conservation affects the ``weight'' of each population in a metapopulation ``portfolio''.
+Here we ask how such a portfolio approach can inform conservation priorities for metapopulations in a changing world.
+We develop a salmon metapopulation simulation in which population-specific productivity is driven by spatially-distributed thermal tolerance and patterns of short- and long-term environmental change.
+We then implement spatial conservation strategies that control subpopulation carrying capacities and evaluate the salmon portfolios along risk-return axes.
+We show that conserving response diversity minimizes risk given environmental stochasticity and ensures persistence given long-term environmental change.
+Further, conserving more subpopulations minimizes risk regardless of response diversity distribution.
+These findings have important implications for how we prioritize metapopulation conservation and emphasize the importance of conserving the processes that promote response diversity.
 
@@
 When an endangered species exists as a metapopulation, we can manage risk at two levels: at the population level or at the metapopulation level.
-Typically we treat sources of risk at the metapopulation level as exogenous and manage risk at the population level by hunting or fishing, implementing reserves, or improving connectivity \citep[e.g.][]{akcakaya2007}.
+Typically we treat sources of risk at the metapopulation level as exogenous and manage risk at the population level by manipulating hunting or fishing levels, implementing reserves, or improving connectivity \citep[e.g.][]{akcakaya2007}.
 
@@
 (3) Salmon are highly valued by society, fishers, conservation groups, and indigenous people \citep{nrc1996}.
-Although civil society allocates extensive resources to conserving salmonids the scale of the problem demands a prioritization of conservation efforts \citep{allendorf1997}.
+Although civil society allocates extensive resources to conserving salmonids, the scale of the problem demands a prioritization of conservation efforts \citep{allendorf1997}.
 
@@
 First, localized habitat features can filter the environment generating unique conditions for subpopulations \citep{schindler2008} (i.e.~Moran effect).
-For salmon, this mechanism acts primarily during freshwater life-history stages and since many salmon species spend more time in the ocean \citep{quinn2005}, a second mechanism is also of key importance --- diversity of response to the environment \citetext{\citealp[i.e.~response diversity,][]{elmqvist2003}; \citealp[and biocomplexity][]{colwell1998}; \citealp{hilborn2003}}.
+For salmon, this mechanism acts primarily during freshwater life-history stages and since many salmon species spend considerable time in the ocean \citep{quinn2005}, a second mechanism is also of key importance --- diversity of response to the environment \citetext{\citealp[i.e.~response diversity,][]{elmqvist2003}; \citealp[and biocomplexity][]{colwell1998}; \citealp{hilborn2003}}.
 This second mechanism can result from unique phenotypic or genetic traits \citep{crozier2008, kovach2012} and be derived from rapid contemporary adaption \citep{stockwell2003, fraser2011} or long-term evolution to historical conditions \citep{eliason2011}.
@@
 Salmon are strongly affected by climate warming and yet show a remarkable diversity of tolerance to temperature \citep{beacham1989, crozier2006, battin2007, crozier2008}.
-In addition to climate warming posing perhaps the greatest threat to global biodiversity \citep{thomas2004}, warming poses a particular threat to riverine species whose ranges are largely confined to existing habitat \citep{thomas2010} and where shallower water can be affected more rapidly by thermal changes than larger water basins \citep{isaak2010} and by snow-melt timing and extreme hydrological events \citep{crozier2008}.
-These consequences, and in particular warmer water itself, can combine to limit salmon productivity through oxygen limitation \citep{portner2007} leading to decreased cardiac performance \citep{eliason2011} and having profound implications for behaviour \citep{goniea2006}, disease resistance (REF), energetic costs (REF Rand 1998 from Pess 2012?), and ultimately survival \citep{peterman1998, eliason2011}.
+In addition to climate warming posing perhaps the greatest threat to global biodiversity \citep{thomas2004}, warming poses a particular threat to riverine species whose ranges are largely confined to existing habitat \citep{thomas2010} and where shallower water, compared to larger water basins, can be affected more rapidly by thermal changes \citep{isaak2010} and by snow-melt timing and extreme hydrological events \citep{crozier2008}.
+These consequences, and in particular warmer water itself, can combine to limit salmon productivity through oxygen limitation \citep{portner2007} leading to decreased cardiac performance \citep{eliason2011} and having profound implications for behaviour \citep{goniea2006}, disease resistance (REF), energetic costs in general (REF Rand 1998 from Pess 2012?), and ultimately survival \citep{peterman1998, eliason2011}.
 With 29\% of Pacific Northwest salmon population having gone extinct in the last XX years and XX\% of remaining stocks classified as at risk by XX, these populations are at risk from any further decline in fitness \citep{crozier2008, mcclure2003}.
@@
 Here, we ask how a portfolio approach to management can inform the conservation of metapopulations in a changing world.
-We ask two primary questions: (1) Given a plausible distribution of response diversity, how does portfolio theory inform spatial approaches to prioritizing metapopulation conservation?
+We ask: (1) Given a plausible distribution of response diversity, how does portfolio theory inform spatial approaches to prioritizing metapopulation conservation?
 (2) If we don't know how response-diversity is distributed, what does portfolio theory tell us about the number of populations to conserve?
@@
 
-\subsubsection{Spawner-return relationship}
-
 We defined the spawner-return relationship with a Ricker model \citep{ricker1954},
@@
 Although we refer to a thermal-tolerance curve because temperature is a dominant driver of salmonid performance \citep{mccullough1999}, our model applies to any environmental tolerance (e.g.~tolerance to stream flow volume or changes in snow melt timing \citep{crozier2008}).
-
-\subsubsection{Straying}
 
@@
 
-\noindent
-Further, to establish a range of spawner-return values and to mimic the start of an open-access fishery, for the first 30 years we drew the fraction of fish harvested randomly from a uniform distribution between 0.1 and 0.9.
+\noindent Further, to establish a range of spawner-return values and to mimic the start of an open-access fishery, for the first 30 years we drew the fraction of fish harvested randomly from a uniform distribution between 0.1 and 0.9.
 We discarded these initial 30 years as a burn-in period.
@@
 In our model, the median variance of metapopulation growth rate was X times lower given balanced response diversity (full range of responses or most stable only vs.~conserving one half or the other).
-This increase in stability occurred despite the portfolio being comprised of warm and cool-thriving populations that were more variable on their own than populations with wide thermal-tolerance curves.We can see the mechanism behind these portfolio properties by inspecting example population time series (Fig.~4c, d).If only the upper or lower half of response diversity is conserved, the portfolio tends to alternate between performing well and poorly depending on the environmental conditions resulting in a risker portfolio (Fig.~4e).
+This increase in stability occurred despite the portfolio being comprised of warm and cool-thriving populations that were more variable on their own than populations with wide thermal-tolerance curves.
+We can see the mechanism behind these portfolio properties by inspecting example population time series (Fig.~4c, d).
+If only the upper or lower half of response diversity is conserved, the portfolio tends to alternate between performing well and poorly depending on the environmental conditions resulting in a risker portfolio (Fig.~4e).
 This risk is buffered with balanced response diversity (Fig.~4c).
@@
 \emph{Given long-term environmental change}, the choice of which populations to conserve affects the ``return'' (metapopulation growth rate) properties of an ecological portfolio (Fig.~4b).
-By conserving balanced response diversity, an ecological manager is hedging his or her bets on the environmental trend and how the populations will respond.The typical metapopulation growth rate for a balanced response diversity strategy was zero --- the metapopulation neither increased or decreased in abundance in the long run.The example metapopulation abundance time series (Fig.~4d, f) illustrate the mechanism.
+By conserving balanced response diversity, an ecological manager is hedging his or her bets on the environmental trend and how the populations will respond.
+The typical metapopulation growth rate for a balanced response diversity strategy was zero --- the metapopulation neither increased or decreased in abundance in the long run.
+The example metapopulation abundance time series (Fig.~4d, f) illustrate the mechanism.
 By conserving response diversity, when one population is doing poorly, another is doing well and the metapopulation abundance remains stationary through time.
@@
 
-In a scenario where we don't know the distribution of population-level response diversity, portfolio optimization informs us that investing in more populations buffers portfolio risk regardless of environmental trend.
-(Fig.~5).
+In a scenario where we don't know the distribution of population-level response diversity, portfolio optimization informs us that investing in more populations buffers portfolio risk regardless of environmental trend (Fig.~5).
 
@@
 We show how maintaining populations with a variety of thermal-tolerances reduces risk caused by short-term environmental stochasticity and optimizes chances for long-term persistence given climate change.
-Further, without knowing how response diversity is distributed, conserving more subpopulations results in less risky metapopulations regardless of environmental trend and response-diversity distribution.
+Further, without knowing how response diversity is distributed, conserving more subpopulations reduces metapopulation risk regardless of environmental trend and response-diversity distribution.
 Combined, these two findings illustrate that focusing on well-performing stocks now at the detriment of others is at best equivalent to a risky but efficient portfolio, and at worst a poorly-performing risky portfolio.
@@
 
-The primary lesson from our simulation model is the importance of response diversity in buffering risk and ensuring long-term persistence in the face of an uncertain world.
-First, our results emphasize the need to make response diversity a primary goal of conservation and recovery efforts \citep{moore2010} and suggest that precautionary management should not just focus on abundance but also be precautionary with respect to response diversity.
-Thus, our results highlight the need to preserve the processes that maintain response diversity, with habitat heterogeneity, which can lead to local adaptation \citep{taylor1991, fraser2011} and hence response diversity, being a key process.
-Second, our results can inform habitat prioritization decisions.
+The primary lesson from our simulation model is the importance of response diversity in buffering risk and ensuring long-term persistence in the face of environmental uncertainty.
+Thus, our results emphasize the need to make the maintenance of response diversity, and the processes that maintain response diversity, a primary goal of conservation and recovery efforts \citep{moore2010}.
+For example, since habitat heterogeneity can lead to local adaptation \citep{taylor1991, fraser2011}, our results emphasize the need to maintain a diversity of habitat.
+Our results can also inform habitat prioritization decisions.
 If closer populations are more correlated with each other and conservation actions must be prioritized, then our model suggests we should focus on populations that aren't spatially contiguous, as long as sufficient straying is maintained, to maximize diversity of response to the environment.
-Third, where possible, avoid structures that artificially remove response diversity.
+Finally, our results demonstrate the advantages of avoiding structures that artificially remove response diversity.
 For salmon, dams are a prominent example \citep{waples2007, pess2012}.
-Dams have a double impact whereby their introduction can selectively eliminate a large swath of habitat, perhaps analogous to our scenario X (Fig.~X), and then mitigation approaches such as hatcheries and small transportation can further reduce population diversity and likely response diversity (REFS).
+Dams can have a double impact whereby their introduction selectively eliminates a large swath of contiguous habitat, perhaps analogous to our upper- or lower-half scenarios in Fig.~4, and then mitigation approaches such as hatcheries and smolt transportation can further reduce population and response diversity (REFS).
 
 Our simulation model emphasizes the importance of thinking at the portfolio level --- considering aggregate risk and performance for ecological portfolios.
-In many ways our results are intuitive, however, we often don't consider aggregate risk despite the many benefits to doing so.
-Only by simultaneously considering aggregate risk and return can we optimally diversify our ecological systems \citep{ando2012}.
+In many ways our results are intuitive, however, we often don't consider aggregate risk and only by simultaneously considering aggregate risk and return can we optimally diversify ecological systems \citep{ando2012}.
 To that end we can think of portfolio risk-return performance as another dimension of conservation.
-This dimension emphasizes the value of a broader scale of management; it emphasizes the value of integrated management planning (REF).
-Further, it emphasizes the need to not just focus on productive populations but also including less productive populations to lower the long-term aggregate risk for the system \citep{hilborn2003}.
-This perspective does not negate an understanding at a finer scale, say of individual stocks.
-Rather, as in finance, it requires an understanding at a finer scale but also an understanding of the covariance between stocks to predict the properties of the portfolio as a whole.
+This dimension emphasizes the value of a broader scale of management and focussing not just on currently productive populations but also on less productive populations to reduce long-term aggregate risk \citep{hilborn2003}.
+This perspective does not negate an understanding at a finer scale, say of individual fish stocks.
+Rather, as in finance, it requires an understanding of stock performance but also an understanding of the covariance between stocks to predict the properties of the portfolio as a whole.
 
-TODO: Across the range of Pacific Northwest salmonids, United States and Canadian guidelines are increasingly well-structured to benefit from the findings of our simulation model.
-In the United States, Evolutionary Significant Units (ESUs) represent\ldots{} and Viable Salmonid Populations (VSPs) represent\ldots{} \citep{mceacheran2000, waples1991, moritz1994}.
-In Canada, the Wild Salmon Policy \citep{dfo2005wsp} structures salmon populations into Conservation Units (CUs).
+Our findings support a number of existing guidelines and management structures for Pacific Northwest salmonids.
+
+Across the range of Pacific Northwest salmonids, United States and Canadian guidelines are increasingly well-structured to benefit from the findings of our simulation model.
+
+In the US, larger-scale Evolutionary Significant Units (ESUs) represent populations that are reproductively isolated with an evolutionary heritage \citep{waples1991, moritz1994} and finer-scale Viable Salmonid Populations (VSPs) represent groups of fish that spawn in a given lake or stream in a particular season \citep{mceacheran2000}.
+In Canada, the rough equivalent to the ESU is a Conservation Unit (CU) as defined in the Wild Salmon Policy.
+A CU consists of a group of salmon that are reproductively isolated and that if lost would be unlikely to recolonize in a reasonable time frame \citep{dfo2005wsp}.
+Both the US and Canadian guidelines recognize still finer-scale population structure.
+
 Our recommendations support a number of VSP and CU guidelines.
-For example\ldots{}
+For example, VSP guidelines suggest\ldots{} and the first objective of CU\ldots{} Therefore, we suggest an ESU or CU could represent a salmon portfolios and the lessons learned from models such as ours are directly applicable to management guidelines in the PNW.
 
-Our model is unique in that it is the first to consider risk for the metapopulation as a whole while identifying the role of response diversity in buffering that risk.
+Our model is unique in that it is the first to consider risk for the metapopulation as a whole while identifying the role of response diversity in buffering risk.
 Our model follows from the work of \citet{cooper1999} who showed the importance of metapopulation habitat structure for salmonid conservation and \citet{moore2010} who linked salmon portfolio performance to asynchrony in population dynamics.
 Our model complements existing simulation-based salmon-habitat prioritization models that emphasize detailed assessment of individual stocks and we compare two of them here.
-One a frequently used simulation-habitat-prioritization model, the Shiraz model \citep{scheuerell2006}, focuses on detailed conditioning of the habitat-population-dynamics relationship for a single salmon population.
+One a frequently used simulation-based habitat-prioritization model, the Shiraz model \citep{scheuerell2006}, focuses on detailed conditioning of the habitat-population-dynamics relationship for a single salmon population.
 While the Shiraz model can be applied to an entire watershed it combines the populations together as a single unit.
 In contrast to our model, the Shiraz model breaks the spawner-recruit model into multiple stages.
-A second prioritization scheme proposes a ranking of priority based on population viability analysis (PVA) and extinction consequences \citep{allendorf1997}.
-They suggests using PVA or surrogate measures of population viability and then considering the genetic consequences of conserving various populations.
-\citet{allendorf1997}, however, also focus on the assessment of individual stocks without considering covariance between them and therefore the performance of the salmon portfolio as a whole.
-Compared to these detailed population-specific models, we emphasize that our model is, at this stage, primarily conceptual, but could be conditioned on real systems or expanded to incorporate greater life-stage realism.
+A second prioritization scheme proposes using population viability analysis (PVA) or surrogate measures of population viability combined with an assessment of the genetic consequences of losing various populations \citep{allendorf1997}.
+\citet{allendorf1997}, however, also focus on the assessment of individual stocks without considering their covariance and therefore the performance of the salmon portfolio as a whole.
+Compared to these detailed population-specific models, our model is at this stage primarily conceptual, but could be conditioned on real systems or expanded to incorporate greater life-stage realism.
 Our intention was to develop a parsimonious model to explore the salient aspects of metapopulation-portfolio performance.
 
-There are number of specific extensions to our simulation model that could be useful.
+Our results suggest a number of specific extensions to our simulation model.
 First our model could be extended beyond thermal response diversity.
-For the example of salmon this might include response diversity to inter-decadal climate signals such as the Pacific Decadal Oscillation \citep{mantua2002}, stream flow, or dissolved oxygen \citep{priede1988}.
-Second, aspects of our model were tailored to salmon but could be extended to metapopulations of other taxa, to ecological communities, and to meta-communities.
-Third, an extension of our model would consider the Moran or environmental-filter concept \citep{schindler2008} where populations further from each other are more likely to experience different environmental forces.
+For salmon this might include response diversity to interdecadal climate signals such as the Pacific Decadal Oscillation \citep{mantua2002}, or to stream flow and dissolved oxygen \citep{priede1988}.
+Second, our model could be extended to metapopulations of other taxa, to ecological communities (after accounting for species interactions), or to meta-communities.
+Third, an extension of our model would consider the Moran or environmental-filter concept \citep{schindler2008} where populations experience increasingly different environmental forces at further distances.
 Fourth, a model could consider the contribution of contemporary evolution \citep{stockwell2003}.
-This rapid adaptation to changes in the environment could strongly affect portfolio performance and emphasize the importance of maintaining genetic diversity.
-Finally, a fruitful extension would be to condition our model on a system of interest, for example a given river basin, and optimize that metapopulation portfolio across conservation and restoration options as part of a formal decision analysis.
+This rapid adaptation to changes in the environment could strongly affect portfolio performance and emphasizes the importance of maintaining genetic diversity.
+Finally, our model could be conditioned on a system of interest --- a particular river basin for example --- and the management metapopulation portfolio could be optimized across conservation and restoration options as part of a formal decision analysis.
 
 Our results suggest three broad research priorities.
-First our work highlights the importance of identifying response diversity \citep[e.g.][]{eliason2011}, spatial asynchrony \citep[e.g.][]{thorson2013} and in particular the management schemes that promote these processes.
-Our work highlights how these processes can contribute to reducing risk and increasing the likelihood of persistence of metapopulations under uncertainty.
-Second, our work suggests the importance of identifying what drives productivity in different scenarios.
-For example, what is the relative effect on productivity of population density, an environment-recruitment relationship, and environmental regime shifts \citep{vert-pre2013} and what management approaches are robust to our uncertainty about these contributions?
-Third, our work suggests research that looks for real-life case studies as in situ experiments testing our conclusions.
-The question remains, does managing metapopulations portfolios using modern portfolio theory result in better outcomes that are in line with the societal goals?
+First, our work highlights the importance of research that identifies response diversity \citep[e.g.][]{eliason2011}, spatial asynchrony \citep[e.g.][]{thorson2013} and in particular the management schemes that promote these processes \citep{moore2010}.
+Our work highlights how these processes can contribute to reducing risk and increasing the likelihood of metapopulation persistence given environmental uncertainty.
+Second, our work suggests the importance of identifying the drivers of productivity across populations and environmental scenarios.
+For example, what is the relative effect on productivity of population density, an environment-recruitment relationship, and environmental regime shifts and what management approaches are robust to our uncertainty about these contributions \citep{vert-pre2013}?
+Third, our work suggests research that identifies real-life case studies as in situ experiments testing our conclusions.
+The question remains, does managing metapopulation portfolios using Modern Portfolio Theory result in better outcomes that are in line with the societal goals?
 
 In conclusion, our results demonstrate the utility of considering metapopulations in a portfolio context and the power of portfolio simulation to consider robust management strategies that minimize metapopulation risk and maximize growth under uncertain environmental conditions.
-Our findings show that response diversity buffers metapopulation risk given short-term environmental stochasticity and ensures persistence given long-term regime changes.
+Our findings show that response diversity buffers metapopulation risk given short-term environmental stochasticity and ensures persistence given longterm regime changes.
 Further, our findings show that, independent of response-diversity, conserving more populations buffers metapopulation risk.
-These results point to the critical importance of choosing management, conservation, and restoration actions that preserve and promote response diversity.
+These results point to the critical importance of choosing management, conservation, and restoration actions that preserve and promote response diversity and are robust to our uncertainty about response diversity distribution and future climatic conditions.
 
 \section{Acknowledgements}
+
+Funding was provided by NSERC (ABC, NKD, SCA), the Canada Research Chairs Program (NKD), Simon Fraser University, the Liber Ero Chair of Coastal Science and Management (JWM), Fulbright Canada (SCA), and a Garfield Weston Foundation/B.C.
+Packers Ltd.
+Graduate Fellowship in Marine Sciences (SCA).
 
@@
 
-\clearpage
+\setlength{\parskip}{6pt} \setlength{\parindent}{0cm} \clearpage
+\includegraphics[height=5.5in]{../examples/simulation-diagram3.pdf}
 
-\begin{figure}[htbp]
-\centering
-\includegraphics[height=5.5in]{../examples/simulation-diagram3.pdf}
-\caption{Flow chart of the salmon-metapopulation simulation.
+Fig.~1: Flow chart of the salmon-metapopulation simulation.
 There are $n$ salmon populations and $t$ generations.
@@
 Grey text indicates parameters that can be set.
-Green text indicates the looping structure of the simulation.}
-\label{f:sim-flow}
-\end{figure}
+Green text indicates the looping structure of the simulation.
 
 \clearpage
+\includegraphics[width=3.0in]{../examples/thermal-curve-scenarios.pdf}
 
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=3.0in]{../examples/thermal-curve-scenarios.pdf}
-\caption{Different ways of prioritizing response-diversity conservation.
-Panel a shows the thermal tolerance cures for ten possible populations and panels b--e show different ways of prioritizing four of those populations.
-The curves describe how productivity varies with the environment for a given population.
-Some populations thrive at low environmental values (cool colours) and some at high (warm colours) values.
+Fig.~2: Different ways of prioritizing response-diversity conservation.
+Panel a shows thermal tolerance cures for ten possible populations and panels b--e show different ways of prioritizing four of those populations.
+The curves describe how productivity varies with temperature for a given population.
+Some populations thrive at low temperatures (cool colours) and some at warm temperatures (warm colours).
 Some are tolerant to a wider range of environmental conditions (yellow-to-green colours) but with a lower maximum productivity.
-The total possible productivity (the area under the curves) is the same for each population.}
-\label{f:curves}
-\end{figure}
+The total possible productivity (the area under the curves) is the same for each population.
 
 \clearpage
+\includegraphics[width=4.0in]{../examples/spatial-arma-sim.pdf}
 
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=4.0in]{../examples/spatial-arma-sim.pdf}
-\caption{[TODO clean up a bit - especially estimated b y-axes] The components of an example metapopulation simulation.
+Fig.~3: The components of an example metapopulation simulation.
 We show, from top to bottom, the environmental signal, the resulting productivity parameter (Ricker $a$), the salmon returns, fisheries catch, salmon escapement, salmon straying from their natal streams, salmon joining from other streams, spawner-return residuals on a log scale, and the estimated $a$ and $b$ parameters in the fitted Ricker curve.
-The colored lines indicate populations that thrive at low (cool colours) to high (warm colours) environmental values.}
-\label{f:sp-eg}
-\end{figure}
+The colored lines indicate populations that thrive at low (cool colours) to high (warm colours) temperatures.
 
 \clearpage
+\includegraphics[width=5.0in]{../examples/spatial-mv.pdf}
 
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=5.0in]{../examples/spatial-mv.pdf}
-\caption{The importance of preserving environmental response diversity through spatial conservation strategies.
-The conservation strategies correspond to Fig.~\ref{f:curves} and represent conserving a range of responses (green), the most stable populations only (orange), or one type of environmental response (purple and pink).
-In risk-return space we show environmental scenarios that are comprised primarily of (a) short-term and (b) long-term environmental fluctuations (see Fig.~X).
+Fig.~4: The importance of preserving environmental response diversity through spatial conservation strategies.
+The conservation strategies correspond to Fig.~2 and represent conserving a range of responses (green), the most stable populations only (orange), or one type of environmental response (purple and pink).
+In risk-return space we show environmental scenarios that are comprised primarily of (a) short-term and (b) long-term environmental fluctuations.
 The dots show simulated metapopulations and the contours show 25\% and 75\% quantiles across 500 simulations per strategy.
-We also show example metapopulation abundance time series for the (c, e) short-term and (d, f) long-term  environmental-fluctuation scenarios.}
-\label{f:sp-mv}
-\end{figure}
+We also show example metapopulation abundance time series for the (c, e) short-term and (d, f) long-term environmental-fluctuation scenarios.
 
 \clearpage
+\includegraphics[width=5.0in]{../examples/cons-plans-n.pdf}
 
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=5.0in]{../examples/cons-plans-n.pdf}
-\caption{[TODO move environmental ts to label, add arrow to high ext.
-risk, add labels to lower panels pointing out result] The importance of preserving as many subpopulations as possible when we don't know how response diversity is distributed.
-In risk-return space we show environmental scenarios that are comprised primarily of (a) short-term and (b) long-term environmental fluctuations (see Fig.~X).
+Fig.~5: The importance of preserving as many subpopulations as possible when we don't know how response diversity is distributed.
+In risk-return space we show environmental scenarios that are comprised primarily of (a) short-term and (b) long-term environmental fluctuations (see Fig.\textasciitilde{}X).
 We show metapopulations in which 2 (red), 4 (orange), 8 (yellow), or 16 (green) populations of random response diversity are conserved.
 The dots show simulated metapopulations and the contours show 25\% and 75\% quantiles across 500 simulations per strategy.
-We also show example metapopulation (c) rate-of-change and (d) abundance time series for the short-term environmental-fluctuation scenario.}
-\label{f:n-mv}
-\end{figure}
-
-\clearpage
+We also show example metapopulation (c) rate-of-change and (d) abundance time series for the short-term environmental-fluctuation scenario.

### 2013-08-07 inst/ms/salmonfolios.tex
Add to ms; move ms to all Pandoc

--- 2013-07-08 inst/ms/salmonfolios.tex
@@
 \documentclass[12pt]{article}
+%\documentclass{aastex}
 \usepackage{geometry}
@@
 \usepackage{booktabs}
-\setlength{\parskip}{3pt} % changes vertical space between paragraphs
+\usepackage[sectionbib]{chapterbib}
+\setlength{\parskip}{9pt} % changes vertical space between paragraphs
+\setlength{\parindent}{0cm} 
 \usepackage[round]{natbib}
@@
 
+\makeatletter
+\renewcommand\@seccntformat[1]{}
+\makeatother
+
 \hyphenation{meta-pop-ulation meta-pop-ulations sub-pop-ulations sub-pop-ulation e-con-o-mist}
 
-\title{Portfolio conservation of metapopulations under climate change}
+%\title{Portfolio conservation of metapopulations\\under climate change}
 
-\author{
-  Sean C. Anderson$^{1*}$, 
-  Jonathan W. Moore$^1$,
-  Trevor A. Branch?,\\
-  Nicholas K. Dulvy$^1$,
-  Michelle M. McClure?,
-  Andrew B. Cooper$^2$
-}
-\date{}
+%\author{
+  %Sean C. Anderson$^{1*}$, 
+  %Jonathan W. Moore$^1$,
+  %Trevor A. Branch?,\\
+  %Nicholas K. Dulvy$^1$,
+  %Michelle M. McClure?,
+  %Andrew B. Cooper$^2$\\(Order to be discussed)
+%}
+%\date{}
 
@@
 \begin{document}
-%\onehalfspacing
-\linenumbers
-\modulolinenumbers[2]
-\begin{spacing}{1.3}
+%\begin{spacing}{1.1}
 %\doublespacing
-\maketitle
+%\maketitle
 
-\noindent
-  \textsuperscript{1}Department of Biological Sciences, Simon Fraser
-  University, Burnaby BC, V5A 1S6, Canada
+%\noindent
+  %\textsuperscript{1}Department of Biological Sciences, Simon Fraser
+  %University, Burnaby BC, V5A 1S6, Canada
 
-\noindent
-	\textsuperscript{2}School of Resource and Environmental Management, Simon
-  Fraser University, Burnaby, BC, V5A 1S6, Canada
+%\noindent
+	%\textsuperscript{2}School of Resource and Environmental Management, Simon
+  %Fraser University, Burnaby, BC, V5A 1S6, Canada
 
-\noindent
-Statement of authorship: 
+%\noindent
+%\textit{Statement of authorship}: SCA, JWM, NKD, and ABC designed the analyses.
+SCA wrote the simulation and analyzed the results.
+SCA and XX \ldots wrote the manuscript.
+All authors participated in conception, discussion of the results, and manuscript preparation.
 %Contributions by authors should be listed on the title page and will be printed at the end of the manuscript.
@@
 
-\noindent
-Running title: 
+%\noindent
+%\textit{Running title}:
+%Portfolio conservation of metapopulations
 %(abbreviated form of title) of less than 45 characters including spaces
 
-\noindent
-Keywords:
-%up to 10
+%\noindent
+%\textit{Keywords}: 
+%biocomplexity, 
+%Modern Portfolio Theory, 
+%portfolio effect, 
+%prioritization, 
+%resource management, 
+%range contraction, 
+%response diversity, 
+%risk assessment, 
+%salmon, 
+%stochastic simulation 
+%(maximum 10)
+%%up to 10
 
-\noindent
-Type of article: Letters
+%\noindent
+%\textit{Type of article}: Letters
 
-\noindent
-Abstract:
+%\noindent
+%\textit{Words in abstract}: (maximum 150)
 
-\noindent
-Main text:
-%(excluding abstract, acknowledgements, references, table and figure legends), and the number of words in each text box
+%\noindent
+%\textit{Words in main text}: (maximum 5000)
+%%(excluding abstract, acknowledgements, references, table and figure legends), and the number of words in each text box
 
-\noindent
-References:
+%\noindent
+%\textit{Number of references}:
 
-\noindent
-Number of figures:
+%\noindent
+%\textit{Number of figures}: 5
 
-\noindent
-Number of tables:
+%\noindent
+%\textit{Number of tables}: 1?
 
  %- the name and complete mailing address (including telephone and fax numbers and e-mail address) of the person to whom correspondence should be sent
-\noindent
-	\textsuperscript{*}Corresponding author: Sean C. Anderson; Department of Biological Sciences, Simon Fraser University, Burnaby BC, V5A 1S6; Phone: 1-778-782-3989; E-mail: sean\_anderson@sfu.ca
+%\noindent
+	%\textsuperscript{*}Corresponding author: Sean C. Anderson; Department of Biological Sciences, Simon Fraser University, Burnaby BC, V5A 1S6; Phone: 1-778-782-3989; E-mail: sean\_anderson@sfu.ca
 
-\clearpage
+%\end{spacing}
+%\clearpage
 
+%\linenumbers
+%\modulolinenumbers[2]
+\begin{spacing}{1.4}
+%\onehalfspacing
 \include{ms}

### 2013-08-17 inst/ms/ms.tex
Make edits to ms

@@
 McClure?, Andrew B.
-Cooper\textsuperscript{2} (Order to be discussed)
+Cooper\textsuperscript{2} (Authors and order to be discussed)
 
@@
 \emph{Statement of authorship}: SCA, JWM, NKD, and ABC designed the analyses.
-SCA wrote the simulation, analyzed the results, and wrote the first draft of the manuscript.
+SCA wrote the simulation code, analyzed the results, and wrote the first draft of the manuscript.
 All authors contributed substantially to revisions.
@@
 
-Managing risk is fundamental to the conservation of an endangered species \citep{burgman2005, iucn2009}.
+Managing risk is fundamental to conserving endangered species \citep{burgman2005, iucn2009}.
 When an endangered species exists as a metapopulation, we can manage risk at two levels: at the population level or at the metapopulation level.
@@
 Economists consider the risk and performance of a financial portfolio based on the weighting of individual investments (called assets) that make up the portfolio.
-Modern Portfolio Theory proposes that there is a set of portfolios that maximizes expected return for a level of expected risk or minimizes expected risk for a level of expected return \citep{markowitz1952, markowitz1959}.
+Modern Portfolio Theory proposes that a set of portfolios maximizes expected return for a level of risk or minimizes risk for a level of return \citep{markowitz1952, markowitz1959}.
 Similarly, expected growth rate and variance of a metapopulation is a function of the variance, covariance, and size of the individual populations \citep{moore2010}.
@@
 Continuing the analogy, predators and fisheries often integrate across multiple populations \citep{hilborn2003, schindler2008}, acting as investors in the salmon portfolio.
-Fisheries managers and conservation agencies then act as portfolio managers through actions such choosing which salmon habitat to prioritize for protection or restoration.
+Fisheries managers and conservation agencies then act as portfolio managers by choosing which salmon habitat to prioritize for protection or restoration.
 (2) Many Pacific salmon metapopulations are highly threatened \citep{mcclure2003, gustafson2007, peterman2012} and will likely become more at risk as threats such as overfishing, dams, roads, logging, and particularly climate change intensify \citep[e.g.][]{lackey2003}.
@@
 A diversified financial portfolio reduces risk by exploiting the covariance between assets --- ideally one stocks rises while another one falls --- and there are two key mechanisms that generate asynchrony in metapopulation dynamics.
-First, localized habitat features can filter the environment generating unique conditions for subpopulations \citep{schindler2008} (i.e.~Moran effect).
-For salmon, this mechanism acts primarily during freshwater life-history stages and since many salmon species spend considerable time in the ocean \citep{quinn2005}, a second mechanism is also of key importance --- diversity of response to the environment \citetext{\citealp[i.e.~response diversity,][]{elmqvist2003}; \citealp[and biocomplexity][]{colwell1998}; \citealp{hilborn2003}}.
+First, localized habitat features can filter the environment generating unique conditions for subpopulations \citep{schindler2008, rogers2008} (\emph{senu} the Moran effect).
+For salmon, this mechanism acts primarily during freshwater life-history stages and since many salmon species spend considerable time in the ocean \citep{quinn2005}, a second mechanism is also of key importance --- diversity of response to a shared environment \citetext{\citealp[i.e.~response diversity,][]{elmqvist2003}; \citealp[and biocomplexity][]{colwell1998}; \citealp{hilborn2003}}.
 This second mechanism can result from unique phenotypic or genetic traits \citep{crozier2008, kovach2012} and be derived from rapid contemporary adaption \citep{stockwell2003, fraser2011} or long-term evolution to historical conditions \citep{eliason2011}.
@@
 In addition to climate warming posing perhaps the greatest threat to global biodiversity \citep{thomas2004}, warming poses a particular threat to riverine species whose ranges are largely confined to existing habitat \citep{thomas2010} and where shallower water, compared to larger water basins, can be affected more rapidly by thermal changes \citep{isaak2010} and by snow-melt timing and extreme hydrological events \citep{crozier2008}.
-These consequences, and in particular warmer water itself, can combine to limit salmon productivity through oxygen limitation \citep{portner2007} leading to decreased cardiac performance \citep{eliason2011} and having profound implications for behaviour \citep{goniea2006}, disease resistance (REF), energetic costs in general (REF Rand 1998 from Pess 2012?), and ultimately survival \citep{peterman1998, eliason2011}.
-With 29\% of Pacific Northwest salmon population having gone extinct in the last XX years and XX\% of remaining stocks classified as at risk by XX, these populations are at risk from any further decline in fitness \citep{crozier2008, mcclure2003}.
-Yet, adverse stream temperatures are already impeding recovery of some PNW salmon populations \citep{mccullough1999} and leading to earlier and more condensed migration of salmon \citep{kovach2013a}\ldots{} finish this.
+These consequences, and in particular warmer water itself, can combine to limit salmon productivity through, for example, oxygen limitation \citep{portner2007} leading to decreased cardiac performance \citep{eliason2011} and having profound implications for behaviour \citep{goniea2006}, disease resistance \citep{crozier2008}, and energetic costs in general \citep{rand1998}, and ultimately survival \citep{peterman1998, eliason2011}.
+With 29\% of Pacific Northwest salmon populations having gone extinct since Euro-American contact \citep{gustafson2007}, these populations are at risk from any further decline in fitness \citep{mcclure2003}.
+Yet, adverse stream temperatures are already impeding recovery of some PNW salmon populations \citep{mccullough1999}.
 
@@
 The term $w_{i(t)}$ represents first-order autocorrelated error.
-Formally, $w_{i(t)} = w_{ti-1} \rho_w + r_{i(t)}$, where $r_{i(t)}$ represents independent and normally distributed error with mean 0 and standard deviation of $\sigma_r$.
-The parameter $\rho_w$ represents the correlation between residuals from subsequent generations.
+Formally, $w_{i(t)} = w_{ti-1} \rho_w + r_{i(t)}$, where $r_{i(t)}$ represents independent and normally distributed error with mean 0 and standard deviation of $\sigma_r$ (0.3; we include the base values in parentheses after relevant parameters).
+The parameter $\rho_w$ (0.4) represents the correlation between residuals from subsequent generations.
 
@@
 \noindent where $W_i$ controls the width of the curve for population $i$, $e_t$ represents the environmental value at generation $t$, $e_i^{\mathrm{opt}}$ represents the optimal temperature for population $i$, and $a_i^{\mathrm{max}}$ represents the maximum possible $a$ value for population $i$.
-We set the $W_i$ parameters (Table S1) and calculated the $a_i^{\mathrm{max}}$ parameters so that the area under each curve $A_i$ was equal.
-We chose parameter values such that some populations were warm-tolerant, some were cold-tolerant, and some had a wider range of thermal-tolerance, but a lower maximum productivity (see Fig.~2a and Appendix S1 in Supplementary Information).
+We set the $W_i$ parameters (evenly spaced values increasing and decreasing between 0.05 and 0.02) and calculated the $a_i^{\mathrm{max}}$ parameters so that the area under each curve was equal (30 units).
+We chose parameter values such that some populations were warm-tolerant, some were cold-tolerant, and some had a wider range of thermal-tolerance, but a lower maximum productivity (see Fig.~2a).
 Although we refer to a thermal-tolerance curve because temperature is a dominant driver of salmonid performance \citep{mccullough1999}, our model applies to any environmental tolerance (e.g.~tolerance to stream flow volume or changes in snow melt timing \citep{crozier2008}).
@@
 We arranged the subpopulations in a line and salmon were more likely to stray to streams near their natal stream.
-Two parameters controlled the straying: the fraction of fish $f_{\mathrm{stray}}$ that stray from their natal stream in any generation and the rate $m$ at which this straying between streams decays with distance.
+Two parameters controlled the straying: the fraction of fish $f_{\mathrm{stray}}$ (0.02) that stray from their natal stream in any generation and the rate $m$ (0.3) at which this straying between streams decays with distance.
 We calculated the number of salmon straying from stream $j$ to stream $i$ as
@@
 \noindent where $\alpha_h$ and $\beta_h$ are the location and shape parameters in a beta distribution.
-They can be calculated from the desired mean $H_{\mathrm{tar}}$ and standard deviation $\sigma_h$ as \citep[p.~97]{morgan1990}
+They can be calculated from the desired mean $H_{\mathrm{tar}}$ and standard deviation $\sigma_h$ (0.05) as \citep[p.~97]{morgan1990}
 
@@
 We evaluated portfolio performance under these two components separately.
-We represented short-term dynamics as a stationary first-order autoregressive process, AR(1), with correlation $\rho_e$
+We represented short-term dynamics as a stationary first-order autoregressive process, AR(1), with correlation $\rho_e$ (0.1)
 
@@
 
-\noindent where $e_t$ represents the environmental value in generation $t$ and $d$ represents normally distributed deviations of mean 0 and standard deviation $\sigma_d$.
+\noindent where $e_t$ represents the environmental value in generation $t$ and $d$ represents normally distributed deviations of mean 0 and standard deviation $\sigma_d$ (2).
 We represented long-term environmental dynamics as a linear shift in the environmental value through time
@@
 
-\noindent where $\beta_e$ represents the slope.
+\noindent where $\beta_e$ (0.114) represents the slope.
 To maintain a balanced response, we centered the trend by subtracting the mean $\overline{\beta_e t}$ so that midway through the simulation (after any burn-in period) the environmental value was at the mean environmental tolerance.
@@
 
-The primary lesson from our simulation model is the importance of response diversity in buffering risk and ensuring long-term persistence in the face of environmental uncertainty.
-Thus, our results emphasize the need to make the maintenance of response diversity, and the processes that maintain response diversity, a primary goal of conservation and recovery efforts \citep{moore2010}.
-For example, since habitat heterogeneity can lead to local adaptation \citep{taylor1991, fraser2011}, our results emphasize the need to maintain a diversity of habitat.
+The primary finding from our simulation model is the importance of response diversity in buffering risk and ensuring long-term persistence in the face of environmental uncertainty.
+Thus, our results emphasize the need to make the maintenance of response diversity, and the processes that promote response diversity, a primary goal of conservation and recovery efforts \citep{moore2010}.
+For example, since habitat heterogeneity can lead to local adaptation \citep{taylor1991, fraser2011}, our results emphasize the need to maintain a diversity of salmon habitat \citep{rogers2008}.
 Our results can also inform habitat prioritization decisions.
-If closer populations are more correlated with each other and conservation actions must be prioritized, then our model suggests we should focus on populations that aren't spatially contiguous, as long as sufficient straying is maintained, to maximize diversity of response to the environment.
+If conservation actions must be prioritized, then our model suggests we should focus on populations that aren't spatially contiguous, as long as sufficient straying is maintained, to maximize diversity of response to the environment.
 Finally, our results demonstrate the advantages of avoiding structures that artificially remove response diversity.
 For salmon, dams are a prominent example \citep{waples2007, pess2012}.
-Dams can have a double impact whereby their introduction selectively eliminates a large swath of contiguous habitat, perhaps analogous to our upper- or lower-half scenarios in Fig.~4, and then mitigation approaches such as hatcheries and smolt transportation can further reduce population and response diversity (REFS).
+Dams can have a double impact whereby their introduction selectively eliminates a large swath of contiguous habitat, perhaps analogous to our upper- or lower-half scenarios in Fig.~4, and then mitigation approaches such as hatcheries can further reduce population and response diversity if not carefully managed \citep{waples1991a, mobrand2005}.
 
 Our simulation model emphasizes the importance of thinking at the portfolio level --- considering aggregate risk and performance for ecological portfolios.
-In many ways our results are intuitive, however, we often don't consider aggregate risk and only by simultaneously considering aggregate risk and return can we optimally diversify ecological systems \citep{ando2012}.
-To that end we can think of portfolio risk-return performance as another dimension of conservation.
-This dimension emphasizes the value of a broader scale of management and focussing not just on currently productive populations but also on less productive populations to reduce long-term aggregate risk \citep{hilborn2003}.
+Although our results may appear intuitive, we often don't often consider aggregate risk and only by simultaneously considering aggregate risk and return can we optimally diversify ecological systems \citep{ando2012}.
+Thus we can think of portfolio risk-return performance as another dimension of conservation.
+This dimension emphasizes the value of a broader scale of management and focussing not just on currently productive populations but also on less productive populations to reduce long-term aggregate risk \citep[e.g.][]{hilborn2003, price2008}.
 This perspective does not negate an understanding at a finer scale, say of individual fish stocks.
-Rather, as in finance, it requires an understanding of stock performance but also an understanding of the covariance between stocks to predict the properties of the portfolio as a whole.
+Rather, as in finance, it requires both an understanding of stock performance and an assessment of their covariance to predict properties of the portfolio as a whole.
 
-Our findings support a number of existing guidelines and management structures for Pacific Northwest salmonids.
+THE FOLLOWING PARAGRAPH IS STILL BEING PIECED TOGETHER
 
-Across the range of Pacific Northwest salmonids, United States and Canadian guidelines are increasingly well-structured to benefit from the findings of our simulation model.
+Our findings support the existence of and inform the implementation of a number of existing guidelines and management structures for PNW salmonids.
+In the US, Evolutionarily Significant Units (ESUs) represent salmon populations that are reproductively isolated with an evolutionary heritage \citep{waples1991, moritz1994} and finer-scale Viable Salmonid Populations (VSPs) represent groups of fish that spawn in a given lake or stream in a particular season \citep{mceacheran2000}.
+In Canada, the rough equivalent to the ESU is a Conservation Unit (CU) as defined in the Wild Salmon Policy \citep{dfo2005wsp}.
+A CU consists of a group of salmon that are reproductively isolated and that if lost would be unlikely to recolonize in a reasonable time frame \citep{dfo2005wsp}.
+Both the US and Canadian guidelines recognize still-finer-scale population structure.
+We suggest that a salmon portfolio in our model could represent an ESU or CU and that the lessons learned from our models are thus directly applicable to management guidelines in the PNW.
 
-In the US, larger-scale Evolutionary Significant Units (ESUs) represent populations that are reproductively isolated with an evolutionary heritage \citep{waples1991, moritz1994} and finer-scale Viable Salmonid Populations (VSPs) represent groups of fish that spawn in a given lake or stream in a particular season \citep{mceacheran2000}.
-In Canada, the rough equivalent to the ESU is a Conservation Unit (CU) as defined in the Wild Salmon Policy.
-A CU consists of a group of salmon that are reproductively isolated and that if lost would be unlikely to recolonize in a reasonable time frame \citep{dfo2005wsp}.
-Both the US and Canadian guidelines recognize still finer-scale population structure.
+VSP guidelines \citep{mceacheran2000} suggest\ldots{}
+
+\begin{itemize}
+\item
+  maintaining diversity of a variety of forms to buffer against short-term environmental changes and adaptation to long-term environmental changes
+\item
+  conservation efforts should not focus exclusively where there is currently an abundance of salmon
+\item
+  metapopulation (multiple populations with some exchange) are important to maintain
+\item
+  some should be close for straying, but some should be far apart to reduce risk from sudden environmental change
+\item
+  important to incorporate uncertainty of ESU-level processes into assessments
+\end{itemize}
+
+CU guidelines \citep{dfo2005wsp} suggest\ldots{}
+
+-
 
@@
 
-Our model is unique in that it is the first to consider risk for the metapopulation as a whole while identifying the role of response diversity in buffering risk.
+END IN PROGRESS TEXT
+
+Our model is unique in that it is the first to consider risk for metapopulations as a whole while identifying the role of response diversity in buffering risk.
 Our model follows from the work of \citet{cooper1999} who showed the importance of metapopulation habitat structure for salmonid conservation and \citet{moore2010} who linked salmon portfolio performance to asynchrony in population dynamics.
@@
 
-Our results suggest a number of specific extensions to our simulation model.
+We suggest a number of specific extensions to our simulation model.
 First our model could be extended beyond thermal response diversity.
@@
 Second, our model could be extended to metapopulations of other taxa, to ecological communities (after accounting for species interactions), or to meta-communities.
-Third, an extension of our model would consider the Moran or environmental-filter concept \citep{schindler2008} where populations experience increasingly different environmental forces at further distances.
+Third, an extension of our model would consider the Moran or environmental-filter concept \citep{schindler2008, rogers2008} where populations experience increasingly different environmental forces at further distances.
 Fourth, a model could consider the contribution of contemporary evolution \citep{stockwell2003}.
@@
 Our results suggest three broad research priorities.
-First, our work highlights the importance of research that identifies response diversity \citep[e.g.][]{eliason2011}, spatial asynchrony \citep[e.g.][]{thorson2013} and in particular the management schemes that promote these processes \citep{moore2010}.
+First, our work highlights the importance of research that identifies response diversity \citep[e.g.][]{eliason2011}, spatial asynchrony \citep[e.g.][]{thorson2013}, and in particular the management schemes that promote these processes \citep{moore2010}.
 Our work highlights how these processes can contribute to reducing risk and increasing the likelihood of metapopulation persistence given environmental uncertainty.
 Second, our work suggests the importance of identifying the drivers of productivity across populations and environmental scenarios.
-For example, what is the relative effect on productivity of population density, an environment-recruitment relationship, and environmental regime shifts and what management approaches are robust to our uncertainty about these contributions \citep{vert-pre2013}?
+For example, what is the relative effect on productivity of population density, an environment-recruitment relationship, and environmental regime shifts, and what management approaches are robust to our uncertainty about these contributions \citep{vert-pre2013}?
 Third, our work suggests research that identifies real-life case studies as in situ experiments testing our conclusions.
@@
 
-\setlength{\parskip}{6pt} \setlength{\parindent}{0cm} \clearpage
-\includegraphics[height=5.5in]{../examples/simulation-diagram3.pdf}
+\setlength{\parskip}{6pt} \setlength{\parindent}{0cm} \includegraphics[height=5.5in]{../examples/simulation-diagram3.pdf}
 
@@
 Fig.~5: The importance of preserving as many subpopulations as possible when we don't know how response diversity is distributed.
-In risk-return space we show environmental scenarios that are comprised primarily of (a) short-term and (b) long-term environmental fluctuations (see Fig.\textasciitilde{}X).
+In risk-return space we show environmental scenarios that are comprised primarily of (a) short-term and (b) long-term environmental fluctuations.
 We show metapopulations in which 2 (red), 4 (orange), 8 (yellow), or 16 (green) populations of random response diversity are conserved.

### 2013-08-17 inst/ms/salmonfolios.tex
Make edits to ms

--- 2013-07-08 inst/ms/salmonfolios.tex
@@
 \usepackage{graphicx}
-\usepackage{lineno}
-\usepackage{fontspec}
-\textheight 20.5cm
 \usepackage{setspace}
-\usepackage{booktabs}
-\setlength{\parskip}{3pt} % changes vertical space between paragraphs
 \usepackage[round]{natbib}
 \bibpunct{(}{)}{;}{a}{}{;}
+\textheight 20.5cm
+\usepackage{lineno}
+\usepackage{xcolor}
+\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
+
+% remove numbers in front of sections:
+\makeatletter
+\renewcommand\@seccntformat[1]{}
+\makeatother
 
@@
 
-\title{Portfolio conservation of metapopulations under climate change}
+\begin{document}
 
-\author{
-  Sean C. Anderson$^{1*}$, 
-  Jonathan W. Moore$^1$,
-  Trevor A. Branch?,\\
-  Nicholas K. Dulvy$^1$,
-  Michelle M. McClure?,
-  Andrew B. Cooper$^2$
-}
-\date{}
-
-%\\Climate 
-  %change and portfolio conservation of metapopulations\\Climate change and 
-  %portfolio conservation\\Portfolio prioritization of salmon 
-  %conservation\\Portfolio prioritization of metapopulation conservation
-
-%-   Prioritizing metapopulation conservation through the lens of
-    %portfolio theory
-
-%-   Portfolio optimization reveals...
-
-%-   Portfolio theory reveals conservation strategies that...
-
-%-   Intelligent tinkering with ecological portfolios: patterns of
-    %environmental response diversity drive salmon conservation
-    %priorities
-
-%-   Keeping every cog and wheel...
-
-%**Ecological Applications**
-
-%-   60 pages including everything
-
-%> Ecological Applications is concerned broadly with the applications of
-%> ecological science to environmental problems. It publishes papers that
-%> develop scientific principles to support environmental
-%> decision-making, as well as papers that discuss the application of
-%> ecological concepts to environmental issues, policy, and management.
-%> Papers may report on experimental tests, actual applications,
-%> scientific decision support techniques, economic analyses, social
-%> implications of environmental issues, or other relevant topics.
-%> Statistical or experimental methods papers that support research and
-%> applications are welcome. Papers submitted to Ecological Applications
-%> should be accessible to both scholars and practitioners.
-
-\begin{document}
-%\onehalfspacing
 \linenumbers
 \modulolinenumbers[2]
-\begin{spacing}{1.3}
-%\doublespacing
-\maketitle
-
-\noindent
-  \textsuperscript{1}Department of Biological Sciences, Simon Fraser
-  University, Burnaby BC, V5A 1S6, Canada
-
-\noindent
-	\textsuperscript{2}School of Resource and Environmental Management, Simon
-  Fraser University, Burnaby, BC, V5A 1S6, Canada
-
-\noindent
-Statement of authorship: 
-%Contributions by authors should be listed on the title page and will be printed at the end of the manuscript.
-This statement should be appropriate to the study described in the manuscript and should clarify who designed the study, who performed the research, who provided new methods or materials, and who wrote the manuscript.
-We encourage concise statements such as “JW performed phylogenetic analyses, MH collected data, performed modeling work and analyzed output data, and PK performed the meta-analysis.
-MH wrote the first draft of the manuscript, and all authors contributed substantially to revisions.”
-
-\noindent
-Running title: 
-%(abbreviated form of title) of less than 45 characters including spaces
-
-\noindent
-Keywords:
-%up to 10
-
-\noindent
-Type of article: Letters
-
-\noindent
-Abstract:
-
-\noindent
-Main text:
-%(excluding abstract, acknowledgements, references, table and figure legends), and the number of words in each text box
-
-\noindent
-References:
-
-\noindent
-Number of figures:
-
-\noindent
-Number of tables:
-
- %- the name and complete mailing address (including telephone and fax numbers and e-mail address) of the person to whom correspondence should be sent
-\noindent
-	\textsuperscript{*}Corresponding author: Sean C. Anderson; Department of Biological Sciences, Simon Fraser University, Burnaby BC, V5A 1S6; Phone: 1-778-782-3989; E-mail: sean\_anderson@sfu.ca
-
-\clearpage
-
+\begin{spacing}{1.5}
 \include{ms}

### 2013-08-17 inst/ms/som.tex
Make substantial additions to main text
Add to ms; move ms to all Pandoc
Make edits to ms

@@
-\section{Supporting online material}
+\section{Supplementary Information}
 
@@
 
-The \texttt{metafolio} \texttt{R} package and documentation.
+The \texttt{metafolio} \texttt{R} package and documentation. Some
+details on what you can do with the package.
 
@@
 populations (indicated by population number). Dark blue indicates a high rate 
-of straying and light blue indicates a low rate of straying. The straying 
-matrix is generated through equation \ref{eq:stray}.}
+of straying and light blue indicates a low rate of straying.}
 \label{f:stray}
@@
 \clearpage
-
-A placeholder: \citep{schindler2010}
-
-\bibliographystyle{apalike}
-
-\bibliography{jshort,som}

### 2013-08-26 inst/ms/ms.tex
Updates to ms; wrap ms

@@
+Overall thoughts and notes:
+
+\begin{itemize}
+\item
+  The introduction may be a bit literature heavy. (We also need to cut 11 references to get to 50.)
+\item
+  The main message may be a bit repetitive throughout.
+\item
+  I wish I didn't have so many paragraphs with numbered lists, but am struggling to work around that.
+\item
+  I think there is room for expansion of the discussion (we still have \textasciitilde{}500 words to play with).
+The discussion could benefit from bringing in more concrete examples of how this ties in with PNW salmon.
+\item
+  The supplementary figures and captions aren't finished but should be quick to generate.
+\item
+  I will write a short vignette describing how to use the R package like I did for the Methods in Ecology and Evolution paper.
+\item
+  I will submit the \texttt{.pdf} version; the \texttt{.docx} version is for you to edit with Track Changes, if you'd like.
+Only the \texttt{.pdf} version contains the figures and if you can't see the equations in the \texttt{.docx} version (Andy?) then see the \texttt{.pdf} version.
+\item
+  We still need to discuss authorship and order of authorship.
+\end{itemize}
+
+\newpage
+
 \section{Portfolio conservation of metapopulations under climate change}
 
-\setlength{\parskip}{6pt} \setlength{\parindent}{0cm}
+\bigskip
 
@@
 Anderson\textsuperscript{1*}, Jonathan W.
-Moore\textsuperscript{1}, Trevor A.
-Branch?, Nicholas K.
-Dulvy\textsuperscript{1}, Michelle M.
-McClure?, Andrew B.
-Cooper\textsuperscript{2} (Authors and order to be discussed)
+Moore\textsuperscript{1}, Nicholas K.
+Dulvy\textsuperscript{1}, Andrew B.
+Cooper\textsuperscript{2} (Michelle M.
+McClure?, Trevor A.
+Branch?, Authors and order to be discussed)
 
@@
 \textsuperscript{*}Corresponding author: Sean C. Anderson; Department of Biological Sciences, Simon Fraser University, Burnaby BC, V5A 1S6; Phone: 1-778-782-3989; E-mail: sean\_anderson@sfu.ca
-
-\setlength{\parskip}{2pt} \setlength{\parindent}{16pt}
 
@@
 
-A diversified financial portfolio reduces risk by exploiting the covariance between assets --- ideally one stocks rises while another one falls --- and there are two key mechanisms that generate asynchrony in metapopulation dynamics.
+A diversified financial portfolio reduces risk by exploiting the covariance between assets --- ideally one stocks rises while another one falls --- and two key mechanisms can generate such asynchrony in metapopulation dynamics.
 First, localized habitat features can filter the environment generating unique conditions for subpopulations \citep{schindler2008, rogers2008} (\emph{senu} the Moran effect).
-For salmon, this mechanism acts primarily during freshwater life-history stages and since many salmon species spend considerable time in the ocean \citep{quinn2005}, a second mechanism is also of key importance --- diversity of response to a shared environment \citetext{\citealp[i.e.~response diversity,][]{elmqvist2003}; \citealp[and biocomplexity][]{colwell1998}; \citealp{hilborn2003}}.
-This second mechanism can result from unique phenotypic or genetic traits \citep{crozier2008, kovach2012} and be derived from rapid contemporary adaption \citep{stockwell2003, fraser2011} or long-term evolution to historical conditions \citep{eliason2011}.
+However, since many salmon species spend considerable time in a shared ocean environment \citep{quinn2005}, a second mechanism is also important --- diversity of response to the environment \citetext{\citealp[i.e.~response diversity,][]{elmqvist2003}; \citealp[and biocomplexity][]{colwell1998}; \citealp{hilborn2003}}.
+This mechanism can result from unique phenotypic or genetic traits \citep{crozier2008, kovach2012} and be derived from rapid contemporary adaption \citep{stockwell2003, fraser2011} or long-term evolution to historical conditions \citep{eliason2011}.
 In this paper we focus on this latter, response-diversity mechanism, and evaluate it in the context of thermal response diversity.
 
-Salmon are strongly affected by climate warming and yet show a remarkable diversity of tolerance to temperature \citep{beacham1989, crozier2006, battin2007, crozier2008}.
+Salmon are strongly affected by climate warming and yet show a remarkable diversity of tolerance to temperature \citep{beacham1989, crozier2006, crozier2008}.
 In addition to climate warming posing perhaps the greatest threat to global biodiversity \citep{thomas2004}, warming poses a particular threat to riverine species whose ranges are largely confined to existing habitat \citep{thomas2010} and where shallower water, compared to larger water basins, can be affected more rapidly by thermal changes \citep{isaak2010} and by snow-melt timing and extreme hydrological events \citep{crozier2008}.
-These consequences, and in particular warmer water itself, can combine to limit salmon productivity through, for example, oxygen limitation \citep{portner2007} leading to decreased cardiac performance \citep{eliason2011} and having profound implications for behaviour \citep{goniea2006}, disease resistance \citep{crozier2008}, and energetic costs in general \citep{rand1998}, and ultimately survival \citep{peterman1998, eliason2011}.
+These consequences, and in particular warmer water itself, can combine to limit salmon productivity through, for example, oxygen limitation \citep{portner2007} leading to decreased cardiac performance \citep{eliason2011} and having profound implications for behaviour \citep{goniea2006}, disease resistance \citep{crozier2008}, energetic costs in general \citep{rand1998}, and ultimately survival \citep{peterman1998, eliason2011}.
 With 29\% of Pacific Northwest salmon populations having gone extinct since Euro-American contact \citep{gustafson2007}, these populations are at risk from any further decline in fitness \citep{mcclure2003}.
-Yet, adverse stream temperatures are already impeding recovery of some PNW salmon populations \citep{mccullough1999}.
+Yet, adverse stream temperatures are already impeding recovery of some PNW salmon populations \citep{mccullough1999} and are expected to make recovery targets more difficult to achieve \citep{battin2007}.
 
@@
 
-\noindent where $i$ represents a population, $t$ a generation time, $R$ the number of returns, $S$ the number of spawners, $a$ the productivity parameter (which can vary with the environment), and $b$ the density-dependent term (which is used as the asset weights in the portfolios).
+where $i$ represents a population, $t$ a generation time, $R$ the number of returns, $S$ the number of spawners, $a$ the productivity parameter (which can vary with the environment), and $b$ the density-dependent term (which is used as the asset weights in the portfolios).
 The term $w_{i(t)}$ represents first-order autocorrelated error.
@@
 
-\[a_{i(t)} =
-  \begin{cases}
-    W_i (e_t - e_i^{\mathrm{opt}})^2 + a_i^{\mathrm{max}},
-      & \text{if } a_{i(t)} > 0\\
-      0, & \text{if } a_{i(t)} \leq 0
-  \end{cases}\]
+\[a_{i(t)} = \begin{cases} W_i (e_t - e_i^{\mathrm{opt}})^2 +
+a_i^{\mathrm{max}}, & \text{if } a_{i(t)} > 0\\ 0, & \text{if } a_{i(t)} \leq
+0 \end{cases}\]
 
-\noindent where $W_i$ controls the width of the curve for population $i$, $e_t$ represents the environmental value at generation $t$, $e_i^{\mathrm{opt}}$ represents the optimal temperature for population $i$, and $a_i^{\mathrm{max}}$ represents the maximum possible $a$ value for population $i$.
+where $W_i$ controls the width of the curve for population $i$, $e_t$ represents the environmental value at generation $t$, $e_i^{\mathrm{opt}}$ represents the optimal temperature for population $i$, and $a_i^{\mathrm{max}}$ represents the maximum possible $a$ value for population $i$.
 We set the $W_i$ parameters (evenly spaced values increasing and decreasing between 0.05 and 0.02) and calculated the $a_i^{\mathrm{max}}$ parameters so that the area under each curve was equal (30 units).
@@
 
-\[\mathrm{strays}_{ij(t)} = f_{\mathrm{stray}} R_{j(t)}
-    \frac{e^{-m \lvert i-j \rvert }}
-      {\displaystyle\sum\limits_{
-        \substack{k = 1 \\ k \neq j}}^{n} 
-        e^{-m \lvert k-j \rvert }}\]
+\[\mathrm{strays}_{ij(t)} = f_{\mathrm{stray}} R_{j(t)} \frac{e^{-m \lvert i-j
+\rvert }} {\displaystyle\sum\limits_{ \substack{k = 1 \\ k \neq j}}^{n} e^{-m
+\lvert k-j \rvert }}\]
 
-\noindent where $R_{j(t)}$ is the number of returning salmon at generation $t$ whose natal stream was stream $j$.
+where $R_{j(t)}$ is the number of returning salmon at generation $t$ whose natal stream was stream $j$.
 The subscript $k$ represents a stream ID and $n$ the number of populations.
@@
 Our simulation did not account for the homogenization of diversity due to straying.
-For example, all salmon in population A maintained the same thermal-tolerance curve regardless of how many salmon it received from population B.
+For example, all salmon in one population maintained the same thermal-tolerance curve regardless of how many salmon it received from another stream.
 However, the low straying rates in our simulation rate should render this effect minimal except for populations at low abundance.
@@
 
-\[H_{\mathrm{tar}} = \frac{A}{b (0.5 - 0.07a)}
-  \label{eq:esc}\]
+\[H_{\mathrm{tar}} = \frac{A}{b (0.5 - 0.07a)} \label{eq:esc}\]
 
-\noindent where $A$ represents the return abundance and $a$ and $b$ represent the Ricker model parameters. We included implementation uncertainty in the actual harvest rate $H_{\mathrm{act}}$ as
+where $A$ represents the return abundance and $a$ and $b$ represent the Ricker model parameters. We included implementation uncertainty in the actual harvest rate $H_{\mathrm{act}}$ as
 
@@
 
-\noindent where $\alpha_h$ and $\beta_h$ are the location and shape parameters in a beta distribution.
+where $\alpha_h$ and $\beta_h$ are the location and shape parameters in a beta distribution.
 They can be calculated from the desired mean $H_{\mathrm{tar}}$ and standard deviation $\sigma_h$ (0.05) as \citep[p.~97]{morgan1990}
 
-\[\begin{aligned}
-  \alpha_h &= H_{\mathrm{tar}}^2
-                \left(
-                \frac{1 - H_{\mathrm{tar}}}{\sigma_h^2} - \frac{1}{H_{\mathrm{tar}}}
-                \right)\\
-   \beta_h &= \alpha \left({\frac{1}{H_{\mathrm{tar}}} - 1}\right).\end{aligned}\]
+\[\begin{aligned} \alpha_h &= H_{\mathrm{tar}}^2 \left( \frac{1 -
+H_{\mathrm{tar}}}{\sigma_h^2} - \frac{1}{H_{\mathrm{tar}}} \right)\\ \beta_h
+&= \alpha \left({\frac{1}{H_{\mathrm{tar}}} - 1}\right).\end{aligned}\]
 
-\noindent Further, to establish a range of spawner-return values and to mimic the start of an open-access fishery, for the first 30 years we drew the fraction of fish harvested randomly from a uniform distribution between 0.1 and 0.9.
+Further, to establish a range of spawner-return values and to mimic the start of an open-access fishery, for the first 30 years we drew the fraction of fish harvested randomly from a uniform distribution between 0.1 and 0.9.
 We discarded these initial 30 years as a burn-in period.
@@
 
-\noindent where $e_t$ represents the environmental value in generation $t$ and $d$ represents normally distributed deviations of mean 0 and standard deviation $\sigma_d$ (2).
+where $e_t$ represents the environmental value in generation $t$ and $d$ represents normally distributed deviations of mean 0 and standard deviation $\sigma_d$ (2).
 We represented long-term environmental dynamics as a linear shift in the environmental value through time
@@
 
-\noindent where $\beta_e$ (0.114) represents the slope.
+where $\beta_e$ (0.114) represents the slope.
 To maintain a balanced response, we centered the trend by subtracting the mean $\overline{\beta_e t}$ so that midway through the simulation (after any burn-in period) the environmental value was at the mean environmental tolerance.
@@
 \def\labelenumi{\arabic{enumi}.}
-\itemsep1pt\parskip0pt\parsep0pt
 \item
@@
 
+\subsection{Implications for salmon conservation}
+
 The primary finding from our simulation model is the importance of response diversity in buffering risk and ensuring long-term persistence in the face of environmental uncertainty.
@@
 
-Our simulation model emphasizes the importance of thinking at the portfolio level --- considering aggregate risk and performance for ecological portfolios.
-Although our results may appear intuitive, we often don't often consider aggregate risk and only by simultaneously considering aggregate risk and return can we optimally diversify ecological systems \citep{ando2012}.
-Thus we can think of portfolio risk-return performance as another dimension of conservation.
-This dimension emphasizes the value of a broader scale of management and focussing not just on currently productive populations but also on less productive populations to reduce long-term aggregate risk \citep[e.g.][]{hilborn2003, price2008}.
-This perspective does not negate an understanding at a finer scale, say of individual fish stocks.
-Rather, as in finance, it requires both an understanding of stock performance and an assessment of their covariance to predict properties of the portfolio as a whole.
-
-THE FOLLOWING PARAGRAPH IS STILL BEING PIECED TOGETHER
-
-Our findings support the existence of and inform the implementation of a number of existing guidelines and management structures for PNW salmonids.
+Existing salmon management structures in the US and Canadian PNW support a portfolio conservation perspective.
 In the US, Evolutionarily Significant Units (ESUs) represent salmon populations that are reproductively isolated with an evolutionary heritage \citep{waples1991, moritz1994} and finer-scale Viable Salmonid Populations (VSPs) represent groups of fish that spawn in a given lake or stream in a particular season \citep{mceacheran2000}.
-In Canada, the rough equivalent to the ESU is a Conservation Unit (CU) as defined in the Wild Salmon Policy \citep{dfo2005wsp}.
+In Canada, the rough equivalent to the ESU is a Conservation Unit (CU) as defined by the Wild Salmon Policy \citep{dfo2005wsp}.
 A CU consists of a group of salmon that are reproductively isolated and that if lost would be unlikely to recolonize in a reasonable time frame \citep{dfo2005wsp}.
@@
 
-VSP guidelines \citep{mceacheran2000} suggest\ldots{}
+\emph{To be incorporated?} Some VSP guidelines \citep{mceacheran2000} that are line with our findings\ldots{}
 
@@
   important to incorporate uncertainty of ESU-level processes into assessments
+\item
+  CU guidelines \citep{dfo2005wsp} suggest\ldots{}
 \end{itemize}
 
-CU guidelines \citep{dfo2005wsp} suggest\ldots{}
+Our model complements other simulation-based salmon-habitat prioritization models.
+While these other models tend to focus on detailed assessment of individual fish stocks, our model is unique as the first to consider the role of response diversity in buffering risk for metapopulations as a whole.
+The Shiraz model \citep{scheuerell2006} is one complementary prioritization scheme.
+It focuses on detailed conditioning of the habitat-population-dynamics relationship at multiple life-history stages for a single salmon population.
+Whereas the Shiraz model can be applied to an entire watershed, it combines the populations together as a single unit.
+A second salmon prioritization model proposes combining population viability measures with an assessment of the genetic consequences of losing particular populations \citep{allendorf1997}.
+This model, however, also focuses on the assessment of individual stocks without considering their covariance and therefore the performance of the salmon portfolio as a whole.
+Our model does not replace these prioritization schemes.
+Rather, it proposes an additional focus on prioritization that optimizes metapopulation risk and growth across plausible environmental scenarios while being robust to numerous sources of uncertainty.
 
--
+\subsection{Broad implications, research extensions, and priorities}
 
-Our recommendations support a number of VSP and CU guidelines.
-For example, VSP guidelines suggest\ldots{} and the first objective of CU\ldots{} Therefore, we suggest an ESU or CU could represent a salmon portfolios and the lessons learned from models such as ours are directly applicable to management guidelines in the PNW.
-
-END IN PROGRESS TEXT
-
-Our model is unique in that it is the first to consider risk for metapopulations as a whole while identifying the role of response diversity in buffering risk.
-Our model follows from the work of \citet{cooper1999} who showed the importance of metapopulation habitat structure for salmonid conservation and \citet{moore2010} who linked salmon portfolio performance to asynchrony in population dynamics.
-Our model complements existing simulation-based salmon-habitat prioritization models that emphasize detailed assessment of individual stocks and we compare two of them here.
-One a frequently used simulation-based habitat-prioritization model, the Shiraz model \citep{scheuerell2006}, focuses on detailed conditioning of the habitat-population-dynamics relationship for a single salmon population.
-While the Shiraz model can be applied to an entire watershed it combines the populations together as a single unit.
-In contrast to our model, the Shiraz model breaks the spawner-recruit model into multiple stages.
-A second prioritization scheme proposes using population viability analysis (PVA) or surrogate measures of population viability combined with an assessment of the genetic consequences of losing various populations \citep{allendorf1997}.
-\citet{allendorf1997}, however, also focus on the assessment of individual stocks without considering their covariance and therefore the performance of the salmon portfolio as a whole.
-Compared to these detailed population-specific models, our model is at this stage primarily conceptual, but could be conditioned on real systems or expanded to incorporate greater life-stage realism.
-Our intention was to develop a parsimonious model to explore the salient aspects of metapopulation-portfolio performance.
+Our results emphasize the importance of thinking at the portfolio level when making conservation decisions --- considering aggregate risk and performance for ecological portfolios.
+Although our results may appear intuitive, we often don't often consider aggregate risk and only by simultaneously considering aggregate risk and return can we optimally diversify ecological systems \citep{ando2012}.
+Thus we can think of portfolio risk-return performance as another dimension of conservation.
+This dimension emphasizes the value of a broader scale of management and focussing not just on currently productive populations but also on less productive populations to reduce long-term aggregate risk \citep[e.g.][]{hilborn2003, price2008}.
+This perspective does not negate an understanding at a finer scale, say of individual fish stocks.
+Rather, as in finance, it requires both an understanding of stock performance and an assessment of their covariance to predict properties of the portfolio as a whole.
 
@@
 This rapid adaptation to changes in the environment could strongly affect portfolio performance and emphasizes the importance of maintaining genetic diversity.
-Finally, our model could be conditioned on a system of interest --- a particular river basin for example --- and the management metapopulation portfolio could be optimized across conservation and restoration options as part of a formal decision analysis.
+Finally, our model could be conditioned on a system of interest --- a particular river basin for example --- and the metapopulation portfolio could be optimized across conservation and restoration options as part of a formal decision analysis.
 
@@
 
-In conclusion, our results demonstrate the utility of considering metapopulations in a portfolio context and the power of portfolio simulation to consider robust management strategies that minimize metapopulation risk and maximize growth under uncertain environmental conditions.
+In conclusion, our results demonstrate the utility of considering metapopulations in a portfolio context and the power of portfolio simulation to consider management strategies that minimize metapopulation risk and maximize growth under uncertain environmental conditions.
 Our findings show that response diversity buffers metapopulation risk given short-term environmental stochasticity and ensures persistence given longterm regime changes.
 Further, our findings show that, independent of response-diversity, conserving more populations buffers metapopulation risk.
-These results point to the critical importance of choosing management, conservation, and restoration actions that preserve and promote response diversity and are robust to our uncertainty about response diversity distribution and future climatic conditions.
+These results point to the critical importance of choosing management, conservation, and restoration actions that preserve and promote response diversity and are robust to our uncertainty about the distribution of response diversity and future climatic conditions.
 
@@
 
-\setlength{\parskip}{6pt} \setlength{\parindent}{0cm} \includegraphics[height=5.5in]{../examples/simulation-diagram3.pdf}
+\includegraphics[height=5.5in]{../examples/simulation-diagram3.pdf}
 
@@
 \clearpage
-\includegraphics[width=3.0in]{../examples/thermal-curve-scenarios.pdf}
+
+\includegraphics[width=2.9in]{../examples/thermal-curve-scenarios.pdf}
 
@@
 \clearpage
+
 \includegraphics[width=4.0in]{../examples/spatial-arma-sim.pdf}
@@
 \clearpage
-\includegraphics[width=5.0in]{../examples/spatial-mv.pdf}
+
+\includegraphics[width=4.7in]{../examples/spatial-mv.pdf}
 
@@
 \clearpage
+
 \includegraphics[width=5.0in]{../examples/cons-plans-n.pdf}

### 2013-08-26 inst/ms/salmonfolios.tex
Updates to ms; wrap ms

@@
 \bibpunct{(}{)}{;}{a}{}{;}
-\textheight 20.5cm
+%\textheight 20.5cm
+%\topmargin -1.43cm
+%\oddsidemargin 1.2cm
+%\evensidemargin 1.2cm
+%\textwidth 14.5cm
 \usepackage{lineno}
@@
 \renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
+
+\usepackage{titlesec}
+\titlespacing\section{0pt}{12pt plus 4pt minus 2pt}{-6pt plus 2pt minus 2pt}
+\titlespacing\subsection{0pt}{12pt plus 4pt minus 2pt}{-6pt plus 2pt minus 2pt}
+
+\usepackage{fontspec}
+\defaultfontfeatures{Ligatures=TeX}
+\setsansfont{Calibri}
+\setmonofont{Inconsolata}
+\setmainfont{Minion Pro}
 
@@
 
-\hyphenation{meta-pop-ulation meta-pop-ulations sub-pop-ulations sub-pop-ulation e-con-o-mist}
+\hyphenation{meta-pop-ulation meta-pop-ulations sub-pop-ulations sub-pop-ulation e-con-o-mist en-vi-ron-men-tal pri-or-i-ti-za-tion}
 
@@
 \modulolinenumbers[2]
-\begin{spacing}{1.5}
+\begin{spacing}{1.4}
+\setlength{\parskip}{12pt}
+\setlength{\parindent}{0cm} 
 \include{ms}

### 2013-08-27 inst/ms/ms.tex
Add notes from Nick

@@
   I think there is room for expansion of the discussion (we still have \textasciitilde{}500 words to play with).
-The discussion could benefit from bringing in more concrete examples of how this ties in with PNW salmon.
+The discussion could benefit from bringing in more concrete examples of how this ties in with PNW salmon and I welcome your ideas here.
+If we have space we could also discuss our results in the context of response diversity literature.
 \item
@@
   We still need to discuss authorship and order of authorship.
+\item
+  Notes from Nick: (1) add something about how many of these concepts are key elements of conservation ecology, but they haven't been tested quantitatively before, (2) consider flipping the discussion around so the broad elements come first and the salmon applications come near the end, (3) consider sticking with ``one half/other half'', or ``lower half/upper half''.
+If lower/upper, need to explain reasoning.
 \end{itemize}

### 2013-08-28 inst/ms/ms.tex
Minor tweaks to ms; mostly in abstract

@@
 \item
-  Notes from Nick: (1) add something about how many of these concepts are key elements of conservation ecology, but they haven't been tested quantitatively before, (2) consider flipping the discussion around so the broad elements come first and the salmon applications come near the end, (3) consider sticking with ``one half/other half'', or ``lower half/upper half''.
+  Notes from Nick: (1) add something about how many of these concepts are key assumptions of conservation ecology, but they haven't been tested quantitatively before, (2) consider flipping the discussion around so the broad elements come first and the salmon applications come near the end, (3) consider sticking with ``one half/other half'', or ``lower half/upper half''.
 If lower/upper, need to explain reasoning.
@@
 
-\emph{Words in abstract and main text}: (maximum 150, 5000)
+\emph{Words in abstract and main text}: 153, 4100 (maximum 150, 5000)
 
-\emph{Number of references}: (maximum 50)
+\emph{Number of references}: 61 (maximum 50)
 
-\emph{Number of tables and figures}: 1, 5
+\emph{Number of tables and figures}: 5 (maximum 6)
 
@@
 
-Managing risk is fundamental to the conservation of an endangered species.
+Managing risk is fundamental to conserving endangered species.
 When an endangered species exists as a metapopulation, we typically manage risk at the subpopulation level.
 However, a portfolio approach might consider how conservation affects the ``weight'' of each population in a metapopulation ``portfolio''.
-Here we ask how such a portfolio approach can inform conservation priorities for metapopulations in a changing world.
-We develop a salmon metapopulation simulation in which population-specific productivity is driven by spatially-distributed thermal tolerance and patterns of short- and long-term environmental change.
+Here we ask how a portfolio approach can inform conservation priorities for metapopulations in a changing world.
+We develop a salmon metapopulation simulation where population-specific productivity is driven by spatially-distributed thermal tolerance and patterns of short- and long-term environmental change.
 We then implement spatial conservation strategies that control subpopulation carrying capacities and evaluate the salmon portfolios along risk-return axes.
@@
 
-Managing risk is fundamental to conserving endangered species \citep{burgman2005, iucn2009}.
+Managing risk is fundamental to the conservation of an endangered species \citep{burgman2005, iucn2009}.
 When an endangered species exists as a metapopulation, we can manage risk at two levels: at the population level or at the metapopulation level.

### 2013-09-04 inst/ms/som.tex
Make a variety of edits to ms; finish SOM figs

@@
 
+\subsection{Supplementary tables}
+
+\input{tables}
+
 \subsection{Supplementary figures}
@@
 \centering
-\includegraphics[width=4.0in]{../examples/figure/plot-various-options-ts-3pops.pdf}
-\caption{The impact of increasing or decreasing various parameter values in the simulations.
-The different lines represent different salmon populations.
-(NEED TO ADD PARAMETER VALUES AND EXPAND THIS SLIGHTLY)}
-\label{f:eg-sens}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=4.0in]{../examples/figure/stray-matrix.pdf}
+\includegraphics[width=3.5in]{../examples/figure/stray-matrix.pdf}
 \caption{An example straying matrix. The rows and columns represent different 
@@
 \centering
-\includegraphics[width=4.5in]{../examples/spatial-arma-sim.pdf}
-\caption{Spatial and short-term environmental fluctuations}
+\includegraphics[width=4.0in]{../examples/plot-various-options-ts-3pops.pdf}
+\caption{The impact of increasing or decreasing various parameter values on metapopulation return abundance.
+The different coloured lines represent three example salmon populations.
+The base case represents the base-case values for the short-term environmental fluctuation scenario.}
+\label{f:eg-sens}
+\end{figure}
+
+\clearpage
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=4.5in]{../examples/spatial-arma-sim-full.pdf}
+\caption{Conserving a \textbf{full range} of response diversity (spatial conservation strategy) with \textbf{short-term} environmental fluctuations.}
 \label{f:eg-sp-arma}
@@
 \centering
-\includegraphics[width=4.5in]{../examples/spatial-linear-sim.pdf}
-\caption{Spatial and long-term environmental fluctuations}
+\includegraphics[width=4.5in]{../examples/spatial-arma-sim-onehalf.pdf}
+\caption{Conserving \textbf{one half} of response diversity (spatial conservation strategy) with \textbf{short-term} environmental fluctuations.}
+\label{f:eg-sp-arma}
+\end{figure}
+
+\clearpage
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=4.5in]{../examples/spatial-linear-sim-full.pdf}
+\caption{Conserving a \textbf{full range} of response diversity (spatial conservation strategy) with \textbf{long-term} environmental fluctuations.}
 \label{f:eg-sp-linear}
@@
 \centering
-\includegraphics[width=4.5in]{../examples/n-arma-sim.pdf}
-\caption{Number and short-term environmental fluctuations}
+\includegraphics[width=4.5in]{../examples/spatial-linear-sim-onehalf.pdf}
+\caption{Conserving \textbf{one half} of response diversity (spatial conservation strategy) with \textbf{long-term} environmental fluctuations.}
+\label{f:eg-sp-linear}
+\end{figure}
+
+\clearpage
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=4.5in]{../examples/n-arma-sim-2.pdf}
+\caption{\textbf{Two populations} conserved with random response diversity and \textbf{short-term} environmental fluctuations.}
 \label{f:eg-n-arma}
@@
 \centering
-\includegraphics[width=4.5in]{../examples/n-linear-sim.pdf}
-\caption{Number and long-term environmental change}
+\includegraphics[width=4.5in]{../examples/n-arma-sim-16.pdf}
+\caption{\textbf{Sixteen populations} conserved with random response diversity and \textbf{short-term} environmental fluctuations.}
+\label{f:eg-n-arma}
+\end{figure}
+
+\clearpage
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=4.5in]{../examples/n-linear-sim-2.pdf}
+\caption{\textbf{Two populations} conserved with random response diversity and \textbf{long-term} environmental change.}
 \label{f:eg-n-linear}
@@
 \clearpage
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=4.5in]{../examples/n-linear-sim-16.pdf}
+\caption{\textbf{Sixteen populations} conserved with random response diversity and \textbf{long-term} environmental change.}
+\label{f:eg-n-linear}
+\end{figure}

### 2013-09-04 inst/ms/tables.tex
Make a variety of edits to ms; finish SOM figs

%\begin{table}[h!]
%\centering
%\small
%\caption{Components of salmon metapopulation portfolios KEEP THIS?}
%\begin{tabular}{p{3.6cm}p{7.5cm}}
%\toprule
%Component          & Definition for the salmon portfolio\\
%\midrule
%Assets             & Stream-level salmon populations; possibly a Viable Salmonid Population\\
%Portfolio          & The salmon metapopulation; possibly an Evolutionarily Significant Unit\\
%Portfolio managers & Salmon managers\\
%Investors          & Salmon managers, conservation agency, or salmon fishers\\
%Asset weights      & Carrying capacity (specifically the $b$ parameters in a Ricker model)\\
%Asset returns      & Rate of change of generation-to-generation salmon metapopulation abundance\\
%Asset risk         & Variance of generation-to-generation salmon metapopulation abundance\\
%\bottomrule
%\end{tabular}
%\label{t:port}
%\end{table}
%\clearpage

\begin{table}[h!]
\centering
\footnotesize
\caption{Input parameters to the salmon metapopulation simulation with default values.}
\begin{tabular}{p{7.7cm}p{1.4cm}p{3.6cm}}
\toprule
Description                                                          & Symbol                & Value \\
\midrule

\textit{Population dynamics parameters}                              &                       & \\
Stock-recruit residual standard deviation (on log scale)             & $\sigma_r$            & 0.3  \\
AR(1) serial correlation of stock-recruit residuals                  & $\rho_w$              & 0.4  \\
Fraction of fish that stray from natal streams                       & $f_{\mathrm{stray}}$  & 0.02  \\
Exponential rate of decay of straying with distance                  & $m$                   & 0.3  \\

\noalign{\vskip 3mm}
\textit{Environmental parameters}                                    &                       & \\
Width of environmental-tolerance curves for populations $i$ 1 to $n$ & $W_i$                 & Sequence of 0.05 to 0.02 to 0.05\\
Optimum environmental value for populations $i$ 1 to $n$             & $e_i^{\mathrm{opt}}$  & Sequence of 13 to 19\\
Area under each environmental-tolerance curve in environmental units & $A$                   & 30\\

Standard deviation of short-term environmental fluctuations          & $\sigma_d$            & 2 \\
AR(1) autocorrelation of short-term environmental fluctuations       & $\rho_e$              & 0.1 \\
Slope of long-term environmental signal                              & $\beta_e$             & 0.114 \\

\noalign{\vskip 3mm}
\textit{Fishery parameters}                                          &                       & \\
Standard deviation of beta distribution for implementation error     & $\sigma_{h}$          & 0.05  \\
Frequency of assessment (years)                                      & $f_{\mathrm{assess}}$ & 5  \\
\bottomrule
\end{tabular}
\label{t:pars}
\end{table}
\clearpage

### 2013-09-04 inst/examples/cons_plans2.R

@@ -35,8 +35,14 @@ arma_env_params <- list(mean_value = 16, ar = 0.1, sigma_env = 2, ma = 0)
     #arma_env_params, env_type = "arma", assess_freq = 5),
   #years_to_show = 100, burn = 0)
 
-pdf("spatial-arma-sim.pdf", width = 5, height = 7)
+pdf("spatial-arma-sim-full.pdf", width = 5, height = 7)
 eg_arma <- meta_sim(b = w[[1]][[1]], n_pop = 10, env_params = arma_env_params,
+  env_type = "arma", assess_freq = 5)
+plot_sim_ts(eg_arma, years_to_show = 70, burn = 30)
+dev.off()
+
+pdf("spatial-arma-sim-onehalf.pdf", width = 5, height = 7)
+eg_arma <- meta_sim(b = w[[4]][[1]], n_pop = 10, env_params = arma_env_params,
   env_type = "arma", assess_freq = 5)
 plot_sim_ts(eg_arma, years_to_show = 70, burn = 30)
 dev.off()
@@ -59,8 +65,14 @@ linear_env_params <- list(min_value = 12, max_value = 20, sigma_env = 0.001,
     #linear_env_params, env_type = "linear", assess_freq = 5),
   #years_to_show = 100, burn = 0)
 
-pdf("spatial-linear-sim.pdf", width = 5, height = 7)
+pdf("spatial-linear-sim-full.pdf", width = 5, height = 7)
 eg_linear <- meta_sim(b = w[[1]][[1]], n_pop = 10, env_params =
+    linear_env_params, env_type = "linear", assess_freq = 5)
+plot_sim_ts(eg_linear, years_to_show = 70, burn = 30)
+dev.off()
+
+pdf("spatial-linear-sim-onehalf.pdf", width = 5, height = 7)
+eg_linear <- meta_sim(b = w[[4]][[1]], n_pop = 10, env_params =
     linear_env_params, env_type = "linear", assess_freq = 5)
 plot_sim_ts(eg_linear, years_to_show = 70, burn = 30)
 dev.off()
@@ -163,3 +175,12 @@ par(xpd = NA)
 mtext("Generation", side = 1, line = 2, outer = FALSE, cex = 0.8, adj = -.2)
 par(xpd = FALSE)
 dev.off()
+
+
+## report summary statistics:
[…]

### 2013-09-04 inst/examples/cons_plans_n_random_resp_div.R

@@ -27,12 +27,20 @@ arma_env_params <- list(mean_value = 16, ar = 0.1, sigma_env = 2, ma = 0)
     #arma_env_params, env_type = "arma", assess_freq = 5),
   #years_to_show = 100, burn = 0)
 
-pdf("n-arma-sim.pdf", width = 5, height = 7)
+pdf("n-arma-sim-2.pdf", width = 5, height = 7)
 eg_arma <- meta_sim(b = w[[1]][[2]], n_pop = 16, env_params =
   arma_env_params, env_type = "arma", assess_freq = 5, max_a =
   thermal_integration(16))
 plot_sim_ts(eg_arma, years_to_show = 70, burn = 30)
 dev.off()
+
+pdf("n-arma-sim-16.pdf", width = 5, height = 7)
+eg_arma <- meta_sim(b = w[[4]][[2]], n_pop = 16, env_params =
+  arma_env_params, env_type = "arma", assess_freq = 5, max_a =
+  thermal_integration(16))
+plot_sim_ts(eg_arma, years_to_show = 70, burn = 30)
+dev.off()
+
 
 # TODO need to fix it here to get max_a integrated based on n_pops
 if(!USE_CACHE) {
@@ -53,7 +61,14 @@ linear_env_params <- list(min_value = 12, max_value = 20, sigma_env = 0.001,
     #linear_env_params, env_type = "linear", assess_freq = 5),
   #years_to_show = 100, burn = 0)
 
-pdf("n-linear-sim.pdf", width = 5, height = 7)
+pdf("n-linear-sim-2.pdf", width = 5, height = 7)
+eg_linear <- meta_sim(b = w[[1]][[1]], n_pop = 16, env_params =
+  linear_env_params, env_type = "linear", assess_freq = 5, max_a =
+  thermal_integration(16))
+plot_sim_ts(eg_linear, years_to_show = 70, burn = 30)
+dev.off()
+
+pdf("n-linear-sim-16.pdf", width = 5, height = 7)
 eg_linear <- meta_sim(b = w[[4]][[1]], n_pop = 16, env_params =
   linear_env_params, env_type = "linear", assess_freq = 5, max_a =
   thermal_integration(16))
@@ -172,3 +187,10 @@ mtext("Metapopulation\nabundance", side = 2, line = 3, outer = FALSE, cex = 0.8)
 par(las =1)
 
 dev.off()
+
[…]

### 2013-09-04 inst/examples/example_returns_3.R

@@ -1,41 +1,153 @@
 #' 
 
+#col_pal <- rev(gg_color_hue(3))
+#n_pop_temp <- 3
+#set.seed(2)
+#eg1 <- meta_sim(sigma_v = 0.1, n_pop = n_pop_temp)
+#set.seed(2)
+#eg2 <- meta_sim(env_params = list(amplitude = .000000001, ang_frequency = 0.2, phase = 0, mean_value = 16, slope = 0, sigma_env = 0), sigma_v = 0.1, n_pop = n_pop_temp)
+#set.seed(2)
+#eg3 <- meta_sim(stray_fraction = 0.00001, sigma_v = 0.1, n_pop = n_pop_temp)
+#set.seed(2)
+#eg10 <- meta_sim(env_params = list(amplitude = 5.0001, ang_frequency = 0.2, phase = 0, mean_value = 16, slope = 0, sigma_env = 0), sigma_v = 0.1, n_pop = n_pop_temp)
+#set.seed(2)
+#eg4 <- meta_sim(sigma_v = 0.000001, n_pop = n_pop_temp)
+#set.seed(2)
+#eg5 <- meta_sim(sigma_v = 0.15, n_pop = n_pop_temp)
+#set.seed(2)
+#eg7 <- meta_sim(max_a = rep(0.9, 10), sigma_v = 0.1, n_pop = n_pop_temp)
+#set.seed(2)
+#eg8 <- meta_sim(max_a = rep(3.0, 10), sigma_v = 0.1, n_pop = n_pop_temp)
+#set.seed(2)
+#eg9 <- meta_sim(stray_fraction = 0.1, sigma_v = 0.1, n_pop = n_pop_temp)
+
+#source("plot_sim_ts_simple.R")
+#par(mfrow = c(9, 1), mar = c(0,0,0,0), oma = c(4, 4, 1, 1), cex = 0.7, las = 1, xpd = FALSE)
+#plot_sim_ts_simple(eg1, pal = col_pal, text = "Base case")
+#plot_sim_ts_simple(eg2, pal = col_pal, text =  "No temperature fluctuations")
+#plot_sim_ts_simple(eg10, pal = col_pal, text =  "Larger temperature fluctuations")
+#plot_sim_ts_simple(eg3, pal = col_pal, text = "No straying")
+#plot_sim_ts_simple(eg9, pal = col_pal, text = "More straying")
[…]

### 2013-09-04 inst/examples/plot_sim_ts_simple.R

@@ -1,24 +1,24 @@
 # plot time series from the simulation
 # just the annual rate of growth and returns in this one
 
-plot_sim_ts_simple <- function(x, pal, burn = 1:60, text = "Returns") {
+plot_sim_ts_simple <- function(x, pal = rev(gg_color_hue(x$n_pop)), burn = 1:60, text = "Returns") {
 # x is output from a simulation run; it's a list
 # pal is the colour palette
 
 to_show <- (max(burn)):(max(burn)+40) # years to show in time series example plots
 
 annotate <- function(text) {
-  mtext(text, side = 3, line = -1.5, cex = 0.7, adj = 0.05)
+  text(par("usr")[1], par("usr")[4] - (par("usr")[4] - par("usr")[3]) * 0.18, text, cex = 1, pos = 4)
 }
-my.axis <- function(side) axis(side, col = "grey50", at = c(800, 1200))
+#my.axis <- function(side) axis(side, col = "grey50", at = c(2000, 4000))
 
 #par(mfrow = c(8, 1), mar = c(0,0,0,0), oma = c(4, 4, 1, 1), cex = 0.7, las = 1, xpd = FALSE)
 # returns:
 
-matplot(x$A[to_show, ], type = "l", col = pal, lty = 1, ylab = "Returns", xlab = "Time", xaxt = "n", axes = FALSE, xaxs = "i", ylim = c(300, 2450))
+matplot(x$A[to_show, ], type = "l", col = pal, lty = 1, ylab = "", xlab = "Time", xaxt = "n", axes = FALSE, xaxs = "i", ylim = c(0, 6000))
 annotate(text)
 box(col = "grey50")
-my.axis(2)
+#my.axis(2)
 
 #my.axis(1)
 

### 2013-09-04 vignette/metafolio.Rnw
Add in-progress vignette

%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{metafolio}

% tw=68
\documentclass[12pt]{article}
\usepackage{geometry}
\usepackage{url}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}

\setlength{\parskip}{12pt}
\setlength{\parindent}{0cm}

\begin{document}

<<setup, include=FALSE, cache=FALSE>>=
library(knitr)
# set global chunk options
opts_chunk$set(fig.path='figure/', fig.align='center', fig.show='hold')
options(replace.assign=TRUE, width=90)
opts_knit$set(out.format = "latex")
opts_chunk$set(warning=FALSE, message=FALSE, comment=NA, tidy=FALSE, 
refresh=TRUE)
#knit_theme$set("solarized-light")
@

\title{A guide to using the \texttt{metafolio} \texttt{R} package}
\author{Sean C. Anderson, and co-authors}
\date{}
\maketitle

The \texttt{metafolio} \texttt{R} package is a tool to simulate salmon
metapopulations and apply financial portfolio optimization concepts
to those metapopulations. This vignette accompanies the paper
\textit{Portfolio conservation of metapopulations under climate
change}. In this vignette we will describe the main functions in the
\texttt{metafolio} package, show how to re-create the main analyses
in our paper, and demonstrate some of the included plotting
functions for exploring the output.

\section{Installing the package}

To install and load the package:

<<my-load-package, include=FALSE>>=
library(metafolio)
set.seed(1)
@

<<load-package, eval=FALSE>>=
install.packages("devtools") # if needed
devtools::install_github("metafolio", "seananderson")
library(metafolio)
@

You can get more detailed installation instructions, read the source
code, and report bugs on GitHub:
\url{https://github.com/seananderson/metafolio}

\section{An example simulation}

The main simulation function is \texttt{meta\_sim}. We'll start by
running a simulation using the base-case parameter values from the
paper. First, we'll set up a list object that contains the argument
values for the stationary environmental stochasticity scenario. The
arguments \texttt{mean\_value}, \texttt{ar}, and \texttt{ma} refer
to the mean, autoregressive (AR), and moving-average (MA) components
of an ARIMA model. See \texttt{?arima} for further explanation of
ARIMA models in R. The argument \texttt{sigma\_env} refers to the
standard deviation.

<<setup-arma>>=
arma_env_params <- list(mean_value = 16, ar = 0.1, sigma_env = 2, 
  ma = 0) 
@

Next we'll run the simulation for one iteration. We'll simulate ten
subpopulations and we'll re-assess the fishery every five years.
Re-assessing the fishery means that the simulation fits a Ricker
curve to the observed spawner-return data and updates the
harvest-level targets.

<<base-case-run, cache=TRUE>>=
base1 <- meta_sim(n_pop = 10, env_params = arma_env_params, 
  env_type = "arma", assess_freq = 5)
@

We can plot the output with the function \texttt{plot\_sim\_ts}. The
output is shown in Figure~\ref{fig:plot-base-case-ts}.

<<plot-base-case-ts, fig.width=5, fig.height=6.5, out.width="4in", fig.cap="An example simulation with stationary environmental stochasticity and the base-case parameter values.">>=
plot_sim_ts(base1, years_to_show = 70, burn = 30)
@

\clearpage

\section{Exploring prioritization strategies}

One of the key elements to the analysis in our paper is the choice
of which populations to prioritize for conservation. We can try
different prioritization strategies by manipulating the "investment
weights" in each stream of salmon. In the case of salmon, we
represent these with the Ricker $b_i$ parameters, which represent
the maximum population capacity of streams $i$ 1 through $n$.

In this example, we'll create a scenario in which we conserve a
variety of balanced response diversity and a scenario in which we
conserve one half of the response diversity. We've set up the
weights carefully so that each metapopulation contains the same
number of subpopulations (10) and the same total capacity. We set
$b$ equal to the nominal level of 5 for the streams that we choose
not to prioritize.

<<spatial-plans>>=
w_plans <- list()
w_plans[["balanced"]] <- matrix(c(5, 1000, 5, 1000, 5, 5, 1000, 5, 
    1000, 5), nrow = 1) 
w_plans[["one_half"]] <- c(rep(1000, 4), rep(5, 6))
w <- list()
for(i in 1:2) { # loop over plans
 w[[i]] <- list()
 for(j in 1:80) { # loop over iterations
   w[[i]][[j]] <- matrix(w_plans[[i]], nrow = 1)
 }
}
@

We've now created a nested list object of stream capacities
(\texttt{w}). The first level of the list contains the two different
plans. The second level of the list contains the $b$ values for each
iteration. Each iteration will have re-sampled process noise and
observation error when run with \texttt{meta\_sim}. Here, we're
keeping the $b$ values the same between iterations, but that might
not be the case if we wanted to stochastically simulate the $b$
values. We're only going to run 80 iterations to reduce the runtime
of the example, but in reality you would likely want to run many
more iterations.

We can now stochastically simulate with these strategies using the
function\\ \texttt{run\_cons\_plans} and plot the output with
\texttt{plot\_cons\_plans} (Figure~\ref{fig:spatial-runs}).

<<spatial-runs, fig.width=5, fig.height=5, out.width="4in", cache=TRUE, fig.cap="Two spatial conservation strategies shown in risk-return space with stationary environmental stochasticity. The dots show simulated metapopulations and the contours show 25\\% and 75\\% quantiles across 80 simulations per strategy. The grey line indicates the efficient frontier across all simulated metapopulations.">>=
set.seed(1)
arma_sp <- run_cons_plans(w, env_type = "arma", env_params =
  arma_env_params)
plot_cons_plans(arma_sp$plans_mv, 
  plans_name = c("Balanced", "One half"), 
  cols = c("#E41A1C", "#377EB8"), xlab = "Variance of growth rate", 
  ylab = "Mean growth rate")
@

\section{Generating alternative environmental time series}

We can use the function \texttt{generate\_env\_ts} to create
alternative environmental time series. The function can generate
five kinds of time series: sine waves, regime shifts, linear
changes, and constant values. Each type has its own set of parameter
arguments that are passed in a list format. See
\texttt{?generate\_env\_ts} for examples of all of these.

We can see demonstrations of all environmental time series types
with the default argument values with the following code
(Figure~\ref{fig:env-eg}).

<<env-eg, fig.width=5, fig.height=6, out.width="4in", fig.cap="Example environmental time series.">>=
types <- c("sine", "arma", "regime", "linear", "constant")
x <- list()
for(i in 1:5) x[[i]] <- generate_env_ts(n_t = 100, type = types[i])
op <- par(mfrow = c(5, 1), mar = c(3,3,1,0), cex = 0.7)
for(i in 1:5) plot(x[[i]], type = "o", main = types[i])
par(op)
@

\section{Extra to be incorporated above}

We can visualize the variability in the Ricker $a$ parameters using the function \texttt{plot\_rickers} (Figure~\ref{fig:plot-rickers}).

<<plot-rickers, fig.width=7, fig.height=4.5, out.width="5in", fig.cap="Ricker curves from a sample of 40 years in the example simulation. Each panel represents a different stream population. Population 1 is cool tolerant and population 10 is warm tolerant. The colour of the Ricker curves represents the relative temperatue in that year (warm: red; cool: blue). The grey shaded area represents the variation in spawners observed within the simulation.">>=
pal <- rep("black", 10) # colours for the population IDs
plot_rickers(base1, pal = pal)
@

We can also use \texttt{metafolio} to sample from possible
investment weights (Figure~\ref{fig:monte-carlo-plot}). This type of
stochastic sampling is sometimes referred to as Monte Carlo
sampling.

<<monte-carlo-eg, cache=TRUE, results="hide">>=
set.seed(1)
weights_matrix <- create_asset_weights(n_pop = 5, n_sims = 500, 
  weight_lower_limit = 0.001)
mc_ports <- monte_carlo_portfolios(weights_matrix = weights_matrix,
  n_sims = 500, mean_b = 1000)
@

<<monte-carlo-plot, fig.width=7, fig.height=4, out.width="5in", fig.cap="Efficient frontier of metapopulation portfolios (red dots). Each dot represents a different set of weights of the Ricker $b$ parameters. The colours in the right panel correspond to the five subpopulations with warm tolerant populations in warmer colours and cool tolerant populations in cooler colours.">>=
col_pal <- rev(gg_color_hue(5))
ef_dat <- plot_efficient_portfolios(port_vals = mc_ports$port_vals,
  pal = col_pal, weights_matrix = weights_matrix)
@

We can look at the correlation between salmon returns in the various
streams using the function
\texttt{plot\_correlation\_between\_returns}
(Figure~\ref{fig:return-correlations}).

<<return-correlations, fig.width=5, fig.height=5, out.width="5in", fig.cap="A plot comparing the log(returns) between each population. The subpopulation IDs are coloured from warm tolerant (warm colours) to cool tolerant (cool colours). Note how populations 1 and 10 have asynchronous returns whereas populations with more similar thermal-tolerance curves (say populations 9 and 10) have more synchronous dynamics. Populations with thermal tolerance curves in the middle (e.g. population 6) are less correlated with other populations. Their population dynamics end up primarily driven by demographic stochasticity and less so by temperature-induced systematic changes in productivity.">>=
plot_correlation_between_returns(base1)
@

\end{document}

### 2013-10-01 R/monte_carlo_portfolios.R
Reduce nsims for monte example to be faster

@@ -17,10 +17,10 @@
 #' @seealso \code{\link{meta_sim}}, \code{\link{create_asset_weights}}
 #' @export
 #' @examples
-#' weights_matrix <- create_asset_weights(n_pop = 5, n_sims = 10,
+#' weights_matrix <- create_asset_weights(n_pop = 5, n_sims = 3,
 #'   weight_lower_limit = 0.001)
 #' mc_ports <- monte_carlo_portfolios(weights_matrix = weights_matrix,
-#'   n_sims = 10, mean_b = 1000)
+#'   n_sims = 3, mean_b = 1000)
 
 monte_carlo_portfolios <- function(weights_matrix, n_sims = 500,
   mean_b = 1000, burn = 1:30, ...) {

### 2013-10-26 vignette/metafolio.Rnw
Remove trailing whitespace in vignette
Make many edits to vignette

@@
 
-% tw=68
+% tw=68 for code
 \documentclass[12pt]{article}
@@
 opts_chunk$set(warning=FALSE, message=FALSE, comment=NA, tidy=FALSE,
-refresh=TRUE)
-#knit_theme$set("solarized-light")
+refresh=TRUE, cache=TRUE, autodep=TRUE)
+#knit_theme$set("print")
 @
 
-\title{A guide to using the \texttt{metafolio} \texttt{R} package}
+\title{An introduction to the \texttt{metafolio} \texttt{R} package}
 \author{Sean C. Anderson, and co-authors}
@@
 
-The \texttt{metafolio} \texttt{R} package is a tool to simulate salmon
-metapopulations and apply financial portfolio optimization concepts
-to those metapopulations. This vignette accompanies the paper
-\textit{Portfolio conservation of metapopulations under climate
-change}. In this vignette we will describe the main functions in the
-\texttt{metafolio} package, show how to re-create the main analyses
-in our paper, and demonstrate some of the included plotting
-functions for exploring the output.
+The \texttt{metafolio} \texttt{R} package is a tool to simulate
+metapopulations and apply financial portfolio optimization concepts to those
+metapopulations. The package is primarily geared towards salmon, but could
+be adapted for other taxonomic groups. This vignette accompanies the paper
+\textit{Portfolio conservation of metapopulations under climate change}. In
+this vignette we will describe the main functions in the \texttt{metafolio}
+package, show how to re-create the main analyses in our paper, and demonstrate
+some of the included plotting functions for exploring the output.
 
@@
 
-You can get more detailed installation instructions, read the source
-code, and report bugs on GitHub:
-\url{https://github.com/seananderson/metafolio}
+You can get more detailed installation instructions, read the source code, and
+report bugs on GitHub: \url{https://github.com/seananderson/metafolio}
 
@@
 
-The main simulation function is \texttt{meta\_sim}. We'll start by
-running a simulation using the base-case parameter values from the
-paper. First, we'll set up a list object that contains the argument
-values for the stationary environmental stochasticity scenario. The
-arguments \texttt{mean\_value}, \texttt{ar}, and \texttt{ma} refer
-to the mean, autoregressive (AR), and moving-average (MA) components
-of an ARIMA model. See \texttt{?arima} for further explanation of
-ARIMA models in R. The argument \texttt{sigma\_env} refers to the
-standard deviation.
+The main simulation function is \texttt{meta\_sim}. We'll start by running a
+simulation using the base-case parameter values from the paper. First, we'll
+set up a list object that contains the argument values for the stationary
+environmental stochasticity scenario. The arguments \texttt{mean\_value},
+\texttt{ar}, and \texttt{ma} refer to the mean, autoregressive (AR), and
+moving-average (MA) components of an ARIMA model. See \texttt{?arima} for
+further explanation of ARIMA models in R. The argument \texttt{sigma\_env}
+refers to the standard deviation of the environmental signal.
 
@@
 Next we'll run the simulation for one iteration. We'll simulate ten
-subpopulations and we'll re-assess the fishery every five years.
-Re-assessing the fishery means that the simulation fits a Ricker
-curve to the observed spawner-return data and updates the
-harvest-level targets.
+populations and re-assess the fishery every five years.  Re-assessing the
+fishery means that the simulation fits a Ricker curve to the observed
+spawner-return data and updates the harvest-level targets.
 
@@
 
-We can plot the output with the function \texttt{plot\_sim\_ts}. The
-output is shown in Figure~\ref{fig:plot-base-case-ts}.
+We can plot the output with the function \texttt{plot\_sim\_ts}. The output is
+shown in Figure~\ref{fig:plot-base-case-ts}.
 
@@
 
-One of the key elements to the analysis in our paper is the choice
-of which populations to prioritize for conservation. We can try
-different prioritization strategies by manipulating the "investment
-weights" in each stream of salmon. In the case of salmon, we
-represent these with the Ricker $b_i$ parameters, which represent
-the maximum population capacity of streams $i$ 1 through $n$.
+One of the key elements to the analysis in our paper is the choice of
+which populations to prioritize for conservation. We can try different
+prioritization strategies by manipulating the ``investment weights'' in each
+stream of salmon. In the case of salmon, we represent these with the Ricker
+$b_i$ parameters, which represent the maximum population capacity of streams
+$i$ 1 through $n$.
 
-In this example, we'll create a scenario in which we conserve a
-variety of balanced response diversity and a scenario in which we
-conserve one half of the response diversity. We've set up the
-weights carefully so that each metapopulation contains the same
-number of subpopulations (10) and the same total capacity. We set
-$b$ equal to the nominal level of 5 for the streams that we choose
-not to prioritize.
+In this example, we'll create a scenario in which we conserve response
+diversity from across the spectrum of possible responses and a scenario
+in which we conserve one half of the response diversity. We've set up the
+weights carefully so that each metapopulation contains the same number of
+populations (10) and the same total capacity. We set the Ricker $b_i$ parameters
+equal to the nominal level of five salmon for the streams that we choose not to
+prioritize.
 
@@
 w_plans <- list()
-w_plans[["balanced"]] <- matrix(c(5, 1000, 5, 1000, 5, 5, 1000, 5, 
-    1000, 5), nrow = 1) 
+w_plans[["balanced"]] <- c(5, 1000, 5, 1000, 5, 5, 1000, 5,
+    1000, 5)
 w_plans[["one_half"]] <- c(rep(1000, 4), rep(5, 6))
@@
 
-We've now created a nested list object of stream capacities
-(\texttt{w}). The first level of the list contains the two different
-plans. The second level of the list contains the $b$ values for each
-iteration. Each iteration will have re-sampled process noise and
-observation error when run with \texttt{meta\_sim}. Here, we're
-keeping the $b$ values the same between iterations, but that might
-not be the case if we wanted to stochastically simulate the $b$
-values. We're only going to run 80 iterations to reduce the runtime
-of the example, but in reality you would likely want to run many
-more iterations.
+We've now created a nested list of stream capacities (\texttt{w}). The
+first level of the list contains the two different plans. The second level
+of the list contains the $b$ values for each iteration. Each iteration
+will have re-sampled process noise and observation error when run with
+\texttt{meta\_sim}. Here, we're keeping the $b$ values the same between
+iterations, but that might not be the case if we wanted to stochastically
+simulate the $b$ values. We're only going to run 80 iterations to reduce the
+runtime of the example, but in reality you would likely want to run many more
+iterations.
 
-We can now stochastically simulate with these strategies using the
-function\\ \texttt{run\_cons\_plans} and plot the output with
-\texttt{plot\_cons\_plans} (Figure~\ref{fig:spatial-runs}).
+We can now stochastically simulate with these strategies using the function\\
+\texttt{run\_cons\_plans} and plot the output with \texttt{plot\_cons\_plans}
+(Figure~\ref{fig:spatial-runs}).
 
@@
 The dots show simulated metapopulations and the contours show 25\\% and 75\\% quantiles across 80 simulations per strategy.
-The grey line indicates the efficient frontier across all simulated metapopulations.">>=
+The grey line indicates the efficient frontier across all simulated metapopulations.
+The efficient frontier represents the minimum expected mean growth rate for a given expected variance in growth rate.">>=
 set.seed(1)
@@
 
-We can use the function \texttt{generate\_env\_ts} to create
-alternative environmental time series. The function can generate
-five kinds of time series: sine waves, regime shifts, linear
-changes, and constant values. Each type has its own set of parameter
-arguments that are passed in a list format. See
-\texttt{?generate\_env\_ts} for examples of all of these.
+We can use the function \texttt{generate\_env\_ts} to create alternative
+environmental time series. The function can generate five kinds of time
+series: sine waves, regime shifts, linear changes, and constant values. Each
+type has its own set of parameter arguments that are passed in a list
+format. See \texttt{?generate\_env\_ts} for examples of all of these.
 
-We can see demonstrations of all environmental time series types
-with the default argument values with the following code
-(Figure~\ref{fig:env-eg}).
+We can see demonstrations of all environmental time series types with the
+default argument values with the following code (Figure~\ref{fig:env-eg}).
 
@@
 for(i in 1:5) x[[i]] <- generate_env_ts(n_t = 100, type = types[i])
-op <- par(mfrow = c(5, 1), mar = c(3,3,1,0), cex = 0.7)
+par(mfrow = c(5, 1), mar = c(3,3,1,0), cex = 0.7)
 for(i in 1:5) plot(x[[i]], type = "o", main = types[i])
-par(op)
 @
 
-\section{Extra to be incorporated above}
+\clearpage
 
-We can visualize the variability in the Ricker $a$ parameters using the function \texttt{plot\_rickers} (Figure~\ref{fig:plot-rickers}).
+\section{Included plot types}
+
+We can visualize the variability in the Ricker $a$ parameters using the
+function \texttt{plot\_rickers} (Figure~\ref{fig:plot-rickers}).
 
@@
 Each panel represents a different stream population.
-Population 1 is cool tolerant and population 10 is warm tolerant.
+Population 1 is more productive in cool conditions and population 10 is more producitive in warm conditions.
 The colour of the Ricker curves represents the relative temperatue in that year (warm: red; cool: blue).
 The grey shaded area represents the variation in spawners observed within the simulation.">>=
-pal <- rep("black", 10) # colours for the population IDs
-plot_rickers(base1, pal = pal)
+plot_rickers(base1, pal = rep("black", 10))
 @
 
-We can also use \texttt{metafolio} to sample from possible
-investment weights (Figure~\ref{fig:monte-carlo-plot}). This type of
-stochastic sampling is sometimes referred to as Monte Carlo
-sampling.
+We can look at the correlation between salmon returns in the various
+streams using the function \texttt{plot\_correlation\_between\_returns}
+(Figure~\ref{fig:return-correlations}).
 
-<<monte-carlo-eg, cache=TRUE, results="hide">>=
-set.seed(1)
-weights_matrix <- create_asset_weights(n_pop = 5, n_sims = 500, 
+<<return-correlations, fig.width=5, fig.height=5, out.width="5in", fig.cap="A plot comparing the log(returns) between each population.
+The population IDs are coloured from warm tolerant (warm colours) to cool tolerant (cool colours).
+Note how populations 1 and 10 have asynchronous returns whereas populations with more similar thermal-tolerance curves (say populations 9 and 10) have more synchronous dynamics.
+Populations with thermal tolerance curves in the middle (e.g.~population 6) are less correlated with other populations.
+Their population dynamics end up primarily driven by demographic stochasticity and less so by temperature-induced systematic changes in productivity.">>=
+plot_correlation_between_returns(base1)
+
+@
+
+\section{Optimizing metapopulation portfolios}
+
+A standard procedure in financial portfolio management is to determine optimal
+weights of the assets in a portfolio. The optimal portfolio weights are
+referred to as the efficient frontier. This efficient frontier describes a
+set of portfolios which have minimal risk for a specified level of return or
+maximum return for a specified level of risk. While it would be complicated
+to determine the optimal metapopulation portfolios via algebra we can do
+so by letting \texttt{metafolio} sample from possible investment weights
+(Figure~\ref{fig:monte-carlo-plot}) --- a form of Monte Carlo sampling.
+
+<<monte-carlo-eg, eval=FALSE>>=
+set.seed(999)
+weights_matrix <- create_asset_weights(n_pop = 5, n_sims = 1000,
   weight_lower_limit = 0.001)
 mc_ports <- monte_carlo_portfolios(weights_matrix = weights_matrix,
-  n_sims = 500, mean_b = 1000)
+  n_sims = 1000, mean_b = 1000)
+@
+
+<<monte-carlo-eg-load, eval=TRUE, echo=FALSE>>=
+# To make the vignette compile more quickly:
+weights_matrix <- create_asset_weights(n_pop = 5, n_sims = 1000,
+  weight_lower_limit = 0.001)
+load("port_vals.rda")
+mc_ports <- list()
+mc_ports$port_vals <- port_vals
 @
@@
 Each dot represents a different set of weights of the Ricker $b$ parameters.
-The colours in the right panel correspond to the five subpopulations with warm tolerant populations in warmer colours and cool tolerant populations in cooler colours.">>=
+The colours in the right panel correspond to the five populations with warm tolerant populations in warmer colours and cool tolerant populations in cooler colours.">>=
 col_pal <- rev(gg_color_hue(5))
@@
 
-We can look at the correlation between salmon returns in the various
-streams using the function
-\texttt{plot\_correlation\_between\_returns}
-(Figure~\ref{fig:return-correlations}).
-
-<<return-correlations, fig.width=5, fig.height=5, out.width="5in", fig.cap="A plot comparing the log(returns) between each population.
-The subpopulation IDs are coloured from warm tolerant (warm colours) to cool tolerant (cool colours).
-Note how populations 1 and 10 have asynchronous returns whereas populations with more similar thermal-tolerance curves (say populations 9 and 10) have more synchronous dynamics.
-Populations with thermal tolerance curves in the middle (e.g.
-population 6) are less correlated with other populations.
-Their population dynamics end up primarily driven by demographic stochasticity and less so by temperature-induced systematic changes in productivity.">>=
-plot_correlation_between_returns(base1)
-@
-
 \end{document}

### 2013-10-26 R/plot_correlation_between_returns.R
Remove wrap in xlab argument value

@@ -14,8 +14,8 @@
 #' @export
 
 plot_correlation_between_returns <- function(x, burn = 1:30, pal =
-  rev(gg_color_hue(x$n_pop)), xlab = "log of return abundance by
-  population", ylab = "log of return abundance by population") {
+  rev(gg_color_hue(x$n_pop)), xlab = "log of return abundance by population", 
+  ylab = "log of return abundance by population") {
 
 par(mfrow = c(10, 10), cex = 0.5, mar = c(0,0,0,0), oma = c(4, 4, 0, 0))
 for (i in 1:10) {

### 2013-10-26 vignette/metafolio.Rnw
Minor vignette edits

@@
 
-The main simulation function is \texttt{meta\_sim}. We'll start by running a
-simulation using the base-case parameter values from the paper. First, we'll
-set up a list object that contains the argument values for the stationary
-environmental stochasticity scenario. The arguments \texttt{mean\_value},
-\texttt{ar}, and \texttt{ma} refer to the mean, autoregressive (AR), and
-moving-average (MA) components of an ARIMA model. See \texttt{?arima} for
-further explanation of ARIMA models in R. The argument \texttt{sigma\_env}
-refers to the standard deviation of the environmental signal.
+The main simulation function in \texttt{metafolio} is
+\texttt{meta\_sim}. We'll start by running a simulation using the base-case
+parameter values from the paper. First, we'll set up a list object that
+contains the argument values for the stationary environmental stochasticity
+scenario.
 
@@
 @
+
+The arguments \texttt{mean\_value}, \texttt{ar}, and \texttt{ma} refer to
+the mean, autoregressive (AR), and moving-average (MA) components of an
+ARIMA model. See \texttt{?arima} for further explanation of ARIMA models in
+R. The argument \texttt{sigma\_env} refers to the standard deviation of the
+environmental signal.
 
@@
 stream of salmon. In the case of salmon, we represent these with the Ricker
-$b_i$ parameters, which represent the maximum population capacity of streams
+$b_i$ parameters, which indicate the maximum population capacity of streams
 $i$ 1 through $n$.

### 2013-11-20 R/plot_sim_ts.R
Fix bug in plot_sim_ts with yticks

@@ -61,7 +61,7 @@ plot_panel_lines <- function(dat, ymin = c("zero", "min"), ystretch = 1.1, ...) 
 
 plot_sim_ts <- function(x, pal = rev(gg_color_hue(x$n_pop)),
   years_to_show = 30, burn = 1:50, shade_years = NULL, adj = 0.02,
-  add_units = FALSE, yticks = NA) {
+  add_units = FALSE, yticks = rep(list(NA), x$n_pop)) {
 
   if(!add_units) {
     ylabs <- rep("", 10)

### 2013-11-27 R/create_asset_weights.R
Wrap roxygen comments at 80 with indent

@@ -3,11 +3,11 @@
 #' Create an asset weight matrix to run through the Monte Carlo
 #' algorithm and test possible portfolios.
 #'
-#' @param  n_pop The number of subpopulations.
-#' @param  n_sims The number of simulations.
-#' @param  weight_lower_limit The lowest fraction allowed for a
-#' subpopulation weight. For example, a value of 0.02 means a
-#' subpopulation will at least be assigned 2\% of the total capacity
+#' @param n_pop The number of subpopulations.
+#' @param n_sims The number of simulations.
+#' @param weight_lower_limit The lowest fraction allowed for a subpopulation
+#'   weight. For example, a value of 0.02 means a subpopulation will at least
+#'   be assigned 2\% of the total capacity
 #' @examples
 #' create_asset_weights(n_pop = 5, n_sims = 10, weight_lower_limit = 0.001)
 #' @export

### 2013-11-27 R/generate_env_ts.R

@@ -7,10 +7,10 @@
 #' @param sine_params Parameters controlling sine wave time series.
 #' @param arma_params Parameters controlling ARMA time series.
 #' @param regime_params Parameters controlling regime-shift time series.
-#' @param linear_params Parameters controlling warming or cooling time
-#' series. Minimum environmental value, maximum environmental value,
-#' environmental standard deviation, and the year to start the linear trend
-#' (useful if you're going to throw out the early years as burn in).
+#' @param linear_params Parameters controlling warming or cooling time series.
+#'   Minimum environmental value, maximum environmental value, environmental
+#'   standard deviation, and the year to start the linear trend (useful if
+#'   you're going to throw out the early years as burn in).
 #' @param constant_params Parameter controlling constant time series.
 #' @export
 #' @examples

### 2013-11-27 R/generate_straying_matrix.R

@@ -4,10 +4,10 @@
 #' Based on Eq. 2 in Cooper and Mangel (1999).
 #'
 #' @param n_pop Number of subpopulations.
-#' @param stray_fraction Fraction of individuals that stray from a
-#' given subpopulation.
-#' @param stray_decay_rate Exponential rate that straying decays with
-#' distance between subpopulations.
+#' @param stray_fraction Fraction of individuals that stray from a given
+#'   subpopulation.
+#' @param stray_decay_rate Exponential rate that straying decays with distance
+#'   between subpopulations.
 #' @export
 #' @references
 #' Cooper, A.B. and Mangel, M. 1999. The dangers of ignoring

### 2013-11-27 R/get_conserv_plans_mv.R

@@ -3,15 +3,15 @@
 #' Run the metapopulation simulation for various conservation
 #' prioritization schemes.
 #'
-#' @return
-#' Returns the portfolio mean and variance values and the simulation
-#' runs.
+#' @return Returns the portfolio mean and variance values and the simulation
+#'   runs.
 #'
-#' @param weights A matrix of habitat weights. Each row corresponds to
-#' another scenario. Each column is a different habitat location.
+
+#' @param weights A matrix of habitat weights. Each row corresponds to another
+#'   scenario. Each column is a different habitat location.
 #' @param reps Number of portfolios to simulate.
-#' @param assess_freq The frequency (in generations) of
-#' spawner-recruit re-assessment. Passed to \code{\link{meta_sim}}.
+#' @param assess_freq The frequency (in generations) of spawner-recruit
+#'   re-assessment. Passed to \code{\link{meta_sim}}.
 #' @param burn Cycles to throw out as burn in.
 #' @param ... Other values to pass to \code{\link{meta_sim}}.
 

### 2013-11-27 R/is_quasi_ext.R

@@ -2,8 +2,8 @@
 #'
 #' @param x A numeric vector.
 #' @param thresh The quasi-extinction threshold in absolute numbers.
-#' @param duration The number of years below the threshold before a
-#' quasi extinction has occurred.
+#' @param duration The number of years below the threshold before a quasi
+#'   extinction has occurred.
 #' @export
 #' @examples
 #' x <- seq(100, 0, length.out = 20)

### 2013-11-27 R/meta_sim.R

@@ -3,33 +3,34 @@
 #' @param n_t The number of years.
 #' @param n_pop Number of (sub)populations
 #' @param stray_decay_rate Rate that straying (exponentially) decays with
-#' distance.
+#'   distance.
 #' @param stray_fraction Fraction of fish that stray from natal streams.
 #' @param b Ricker density-dependent parameter. Should be one per
-#' (sub)population.
+#'   (sub)population.
 #' @param spawners_0 A vector of spawner abundances at the start of the
-#' simulation. Length of the vector should equal the number of subpopulations.
+#'   simulation. Length of the vector should equal the number of
+#'   subpopulations.
 #' @param sigma_v Stock-recruit residual standard deviation. Will be
-#' exponentiated.
+#'   exponentiated.
 #' @param v_rho AR1 serial correlation of stock-recruit residuals.
 #' @param a_width_param Width of the thermal curves by (sub)population.
 #' @param optim_temp Optimal temperatures by (sub)population.
 #' @param max_a Maximum Ricker productivity parameters (a) by (sub)population.
-#' The value obtained at the optimum temperature.
+#'   The value obtained at the optimum temperature.
 #' @param env_type The type of environmental time series to generate.  One of
-#' "sine", "arma", "regime", "linear", or "constant".
+#'   "sine", "arma", "regime", "linear", or "constant".
 #' @param env_params Parameters to pass on to \code{generate_env_ts}. You must
-#' provide the appropriate list given your chosen type of environmental signal.
-#' @param start_assessment Generation to start estimating SR
-#' relationship for escapement targets
-#' @param assessment_window Number of generations to use when fitting
-#' SR relationship for escapement targets; must be bigger than
-#' start_assessment
+#'   provide the appropriate list given your chosen type of environmental
+#'   signal.
+#' @param start_assessment Generation to start estimating SR relationship for
+#'   escapement targets
+#' @param assessment_window Number of generations to use when fitting SR
+#'   relationship for escapement targets; must be bigger than start_assessment
 #' @param sigma_impl implementation sd for beta distribution
 #' @param assess_freq How many generations before re-assessing a and b
-#' parameters.
+#'   parameters.
 #' @param use_cache Use the stochastically generated values (SR residuals and
-#' possibly environmental time series) from the previous run?
+#'   possibly environmental time series) from the previous run?
 #' @param add_straying Implement straying between (sub)populations?
 #' @param add_impl_error Add implementation error?
 #' @export

### 2013-11-27 R/monte_carlo_portfolios.R

@@ -3,17 +3,16 @@
 #' Monte Carlo the asset weights into portfolios and record the
 #' simulation output and portfolio metrics (mean and variance).
 #'
-#' @param weights_matrix A matrix of asset weights. The columns
-#' correspond to the different assets and the rows correspond to the
-#' simulation iterations.
+#' @param weights_matrix A matrix of asset weights. The columns correspond to
+#'   the different assets and the rows correspond to the simulation iterations.
 #' @param n_sims The number of simulations to run.
 #' @param mean_b The mean Ricker capacity value.
 #' @param burn The number of years to discard as burn in.
-#' @return A list object with three elements: \code{port_vals} (a
-#' matrix with a column of mean rate of change and variance of rate of
-#' change), \code{n_sims} (the number of simulations ran), and
-#' \code{sims_out} (a list in which each element corresponds to the
-#' output from the run of \code{\link{meta_sim}}.
+#' @return A list object with three elements: \code{port_vals} (a matrix with a
+#'   column of mean rate of change and variance of rate of change),
+#'   \code{n_sims} (the number of simulations ran), and \code{sims_out} (a list
+#'   in which each element corresponds to the output from the run of
+#'   \code{\link{meta_sim}}.
 #' @seealso \code{\link{meta_sim}}, \code{\link{create_asset_weights}}
 #' @export
 #' @examples

### 2013-11-27 R/plot_cons_plans.R

@@ -12,7 +12,7 @@ get_quantile_contour <- function(x, alpha = 0.8) {
 add_dens_polygon <- function(x, y, col, lwd = 1.7, alpha = 0.75, add_pts = FALSE) {
   k <- get_quantile_contour(MASS::kde2d(x,y), alpha = 0.75)
   polygon(k$x, k$y, border = col, col = paste(col, "30", sep = ""), lwd = lwd)
-  k <- get_quantile_contour(MASS::kde2d(x,y), alpha = 0.25)
+  k <- get_quantile_contour(MASS::kde3d(x,y), alpha = 0.25)
   polygon(k$x, k$y, border = col, col = paste(col, "70", sep = ""), lwd = lwd)
   if(add_pts) points(x, y, pch = 21, col = paste(col, "30", sep = ""), cex = 0.7)
 }

### 2013-11-27 R/plot_efficient_portfolios.R

@@ -1,16 +1,15 @@
 #' Basic plot of efficient portfolio and asset contributions
 #'
-#' This function creates a mean-variance plot of the portfolios across
-#' possible asset weights, colour the efficient frontier, and show the
-#' contribution of the different stocks/assets. It also (invisibly)
-#' returns the values that make up the plot so you can create your own
-#' custom plots with the data. See the Returns section for more
-#' details.
-#' @param port_vals A matrix of means and variances (down the two
-#' columns). This likely comes from the output of
-#' \code{\link{monte_carlo_portfolios}}.
+#' This function creates a mean-variance plot of the portfolios across possible
+#' asset weights, colour the efficient frontier, and show the contribution of
+#' the different stocks/assets. It also (invisibly) returns the values that make
+#' up the plot so you can create your own custom plots with the data. See the
+#' Returns section for more details.
+#'
+#' @param port_vals A matrix of means and variances (down the two columns).
+#'   This likely comes from the output of \code{\link{monte_carlo_portfolios}}.
 #' @param weights_matrix The same weight matrix that was passed to
-#' \code{\link{monte_carlo_portfolios}}.
+#'   \code{\link{monte_carlo_portfolios}}.
 #' @param pal Colour palette for the stocks/assets in the barplot.
 #' @param plot Logical: should the plots be made?
 #' @param ylab_dots Y axis label for the mean-variance scatterplot.
@@ -18,18 +17,17 @@
 #' @param ylab_bars Y axis label for the barplot.
 #' @param xlab_bars X axis label for the barplot.
 #' @param port_cols Colours for the dots. A vector of colours for the
-#' non-efficient and efficient portfolios.
+#'   non-efficient and efficient portfolios.
 #' @param pch Dot type
 #' @param ... Anything else to pass to both \code{plot.default} and
-#' \code{barplot}.
+#'   \code{barplot}.
 #' @return
-#' A two panel plot and an (invisible) list of values calculated
-#' within the function. This list contains \code{pv} (mean, variance,
-#' and whether it was part of the efficient frontier);
-#' \code{ef_port_ids} (the portfolio IDs [run numbers] that are part
-#' of the efficient frontier; \code{min_var_port_id} (the portfolio ID
-#' for the minimum-variance portfolio); \code{ef_weights} (the weights
-#' of the portfolios on the efficient frontier).
+#' A two panel plot and an (invisible) list of values calculated within the
[…]

### 2013-11-27 R/plot_rickers.R

@@ -1,21 +1,20 @@
 #' Plot sample Ricker curves for each stock
 #'
-#' Make a plot of Ricker curves for each stock. Can be useful for
-#' visualizing how the simulation parameters are impacting the Ricker
-#' curves and how these vary with temperature across stocks. The
-#' colour of the lines corresponds to the relative thermal tolerance
-#' of that stock. The shaded region shows the range of spawners
-#' observed throughout the simulations.
+#' Make a plot of Ricker curves for each stock. Can be useful for visualizing
+#' how the simulation parameters are impacting the Ricker curves and how these
+#' vary with temperature across stocks. The colour of the lines corresponds to
+#' the relative thermal tolerance of that stock. The shaded region shows the
+#' range of spawners observed throughout the simulations.
 #'
 #' @param x Output list from \code{\link{meta_sim}}.
 #' @param pal Colours for stocks.
 #' @param n_samples Number of sample lines to draw from the \code{a}
-#' parameters.
+#'   parameters.
 #' @param add_y_axes_pops Panels to add y axes on.
 #' @param add_x_axes_pops Panels to add x axes on.
 #' @param burn Number of initial years to throw out as burn in.
-#' @param add_shading Logical: add the light grey shading for the
-#' range of observed spawner abundance?
+#' @param add_shading Logical: add the light grey shading for the range of
+#'   observed spawner abundance?
 #' @param ... Anything else to pass to \code{plot}
 #' @export
 #' @examples

### 2013-11-27 R/ricker.R

@@ -1,8 +1,7 @@
 #' A simple Ricker model
 #'
 #' @param spawners Spawner abundance
-#' @param a Ricker productivity parameter. Recruits are e^a at the
-#' origin.
+#' @param a Ricker productivity parameter. Recruits are e^a at the origin.
 #' @param b Ricker density dependent parameter.
 #' @return Returns the number of recruits.
 #' @export

### 2013-11-27 R/ricker_escapement.R

@@ -5,9 +5,8 @@
 #'
 #' @param a Ricker productivity parameter.
 #' @param b Ricker density-dependent parameter.
-#' @references Hilborn, R.W. and Walters, C. 1992. Quantitative
-#' fisheries stock assessment: Choice, dynamics, and uncertainty.
-#' Chapman and Hall, London.
+#' @references Hilborn, R.W. and Walters, C. 1992. Quantitative fisheries stock
+#'   assessment: Choice, dynamics, and uncertainty.  Chapman and Hall, London.
 #' @export
 #' @examples
 #' ricker_escapement(1.1, 1000)

### 2013-11-27 R/ricker_v_t.R

@@ -3,12 +3,12 @@
 #' @param spawners Spawner abundance
 #' @param a Ricker productivity parameter. Recruits are e^a at the origin.
 #' @param b Ricker density dependent parameter.
-#' @param d Depensation parameter. A value of 1 means no depensation.
-#' Larger values indicate depensation.
-#' @param v_t Residual on the curve. Will be exponentiated. Note that
-#' we are *not* bias correcting within this function (subtracting half
-#' the variance squared) and so the deviations will not be mean
-#' unbiased unless they were bias corrected previously.
+#' @param d Depensation parameter. A value of 1 means no depensation. Larger
+#'   values indicate depensation.
+#' @param v_t Residual on the curve. Will be exponentiated. Note that we are
+#'   *not* bias correcting within this function (subtracting half the variance
+#'   squared) and so the deviations will not be mean unbiased unless they were
+#'   bias corrected previously.
 #' @return Returns a vector of recruits.
 #' @export
 #' @examples

### 2013-11-27 R/run_cons_plans.R

@@ -1,37 +1,33 @@
-#' Run conservation plans and return the portfolio mean and variance
-#' values
+#' Run conservation plans and return the portfolio mean and variance values
 #'
-#' This function takes a set of weights representing different
-#' conservation plans and gets the mean and variance in portfolio
-#' space. This function allows a maximally complicated set of weights
-#' to accommodate all possible scenarios. It can accommodate different
-#' spatial strategies of conservation, conserving different numbers of
-#' populations, and a lack of knowledge. You can do this by how you
-#' set your \code{w} weight object. See the example.
+#' This function takes a set of weights representing different conservation
+#' plans and gets the mean and variance in portfolio space. This function allows
+#' a maximally complicated set of weights to accommodate all possible scenarios.
+#' It can accommodate different spatial strategies of conservation, conserving
+#' different numbers of populations, and a lack of knowledge. You can do this by
+#' how you set your \code{w} weight object. See the example.
 #'
 #' @param w A (nested) list of weights. The first list level contains the
-#' different plans. The next level contains repetitions for a given
-#' plan. E.g. \code{cp[[2]][[1]]} contains the first iteration of the
-#' second conservation plan. Each end element should be a matrix of
-#' weights with one row and the number of columns equal to the number
-#' of subpopulations.
+#'   different plans. The next level contains repetitions for a given plan.
+#'   E.g. \code{cp[[2]][[1]]} contains the first iteration of the second
+#'   conservation plan. Each end element should be a matrix of weights with one
+#'   row and the number of columns equal to the number of subpopulations.
 #' @param env_type The environmental type to pass to
-#' \code{\link{generate_env_ts}}
+#'   \code{\link{generate_env_ts}}
 #' @param env_params The environmental parameters to pass to
-#' \code{\link{generate_env_ts}}
+#'   \code{\link{generate_env_ts}}
 #' @param show_progress Logical: show an indication of progress?
 #' @param burn Cycles to throw out as burn in
-#' @param assess_freq How frequently (in years) to re-assess the
-#' Ricker a and b values.
+#' @param assess_freq How frequently (in years) to re-assess the Ricker a and b
+#'   values.
 #' @param ... Other values to pass to \code{\link{meta_sim}}
 #' @export
-#' @return A list with two high-level elements: the mean variance
-#' output (\code{plans_mv}) and the raw simulation output
-#' (\code{plans_port}). Within \code{plans_mv}, each element of the
[…]

### 2013-11-27 R/thermal_curve_a.R

@@ -1,18 +1,17 @@
 #' Create thermal tolerance curves.
 #'
-#' Creates a quadratic thermal tolerance curve of the form:
-#' width_param * (temp - optim_temp)^2 + max_a
-#' Negative values are *not* returned as 0 for speed of computation.
-#' You should check for this after.
+#' Creates a quadratic thermal tolerance curve of the form: width_param * (temp
+#' - optim_temp)^2 + max_a Negative values are *not* returned as 0 for speed of
+#' computation. You should check for this after.
 #'
 #' @param temp The input temperature value.
 #' @param optim_temp The optimal temperature.
-#' @param max_a The maximum productivity parameter `a` from a Ricker
-#' model (or whatever the y-axis value is you want to return).
-#' @param width_param A parameter to control the width of the
-#' parabola. Smaller numbers make wider parabolas.
-#' @return A productivity parameter given the location on a thermal
-#' tolerance curve.
+#' @param max_a The maximum productivity parameter `a` from a Ricker model (or
+#'   whatever the y-axis value is you want to return).
+#' @param width_param A parameter to control the width of the parabola. Smaller
+#'   numbers make wider parabolas.
+#' @return A productivity parameter given the location on a thermal tolerance
+#'   curve.
 #' @export
 #' @examples
 #' x <- seq(5, 30, length.out = 200)

### 2013-11-27 R/thermal_integration.R

@@ -25,14 +25,14 @@ optim_thermal <- function(optim_temp, width_param, desired_area) {
 
 #' Integrate thermal tolerance curves to get maximum Ricker a values
 #'
-#' Get maximum Ricker a values for a given number of populations.
-#' Useful for assembling multiple thermal tolerance curves in which
-#' each has the same total area under it.
+#' Get maximum Ricker a values for a given number of populations. Useful for
+#' assembling multiple thermal tolerance curves in which each has the same total
+#' area under it.
 #'
 #' @param n_pop The number of populations.
 #' @param width_params Desired widths of the thermal tolerance curves.
-#' @param optim_temps Temperature value at which to reach the peak of
-#' each thermal tolerance curve.
+#' @param optim_temps Temperature value at which to reach the peak of each
+#'   thermal tolerance curve.
 #' @param desired_area Desired area under each curve.
 #' @export
 #' @examples
@@ -64,4 +64,3 @@ thermal_integration <- function(n_pop, width_params = c(seq(0.05,
   }
   mapply(optim_thermal, optim_temps, width_params, desired_area)
 }
-

### 2013-11-27 inst/ms/salmonfolios-letter.tex
Add early draft of letter

\documentclass[letterpaper,12pt]{letter}
\usepackage{geometry}
\geometry{letterpaper}
\usepackage{graphicx}
\usepackage{fontspec}
\usepackage{pdfsync}
\defaultfontfeatures{Ligatures=TeX}
\setsansfont{Calibri}
\setmonofont{Inconsolata}
\setmainfont{Linux Libertine}
% To reduce the vertical space at the top:
\topmargin -3.3cm
\textheight 28cm
%\oddsidemargin 0.3cm
%\evensidemargin 0.3cm
%\textwidth 16.1cm

\address{Sean C. Anderson\\Biological Sciences\\Simon Fraser University\\Burnaby, BC, V5A 1S6\\Canada\\sean@seananderson.ca}

\begin{document}
\begin{letter}{}
\pagestyle{empty}
\opening{Dear D. Holyoak} % eg Hello.

Please find enclosed a manuscript entitled ``Portfolio conservation of metapopulations under climate change'' that we wish to submit for possible publication as a Letter in Ecology Letters.

XX

We confirm that XX. All authors have approved the manuscript and agree with
its submission to Ecology Letters.

This work is timely given the large and expanding interest in XX, the unusual super-abundance of American lobster, and the largely unknown mechanisms driving XX. The results will be of broad interest to Ecology Letters readers due to XX.

This work is entirely original; none of it has been previously published and is not under consideration elsewhere. This manuscript has been approved by all authors. No other recent work of ours deals with the topic presented in this paper. MEE Paper? Ecol Lett paper Yeakel?

We would like to suggest the following five independent referees, listed alphabetically:

Thank you for considering our manuscript for publication and we look forward to hearing from you.

\vspace{14mm}
\closing{Sincerely,\\
\fromname{Sean Anderson, Jonathan Moore, Michelle McClure, Nicholas Dulvy, Andrew Cooper}
}

\end{letter}
\end{document}

### 2013-11-27 R/metafolio.R
Add early draft of metafolio doc

@@ -1,0 +1,41 @@
+#' metafolio: An \R package to simulate salmon metapopulation for portfolio
+#' optimization
+#'
+#' The metafolio \R package is a tool to simulate metapopulations and apply
+#' financial portfolio optimization concepts to those metapopulations. The
+#' package is primarily intended for salmon simulations, but could be adapted
+#' for other taxonomic groups.
+#'
+#' The main functions are:
+#'
+#' @section Running a simulation once:
+#'
+#' \code{\link{meta_sim}}
+#' \code{\link{plot_sim_ts}}
+#'
+#' @section Assessing multiple conservation scenarios:
+#'
+#' \code{\link{run_cons_plans}}
+#' \code{\link{plot_cons_plans}}
+#'
+#' @section Specifying environmental patterns:
+#'
+#' \code{\link{generate_env_ts}}
+#'
+#' @section Diagnostic plots:
+#'
+#' \code{\link{plot_rickers}}
+#' \code{\link{plot_correlation_between_returns}}
+#'
+#' @section Portfolio optimization:
+#'
+#' \code{\link{create_asset_weights}}
+#' \code{\link{monte_carlo_portfolios}}
+#' \code{\link{plot_efficient_portfolios}}
+#'
+#' See the package vignette \code{vignette("metafolio")} for more extensive
+#' explanation of how to use the \code{metafolio} \R package.
+#'
+#' @docType package
+#' @name metafolio
+NULL

### 2013-12-03 R/meta_sim.R
Add warnings in meta_sim

@@ -177,12 +177,24 @@ meta_sim <- function(
           spawners <- E[2:(i - 1), j]
           rick <- fit_ricker(R = recruits, S = spawners)
           # bounds for sanity:
-          if (rick$a > 4) rick$a <- 4
-          if (rick$a < 0.02) rick$a <- 0.02
+          if (rick$a > 4) {
+            warning("Ricker a was estimated at greater than 4. Setting to 4.")
+            rick$a <- 4
+          }
+          if (rick$a < 0.02) {
+            warning("Ricker a was estimated at less than 0.02. Setting to 0.02.")
+            rick$a <- 0.02
+          }
           # at most increase b by 50%
-          if (rick$b > Est_b[i - 1, j] * 1.5) rick$b <- Est_b[i - 1, j] * 1.5
+          if (rick$b > Est_b[i - 1, j] * 1.5) {
+            warning("Jump in Ricker b was too large. Setting to 150% of previous value.")
+            rick$b <- Est_b[i - 1, j] * 1.5
+          }
           # at most decrease b by 50%
-          if (rick$b < Est_b[i - 1, j] * 0.5) rick$b <- Est_b[i - 1, j] * 0.5
+          if (rick$b < Est_b[i - 1, j] * 0.5) {
+            warning("Jump in Ricker b was too large. Setting to 50% of previous value.")
+            rick$b <- Est_b[i - 1, j] * 0.5
+          }
           Est_a[i,j]<-rick$a
           Est_b[i,j]<-rick$b
         }

### 2013-12-12 inst/ms/salmonfolios.tex
Make a variety of edits to ms; finish SOM figs
Switch to Computer Modern
Switch to non computer modern fonts
Switch to Times math
Switch to Computer Modern and size 11
Toggle spacing in SOM
Move Fig and Table S code to som.md
Tweaks to ms: remove refs, tweak abstract, formatting

@@
-\documentclass[12pt]{article}
+\documentclass[11pt]{article}
 \usepackage{geometry}
@@
 \bibpunct{(}{)}{;}{a}{}{;}
-%\textheight 20.5cm
+\textheight 20.5cm
 %\topmargin -1.43cm
@@
 \renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
+\usepackage{booktabs}
+
+\widowpenalty=1000
+\clubpenalty=1000
 
 \usepackage{titlesec}
-\titlespacing\section{0pt}{12pt plus 4pt minus 2pt}{-6pt plus 2pt minus 2pt}
-\titlespacing\subsection{0pt}{12pt plus 4pt minus 2pt}{-6pt plus 2pt minus 2pt}
+\titlespacing\section{0pt}{6pt plus 4pt minus 2pt}{-8pt plus 2pt minus 2pt}
+\titlespacing\subsection{0pt}{6pt plus 4pt minus 2pt}{-8pt plus 2pt minus 2pt}
 
-\usepackage{fontspec}
-\defaultfontfeatures{Ligatures=TeX}
-\setsansfont{Calibri}
-\setmonofont{Inconsolata}
-\setmainfont{Minion Pro}
+%\usepackage{fontspec}
+%\defaultfontfeatures{Ligatures=TeX}
+%\setsansfont{Calibri}
+%\setmonofont{Inconsolata}
+%\usepackage{mathptmx}
+%\setmainfont{Minion Pro}
 
@@
 \begin{document}
+\raggedright
 
@@
 \modulolinenumbers[2]
+\begin{spacing}{1.6}
+\setlength{\parindent}{0cm}
+\input{ms}
+
+\bibliographystyle{ecologyletters2}
+%\bibliography{jshort,ms}
+
+\begin{thebibliography}{50}
+\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
+\expandafter\ifx\csname url\endcsname\relax
+  \def\url#1{\texttt{#1}}\fi
+\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
+
+\bibitem[{Allendorf \emph{et~al.}(1997)Allendorf, Bayles, Bottom, Currens,
+  Frissell, Hankin, Lichatowich, Nehlsen, Trotter \& Williams}]{allendorf1997}
+1.\\
+Allendorf, F.~W., Bayles, D., Bottom, D.~L., Currens, K.~P., Frissell, C.~A.,
+  Hankin, D., Lichatowich, J.~A., Nehlsen, W., Trotter, P.~C. \& Williams,
+  T.~H. (1997).
+\newblock Prioritizing {Pacific} salmon stocks for conservation.
+\newblock \emph{Conserv. Biol.}, 11, 140--152.
+
+\bibitem[{Anderson \emph{et~al.}(2013)Anderson, Cooper \& Dulvy}]{anderson2013}
+2.\\
+Anderson, S.~C., Cooper, A.~B. \& Dulvy, N.~K. (2013).
+\newblock Ecological prophets: Quantifying metapopulation portfolio effects.
+\newblock \emph{Methods Ecol. Evol.}, 4, 971--981.
+
+\bibitem[{Ando \& Mallory(2012)}]{ando2012}
+3.\\
+Ando, A.~W. \& Mallory, M.~L. (2012).
+\newblock Optimal portfolio design to reduce climate-related conservation
+  uncertainty in the {Prairie Pothole Region}.
+\newblock \emph{Proc. Natl. Acad. Sci. U.S.A.}, 109, 6484--6489.
+
+\bibitem[{Balvanera \emph{et~al.}(2006)Balvanera, Pfisterer, Buchmann, He,
+  Nakashizuka, Raffaelli \& Schmid}]{balvanera2006}
+4.\\
+Balvanera, P., Pfisterer, A.~B., Buchmann, N., He, J.-S., Nakashizuka, T.,
+  Raffaelli, D. \& Schmid, B. (2006).
+\newblock Quantifying the evidence for biodiversity effects on ecosystem
+  functioning and services.
+\newblock \emph{Ecol. Lett.}, 9, 1146--1156.
+
+\bibitem[{Battin \emph{et~al.}(2007)Battin, Wiley, Ruckelshaus, Palmer, Korb,
+  Bartz \& Imaki}]{battin2007}
+5.\\
+Battin, J., Wiley, M.~W., Ruckelshaus, M.~H., Palmer, R.~N., Korb, E., Bartz,
+  K.~K. \& Imaki, H. (2007).
+\newblock Projected impacts of climate change on salmon habitat restoration.
+\newblock \emph{Proc. Natl. Acad. Sci. U.S.A.}, 104, 6720--6725.
+
+\bibitem[{Beechie \emph{et~al.}(2013)Beechie, Imaki, Greene, Wade, Wu, Pess,
+  Roni, Kimball, Stanford, Kiffney \& Mantua}]{beechie2013}
+6.\\
+Beechie, T., Imaki, H., Greene, J., Wade, A., Wu, H., Pess, G., Roni, P.,
+  Kimball, J., Stanford, J., Kiffney, P. \& Mantua, N. (2013).
+\newblock Restoring salmon habitat for a changing climate.
+\newblock \emph{River Res. Applic.}, 29, 939--960.
+
+\bibitem[{Cooper \& Mangel(1999)}]{cooper1999}
+7.\\
+Cooper, A.~B. \& Mangel, M. (1999).
+\newblock The dangers of ignoring metapopulation structure for the conservation
+  of salmonids.
+\newblock \emph{Fish. Bull.}, 97, 213--226.
+
+\bibitem[{Crozier \emph{et~al.}(2008)Crozier, Hendry, Lawson, Quinn, Mantua,
+  Battin, Shaw \& Huey}]{crozier2008}
+8.\\
+Crozier, L.~G., Hendry, A.~P., Lawson, P.~W., Quinn, T., Mantua, N.~J., Battin,
+  J., Shaw, R.~G. \& Huey, R.~B. (2008).
+\newblock Potential responses to climate change in organisms with complex life
+  histories: evolution and plasticity in {Pacific} salmon.
+\newblock \emph{Evol. Appl.}, 1, 252--270.
+
+\bibitem[{de~Mazancourt \emph{et~al.}(2013)de~Mazancourt, Isbell, Larocque,
+  Berendse, De~Luca, Grace, Haegeman, Wayne~Polley, Roscher, Schmid, Tilman,
+  van Ruijven, Weigelt, Wilsey \& Loreau}]{demazancourt2013}
+9.\\
+de~Mazancourt, C., Isbell, F., Larocque, A., Berendse, F., De~Luca, E., Grace,
+  J.~B., Haegeman, B., Wayne~Polley, H., Roscher, C., Schmid, B., Tilman, D.,
+  van Ruijven, J., Weigelt, A., Wilsey, B.~J. \& Loreau, M. (2013).
+\newblock Predicting ecosystem stability from community composition and
+  biodiversity.
+\newblock \emph{Ecol. Lett.}, 16, 617--625.
+
+\bibitem[{DFO(2005)}]{dfo2005wsp}
+10.\\
+DFO (2005).
+\newblock Canada's policy for conservation of wild {Pacific} salmon.
+\newblock Tech. rep., Fisheries and Oceans Canada, Vancouver, BC.
+
+\bibitem[{Eliason \emph{et~al.}(2011)Eliason, Clark, Hague, Hanson, Gallagher,
+  Jeffries, Gale, Patterson, Hinch \& Farrell}]{eliason2011}
+11.\\
+Eliason, E.~J., Clark, T.~D., Hague, M.~J., Hanson, L.~M., Gallagher, Z.~S.,
+  Jeffries, K.~M., Gale, M.~K., Patterson, D.~A., Hinch, S.~G. \& Farrell,
+  A.~P. (2011).
+\newblock Differences in thermal tolerance among sockeye salmon populations.
+\newblock \emph{Science}, 332, 109--112.
+
+\bibitem[{Elmqvist \emph{et~al.}(2003)Elmqvist, Folke, Nystr{\"o}m, Peterson,
+  Bengtsson, Walker \& Norberg}]{elmqvist2003}
+12.\\
+Elmqvist, T., Folke, C., Nystr{\"o}m, M., Peterson, G., Bengtsson, J., Walker,
+  B. \& Norberg, J. (2003).
+\newblock Response diversity, ecosystem change, and resilience.
+\newblock \emph{Front. Ecol. Environ.}, 1, 488--494.
+
+\bibitem[{Figge(2004)}]{figge2004}
+13.\\
+Figge, F. (2004).
+\newblock Bio-folio: applying portfolio theory to biodiversity.
+\newblock \emph{Biodivers. Conserv.}, 13, 827--849.
+
+\bibitem[{Fraser \emph{et~al.}(2011)Fraser, Weir, Bernatchez, Hansen \&
+  Taylor}]{fraser2011}
+14.\\
+Fraser, D.~J., Weir, L.~K., Bernatchez, L., Hansen, M.~M. \& Taylor, E.~B.
+  (2011).
+\newblock Extent and scale of local adaptation in salmonid fishes: review and
+  meta-analysis.
+\newblock \emph{Heredity}, 106, 404--420.
+
+\bibitem[{Gustafson \emph{et~al.}(2007)Gustafson, Waples, Myers, Weitkamp,
+  Bryant, Johnson \& Hard}]{gustafson2007}
+15.\\
+Gustafson, R.~G., Waples, R.~S., Myers, J.~M., Weitkamp, L.~A., Bryant, G.~J.,
+  Johnson, O.~W. \& Hard, J.~J. (2007).
+\newblock Pacific salmon extinctions: quantifying lost and remaining diversity.
+\newblock \emph{Conserv. Biol.}, 21, 1009--1020.
+
+\bibitem[{Haak \& Williams(2012)}]{haak2012}
+16.\\
+Haak, A.~L. \& Williams, J.~E. (2012).
+\newblock Spreading the risk: native trout management in a warmer and
+  less-certain future.
+\newblock \emph{N. Am. J. Fish. Manage.}, 32, 387--401.
+
+\bibitem[{Hannah \emph{et~al.}(2002)Hannah, Midgley, Lovejoy, Bond, Bush,
+  Lovett, Scott \& Woodward}]{hannah2002a}
+17.\\
+Hannah, L., Midgley, G.~F., Lovejoy, T., Bond, W.~J., Bush, M., Lovett, J.~C.,
+  Scott, D. \& Woodward, F.~I. (2002).
+\newblock Conservation of biodiversity in a changing climate.
+\newblock \emph{Conserv. Biol.}, 16, 264--268.
+
+\bibitem[{Hilborn \emph{et~al.}(2003)Hilborn, Quinn, Schindler \&
+  Rogers}]{hilborn2003}
+18.\\
+Hilborn, R.~W., Quinn, T., Schindler, D. \& Rogers, D. (2003).
+\newblock Biocomplexity and fisheries sustainability.
+\newblock \emph{Proc. Natl. Acad. Sci. U.S.A.}, 100, 6564--6568.
+
+\bibitem[{Hilborn \& Walters(1992)}]{hilborn1992}
+19.\\
+Hilborn, R.~W. \& Walters, C. (1992).
+\newblock \emph{Quantitative Fisheries Stock Assessment: Choice, Dynamics, and
+  Uncertainty}.
+\newblock Chapman and Hall, London.
+
+\bibitem[{Ives \& Carpenter(2007)}]{ives2007}
+20.\\
+Ives, A.~R. \& Carpenter, S.~R. (2007).
+\newblock Stability and diversity of ecosystems.
+\newblock \emph{Science}, 317, 58--62.
+
+\bibitem[{Johnson \& Schindler(2013)}]{johnson2013a}
+21.\\
+Johnson, S.~P. \& Schindler, D.~E. (2013).
+\newblock Marine trophic diversity in an anadromous fish is linked to its
+  life-history variation in fresh water.
+\newblock \emph{Biol. Lett.}, 9, 20120824.
+
+\bibitem[{Koellner \& Schmitz(2006)}]{koellner2006}
+22.\\
+Koellner, T. \& Schmitz, O.~J. (2006).
+\newblock Biodiversity, ecosystem function, and investment risk.
+\newblock \emph{BioScience}, 56, 977--985.
+
+\bibitem[{Lackey(2003)}]{lackey2003}
+23.\\
+Lackey, R.~T. (2003).
+\newblock {Pacific Northwest} salmon: Forecasting their status in 2100.
+\newblock \emph{Rev. Fish. Sci.}, 11, 35--88.
+
+\bibitem[{Margules \& Pressey(2000)}]{margules2000}
+24.\\
+Margules, C.~R. \& Pressey, R.~L. (2000).
+\newblock Systematic conservation planning.
+\newblock \emph{Nature}, 405, 243--253.
+
+\bibitem[{Markowitz(1952)}]{markowitz1952}
+25.\\
+Markowitz, H. (1952).
+\newblock Portfolio selection.
+\newblock \emph{J. Finance}, 7, 77--91.
+
+\bibitem[{McClure \emph{et~al.}(2008{\natexlab{a}})McClure, Carlson, Beechie,
+  Pess, Jorgensen, Sogard, Sultan, Holzer, Travis, Sanderson, Power \&
+  Carmichael}]{mcclure2008a}
+26.\\
+McClure, M.~M., Carlson, S.~M., Beechie, T.~J., Pess, G.~R., Jorgensen, J.~C.,
+  Sogard, S.~M., Sultan, S.~E., Holzer, D.~M., Travis, J., Sanderson, B.~L.,
+  Power, M.~E. \& Carmichael, R.~W. (2008{\natexlab{a}}).
+\newblock Evolutionary consequences of habitat loss for {Pacific} anadromous
+  salmonids.
+\newblock \emph{Evol. Appl.}, 1, 300--318.
+
+\bibitem[{McClure \emph{et~al.}(2008{\natexlab{b}})McClure, Utter, Baldwin,
+  Carmichael, Hassemer, Howell, Spruell, Cooney, Schaller \&
+  Petrosky}]{mcclure2008b}
+27.\\
+McClure, M.~M., Utter, F.~M., Baldwin, C., Carmichael, R.~W., Hassemer, P.~F.,
+  Howell, P.~J., Spruell, P., Cooney, T.~D., Schaller, H.~A. \& Petrosky, C.~E.
+  (2008{\natexlab{b}}).
+\newblock Evolutionary effects of alternative artificial propagation programs:
+  implications for viability of endangered anadromous salmonids.
+\newblock \emph{Evol. Appl.}, 1, 356--375.
+
+\bibitem[{McCullough(1999)}]{mccullough1999}
+28.\\
+McCullough, D.~A. (1999).
+\newblock A review and synthesis of effects of alterations to the water
+  temperature regime on freshwater life stages of salmonids, with special
+  reference to {Chinook} salmon.
+\newblock EPA 910-R-99-010, US Environmental Protection Agency.
+
+\bibitem[{McElhany \emph{et~al.}(2000)McElhany, Ruckelshaus, Ford, Wainwright
+  \& Bjorkstedt}]{mcelhany2000}
+29.\\
+McElhany, P., Ruckelshaus, M.~H., Ford, M.~J., Wainwright, T.~C. \& Bjorkstedt,
+  E.~P. (2000).
+\newblock {Viable Salmonid Populations} and the recovery of {Evolutionarily
+  Significant Units}.
+\newblock Tech. Rep. NOAA Tech. Memo. NMFS-NWFSC-42, U.S. Deptartment of
+  Commerce.
+
+\bibitem[{Moore \emph{et~al.}(2010)Moore, Mcclure, Rogers \&
+  Schindler}]{moore2010}
+30.\\
+Moore, J.~W., Mcclure, M., Rogers, L.~A. \& Schindler, D.~E. (2010).
+\newblock Synchronization and portfolio performance of threatened salmon.
+\newblock \emph{Conserv. Lett.}, 3, 340--348.
+
+\bibitem[{Mori \emph{et~al.}(2013)Mori, Furukawa \& Sasaki}]{mori2013}
+31.\\
+Mori, A.~S., Furukawa, T. \& Sasaki, T. (2013).
+\newblock Response diversity determines the resilience of ecosystems to
+  environmental change.
+\newblock \emph{Biol. Rev.}, 88, 349--364.
+
+\bibitem[{Patterson \emph{et~al.}(2007)Patterson, Skibo, Barnes, Hills \&
+  Macdonald}]{patterson2007}
+32.\\
+Patterson, D.~A., Skibo, K.~M., Barnes, D.~P., Hills, J.~A. \& Macdonald, J.~S.
+  (2007).
+\newblock The influence of water temperature on time to surface for adult
+  sockeye salmon carcasses and the limitations in estimating salmon carcasses
+  in the {Fraser River, British Columbia}.
+\newblock \emph{N. Am. J. Fish. Manage.}, 27, 878--884.
+
+\bibitem[{Pressey \emph{et~al.}(2007)Pressey, Cabeza, Watts, Cowling \&
+  Wilson}]{pressey2007}
+33.\\
+Pressey, R.~L., Cabeza, M., Watts, M.~E., Cowling, R.~M. \& Wilson, K.~A.
+  (2007).
+\newblock Conservation planning in a changing world.
+\newblock \emph{Trends Ecol. Evolut.}, 22, 583--592.
+
+\bibitem[{Quinn(2005)}]{quinn2005}
+34.\\
+Quinn, T.~P. (2005).
+\newblock \emph{The Behaviour and Ecology of Pacific Salmon and Trout}.
+\newblock American Fisheries Society, Bethesda, Maryland.
+
+\bibitem[{{\texttt{R} Core Team}(2013)}]{r2013}
+35.\\
+{\texttt{R} Core Team} (2013).
+\newblock \emph{\texttt{R}: {A} language and environment for statistical
+  computing}.
+\newblock \texttt{R} Foundation for Statistical Computing, Vienna, Austria.
+
+\bibitem[{Rogers \& Schindler(2008)}]{rogers2008}
+36.\\
+Rogers, L.~A. \& Schindler, D.~E. (2008).
+\newblock Asynchrony in population dynamics of sockeye salmon in southwest
+  {Alaska}.
+\newblock \emph{Oikos}, 117, 1578--1586.
+
+\bibitem[{Ruckelshaus \emph{et~al.}(2002)Ruckelshaus, Levin, Johnson \&
+  Kareiva}]{ruckelshaus2002}
+37.\\
+Ruckelshaus, M.~H., Levin, P., Johnson, J.~B. \& Kareiva, P.~M. (2002).
+\newblock The {Pacific} salmon wars: what science brings to the challenge of
+  recovering species.
+\newblock \emph{Annu. Rev. Ecol. Syst.}, 33, 665--706.
+
+\bibitem[{Scheuerell \emph{et~al.}(2006)Scheuerell, Hilborn, Ruckelshaus,
+  Bartz, Lagueux, Haas \& Rawson}]{scheuerell2006}
+38.\\
+Scheuerell, M.~D., Hilborn, R., Ruckelshaus, M.~H., Bartz, K.~K., Lagueux,
+  K.~M., Haas, A.~D. \& Rawson, K. (2006).
+\newblock The {Shiraz} model: a tool for incorporating anthropogenic effects
+  and fish--habitat relationships in conservation planning.
+\newblock \emph{Can. J. Fish. Aquat. Sci.}, 63, 1596--1607.
+
+\bibitem[{Schindler \emph{et~al.}(2008)Schindler, Augerot, Fleishman, Mantua,
+  Riddell, Ruckelshaus, Seeb \& Webster}]{schindler2008}
+39.\\
+Schindler, D.~E., Augerot, X., Fleishman, E., Mantua, N.~J., Riddell, B.,
+  Ruckelshaus, M., Seeb, J. \& Webster, M. (2008).
+\newblock Climate change, ecosystem impacts, and management for {Pacific}
+  salmon.
+\newblock \emph{Fisheries}, 33, 502--506.
+
+\bibitem[{Schindler \emph{et~al.}(2010)Schindler, Hilborn, Chasco, Boatright,
+  Quinn, Rogers \& Webster}]{schindler2010}
+40.\\
+Schindler, D.~E., Hilborn, R., Chasco, B., Boatright, C.~P., Quinn, T.~P.,
+  Rogers, L.~A. \& Webster, M.~S. (2010).
+\newblock Population diversity and the portfolio effect in an exploited
+  species.
+\newblock \emph{Nature}, 465, 609--612.
+
+\bibitem[{Schtickzelle \& Quinn(2007)}]{schtickzelle2007}
+41.\\
+Schtickzelle, N. \& Quinn, T.~P. (2007).
+\newblock A metapopulation perspective for salmon and other anadromous fish.
+\newblock \emph{Fish Fish.}, 8, 297--314.
+
+\bibitem[{Stockwell \emph{et~al.}(2003)Stockwell, Hendry \&
+  Kinnison}]{stockwell2003}
+42.\\
+Stockwell, C., Hendry, A. \& Kinnison, M. (2003).
+\newblock Contemporary evolution meets conservation biology.
+\newblock \emph{Trends. Ecol. Evol.}, 18, 94--101.
+
+\bibitem[{Thibaut \emph{et~al.}(2012)Thibaut, Connolly \&
+  Sweatman}]{thibaut2012}
+43.\\
+Thibaut, L., Connolly, S.~R. \& Sweatman, H. P.~A. (2012).
+\newblock Diversity and stability of herbivorous fishes on coral reefs.
+\newblock \emph{Ecology}, 93, 891--901.
+
+\bibitem[{Thomas(2010)}]{thomas2010}
+44.\\
+Thomas, C.~D. (2010).
+\newblock Climate, climate change and range boundaries.
+\newblock \emph{Divers. Distrib.}, 16, 488--495.
+
+\bibitem[{Thomas \emph{et~al.}(2004)Thomas, Cameron, Green, Bakkenes, Beaumont,
+  Collingham, Erasmus, de~Siqueira, Grainger, Hannah, Hughes, Huntley, van
+  Jaarsveld, Midgley, Miles, Ortega-Huerta, Townsend~Peterson, Phillips \&
+  Williams}]{thomas2004}
+45.\\
+Thomas, C.~D., Cameron, A., Green, R.~E., Bakkenes, M., Beaumont, L.~J.,
+  Collingham, Y.~C., Erasmus, B. F.~N., de~Siqueira, M.~F., Grainger, A.,
+  Hannah, L., Hughes, L., Huntley, B., van Jaarsveld, A.~S., Midgley, G.~F.,
+  Miles, L., Ortega-Huerta, M.~A., Townsend~Peterson, A., Phillips, O.~L. \&
+  Williams, S.~E. (2004).
+\newblock Extinction risk from climate change.
+\newblock \emph{Nature}, 427, 145--148.
+
+\bibitem[{Thorson \emph{et~al.}(2013)Thorson, Scheuerell, Buhle \&
+  Copeland}]{thorson2013}
+46.\\
+Thorson, J.~T., Scheuerell, M.~D., Buhle, E.~R. \& Copeland, T. (2013).
+\newblock Spatial variation buffers temporal fluctuations in early juvenile
+  survival for an endangered {Pacific} salmon.
+\newblock \emph{J. Anim. Ecol.}, In press. doi:10.1111/1365-2656.12117.
+
+\bibitem[{Vert-pre \emph{et~al.}(2013)Vert-pre, Amoroso, Jensen \&
+  Hilborn}]{vert-pre2013}
+47.\\
+Vert-pre, K.~A., Amoroso, R.~O., Jensen, O.~P. \& Hilborn, R. (2013).
+\newblock Frequency and intensity of productivity regime shifts in marine fish
+  stocks.
+\newblock \emph{Proc. Natl. Acad. Sci. U.S.A.}, 110, 1779--1784.
+
+\bibitem[{Walther \emph{et~al.}(2002)Walther, Post, Convey, Menzel, Parmesan,
+  Beebee, Fromentin, Hoegh-Guldberg \& Bairlein}]{walther2002}
+48.\\
+Walther, G.-R., Post, E., Convey, P., Menzel, A., Parmesan, C., Beebee, T.
+  J.~C., Fromentin, J.-M., Hoegh-Guldberg, O. \& Bairlein, F. (2002).
+\newblock Ecological responses to recent climate change.
+\newblock \emph{Nature}, 416, 389--395.
+
+\bibitem[{Waples \emph{et~al.}(2009)Waples, Beechie \& Pess}]{waples2009}
+49.\\
+Waples, R., Beechie, T. \& Pess, G.~R. (2009).
+\newblock Evolutionary history, habitat disturbance regimes, and anthropogenic
+  changes: what do these mean for resilience of {Pacific} salmon populations?
+\newblock \emph{Ecol. Soc.}, 14, 3.
+
+\bibitem[{Yeakel \emph{et~al.}(2013)Yeakel, Moore, Guimar{\~a}es \&
+  de~Aguiar}]{yeakel2013}
+50.\\
+Yeakel, J.~D., Moore, J.~W., Guimar{\~a}es, P.~R. \& de~Aguiar, M. A.~M.
+  (2013).
+\newblock Synchronisation and stability in river metapopulation networks.
+\newblock \emph{Ecol. Lett.}, In press. doi: 10.1111/ele.12228.
+
+\end{thebibliography}
+
+\end{spacing}
+
 \begin{spacing}{1.4}
-\setlength{\parskip}{12pt}
+
+\clearpage
+
+\section{Figures}
+
+\begin{center}
+\includegraphics[height=5.5in]{../examples/simulation-diagram3.pdf}
+\end{center}
+Fig.\ 1: Flow chart of the salmon-metapopulation simulation.
+There are $n$ salmon populations and $t$ generations.
+Blue text indicates values that are generated before the simulation progresses through time.
+Red text indicates steps in which calculations are performed through time.
+Black text indicates values that are calculated.
+Grey text indicates parameters that can be set.
+Green text indicates the looping structure of the simulation.
+
+\clearpage
+
+\begin{center}
+\includegraphics[width=2.9in]{../examples/thermal-curve-scenarios.pdf}
+\end{center}
+Fig.\ 2: Different ways of prioritizing response-diversity conservation.
+Panel a shows thermal tolerance curves for ten possible populations and panels b--e show different ways of prioritizing four of those populations.
+The curves describe how productivity varies with temperature for a given population.
+Some populations thrive at low temperatures (cool colours) and some at warm temperatures (warm colours).
+Some are tolerant to a wider range of environmental conditions (yellow-to-green colours) but with a lower maximum productivity.
+The total possible productivity (the area under the curves) is the same for each population.
+
+\clearpage
+
+\begin{center}
+\includegraphics[width=4.0in]{../examples/spatial-arma-sim-full.pdf}
+\end{center}
+Fig.\ 3: The components of an example metapopulation simulation.
+We show, from top to bottom, the temperature signal, the resulting productivity parameter (Ricker $a$), the salmon returns, fisheries catch, salmon escapement, salmon straying from their natal streams, salmon joining from other streams, spawner-return residuals on a log scale, and the estimated $a$ and $b$ parameters in the fitted Ricker curve.
+The colored lines indicate populations that thrive at low (cool colours) to high (warm colours) temperatures.
+
+\clearpage
+
+\begin{center}
+\includegraphics[width=4.5in]{../examples/spatial-mv.pdf}
+\end{center}
+Fig.\ 4: The importance of preserving environmental response diversity through spatial conservation strategies.
+The conservation strategies correspond to Fig.\ 2 and represent conserving a range of responses (green), the most stable populations only (orange), or one type of environmental response (purple and pink).
+In risk-return space we show environmental scenarios that are comprised primarily of (a) short-term and (b) long-term environmental fluctuations.
+The dots show simulated metapopulations and the contours show 25\% and 75\% quantiles across 500 simulations per strategy.
+We also show example metapopulation abundance time series for the (c, e) short-term and (d, f) long-term environmental-fluctuation scenarios.
+The grey line (a, b) indicates the efficient frontier across all simulated metapopulations --- metapopulations with the minimum variability for a given level of growth rate.
+
+\clearpage
+
+\begin{center}
+\includegraphics[width=4.5in]{../examples/cons-plans-n.pdf}
+\end{center}
+Fig.\ 5: The importance of preserving as many populations as possible when we don't know how response diversity is distributed.
+In risk-return space we show environmental scenarios that are comprised primarily of (a) short-term and (b) long-term environmental fluctuations.
+We show metapopulations in which 2 (red), 4 (orange), 8 (yellow), or 16 (green) populations of random response diversity are conserved.
+The dots show simulated metapopulations and the contours show 25\% and 75\% quantiles across 500 simulations per strategy.
+We also show example metapopulation (c) rate-of-change and (d) abundance time series for the short-term environmental-fluctuation scenario.
+The grey line (a, b) indicates the efficient frontier across all simulated metapopulations --- metapopulations with the minimum variability for a given level of growth rate
+
+\end{spacing}
+
+\setlength{\parskip}{8pt}
 \setlength{\parindent}{0cm}
-\include{ms}
 
-% redefine the command that creates the Table number.
-\renewcommand{\thetable}{S\arabic{table}}
-\setcounter{table}{0} % reset counter
-
-% redefine the command that creates the Figure number.
-\renewcommand{\thefigure}{S\arabic{figure}}
-\renewcommand{\figurename}{Fig.}
-\setcounter{figure}{0}  % reset counter
+\begin{spacing}{1.2}
 
@@
 \end{spacing}
+
 \end{document}

### 2013-12-12 inst/examples/README.R
Clean up readme

@@ -2,17 +2,22 @@
 
 # Set your R working directory to this folder (the examples folder).
 
-# install and load the package:
-# devtools::install_github("metafolio", "seananderson")
-# library(metafolio)
+# Install and load the package and then:
+library(metafolio)
 
 # or, for rapid development:
-devtools::load_all("../../")
+# devtools::load_all("../../")
 
 # ggplot like colour palette to use throughout for stream ID:
 col_pal <- rev(gg_color_hue(10))
 
-# Figure showing different number rules of thumb:
+# Warning: these first two sourced files will take a while to run.
+# Once they've be run once, you can go in and turn USE_CACHE <- TRUE
+# near the top of each file if you want to use the previously
+# generated output.
+#
+# Figure showing random response diversity but different
+# numbers of populations conserved:
 source("cons_plans_n_random_resp_div.R")
 
 # Figure showing different spatial rules of thumb:

### 2013-12-12 inst/examples/cons_plans2.R
Clean up fig making scripts

@@ -1,21 +1,15 @@
-# look at conservation rules of thumb
-# This file looks at different spatial strategies of conservation
-# This version uses the new package functions
+# This file looks at different spatial conservation strategies
 
 set.seed(2)
-USE_CACHE <- TRUE
+USE_CACHE <- FALSE
 
-w_plans <- list() 
+w_plans <- list()
 w_plans[[1]] <- c(5, 1000, 5, 1000, 5, 5, 1000, 5, 1000, 5)
-#w_plans[[1]] <- c(1000, 5, 1000, 5, 1000, 1000, 5, 1000, 5, 1000)
 w_plans[[2]] <- c(5, 5, 5, 1000, 1000, 1000, 1000, 5, 5, 5)
-#w_plans[[2]] <- c(5, 5, 1000, 1000, 1000, 1000, 1000, 1000, 5, 5)
 w_plans[[3]] <- c(rep(1000, 4), rep(5, 6))
-#w_plans[[3]] <- c(rep(1000, 6), rep(5, 4))
 w_plans[[4]] <- rev(w_plans[[3]])
-#w_plans[[5]] <- c(1000, 1000, 5, 5, 5, 5, 5, 5, 1000, 1000)
-#plans_name_sp <- c("Full range of responses", "Most stable only", "One half", "One half", "Most asynchronous only")
-plans_name_sp <- c("Full range of responses", "Most stable only", "Lower half", "Upper half")
+plans_name_sp <- c("Full range of responses", "Most stable only",
+  "Lower half", "Upper half")
 
 n_trials <- 500 # number of trials at each n conservation plan
 num_pops <- c(10, 10, 10, 10) # n pops to conserve
@@ -31,14 +25,11 @@ for(i in 1:n_plans) { # loop over plans
 ## ARMA:
 arma_env_params <- list(mean_value = 16, ar = 0.1, sigma_env = 2, ma = 0)
 
-#plot_sim_ts(meta_sim(b = rep(1000, 10), n_pop = 10, env_params =
-    #arma_env_params, env_type = "arma", assess_freq = 5),
-  #years_to_show = 100, burn = 0)
-
 pdf("spatial-arma-sim-full.pdf", width = 5, height = 7)
 eg_arma <- meta_sim(b = w[[1]][[1]], n_pop = 10, env_params = arma_env_params,
[…]

### 2013-12-12 inst/examples/cons_plans_n_random_resp_div.R

@@ -4,7 +4,7 @@
 # In this version, the response diversity is randomly drawn
 
 set.seed(1)
-USE_CACHE <- TRUE
+USE_CACHE <- FALSE
 
 # moved this out of the package to avoid Teaching Demos import
 add_inset_env <- function(env, x = 0.12, y = -0.013, size = c(1, .5), ...) {
@@ -16,7 +16,6 @@ add_inset_env <- function(env, x = 0.12, y = -0.013, size = c(1, .5), ...) {
   par(xpd = FALSE)
 }
 
-
 # in this version, the pops are wiped out; total abundance changes
 n_trials <- 800 # number of trials at each n conservation plan
 num_pops <- c(2, 4, 8, 16) # n pops to conserve
@@ -26,17 +25,14 @@ for(i in 1:n_plans) { # loop over number conserved
  w[[i]] <- list()
  for(j in 1:n_trials) { # loop over trials
    w[[i]][[j]] <- matrix(rep(1000, 16), nrow = 1)
-   w[[i]][[j]][-sample(1:16, num_pops[i])] <- 5 # conserve num_pops[i] populations; wipe out rest
+   # conserve num_pops[i] populations; wipe out rest:
+   w[[i]][[j]][-sample(1:16, num_pops[i])] <- 5
  }
 }
 plans_name_n <- paste(num_pops, "populations")
 
 ## ARMA:
 arma_env_params <- list(mean_value = 16, ar = 0.1, sigma_env = 2, ma = 0)
-
-#plot_sim_ts(meta_sim(b = rep(1000, 10), n_pop = 10, env_params =
-    #arma_env_params, env_type = "arma", assess_freq = 5),
-  #years_to_show = 100, burn = 0)
 
 pdf("n-arma-sim-2.pdf", width = 5, height = 7)
 eg_arma <- meta_sim(b = w[[1]][[2]], n_pop = 16, env_params =
@@ -52,8 +48,6 @@ eg_arma <- meta_sim(b = w[[4]][[2]], n_pop = 16, env_params =
 plot_sim_ts(eg_arma, years_to_show = 70, burn = 30)
 dev.off()
 
-
-# TODO need to fix it here to get max_a integrated based on n_pops
 if(!USE_CACHE) {
[…]

### 2013-12-12 inst/examples/figs2eps.R
Add functions to make eps files

@@ -1,0 +1,9 @@
+f <- list.files(pattern = "*.pdf")
+# these have transparency, need to use AI
+f <- f[-which(f == "spatial-mv.pdf")]
+f <- f[-which(f == "plot-various-options-ts-3pops.pdf")]
+f <- f[-which(f == "cons-plans-n.pdf")]
+for(i in 1:length(f)) {
+  system(paste0("pdftops -eps ", f[i]))
+}
+

### 2013-12-12 inst/ms/salmonfolios.tex
Switch to local extensionless figs

@@
 \begin{center}
-\includegraphics[height=5.5in]{../examples/simulation-diagram3.pdf}
+\includegraphics[height=5.5in]{simulation-diagram3}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=2.9in]{../examples/thermal-curve-scenarios.pdf}
+\includegraphics[width=2.9in]{thermal-curve-scenarios}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=4.0in]{../examples/spatial-arma-sim-full.pdf}
+\includegraphics[width=4.0in]{spatial-arma-sim-full}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=4.5in]{../examples/spatial-mv.pdf}
+\includegraphics[width=4.5in]{spatial-mv}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=4.5in]{../examples/cons-plans-n.pdf}
+\includegraphics[width=4.5in]{cons-plans-n}
 \end{center}

### 2013-12-13 inst/ms/salmonfolios.tex
Clean main .tex file, fix numbers, declare file extension

@@
 \textheight 20.5cm
-%\topmargin -1.43cm
-%\oddsidemargin 1.2cm
-%\evensidemargin 1.2cm
-%\textwidth 14.5cm
 \usepackage{lineno}
@@
 \usepackage{booktabs}
+
+\DeclareGraphicsExtensions{.eps}
 
@@
 \titlespacing\subsection{0pt}{6pt plus 4pt minus 2pt}{-8pt plus 2pt minus 2pt}
-
-%\usepackage{fontspec}
-%\defaultfontfeatures{Ligatures=TeX}
-%\setsansfont{Calibri}
-%\setmonofont{Inconsolata}
-%\usepackage{mathptmx}
-%\setmainfont{Minion Pro}
 
@@
 
+\bibitem[{Carlson \& Satterthwaite(2011)}]{carlson2011}
+7.\\
+Carlson, S.~M. \& Satterthwaite, W.~H. (2011).
+\newblock Weakened portfolio effect in a collapsed salmon population complex.
+\newblock \emph{Can. J. Fish. Aquat. Sci.}, 68, 1579--1589.
+
 \bibitem[{Cooper \& Mangel(1999)}]{cooper1999}
-7.\\
+8.\\
 Cooper, A.~B. \& Mangel, M. (1999).
@@
   Battin, Shaw \& Huey}]{crozier2008}
-8.\\
+9.\\
 Crozier, L.~G., Hendry, A.~P., Lawson, P.~W., Quinn, T., Mantua, N.~J., Battin,
@@
   van Ruijven, Weigelt, Wilsey \& Loreau}]{demazancourt2013}
-9.\\
+10.\\
 de~Mazancourt, C., Isbell, F., Larocque, A., Berendse, F., De~Luca, E., Grace,
@@
 \bibitem[{DFO(2005)}]{dfo2005wsp}
-10.\\
+11.\\
 DFO (2005).
@@
   Jeffries, Gale, Patterson, Hinch \& Farrell}]{eliason2011}
-11.\\
+12.\\
 Eliason, E.~J., Clark, T.~D., Hague, M.~J., Hanson, L.~M., Gallagher, Z.~S.,
@@
   Bengtsson, Walker \& Norberg}]{elmqvist2003}
-12.\\
+13.\\
 Elmqvist, T., Folke, C., Nystr{\"o}m, M., Peterson, G., Bengtsson, J., Walker,
@@
 \bibitem[{Figge(2004)}]{figge2004}
-13.\\
+14.\\
 Figge, F. (2004).
@@
   Taylor}]{fraser2011}
-14.\\
+15.\\
 Fraser, D.~J., Weir, L.~K., Bernatchez, L., Hansen, M.~M. \& Taylor, E.~B.
@@
   Bryant, Johnson \& Hard}]{gustafson2007}
-15.\\
+16.\\
 Gustafson, R.~G., Waples, R.~S., Myers, J.~M., Weitkamp, L.~A., Bryant, G.~J.,
@@
 \bibitem[{Haak \& Williams(2012)}]{haak2012}
-16.\\
+17.\\
 Haak, A.~L. \& Williams, J.~E. (2012).
@@
   Lovett, Scott \& Woodward}]{hannah2002a}
-17.\\
+18.\\
 Hannah, L., Midgley, G.~F., Lovejoy, T., Bond, W.~J., Bush, M., Lovett, J.~C.,
@@
   Rogers}]{hilborn2003}
-18.\\
+19.\\
 Hilborn, R.~W., Quinn, T., Schindler, D. \& Rogers, D. (2003).
@@
 \bibitem[{Hilborn \& Walters(1992)}]{hilborn1992}
-19.\\
+20.\\
 Hilborn, R.~W. \& Walters, C. (1992).
@@
 \bibitem[{Ives \& Carpenter(2007)}]{ives2007}
-20.\\
+21.\\
 Ives, A.~R. \& Carpenter, S.~R. (2007).
@@
 \bibitem[{Johnson \& Schindler(2013)}]{johnson2013a}
-21.\\
+22.\\
 Johnson, S.~P. \& Schindler, D.~E. (2013).
@@
 \bibitem[{Koellner \& Schmitz(2006)}]{koellner2006}
-22.\\
+23.\\
 Koellner, T. \& Schmitz, O.~J. (2006).
@@
 \bibitem[{Lackey(2003)}]{lackey2003}
-23.\\
+24.\\
 Lackey, R.~T. (2003).
@@
 \bibitem[{Margules \& Pressey(2000)}]{margules2000}
-24.\\
+25.\\
 Margules, C.~R. \& Pressey, R.~L. (2000).
@@
 \bibitem[{Markowitz(1952)}]{markowitz1952}
-25.\\
+26.\\
 Markowitz, H. (1952).
@@
   Carmichael}]{mcclure2008a}
-26.\\
+27.\\
 McClure, M.~M., Carlson, S.~M., Beechie, T.~J., Pess, G.~R., Jorgensen, J.~C.,
@@
   Petrosky}]{mcclure2008b}
-27.\\
+28.\\
 McClure, M.~M., Utter, F.~M., Baldwin, C., Carmichael, R.~W., Hassemer, P.~F.,
@@
 \bibitem[{McCullough(1999)}]{mccullough1999}
-28.\\
+29.\\
 McCullough, D.~A. (1999).
@@
   \& Bjorkstedt}]{mcelhany2000}
-29.\\
+30.\\
 McElhany, P., Ruckelshaus, M.~H., Ford, M.~J., Wainwright, T.~C. \& Bjorkstedt,
@@
   Schindler}]{moore2010}
-30.\\
+31.\\
 Moore, J.~W., Mcclure, M., Rogers, L.~A. \& Schindler, D.~E. (2010).
@@
 \bibitem[{Mori \emph{et~al.}(2013)Mori, Furukawa \& Sasaki}]{mori2013}
-31.\\
+32.\\
 Mori, A.~S., Furukawa, T. \& Sasaki, T. (2013).
@@
   Macdonald}]{patterson2007}
-32.\\
+33.\\
 Patterson, D.~A., Skibo, K.~M., Barnes, D.~P., Hills, J.~A. \& Macdonald, J.~S.
@@
   Wilson}]{pressey2007}
-33.\\
+34.\\
 Pressey, R.~L., Cabeza, M., Watts, M.~E., Cowling, R.~M. \& Wilson, K.~A.
@@
 \bibitem[{Quinn(2005)}]{quinn2005}
-34.\\
+35.\\
 Quinn, T.~P. (2005).
@@
 \bibitem[{{\texttt{R} Core Team}(2013)}]{r2013}
-35.\\
+36.\\
 {\texttt{R} Core Team} (2013).
@@
 \bibitem[{Rogers \& Schindler(2008)}]{rogers2008}
-36.\\
+37.\\
 Rogers, L.~A. \& Schindler, D.~E. (2008).
@@
   Kareiva}]{ruckelshaus2002}
-37.\\
+38.\\
 Ruckelshaus, M.~H., Levin, P., Johnson, J.~B. \& Kareiva, P.~M. (2002).
@@
   Bartz, Lagueux, Haas \& Rawson}]{scheuerell2006}
-38.\\
+39.\\
 Scheuerell, M.~D., Hilborn, R., Ruckelshaus, M.~H., Bartz, K.~K., Lagueux,
@@
   Riddell, Ruckelshaus, Seeb \& Webster}]{schindler2008}
-39.\\
+40.\\
 Schindler, D.~E., Augerot, X., Fleishman, E., Mantua, N.~J., Riddell, B.,
@@
   Quinn, Rogers \& Webster}]{schindler2010}
-40.\\
+41.\\
 Schindler, D.~E., Hilborn, R., Chasco, B., Boatright, C.~P., Quinn, T.~P.,
@@
 \bibitem[{Schtickzelle \& Quinn(2007)}]{schtickzelle2007}
-41.\\
+42.\\
 Schtickzelle, N. \& Quinn, T.~P. (2007).
@@
   Kinnison}]{stockwell2003}
-42.\\
+43.\\
 Stockwell, C., Hendry, A. \& Kinnison, M. (2003).
@@
   Sweatman}]{thibaut2012}
-43.\\
+44.\\
 Thibaut, L., Connolly, S.~R. \& Sweatman, H. P.~A. (2012).
@@
 \bibitem[{Thomas(2010)}]{thomas2010}
-44.\\
+45.\\
 Thomas, C.~D. (2010).
@@
   Williams}]{thomas2004}
-45.\\
+46.\\
 Thomas, C.~D., Cameron, A., Green, R.~E., Bakkenes, M., Beaumont, L.~J.,
@@
   Copeland}]{thorson2013}
-46.\\
+47.\\
 Thorson, J.~T., Scheuerell, M.~D., Buhle, E.~R. \& Copeland, T. (2013).
@@
 \newblock \emph{J. Anim. Ecol.}, In press. doi:10.1111/1365-2656.12117.
-
-\bibitem[{Vert-pre \emph{et~al.}(2013)Vert-pre, Amoroso, Jensen \&
-  Hilborn}]{vert-pre2013}
-47.\\
-Vert-pre, K.~A., Amoroso, R.~O., Jensen, O.~P. \& Hilborn, R. (2013).
-\newblock Frequency and intensity of productivity regime shifts in marine fish
-  stocks.
-\newblock \emph{Proc. Natl. Acad. Sci. U.S.A.}, 110, 1779--1784.
 

### 2014-01-08 inst/ms/salmonfolios.tex
Update ms and salmonfolios for prsb

@@
 
-\DeclareGraphicsExtensions{.eps}
+\DeclareGraphicsExtensions{.pdf}
 
@@
 
-\bibliographystyle{ecologyletters2}
-%\bibliography{jshort,ms}
-
-\begin{thebibliography}{50}
-\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
-\expandafter\ifx\csname url\endcsname\relax
-  \def\url#1{\texttt{#1}}\fi
-\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
-
-\bibitem[{Allendorf \emph{et~al.}(1997)Allendorf, Bayles, Bottom, Currens,
-  Frissell, Hankin, Lichatowich, Nehlsen, Trotter \& Williams}]{allendorf1997}
-1.\\
-Allendorf, F.~W., Bayles, D., Bottom, D.~L., Currens, K.~P., Frissell, C.~A.,
-  Hankin, D., Lichatowich, J.~A., Nehlsen, W., Trotter, P.~C. \& Williams,
-  T.~H. (1997).
-\newblock Prioritizing {Pacific} salmon stocks for conservation.
-\newblock \emph{Conserv. Biol.}, 11, 140--152.
-
-\bibitem[{Anderson \emph{et~al.}(2013)Anderson, Cooper \& Dulvy}]{anderson2013}
-2.\\
-Anderson, S.~C., Cooper, A.~B. \& Dulvy, N.~K. (2013).
-\newblock Ecological prophets: Quantifying metapopulation portfolio effects.
-\newblock \emph{Methods Ecol. Evol.}, 4, 971--981.
-
-\bibitem[{Ando \& Mallory(2012)}]{ando2012}
-3.\\
-Ando, A.~W. \& Mallory, M.~L. (2012).
-\newblock Optimal portfolio design to reduce climate-related conservation
-  uncertainty in the {Prairie Pothole Region}.
-\newblock \emph{Proc. Natl. Acad. Sci. U.S.A.}, 109, 6484--6489.
-
-\bibitem[{Balvanera \emph{et~al.}(2006)Balvanera, Pfisterer, Buchmann, He,
-  Nakashizuka, Raffaelli \& Schmid}]{balvanera2006}
-4.\\
-Balvanera, P., Pfisterer, A.~B., Buchmann, N., He, J.-S., Nakashizuka, T.,
-  Raffaelli, D. \& Schmid, B. (2006).
-\newblock Quantifying the evidence for biodiversity effects on ecosystem
-  functioning and services.
-\newblock \emph{Ecol. Lett.}, 9, 1146--1156.
-
-\bibitem[{Battin \emph{et~al.}(2007)Battin, Wiley, Ruckelshaus, Palmer, Korb,
-  Bartz \& Imaki}]{battin2007}
-5.\\
-Battin, J., Wiley, M.~W., Ruckelshaus, M.~H., Palmer, R.~N., Korb, E., Bartz,
-  K.~K. \& Imaki, H. (2007).
-\newblock Projected impacts of climate change on salmon habitat restoration.
-\newblock \emph{Proc. Natl. Acad. Sci. U.S.A.}, 104, 6720--6725.
-
-\bibitem[{Beechie \emph{et~al.}(2013)Beechie, Imaki, Greene, Wade, Wu, Pess,
-  Roni, Kimball, Stanford, Kiffney \& Mantua}]{beechie2013}
-6.\\
-Beechie, T., Imaki, H., Greene, J., Wade, A., Wu, H., Pess, G., Roni, P.,
-  Kimball, J., Stanford, J., Kiffney, P. \& Mantua, N. (2013).
-\newblock Restoring salmon habitat for a changing climate.
-\newblock \emph{River Res. Applic.}, 29, 939--960.
-
-\bibitem[{Carlson \& Satterthwaite(2011)}]{carlson2011}
-7.\\
-Carlson, S.~M. \& Satterthwaite, W.~H. (2011).
-\newblock Weakened portfolio effect in a collapsed salmon population complex.
-\newblock \emph{Can. J. Fish. Aquat. Sci.}, 68, 1579--1589.
-
-\bibitem[{Cooper \& Mangel(1999)}]{cooper1999}
-8.\\
-Cooper, A.~B. \& Mangel, M. (1999).
-\newblock The dangers of ignoring metapopulation structure for the conservation
-  of salmonids.
-\newblock \emph{Fish. Bull.}, 97, 213--226.
-
-\bibitem[{Crozier \emph{et~al.}(2008)Crozier, Hendry, Lawson, Quinn, Mantua,
-  Battin, Shaw \& Huey}]{crozier2008}
-9.\\
-Crozier, L.~G., Hendry, A.~P., Lawson, P.~W., Quinn, T., Mantua, N.~J., Battin,
-  J., Shaw, R.~G. \& Huey, R.~B. (2008).
-\newblock Potential responses to climate change in organisms with complex life
-  histories: evolution and plasticity in {Pacific} salmon.
-\newblock \emph{Evol. Appl.}, 1, 252--270.
-
-\bibitem[{de~Mazancourt \emph{et~al.}(2013)de~Mazancourt, Isbell, Larocque,
-  Berendse, De~Luca, Grace, Haegeman, Wayne~Polley, Roscher, Schmid, Tilman,
-  van Ruijven, Weigelt, Wilsey \& Loreau}]{demazancourt2013}
-10.\\
-de~Mazancourt, C., Isbell, F., Larocque, A., Berendse, F., De~Luca, E., Grace,
-  J.~B., Haegeman, B., Wayne~Polley, H., Roscher, C., Schmid, B., Tilman, D.,
-  van Ruijven, J., Weigelt, A., Wilsey, B.~J. \& Loreau, M. (2013).
-\newblock Predicting ecosystem stability from community composition and
-  biodiversity.
-\newblock \emph{Ecol. Lett.}, 16, 617--625.
-
-\bibitem[{DFO(2005)}]{dfo2005wsp}
-11.\\
-DFO (2005).
-\newblock Canada's policy for conservation of wild {Pacific} salmon.
-\newblock Tech. rep., Fisheries and Oceans Canada, Vancouver, BC.
-
-\bibitem[{Eliason \emph{et~al.}(2011)Eliason, Clark, Hague, Hanson, Gallagher,
-  Jeffries, Gale, Patterson, Hinch \& Farrell}]{eliason2011}
-12.\\
-Eliason, E.~J., Clark, T.~D., Hague, M.~J., Hanson, L.~M., Gallagher, Z.~S.,
-  Jeffries, K.~M., Gale, M.~K., Patterson, D.~A., Hinch, S.~G. \& Farrell,
-  A.~P. (2011).
-\newblock Differences in thermal tolerance among sockeye salmon populations.
-\newblock \emph{Science}, 332, 109--112.
-
-\bibitem[{Elmqvist \emph{et~al.}(2003)Elmqvist, Folke, Nystr{\"o}m, Peterson,
-  Bengtsson, Walker \& Norberg}]{elmqvist2003}
-13.\\
-Elmqvist, T., Folke, C., Nystr{\"o}m, M., Peterson, G., Bengtsson, J., Walker,
-  B. \& Norberg, J. (2003).
-\newblock Response diversity, ecosystem change, and resilience.
-\newblock \emph{Front. Ecol. Environ.}, 1, 488--494.
-
-\bibitem[{Figge(2004)}]{figge2004}
-14.\\
-Figge, F. (2004).
-\newblock Bio-folio: applying portfolio theory to biodiversity.
-\newblock \emph{Biodivers. Conserv.}, 13, 827--849.
-
-\bibitem[{Fraser \emph{et~al.}(2011)Fraser, Weir, Bernatchez, Hansen \&
-  Taylor}]{fraser2011}
-15.\\
-Fraser, D.~J., Weir, L.~K., Bernatchez, L., Hansen, M.~M. \& Taylor, E.~B.
-  (2011).
-\newblock Extent and scale of local adaptation in salmonid fishes: review and
-  meta-analysis.
-\newblock \emph{Heredity}, 106, 404--420.
-
-\bibitem[{Gustafson \emph{et~al.}(2007)Gustafson, Waples, Myers, Weitkamp,
-  Bryant, Johnson \& Hard}]{gustafson2007}
-16.\\
-Gustafson, R.~G., Waples, R.~S., Myers, J.~M., Weitkamp, L.~A., Bryant, G.~J.,
-  Johnson, O.~W. \& Hard, J.~J. (2007).
-\newblock Pacific salmon extinctions: quantifying lost and remaining diversity.
-\newblock \emph{Conserv. Biol.}, 21, 1009--1020.
-
-\bibitem[{Haak \& Williams(2012)}]{haak2012}
-17.\\
-Haak, A.~L. \& Williams, J.~E. (2012).
-\newblock Spreading the risk: native trout management in a warmer and
-  less-certain future.
-\newblock \emph{N. Am. J. Fish. Manage.}, 32, 387--401.
-
-\bibitem[{Hannah \emph{et~al.}(2002)Hannah, Midgley, Lovejoy, Bond, Bush,
-  Lovett, Scott \& Woodward}]{hannah2002a}
-18.\\
-Hannah, L., Midgley, G.~F., Lovejoy, T., Bond, W.~J., Bush, M., Lovett, J.~C.,
-  Scott, D. \& Woodward, F.~I. (2002).
-\newblock Conservation of biodiversity in a changing climate.
-\newblock \emph{Conserv. Biol.}, 16, 264--268.
-
-\bibitem[{Hilborn \emph{et~al.}(2003)Hilborn, Quinn, Schindler \&
-  Rogers}]{hilborn2003}
-19.\\
-Hilborn, R.~W., Quinn, T., Schindler, D. \& Rogers, D. (2003).
-\newblock Biocomplexity and fisheries sustainability.
-\newblock \emph{Proc. Natl. Acad. Sci. U.S.A.}, 100, 6564--6568.
-
-\bibitem[{Hilborn \& Walters(1992)}]{hilborn1992}
-20.\\
-Hilborn, R.~W. \& Walters, C. (1992).
-\newblock \emph{Quantitative Fisheries Stock Assessment: Choice, Dynamics, and
-  Uncertainty}.
-\newblock Chapman and Hall, London.
-
-\bibitem[{Ives \& Carpenter(2007)}]{ives2007}
-21.\\
-Ives, A.~R. \& Carpenter, S.~R. (2007).
-\newblock Stability and diversity of ecosystems.
-\newblock \emph{Science}, 317, 58--62.
-
-\bibitem[{Johnson \& Schindler(2013)}]{johnson2013a}
-22.\\
-Johnson, S.~P. \& Schindler, D.~E. (2013).
-\newblock Marine trophic diversity in an anadromous fish is linked to its
-  life-history variation in fresh water.
-\newblock \emph{Biol. Lett.}, 9, 20120824.
-
-\bibitem[{Koellner \& Schmitz(2006)}]{koellner2006}
-23.\\
-Koellner, T. \& Schmitz, O.~J. (2006).
-\newblock Biodiversity, ecosystem function, and investment risk.
-\newblock \emph{BioScience}, 56, 977--985.
-
-\bibitem[{Lackey(2003)}]{lackey2003}
-24.\\
-Lackey, R.~T. (2003).
-\newblock {Pacific Northwest} salmon: Forecasting their status in 2100.
-\newblock \emph{Rev. Fish. Sci.}, 11, 35--88.
-
-\bibitem[{Margules \& Pressey(2000)}]{margules2000}
-25.\\
-Margules, C.~R. \& Pressey, R.~L. (2000).
-\newblock Systematic conservation planning.
-\newblock \emph{Nature}, 405, 243--253.
-
-\bibitem[{Markowitz(1952)}]{markowitz1952}
-26.\\
-Markowitz, H. (1952).
-\newblock Portfolio selection.
-\newblock \emph{J. Finance}, 7, 77--91.
-
-\bibitem[{McClure \emph{et~al.}(2008{\natexlab{a}})McClure, Carlson, Beechie,
-  Pess, Jorgensen, Sogard, Sultan, Holzer, Travis, Sanderson, Power \&
-  Carmichael}]{mcclure2008a}
-27.\\
-McClure, M.~M., Carlson, S.~M., Beechie, T.~J., Pess, G.~R., Jorgensen, J.~C.,
-  Sogard, S.~M., Sultan, S.~E., Holzer, D.~M., Travis, J., Sanderson, B.~L.,
-  Power, M.~E. \& Carmichael, R.~W. (2008{\natexlab{a}}).
-\newblock Evolutionary consequences of habitat loss for {Pacific} anadromous
-  salmonids.
-\newblock \emph{Evol. Appl.}, 1, 300--318.
-
-\bibitem[{McClure \emph{et~al.}(2008{\natexlab{b}})McClure, Utter, Baldwin,
-  Carmichael, Hassemer, Howell, Spruell, Cooney, Schaller \&
-  Petrosky}]{mcclure2008b}
-28.\\
-McClure, M.~M., Utter, F.~M., Baldwin, C., Carmichael, R.~W., Hassemer, P.~F.,
-  Howell, P.~J., Spruell, P., Cooney, T.~D., Schaller, H.~A. \& Petrosky, C.~E.
-  (2008{\natexlab{b}}).
-\newblock Evolutionary effects of alternative artificial propagation programs:
-  implications for viability of endangered anadromous salmonids.
-\newblock \emph{Evol. Appl.}, 1, 356--375.
-
-\bibitem[{McCullough(1999)}]{mccullough1999}
-29.\\
-McCullough, D.~A. (1999).
-\newblock A review and synthesis of effects of alterations to the water
-  temperature regime on freshwater life stages of salmonids, with special
-  reference to {Chinook} salmon.
-\newblock EPA 910-R-99-010, US Environmental Protection Agency.
-
-\bibitem[{McElhany \emph{et~al.}(2000)McElhany, Ruckelshaus, Ford, Wainwright
-  \& Bjorkstedt}]{mcelhany2000}
-30.\\
-McElhany, P., Ruckelshaus, M.~H., Ford, M.~J., Wainwright, T.~C. \& Bjorkstedt,
-  E.~P. (2000).
-\newblock {Viable Salmonid Populations} and the recovery of {Evolutionarily
-  Significant Units}.
-\newblock Tech. Rep. NOAA Tech. Memo. NMFS-NWFSC-42, U.S. Deptartment of
-  Commerce.
-
-\bibitem[{Moore \emph{et~al.}(2010)Moore, Mcclure, Rogers \&
-  Schindler}]{moore2010}
-31.\\
-Moore, J.~W., Mcclure, M., Rogers, L.~A. \& Schindler, D.~E. (2010).
-\newblock Synchronization and portfolio performance of threatened salmon.
-\newblock \emph{Conserv. Lett.}, 3, 340--348.
-
-\bibitem[{Mori \emph{et~al.}(2013)Mori, Furukawa \& Sasaki}]{mori2013}
-32.\\
-Mori, A.~S., Furukawa, T. \& Sasaki, T. (2013).
-\newblock Response diversity determines the resilience of ecosystems to
-  environmental change.
-\newblock \emph{Biol. Rev.}, 88, 349--364.
-
-\bibitem[{Patterson \emph{et~al.}(2007)Patterson, Skibo, Barnes, Hills \&
-  Macdonald}]{patterson2007}
-33.\\
-Patterson, D.~A., Skibo, K.~M., Barnes, D.~P., Hills, J.~A. \& Macdonald, J.~S.
-  (2007).
-\newblock The influence of water temperature on time to surface for adult
-  sockeye salmon carcasses and the limitations in estimating salmon carcasses
-  in the {Fraser River, British Columbia}.
-\newblock \emph{N. Am. J. Fish. Manage.}, 27, 878--884.
-
-\bibitem[{Pressey \emph{et~al.}(2007)Pressey, Cabeza, Watts, Cowling \&
-  Wilson}]{pressey2007}
-34.\\
-Pressey, R.~L., Cabeza, M., Watts, M.~E., Cowling, R.~M. \& Wilson, K.~A.
-  (2007).
-\newblock Conservation planning in a changing world.
-\newblock \emph{Trends Ecol. Evolut.}, 22, 583--592.
-
-\bibitem[{Quinn(2005)}]{quinn2005}
-35.\\
-Quinn, T.~P. (2005).
-\newblock \emph{The Behaviour and Ecology of Pacific Salmon and Trout}.
-\newblock American Fisheries Society, Bethesda, Maryland.
-
-\bibitem[{{\texttt{R} Core Team}(2013)}]{r2013}
-36.\\
-{\texttt{R} Core Team} (2013).
-\newblock \emph{\texttt{R}: {A} language and environment for statistical
-  computing}.
-\newblock \texttt{R} Foundation for Statistical Computing, Vienna, Austria.
-
-\bibitem[{Rogers \& Schindler(2008)}]{rogers2008}
-37.\\
-Rogers, L.~A. \& Schindler, D.~E. (2008).
-\newblock Asynchrony in population dynamics of sockeye salmon in southwest
-  {Alaska}.
-\newblock \emph{Oikos}, 117, 1578--1586.
-
-\bibitem[{Ruckelshaus \emph{et~al.}(2002)Ruckelshaus, Levin, Johnson \&
-  Kareiva}]{ruckelshaus2002}
-38.\\
-Ruckelshaus, M.~H., Levin, P., Johnson, J.~B. \& Kareiva, P.~M. (2002).
-\newblock The {Pacific} salmon wars: what science brings to the challenge of
-  recovering species.
-\newblock \emph{Annu. Rev. Ecol. Syst.}, 33, 665--706.
-
-\bibitem[{Scheuerell \emph{et~al.}(2006)Scheuerell, Hilborn, Ruckelshaus,
-  Bartz, Lagueux, Haas \& Rawson}]{scheuerell2006}
-39.\\
-Scheuerell, M.~D., Hilborn, R., Ruckelshaus, M.~H., Bartz, K.~K., Lagueux,
-  K.~M., Haas, A.~D. \& Rawson, K. (2006).
-\newblock The {Shiraz} model: a tool for incorporating anthropogenic effects
-  and fish--habitat relationships in conservation planning.
-\newblock \emph{Can. J. Fish. Aquat. Sci.}, 63, 1596--1607.
-
-\bibitem[{Schindler \emph{et~al.}(2008)Schindler, Augerot, Fleishman, Mantua,
-  Riddell, Ruckelshaus, Seeb \& Webster}]{schindler2008}
-40.\\
-Schindler, D.~E., Augerot, X., Fleishman, E., Mantua, N.~J., Riddell, B.,
-  Ruckelshaus, M., Seeb, J. \& Webster, M. (2008).
-\newblock Climate change, ecosystem impacts, and management for {Pacific}
-  salmon.
-\newblock \emph{Fisheries}, 33, 502--506.
-
-\bibitem[{Schindler \emph{et~al.}(2010)Schindler, Hilborn, Chasco, Boatright,
-  Quinn, Rogers \& Webster}]{schindler2010}
-41.\\
-Schindler, D.~E., Hilborn, R., Chasco, B., Boatright, C.~P., Quinn, T.~P.,
-  Rogers, L.~A. \& Webster, M.~S. (2010).
-\newblock Population diversity and the portfolio effect in an exploited
-  species.
-\newblock \emph{Nature}, 465, 609--612.
-
-\bibitem[{Schtickzelle \& Quinn(2007)}]{schtickzelle2007}
-42.\\
-Schtickzelle, N. \& Quinn, T.~P. (2007).
-\newblock A metapopulation perspective for salmon and other anadromous fish.
-\newblock \emph{Fish Fish.}, 8, 297--314.
-
-\bibitem[{Stockwell \emph{et~al.}(2003)Stockwell, Hendry \&
-  Kinnison}]{stockwell2003}
-43.\\
-Stockwell, C., Hendry, A. \& Kinnison, M. (2003).
-\newblock Contemporary evolution meets conservation biology.
-\newblock \emph{Trends. Ecol. Evol.}, 18, 94--101.
-
-\bibitem[{Thibaut \emph{et~al.}(2012)Thibaut, Connolly \&
-  Sweatman}]{thibaut2012}
-44.\\
-Thibaut, L., Connolly, S.~R. \& Sweatman, H. P.~A. (2012).
-\newblock Diversity and stability of herbivorous fishes on coral reefs.
-\newblock \emph{Ecology}, 93, 891--901.
-
-\bibitem[{Thomas(2010)}]{thomas2010}
-45.\\
-Thomas, C.~D. (2010).
-\newblock Climate, climate change and range boundaries.
-\newblock \emph{Divers. Distrib.}, 16, 488--495.
-
-\bibitem[{Thomas \emph{et~al.}(2004)Thomas, Cameron, Green, Bakkenes, Beaumont,
-  Collingham, Erasmus, de~Siqueira, Grainger, Hannah, Hughes, Huntley, van
-  Jaarsveld, Midgley, Miles, Ortega-Huerta, Townsend~Peterson, Phillips \&
-  Williams}]{thomas2004}
-46.\\
-Thomas, C.~D., Cameron, A., Green, R.~E., Bakkenes, M., Beaumont, L.~J.,
-  Collingham, Y.~C., Erasmus, B. F.~N., de~Siqueira, M.~F., Grainger, A.,
-  Hannah, L., Hughes, L., Huntley, B., van Jaarsveld, A.~S., Midgley, G.~F.,
-  Miles, L., Ortega-Huerta, M.~A., Townsend~Peterson, A., Phillips, O.~L. \&
-  Williams, S.~E. (2004).
-\newblock Extinction risk from climate change.
-\newblock \emph{Nature}, 427, 145--148.
-
-\bibitem[{Thorson \emph{et~al.}(2013)Thorson, Scheuerell, Buhle \&
-  Copeland}]{thorson2013}
-47.\\
-Thorson, J.~T., Scheuerell, M.~D., Buhle, E.~R. \& Copeland, T. (2013).
-\newblock Spatial variation buffers temporal fluctuations in early juvenile
-  survival for an endangered {Pacific} salmon.
-\newblock \emph{J. Anim. Ecol.}, In press. doi:10.1111/1365-2656.12117.
-
-\bibitem[{Walther \emph{et~al.}(2002)Walther, Post, Convey, Menzel, Parmesan,
-  Beebee, Fromentin, Hoegh-Guldberg \& Bairlein}]{walther2002}
-48.\\
-Walther, G.-R., Post, E., Convey, P., Menzel, A., Parmesan, C., Beebee, T.
-  J.~C., Fromentin, J.-M., Hoegh-Guldberg, O. \& Bairlein, F. (2002).
-\newblock Ecological responses to recent climate change.
-\newblock \emph{Nature}, 416, 389--395.
-
-\bibitem[{Waples \emph{et~al.}(2009)Waples, Beechie \& Pess}]{waples2009}
-49.\\
-Waples, R., Beechie, T. \& Pess, G.~R. (2009).
-\newblock Evolutionary history, habitat disturbance regimes, and anthropogenic
-  changes: what do these mean for resilience of {Pacific} salmon populations?
-\newblock \emph{Ecol. Soc.}, 14, 3.
-
-\bibitem[{Yeakel \emph{et~al.}(2013)Yeakel, Moore, Guimar{\~a}es \&
-  de~Aguiar}]{yeakel2013}
-50.\\
-Yeakel, J.~D., Moore, J.~W., Guimar{\~a}es, P.~R. \& de~Aguiar, M. A.~M.
-  (2013).
-\newblock Synchronisation and stability in river metapopulation networks.
-\newblock \emph{Ecol. Lett.}, In press. doi: 10.1111/ele.12228.
-
-\end{thebibliography}
-
-\end{spacing}
-
-\begin{spacing}{1.4}
+\bibliographystyle{gcb}
+\bibliography{jshort,ms}
 
@@
 
-\section{Figures}
+\section{Figure legends}
 
-\begin{center}
-\includegraphics[height=5.5in]{simulation-diagram3}
-\end{center}
-Fig.\ 1: Flow chart of the salmon-metapopulation simulation.
+%\begin{center}
+%\includegraphics[height=5.5in]{../examples/simulation-diagram3}
+%\end{center}
+\textbf{Fig.\ 1} Flow chart of the salmon-metapopulation simulation.
 There are $n$ salmon populations and $t$ generations.
@@
 
-\clearpage
+%\clearpage
 
-\begin{center}
-\includegraphics[width=2.9in]{thermal-curve-scenarios}
-\end{center}
-Fig.\ 2: Different ways of prioritizing response-diversity conservation.
+%\begin{center}
+%\includegraphics[width=2.9in]{../examples/thermal-curve-scenarios}
+%\end{center}
+\textbf{Figure 2.} Different ways of prioritizing response-diversity conservation.
 Panel a shows thermal tolerance curves for ten possible populations and panels b--e show different ways of prioritizing four of those populations.
@@
 
-\clearpage
+%\clearpage
 
-\begin{center}
-\includegraphics[width=4.0in]{spatial-arma-sim-full}
-\end{center}
-Fig.\ 3: The components of an example metapopulation simulation.
+%\begin{center}
+%\includegraphics[width=4.0in]{../examples/spatial-arma-sim-full}
+%\end{center}
+\textbf{Figure 3.} The components of an example metapopulation simulation.
 We show, from top to bottom, the temperature signal, the resulting productivity parameter (Ricker $a$), the salmon returns, fisheries catch, salmon escapement, salmon straying from their natal streams, salmon joining from other streams, spawner-return residuals on a log scale, and the estimated $a$ and $b$ parameters in the fitted Ricker curve.
@@
 
-\clearpage
+%\clearpage
 
-\begin{center}
-\includegraphics[width=4.5in]{spatial-mv}
-\end{center}
-Fig.\ 4: The importance of preserving environmental response diversity through spatial conservation strategies.
-The conservation strategies correspond to Fig.\ 2 and represent conserving a range of responses (green), the most stable populations only (orange), or one type of environmental response (purple and pink).
+%\begin{center}
+%\includegraphics[width=4.5in]{../examples/spatial-mv}
+%\end{center}
+\textbf{Figure 4.} The importance of preserving environmental response diversity through spatial conservation strategies.
+The conservation strategies correspond to Figure 2 and represent conserving a range of responses (green), the most stable populations only (orange), or one type of environmental response (purple and pink).
 In risk-return space we show environmental scenarios that are comprised primarily of (a) short-term and (b) long-term environmental fluctuations.
@@
 
-\clearpage
+%\clearpage
 
-\begin{center}
-\includegraphics[width=4.5in]{cons-plans-n}
-\end{center}
-Fig.\ 5: The importance of preserving as many populations as possible when we don't know how response diversity is distributed.
+%\begin{center}
+%\includegraphics[width=4.5in]{../examples/cons-plans-n}
+%\end{center}
+\textbf{Figure 5.} The importance of preserving as many populations as possible when we don't know how response diversity is distributed.
 In risk-return space we show environmental scenarios that are comprised primarily of (a) short-term and (b) long-term environmental fluctuations.
@@
 
+\clearpage
+
+\section{Supporting information}
+
+\textbf{Supporting code} Instructions for installing and using the \texttt{metafolio} \texttt{R} package.
+
+\textbf{Table \ref{t:pars}}
+Input parameters to the salmon metapopulation simulation with default values.
+
+\textbf{Fig.\ \ref{f:stray}}
+An example straying matrix.
+
+\textbf{Fig.\ \ref{f:eg-sens}}
+The impact of increasing or decreasing various parameter values on metapopulation return abundance.
+
+\textbf{Fig.\ \ref{f:eg-sp-arma-full}}
+Conserving a full range of response diversity (spatial conservation strategy) with short-term environmental fluctuations.
+
+\textbf{Fig.\ \ref{f:eg-sp-arma-half}}
+Conserving one half of response diversity (spatial conservation strategy) with short-term environmental fluctuations.
+
+\textbf{Fig.\ \ref{f:ret-corr}}
+A comparison of the log(returns) between populations.
+
+\textbf{Fig.\ \ref{f:eg-sp-linear-full}}
+Conserving a full range of response diversity (spatial conservation strategy) with long-term environmental change.
+
+\textbf{Fig.\ \ref{f:eg-sp-linear-half}}
+Conserving one half of response diversity (spatial
+conservation strategy) with long-term environmental change.
+
+\textbf{Fig.\ \ref{f:eg-n-arma-two}}
+Two populations conserved with random response diversity and
+short-term environmental fluctuations.
+
+\textbf{Fig.\ \ref{f:eg-n-arma-sixteen}}
+Sixteen populations conserved with random response diversity
+and short-term environmental fluctuations.
+
+\textbf{Fig.\ \ref{f:eg-n-linear-two}}
+Two populations conserved with random response diversity and
+long-term environmental change.
+
+\textbf{Fig.\ \ref{f:eg-n-linear-sixteen}}
+Sixteen populations conserved with random response diversity
+and long-term environmental change.
+
 \end{spacing}
@@
 
-\begin{spacing}{1.2}
+\begin{spacing}{1.0}
 

### 2014-01-08 inst/ms/salmonfolios.tex
Updates for prsb
A variety of changes prior to prsb submission

@@
 \usepackage{setspace}
-\usepackage[round]{natbib}
-\bibpunct{(}{)}{;}{a}{}{;}
-\textheight 20.5cm
+%\usepackage[round]{natbib}
+%\bibpunct{(}{)}{;}{a}{}{;}
+\textheight 20.0cm
 \usepackage{lineno}
@@
 \usepackage{booktabs}
+\usepackage{cite}
 
@@
 
-\widowpenalty=1000
-\clubpenalty=1000
+\widowpenalty=10000
+\clubpenalty=10000
 
@@
 \begin{document}
-\raggedright
+%\raggedright
 
@@
 
-\bibliographystyle{gcb}
+\bibliographystyle{prsb}
 \bibliography{jshort,ms}
@@
 
-\section{Supporting information}
-
-\textbf{Supporting code} Instructions for installing and using the \texttt{metafolio} \texttt{R} package.
-
-\textbf{Table \ref{t:pars}}
-Input parameters to the salmon metapopulation simulation with default values.
-
-\textbf{Fig.\ \ref{f:stray}}
-An example straying matrix.
-
-\textbf{Fig.\ \ref{f:eg-sens}}
-The impact of increasing or decreasing various parameter values on metapopulation return abundance.
-
-\textbf{Fig.\ \ref{f:eg-sp-arma-full}}
-Conserving a full range of response diversity (spatial conservation strategy) with short-term environmental fluctuations.
-
-\textbf{Fig.\ \ref{f:eg-sp-arma-half}}
-Conserving one half of response diversity (spatial conservation strategy) with short-term environmental fluctuations.
-
-\textbf{Fig.\ \ref{f:ret-corr}}
-A comparison of the log(returns) between populations.
-
-\textbf{Fig.\ \ref{f:eg-sp-linear-full}}
-Conserving a full range of response diversity (spatial conservation strategy) with long-term environmental change.
-
-\textbf{Fig.\ \ref{f:eg-sp-linear-half}}
-Conserving one half of response diversity (spatial
-conservation strategy) with long-term environmental change.
-
-\textbf{Fig.\ \ref{f:eg-n-arma-two}}
-Two populations conserved with random response diversity and
-short-term environmental fluctuations.
-
-\textbf{Fig.\ \ref{f:eg-n-arma-sixteen}}
-Sixteen populations conserved with random response diversity
-and short-term environmental fluctuations.
-
-\textbf{Fig.\ \ref{f:eg-n-linear-two}}
-Two populations conserved with random response diversity and
-long-term environmental change.
-
-\textbf{Fig.\ \ref{f:eg-n-linear-sixteen}}
-Sixteen populations conserved with random response diversity
-and long-term environmental change.
-
 \end{spacing}

### 2014-01-13 src/metasim_base.cpp
Fix examples and exports and dynlib in .cpp file

@@ -27,6 +27,7 @@ NumericVector est_beta_params(double mu, double var) {
 //' A single numeric values representing a sample from a beta
 //' distribution with the specified mean and standard deviation.
 //'
+//' @export
 //' @references
 //' Morgan, M. G. & Henrion, M. (1990). Uncertainty: A Guide to Dealing
 //' with Uncertainty in Quantitative Risk and Policy Analysis.
@@ -64,29 +65,33 @@ NumericVector impl_error(NumericVector mu, double sigma_impl) {
 
 //' Ricker stock-recruit function with specified error
 //'
-//' @param spawners Spawner abundance
+//' @param spawners A single spawner abundance
 //' @param a Ricker productivity parameter. Recruits are e^a at the origin.
 //' @param b Ricker density dependent parameter.
 //' @param d Depensation parameter. A value of 1 means no depensation. Larger
 //'   values indicate depensation.
-//' @param v_t Residual on the curve. Will be exponentiated. Note that we are
+//' @param v_t A single residual on the curve. Will be exponentiated. Note that we are
 //'   *not* bias correcting within this function (subtracting half the variance
 //'   squared) and so the deviations will not be mean unbiased unless they were
 //'   bias corrected previously.
+//' @export
 //' @return Returns a vector of recruits.
 //' @examples
-//' x <- seq(1, 100)
-//' v_t <- as.numeric(arima.sim(n = 100, model = list(ar = 0.3), sd =
-//'     0.1, mean = 0))
-//' plot(x, ricker_v_t(spawners = x, a = 1.1, b = 60, d = 1, v_t), xlab =
-//'   "Spawners", ylab = "Returns")
-//'
+//' plot(1, 1, xlim = c(1, 100), ylim = c(0, 90), type = "n", xlab = "Spawners",
+//'   ylab = "Returns")
+//' for(i in 1:100) {
+//' points(i, ricker_v_t(i, a = 1.1, b = 60, d = 1, v_t = rnorm(1, mean =
+//'   -(0.1^2)/2, sd = 0.1)))
+//' }
 // [[Rcpp::export]]
 double ricker_v_t(double spawners, double a, double b, double d, double v_t) {
   return pow(spawners, d) * exp(a * (1 - pow(spawners, d) / b) + v_t);
 }
 
-//' Is element
+//' Check if x is an element of y.
+//'
[…]

### 2014-01-13 R/plot_sim_ts.R
Minor fixes to R function examples

@@ -1,4 +1,5 @@
 #' Add annotations to panel
+#'
 #' @param label The text to add as a label
 #' @param xfrac Fraction over from the left
 #' @param yfrac Fraction down from the top
@@ -13,6 +14,7 @@ annotate <- function(label, xfrac = 0.008, yfrac = 0.18, pos = 4, cex = 0.9, ...
 }
 
 #' Add a pretty axis
+#'
 #' @param side Number indicating the side to add an axis (as in the side number
 #' passed to \code{\link[graphics]{axis}}).
 #' @param shade_years An optional numerical vector of length two giving the
@@ -56,14 +58,13 @@ plot_panel_lines <- function(dat, ymin = c("zero", "min"), ystretch = 1.1, ...) 
 
 #' Plot various time series from a simulation run
 #'
-#' This function is rather crude at the moment. It lets you quickly visualize
-#' the time series of output from a simulation run, but it isn't very flexible
-#' yet.
+#' This function lets you quickly visualize the time series of output from a
+#' simulation run.
 #'
 #' @param x A list output object from a simulation run of
 #'   \code{link{meta_sim}}.
-#' @param pal A colour palette for the lines. One colour per line (population
-#' time series).
+#' @param pal A colour palette for the lines. One colour per line (each
+#' line is a population time series).
 #' @param years_to_show How many years to plot after the burn in period.
 #' @param burn The number of years to discard as burn in at the beginning of
 #'   the time series.

### 2014-01-17 inst/ms/anderson-etal-metafolio.tex
Add anderson-etal-metafolio, which was ignored

--- 2014-01-08 inst/ms/salmonfolios.tex
@@
 \usepackage{cite}
+
+\newlabel{t:pars}{{S1}{2}}
+\newlabel{f:stray}{{S1}{3}}
+\newlabel{f:eg-sim-ts}{{S2}{4}}
+\newlabel{f:eg-sens}{{S3}{5}}
+\newlabel{f:eg-sp-arma-full}{{S4}{6}}
+\newlabel{f:eg-sp-arma-half}{{S5}{7}}
+\newlabel{f:ret-corr}{{S6}{8}}
+\newlabel{f:eg-sp-linear-full}{{S7}{9}}
+\newlabel{f:eg-sp-linear-half}{{S8}{10}}
+\newlabel{f:eg-n-arma-two}{{S9}{11}}
+\newlabel{f:eg-n-arma-sixteen}{{S10}{12}}
+\newlabel{f:eg-n-linear-two}{{S11}{13}}
+\newlabel{f:eg-n-linear-sixteen}{{S12}{14}}
 
@@
 \bibliographystyle{prsb}
-\bibliography{jshort,ms}
+%\bibliography{jshort,ms}
+\input{anderson-etal-metafolio.bbl}
 
@@
 %\end{center}
-\textbf{Fig.\ 1} Flow chart of the salmon-metapopulation simulation.
+%\textbf{Figure 1.} Flow chart of the salmon-metapopulation simulation.
 There are $n$ salmon populations and $t$ generations.
@@
 %\end{center}
-\textbf{Figure 2.} Different ways of prioritizing response-diversity conservation.
+\textbf{Figure 1.} Different ways of prioritizing response-diversity conservation.
 Panel a shows thermal tolerance curves for ten possible populations and panels b--e show different ways of prioritizing four of those populations.
@@
 %\end{center}
-\textbf{Figure 3.} The components of an example metapopulation simulation.
+%\textbf{Figure 3.} The components of an example metapopulation simulation.
 We show, from top to bottom, the temperature signal, the resulting productivity parameter (Ricker $a$), the salmon returns, fisheries catch, salmon escapement, salmon straying from their natal streams, salmon joining from other streams, spawner-return residuals on a log scale, and the estimated $a$ and $b$ parameters in the fitted Ricker curve.
@@
 %\end{center}
-\textbf{Figure 4.} The importance of preserving environmental response diversity through spatial conservation strategies.
-The conservation strategies correspond to Figure 2 and represent conserving a range of responses (green), the most stable populations only (orange), or one type of environmental response (purple and pink).
+\textbf{Figure 2.} The importance of preserving environmental response diversity through spatial conservation strategies.
+The conservation strategies correspond to figure 2 and represent conserving a range of responses (green), the most stable populations only (orange), or one type of environmental response (purple and pink).
 In risk-return space we show environmental scenarios that are comprised primarily of (a) short-term and (b) long-term environmental fluctuations.
@@
 %\end{center}
-\textbf{Figure 5.} The importance of preserving as many populations as possible when we don't know how response diversity is distributed.
+\textbf{Figure 3.} The importance of preserving as many populations as possible when we don't know how response diversity is distributed.
 In risk-return space we show environmental scenarios that are comprised primarily of (a) short-term and (b) long-term environmental fluctuations.
@@
 We also show example metapopulation (c) rate-of-change and (d) abundance time series for the short-term environmental-fluctuation scenario.
-The grey line (a, b) indicates the efficient frontier across all simulated metapopulations --- metapopulations with the minimum variability for a given level of growth rate
+The grey line (a, b) indicates the efficient frontier across all simulated metapopulations --- metapopulations with the minimum variability for a given level of growth rate.
 
@@
 
-\setlength{\parskip}{8pt}
-\setlength{\parindent}{0cm}
+%\setlength{\parskip}{8pt}
+%\setlength{\parindent}{0cm}
 
-\begin{spacing}{1.0}
-
-\include{som}
-
-\end{spacing}
+% \begin{spacing}{1.1}
+%
+% \setcounter{page}{1}
+% \nolinenumbers
+% \include{som}
+%
+% \end{spacing}
 

### 2014-02-06 inst/ms/anderson-etal-metafolio.tex
Final edits before branching
Update files for Ecol Appl

--- 2014-01-08 inst/ms/salmonfolios.tex
@@
-\documentclass[11pt]{article}
+\documentclass[12pt]{article}
 \usepackage{geometry}
+\geometry{verbose,letterpaper,tmargin=2.54cm,bmargin=2.54cm,lmargin=2.54cm,rmargin=2.54cm} 
 \geometry{letterpaper}
@@
 \usepackage{setspace}
-%\usepackage[round]{natbib}
-%\bibpunct{(}{)}{;}{a}{}{;}
-\textheight 20.0cm
+\usepackage[round]{natbib}
+\bibpunct{(}{)}{;}{a}{}{;}
+\textheight 22.0cm
 \usepackage{lineno}
@@
 
+\newlabel{f:flowchart}{{1}{999}}
+\newlabel{f:curves}{{2}{999}}
+\newlabel{f:ts}{{3}{999}}
+\newlabel{f:sp}{{4}{999}}
+\newlabel{f:n}{{5}{999}}
+
+\newcommand{\somR}{Appendix A}
+\newcommand{\somparam}{Appendix B}
+\newcommand{\somstray}{Appendix C}
+\newcommand{\somsens}{Appendix D}
+\newcommand{\somcor}{Appendix E}
+\newcommand{\somts}{Appendix F}
+
+% SOM:
+\newlabel{t:pars}{{B1}{2}}
+\newlabel{f:stray}{{B1}{3}}
+\newlabel{f:eg-sens}{{D1}{4}}
+\newlabel{f:ret-corr}{{E1}{5}}
+\newlabel{f:eg-sp-arma-full}{{F1}{6}}
+\newlabel{f:eg-sp-arma-half}{{F2}{7}}
+\newlabel{f:eg-sp-linear-full}{{F3}{8}}
+\newlabel{f:eg-sp-linear-half}{{F4}{9}}
+\newlabel{f:eg-n-arma-two}{{F5}{10}}
+\newlabel{f:eg-n-arma-sixteen}{{F6}{11}}
+\newlabel{f:eg-n-linear-two}{{F7}{12}}
+\newlabel{f:eg-n-linear-sixteen}{{F8}{13}}
+
 \DeclareGraphicsExtensions{.pdf}
@@
 
-\usepackage{titlesec}
-\titlespacing\section{0pt}{6pt plus 4pt minus 2pt}{-8pt plus 2pt minus 2pt}
-\titlespacing\subsection{0pt}{6pt plus 4pt minus 2pt}{-8pt plus 2pt minus 2pt}
+%\usepackage{titlesec}
+%\titlespacing\section{0pt}{6pt plus 4pt minus 2pt}{-8pt plus 2pt minus 2pt}
+%\titlespacing\subsection{0pt}{6pt plus 4pt minus 2pt}{-8pt plus 2pt minus 2pt}
+
+\title{Portfolio conservation of metapopulations under\\climate change}
+
+\author{
+Sean C. Anderson$^{1\ast}$ \and Jonathan W. Moore$^{1,2}$ \and Michelle M. McClure$^3$ \and
+Nicholas K. Dulvy$^1$ \and Andrew B. Cooper$^2$
+}
+\date{}
 
@@
 \begin{document}
-%\raggedright
+\raggedright
 
 \linenumbers
-\modulolinenumbers[2]
-\begin{spacing}{1.6}
-\setlength{\parindent}{0cm}
+
+%\modulolinenumbers[2]
+\begin{spacing}{1.9}
+\setlength{\parindent}{0.9cm}
+
+\maketitle
+
 \input{ms}
 
-\bibliographystyle{prsb}
+\bibliographystyle{ecology3}
 \bibliography{jshort,ms}
+%\input{anderson-etal-metafolio.bbl}
+
+\clearpage
+
+\section{Supplemental Material}
+
+\noindent
+\textsc{Appendix A.} The \texttt{metafolio} \textsf{R} package.
+
+\noindent
+\textsc{Appendix B.} Simulation input parameters and default values.
+
+\noindent
+\textsc{Appendix C.} An example straying matrix.
+
+\noindent
+\textsc{Appendix D.} Sensitivity illustration with alternative parameter values.
+
+\noindent
+\textsc{Appendix E.} An illustration of the correlation between populations.
+
+\noindent
+\textsc{Appendix F.} Example simulated time series from alternative conservation scenarios.
 
@@
 %\begin{center}
-%\includegraphics[height=5.5in]{../examples/simulation-diagram3}
 %\end{center}
-\textbf{Fig.\ 1} Flow chart of the salmon-metapopulation simulation.
+\textsc{Fig.
+1}.
+Flow chart of the salmon-metapopulation simulation.
 There are $n$ salmon populations and $t$ generations.
@@
 %\end{center}
-\textbf{Figure 2.} Different ways of prioritizing response-diversity conservation.
+\bigskip
+\noindent
+\textsc{Fig.
+2}.
+Different ways of prioritizing response-diversity conservation.
 Panel a shows thermal tolerance curves for ten possible populations and panels b--e show different ways of prioritizing four of those populations.
@@
 %\end{center}
-\textbf{Figure 3.} The components of an example metapopulation simulation.
+\bigskip
+\noindent
+\textsc{Fig.
+\ref{f:ts}}.
+The components of an example metapopulation simulation.
 We show, from top to bottom, the temperature signal, the resulting productivity parameter (Ricker $a$), the salmon returns, fisheries catch, salmon escapement, salmon straying from their natal streams, salmon joining from other streams, spawner-return residuals on a log scale, and the estimated $a$ and $b$ parameters in the fitted Ricker curve.
@@
 %\end{center}
-\textbf{Figure 4.} The importance of preserving environmental response diversity through spatial conservation strategies.
-The conservation strategies correspond to Figure 2 and represent conserving a range of responses (green), the most stable populations only (orange), or one type of environmental response (purple and pink).
+\bigskip
+\noindent
+\textsc{Fig.
+4}.
+The importance of preserving environmental response diversity through spatial conservation strategies.
+The conservation strategies correspond to figure 2 and represent conserving a range of responses (green), the most stable populations only (orange), or one type of environmental response (purple and pink).
 In risk-return space we show environmental scenarios that are comprised primarily of (a) short-term and (b) long-term environmental fluctuations.
@@
 %\end{center}
-\textbf{Figure 5.} The importance of preserving as many populations as possible when we don't know how response diversity is distributed.
+\bigskip
+\noindent
+\textsc{Fig.
+5}.
+The importance of preserving as many populations as possible when we don't know how response diversity is distributed.
 In risk-return space we show environmental scenarios that are comprised primarily of (a) short-term and (b) long-term environmental fluctuations.
@@
 
+\section{Figures}
+
+\begin{center}
+\includegraphics[height=5.5in]{../examples/simulation-diagram3}\\
+\textsc{Fig.} 1
+\clearpage
+\includegraphics[width=2.9in]{../examples/thermal-curve-scenarios}\\
+\textsc{Fig.} 2
+\clearpage
+\includegraphics[width=4.0in]{../examples/spatial-arma-sim-full}\\
+\textsc{Fig.} 3
+\clearpage
+\includegraphics[width=4.5in]{../examples/spatial-mv}\\
+\textsc{Fig.} 4
+\clearpage
+\includegraphics[width=4.5in]{../examples/cons-plans-n}\\
+\textsc{Fig.} 5
+\clearpage
+\end{center}
+
 \end{spacing}
 
-\setlength{\parskip}{8pt}
-\setlength{\parindent}{0cm}
+%\setlength{\parskip}{8pt}
+%\setlength{\parindent}{0cm}
 
-\begin{spacing}{1.0}
+ \begin{spacing}{1.1}
 
+ \setcounter{page}{1}
+ \nolinenumbers
  \include{som}

### 2014-02-06 inst/examples/cons_plans2.R

@@ -26,7 +26,7 @@ for(i in 1:n_plans) { # loop over plans
 arma_env_params <- list(mean_value = 16, ar = 0.1, sigma_env = 2, ma = 0)
 
 pdf_eps("spatial-arma-sim-full", width = 5, height = 7, type = TYPE)
-par(family = "Times")
+#par(family = "Times")
 eg_arma <- meta_sim(b = w[[1]][[1]], n_pop = 10, env_params = arma_env_params,
   env_type = "arma", assess_freq = 5)
 plot_sim_ts(eg_arma, years_to_show = 70, burn = 30, yticks = list(NA, NA, NA,
@@ -95,7 +95,7 @@ xlim <- c(0.02, 0.15)
 ylim <- c(-0.017, 0.017)
 par(las = 1, cex = 0.8, mar = c(0, 0, 0, 0), oma = c(4, 5.2, 1.5, .5),
   tck = -0.02, mgp = c(2, .6, 0))
-par(family = "Times")
+#par(family = "Times")
 plot_cons_plans(x_arma_sp$plans_mv, plans_name = plans_name_sp, cols = cols,
   add_all_efs = FALSE, xlim = xlim, ylim = ylim, add_legend = FALSE)
 add_inset_env(eg_arma$env_ts[-c(1:30)], x = 0.12, y = -0.013, size = c(1, .5))

### 2014-02-06 inst/examples/cons_plans_n_random_resp_div.R

@@ -104,7 +104,7 @@ layout(rbind(
 
 xlim <- c(0.008, 0.15)
 ylim <- c(-0.017, 0.017)
-par(family = "Times")
+#par(family = "Times")
 par(las = 1, cex = 0.8, mar = c(0, 0, 0, 0), oma = c(4, 5.2, 1.5, .5),
   tck = -0.02, mgp = c(2, .6, 0))
 plot_cons_plans(x_arma_n$plans_mv, plans_name = plans_name_n, cols = cols,
@@ -127,7 +127,7 @@ add_inset_env(eg_linear$env_ts[-c(1:30)], x = 0.12, y = -0.013, size = c(1, .5))
 mtext("(b) Long-term environmental change", side = 3, line = 0.2, cex = 0.8,
   adj = 0.05)
 mtext("Variance of metapopulation growth rate", side = 1, line = 2.25,
-  outer = FALSE, cex = 0.8, adj = -2.1)
+  outer = FALSE, cex = 0.8, adj = -3)
 
 ## time series plots:
 par(tck = -0.035)

### 2014-02-06 inst/examples/plot_thermal_curves_scenarios.R

@@ -3,7 +3,7 @@
 pdf_eps("thermal-curve-scenarios", width = 2.6, height = 4.7, type = TYPE)
 par(cex = 0.8, mgp = c(2.4, 0.55, 0), tck = -0.04, mfrow = c(5,1), mar = c(0,3,0,.5), oma = c(3, 0, .5, 0))
     par(las = 1)
-par(family = "Times")
+#par(family = "Times")
 
 par(xpd = FALSE)
 optim_temps <- seq(13, 19, length.out = 10)

### 2014-06-18 inst/examples/README.R
Annotate README a bit more

@@ -2,7 +2,10 @@
 
 # Set your R working directory to this folder (the examples folder).
 
-# Install and load the package and then:
+# Install:
+# devtools::install("../../")
+
+# and load the package:
 library(metafolio)
 
 # or, for rapid development:
@@ -22,10 +25,14 @@ col_pal <- rev(gg_color_hue(10))
 source("pdf_eps.R")
 TYPE <- "pdf" # or "pdf" for plot output
 
+# Conserve random response diversity; control N:
 source("cons_plans_n_random_resp_div.R")
 
 # Figure showing different spatial rules of thumb:
 source("cons_plans2.R")
+
+# Conservation plans with absolute reductions in capacity:
+source("cons_squeeze.R")
 
 # Figure showing thermal curves for spatial scenarios:
 source("plot_thermal_curves_scenarios.R")

### 2014-06-18 inst/ms/anderson-etal-metafolio.tex
Updates to main .tex wrapper

@@
-\documentclass[12pt]{article}
+\documentclass[12pt,english]{article}
 \usepackage{geometry}
@@
 \usepackage{xcolor}
-\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
+%\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
 \usepackage{booktabs}
 \usepackage{cite}
+\usepackage[originalparameters]{ragged2e}
+\usepackage{array}
+%\usepackage{mathptmx}
 
@@
 \newlabel{f:n}{{5}{999}}
+\newlabel{f:squeeze}{{6}{999}}
 
@@
 
+\newcommand{\R}[1]{\label{#1}\linelabel{#1}}
+\newcommand{\lr}[1]{line~\lineref{#1}}
+
 \begin{document}
@@
 
+\begin{spacing}{1.1}
+\setlength\parskip{0.11in}
+\setlength\parindent{0in}
+\include{response}
+\setlength\parskip{0in}
+\end{spacing}
+
+\setcounter{page}{1}
 \linenumbers
@@
 \begin{spacing}{1.9}
+%\begin{spacing}{1.2}
 \setlength{\parindent}{0.9cm}
@@
 %\end{center}
-\textsc{Fig.
-1}.
-Flow chart of the salmon-metapopulation simulation.
-There are $n$ salmon populations and $t$ generations.
-Blue text indicates values that are generated before the simulation progresses through time.
-Red text indicates steps in which calculations are performed through time.
-Black text indicates values that are calculated.
-Grey text indicates parameters that can be set.
-Green text indicates the looping structure of the simulation.
+\textsc{Fig. 1}. Flow chart of the salmon-metapopulation simulation.
 
@@
 The curves describe how productivity varies with temperature for a given population.
-Some populations thrive at low temperatures (cool colours) and some at warm temperatures (warm colours).
-Some are tolerant to a wider range of environmental conditions (yellow-to-green colours) but with a lower maximum productivity.
+Some populations thrive at low temperatures (light greys) and some at warm temperatures (dark greys).
+Some are tolerant to a wider range of environmental conditions (mid greys) but with a lower maximum productivity.
 The total possible productivity (the area under the curves) is the same for each population.
@@
 We show, from top to bottom, the temperature signal, the resulting productivity parameter (Ricker $a$), the salmon returns, fisheries catch, salmon escapement, salmon straying from their natal streams, salmon joining from other streams, spawner-return residuals on a log scale, and the estimated $a$ and $b$ parameters in the fitted Ricker curve.
-The colored lines indicate populations that thrive at low (cool colours) to high (warm colours) temperatures.
+The shaded lines indicate populations that thrive at low (light grey) to high (dark grey) temperatures.
 
@@
 We also show example metapopulation abundance time series for the (c, e) short-term and (d, f) long-term environmental-fluctuation scenarios.
-The grey line (a, b) indicates the efficient frontier across all simulated metapopulations --- metapopulations with the minimum variability for a given level of growth rate.
+The thick grey line (a, b) indicates the efficient frontier across all simulated metapopulations --- metapopulations with the minimum variability for a given level of growth rate.
 
@@
 In risk-return space we show environmental scenarios that are comprised primarily of (a) short-term and (b) long-term environmental fluctuations.
-We show metapopulations in which 2 (red), 4 (orange), 8 (yellow), or 16 (green) populations of random response diversity are conserved.
+We show metapopulations in which two through 16 populations of random response diversity are conserved.
 The dots show simulated metapopulations and the contours show 25\% and 75\% quantiles across 500 simulations per strategy.
 We also show example metapopulation (c) rate-of-change and (d) abundance time series for the short-term environmental-fluctuation scenario.
-The grey line (a, b) indicates the efficient frontier across all simulated metapopulations --- metapopulations with the minimum variability for a given level of growth rate.
+The thick grey line (a, b) indicates the efficient frontier across all simulated metapopulations --- metapopulations with the minimum variability for a given level of growth rate.
+
+\bigskip
+\noindent
+\textsc{Fig.
+6}.
+Risk-return trade-off in the case where habitat is lost over time through stream flow reduction.
+The temperature follows both short-term fluctuations and a long-term increase.
+Response diversity is randomly conserved.
+Shading indicates conservation plans where two through 16 populations are conserved.
+(a) Conserving more populations decreases expected variance but also decreases expected growth rate.
+Dots show simulated metapopulations and contours show 25\% and 75\% quantiles across 500 simulations per strategy.
+The thick grey line indicates the efficient frontier across all simulated metapopulations --- metapopulations with the minimum variability for a given level of growth rate.
+Also shown are (b) example metapopulation growth rate and (c) abundance time series from the 2 and 16 population scenarios.
+Regression lines in (b) illustrate a decreasing growth rate through time.
 
@@
 \begin{center}
-\includegraphics[height=5.5in]{../examples/simulation-diagram3}\\
+\includegraphics[height=5.5in]{../examples/simulation-diagram3-sansserif-bw}\\
 \textsc{Fig.} 1
@@
 \textsc{Fig.} 5
+\clearpage
+\includegraphics[width=3.0in]{../examples/cons-plans-squeeze}\\
+\textsc{Fig.} 6
 \clearpage

### 2014-06-24 inst/ms/anderson-etal-metafolio.tex

@@
 2}.
-Different ways of prioritizing response-diversity conservation.
-Panel a shows thermal tolerance curves for ten possible populations and panels b--e show different ways of prioritizing four of those populations.
+Different ways of prioritizing thermal-tolerance conservation.
+Panel a shows thermal-tolerance curves for ten possible populations and panels b--e show different ways of prioritizing four of those populations.
 The curves describe how productivity varies with temperature for a given population.
@@
 4}.
-The importance of preserving environmental response diversity through spatial conservation strategies.
+The importance of preserving thermal-tolerance diversity through spatial conservation strategies.
 The conservation strategies correspond to figure 2 and represent conserving a range of responses (green), the most stable populations only (orange), or one type of environmental response (purple and pink).
@@
 5}.
-The importance of preserving as many populations as possible when we don't know how response diversity is distributed.
+The importance of preserving as many populations as possible when we do not know how thermal-tolerance is distributed.
 In risk-return space we show environmental scenarios that are comprised primarily of (a) short-term and (b) long-term environmental fluctuations.
-We show metapopulations in which two through 16 populations of random response diversity are conserved.
+We show metapopulations in which two through 16 populations are conserved.
 The dots show simulated metapopulations and the contours show 25\% and 75\% quantiles across 500 simulations per strategy.
@@
 The temperature follows both short-term fluctuations and a long-term increase.
-Response diversity is randomly conserved.
+Thermal tolerance is randomly conserved.
 Shading indicates conservation plans where two through 16 populations are conserved.
@@
 \clearpage
-\includegraphics[width=2.9in]{../examples/thermal-curve-scenarios}\\
+\includegraphics[width=2.4in]{../examples/thermal-curve-scenarios}\\
 \textsc{Fig.} 2
 \clearpage
-\includegraphics[width=4.0in]{../examples/spatial-arma-sim-full}\\
+\includegraphics[width=4.2in]{../examples/spatial-arma-sim-full}\\
 \textsc{Fig.} 3

### 2014-06-24 inst/ms/salmonfolios.tex

@@
 \usepackage{cite}
+
+\newlabel{t:pars}{{S1}{3}}
+\newlabel{f:stray}{{S1}{4}}
+\newlabel{f:eg-sens}{{S2}{5}}
+\newlabel{f:eg-sp-arma-full}{{S3}{6}}
+\newlabel{f:eg-sp-arma-half}{{S4}{7}}
+\newlabel{f:ret-corr}{{S5}{8}}
+\newlabel{f:eg-sp-linear-full}{{S6}{9}}
+\newlabel{f:eg-sp-linear-half}{{S7}{10}}
+\newlabel{f:eg-n-arma-two}{{S8}{11}}
+\newlabel{f:eg-n-arma-sixteen}{{S9}{12}}
+\newlabel{f:eg-n-linear-two}{{S10}{13}}
+\newlabel{f:eg-n-linear-sixteen}{{S11}{14}}
 
@@
 \bibliography{jshort,ms}
+%\input{salmonfolios.bbl}
 
@@
 \textbf{Figure 4.} The importance of preserving environmental response diversity through spatial conservation strategies.
-The conservation strategies correspond to Figure 2 and represent conserving a range of responses (green), the most stable populations only (orange), or one type of environmental response (purple and pink).
+The conservation strategies correspond to figure 2 and represent conserving a range of responses (green), the most stable populations only (orange), or one type of environmental response (purple and pink).
 In risk-return space we show environmental scenarios that are comprised primarily of (a) short-term and (b) long-term environmental fluctuations.
@@
 
-\begin{spacing}{1.0}
+\begin{spacing}{1.1}
 
+\setcounter{page}{1}
+\nolinenumbers
 \include{som}

### 2014-06-24 inst/ms/tables.tex
Update files for Ecol Appl

@@
 
+% TODO check that these values are all still correct - some aren't
+
 \begin{table}[h!]
@@
 \caption{Input parameters to the salmon metapopulation simulation with default values.}
-\begin{tabular}{p{7.7cm}p{1.4cm}p{3.6cm}}
+\smallskip
+\begin{tabular}{>{\RaggedRight}p{7.8cm}p{1.1cm}p{2.5cm}>{\RaggedRight}p{3.3cm}}
 \toprule
-Description                                                          & Symbol                & Value \\
+Description                                                          & Symbol                & Value                  & Reference  \\
 \midrule
 
-\textit{Population dynamics parameters}                              &                       & \\
-Stock-recruit residual standard deviation (on log scale)             & $\sigma_r$            & 0.3  \\
-AR(1) serial correlation of stock-recruit residuals                  & $\rho_w$              & 0.4  \\
-Fraction of fish that stray from natal streams                       & $f_{\mathrm{stray}}$  & 0.02  \\
-Exponential rate of decay of straying with distance                  & $m$                   & 0.3  \\
+\bibpunct{}{}{;}{a}{}{}
+\textit{Population dynamics parameters}                              &                       &                        &             \\
+Stock-recruit residual standard deviation (on log scale)             & $\sigma_r$            & 0.7                    & \citep{thorson2014a}  \\
+AR(1) serial correlation of stock-recruit residuals                  & $\rho_w$              & 0.4                    & \citep{thorson2014a}  \\
+Fraction of fish that stray from natal streams                       & $f_{\mathrm{stray}}$  & 0.02                   & \citep{quinn2005} and references therin  \\
+Exponential rate of decay of straying with distance                  & $m$                   & 0.1                    & \citep{cooper1999}      \\
+Range of maximum productivities                                      & $a_i^{\mathrm{max}}$  & 2.2--2.9             &  \citep{dorner2008}   \\
 
 \noalign{\vskip 3mm}
-\textit{Environmental parameters}                                    &                       & \\
-Width of environmental-tolerance curves for populations $i$ 1 to $n$ & $W_i$                 & Sequence of 0.05 to 0.02 to 0.05\\
-Optimum environmental value for populations $i$ 1 to $n$             & $e_i^{\mathrm{opt}}$  & Sequence of 13 to 19\\
-Area under each environmental-tolerance curve in environmental units & $A$                   & 30\\
+\textit{Environmental parameters}                                    &                       &                        &             \\
+Width parameter for thermal-tolerance curves for populations $i$ 1 to $n$  (values generate widths in line with listed references) & $W_i$                 & 0.08--0.04--0.08       &  \citep{brett1952, eliason2011}           \\
+Optimum environmental value for populations $i$ 1 to $n$             & $e_i^{\mathrm{opt}}$  & 13--19                 &  \citep{eliason2011}           \\
+%Area under each environmental-tolerance curve in environmental units & $A$                  & 30                     &             \\
+Standard deviation of annual temperature fluctuations          & $\sigma_d$            & 2                      & \citep{eliason2011}      \\
+AR(1) autocorrelation of annual temperature fluctuations       & $\rho_e$              & 0.1                    &             \\
+Annual increase in stream temperature in degrees Celcius             & $\beta_e$             & 0.04                   & \citep{mantua2010}     \\
 
-Standard deviation of short-term environmental fluctuations          & $\sigma_d$            & 2 \\
-AR(1) autocorrelation of short-term environmental fluctuations       & $\rho_e$              & 0.1 \\
-Slope of long-term environmental signal                              & $\beta_e$             & 0.114 \\
+%TODO cite Mantua 2010 in main text
 
 \noalign{\vskip 3mm}
-\textit{Fishery parameters}                                          &                       & \\
-Standard deviation of beta distribution for implementation error     & $\sigma_{h}$          & 0.05  \\
-Frequency of assessment (years)                                      & $f_{\mathrm{assess}}$ & 5  \\
+\textit{Fishery parameters}                                          &                       &                        &             \\
+Standard deviation of beta distribution for implementation error     & $\sigma_{h}$          & 0.1                    & \citep{pestes2008}            \\
+Frequency of assessment (years)                                      & $f_{\mathrm{assess}}$ & 5                      &             \\
 \bottomrule
@@
 \clearpage
+\bibpunct{(}{)}{;}{a}{}{}

### 2014-07-02 inst/examples/cons_plans2.R
Add more illustrative case for Appendix E

@@ -45,6 +45,9 @@ plot_sim_ts(eg_arma, years_to_show = 100, burn = 30, yticks = list(NA, NA, NA,
     NA, NA, NA, NA, c(2, 0, 2), NA, NA))
 dev.off()
 
+eg_arma <- meta_sim(b = rep(1000, 10), n_pop = 10, env_params = arma_env_params,
+  env_type = "arma", assess_freq = 5)
+
 pdf_eps("example-return-correlations", width = 5, height = 5, type = TYPE)
 plot_correlation_between_returns(eg_arma)
 dev.off()

### 2014-07-02 vignettes/metafolio.R
Remove extra vignette files

@@ -1,120 +1,0 @@
-
-## @knitr setup, include=FALSE, cache=FALSE
-library(knitr)
-# set global chunk options
-opts_chunk$set(fig.path='figure/', fig.align='center', fig.show='hold')
-options(replace.assign=TRUE, width=90)
-opts_knit$set(out.format = "latex")
-opts_chunk$set(warning=FALSE, message=FALSE, comment=NA, tidy=FALSE,
-refresh=TRUE, cache=FALSE, autodep=TRUE)
-thm = knit_theme$get("github.css")
-knit_theme$set(thm)
-
-
-## @knitr my-load-package, include=FALSE
-library(metafolio)
-set.seed(1)
-
-
-## @knitr install-source, eval=FALSE
-## # OS X:
-## install.packages("metafolio_0.3.0.tgz", repos = NULL)
-## 
-## # Windows:
-## install.packages("metafolio_0.3.0.zip", repos = NULL)
-## 
-## # Or, install from source if you have a C++ compiler installed:
-## install.packages("metafolio_0.3.0.tar.gz", repos = NULL,
-##   type = "source")
-
-
-## @knitr load-package, eval=FALSE
-## library(metafolio)
-
-
-## @knitr help, eval=FALSE
-## vignette("metafolio")
-## ?metafolio
-## help(package = "metafolio")
-
-
-## @knitr setup-arma
-arma_env_params <- list(mean_value = 16, ar = 0.1, sigma_env = 2,
-  ma = 0)
-
-
-## @knitr base-case-run, cache=FALSE
-base1 <- meta_sim(n_pop = 10, env_params = arma_env_params,
-  env_type = "arma", assess_freq = 5)
-
-
-## @knitr plot-base-case-ts, fig.width=5, fig.height=6.5, out.width="4in", fig.cap="An example simulation with stationary environmental stochasticity and the base-case parameter values."
-plot_sim_ts(base1, years_to_show = 70, burn = 30)
-
-
-## @knitr spatial-plans
-w_plans <- list()
-w_plans[["balanced"]] <- c(5, 1000, 5, 1000, 5, 5, 1000, 5,
-    1000, 5)
-w_plans[["one_half"]] <- c(rep(1000, 4), rep(5, 6))
-w <- list()
-for(i in 1:2) { # loop over plans
- w[[i]] <- list()
- for(j in 1:80) { # loop over iterations
-   w[[i]][[j]] <- matrix(w_plans[[i]], nrow = 1)
- }
-}
-
-
[…]

### 2014-07-02 inst/examples/pe.R
Delete extra vignette files

@@ -1,39 +1,0 @@
-
-library(metafolio)
-
-arma_env_params <- list(mean_value = 16, ar = 0.1, sigma_env = 2, ma = 0)
-
-base1 <- meta_sim(n_pop = 10, env_params = arma_env_params,
-  env_type = "arma", assess_freq = 5)
-
-out <- sapply(1:30, function(x) {
-  base1 <- meta_sim(n_pop = 10, stray_fraction = 0,
-    env_params = arma_env_params, a_width_param = rep(0.05, 10),
-    optim_temp = rep(16, 10),  max_a = rep(2.8, 10),
-  env_type = "arma", assess_freq = 5
-  )
-  cv <- sd(base1$A) / mean(base1$A)
-  cv
-  })
-
-out1 <- sapply(1:30, function(x) {
-  base1 <- meta_sim(n_pop = 10, stray_fraction = 0.2,
-    env_params = arma_env_params, a_width_param = rep(0.05, 10),
-    optim_temp = rep(16, 10), max_a = rep(2.8, 10),
-  env_type = "arma", assess_freq = 5
-  )
-  cv <- sd(base1$A) / mean(base1$A)
-  cv
-  })
-
-
-out2 <- sapply(1:30, function(x) {
-  base1 <- meta_sim(n_pop = 10, stray_fraction = 0.2,
-    env_params = arma_env_params, a_width_param = rep(0.05, 10),
-    optim_temp = rep(16, 10), max_a = rep(2.8, 10),
-  env_type = "arma", assess_freq = 5
-  )
-  cv <- sd(base1$A) / mean(base1$A)
-  cv
-  })
-

### 2014-07-02 R/plot_cons_plans.R
Add missing word in docs

@@ -72,7 +72,7 @@ get_efficient_frontier <- function(m, v) {
 #'
 #' This makes a mean-variance plot of the portfolio output. It can take care of:
 #' plotting the individual portfolios, adding 2D kernel density polygons at two
-#' quantile levels, adding an efficient frontier.
+#' quantile levels, and adding an efficient frontier.
 #'
 #' @param plans_mv The \code{plans_mv} element of the output from
 #'   \code{\link{run_cons_plans}}.

### 2014-07-10 inst/ms/appendices/Appendix-A.tex
Add separate appendix files

\documentclass[12pt]{article}

\setlength\parindent{0in}

\begin{document}

\textbf{Sean C. Anderson, Jonathan W. Moore, Michelle M. McClure, Nicholas K. Dulvy, Andrew B. Cooper. 2014. Portfolio conservation of metapopulations under climate change. Ecological Applications. VOL:pp--pp.}

\bigskip

\textsc{Appendix A.} The \texttt{metafolio} \textsf{R} package.

\bigskip

The \texttt{metafolio} \textsf{R} package contains the functions and code to carry out the analyses in our paper. \texttt{metafolio}. The package can be installed from CRAN with:

\begin{verbatim}
install.packages("metafolio")
\end{verbatim}

%Alternatively, you can view the code and install the package from\\ \texttt{http://github.com/seananderson/metafolio}
%\bigskip

The included vignette describes the package and illustrates some example simulations. You can view the vignette with:

\begin{verbatim}
vignette("metafolio")
\end{verbatim}

You can view the help for the package with:

\begin{verbatim}
?metafolio
help(package = "metafolio")
\end{verbatim}

The figures from this paper can be re-created by downloading the code from GitHub and sourcing the file \texttt{README.R} in the \texttt{inst/examples} folder:

\begin{verbatim}
setwd("metafolio/inst/examples")
source("README.R")
\end{verbatim}

\end{document}

### 2014-07-10 inst/ms/appendices/Appendix-B.tex
Add separate appendix files

\documentclass[12pt]{article}

\usepackage[round]{natbib}
\bibpunct{}{}{;}{a}{}{}
\setlength\parindent{0in}
\usepackage[originalparameters]{ragged2e}
\usepackage{array}
\usepackage{booktabs}

\begin{document}

\textbf{Sean C. Anderson, Jonathan W. Moore, Michelle M. McClure, Nicholas K. Dulvy, Andrew B. Cooper. 2014. Portfolio conservation of metapopulations under climate change. Ecological Applications. VOL:pp--pp.}

\bigskip

\textsc{Appendix B.} Simulation input parameters and default values.

\bigskip

\begin{table}[h!]
\centering
\footnotesize
\caption{Input parameters to the salmon metapopulation simulation with default values.}
\smallskip
\begin{tabular}{>{\RaggedRight}p{7.8cm}p{1.1cm}p{2.5cm}>{\RaggedRight}p{3.3cm}}
\toprule
Description                                                          & Symbol                & Value                  & Reference  \\
\midrule

\textit{Population dynamics parameters}                              &                       &                        &             \\
Stock-recruit residual standard deviation (on log scale)             & $\sigma_r$            & 0.7                    & \citep{thorson2014a}  \\
AR(1) serial correlation of stock-recruit residuals                  & $\rho_w$              & 0.4                    & \citep{thorson2014a}  \\
Fraction of fish that stray from natal streams                       & $f_{\mathrm{stray}}$  & 0.02                   & \citep{quinn2005} and references therein  \\
Exponential rate of decay of straying with distance                  & $m$                   & 0.1                    & \citep{cooper1999}      \\
Range of maximum productivities                                      & $a_i^{\mathrm{max}}$  & 2.2--2.9             &  \citep{dorner2008}   \\

\noalign{\vskip 3mm}
\textit{Environmental parameters}                                    &                       &                        &             \\
Width parameter for thermal-tolerance curves for populations $i$ 1 to $n$  (values generate widths in line with listed references) & $W_i$                 & 0.08--0.04--0.08       &  \citep{brett1952, eliason2011}           \\
Optimum environmental value for populations $i$ 1 to $n$             & $e_i^{\mathrm{opt}}$  & 13--19                 &  \citep{eliason2011}           \\
Standard deviation of annual temperature fluctuations          & $\sigma_d$            & 2                      & \citep{eliason2011}      \\
AR(1) autocorrelation of annual temperature fluctuations       & $\rho_e$              & 0.1                    &             \\
Annual increase in stream temperature in degrees Celcius             & $\beta_e$             & 0.04                   & \citep{mantua2010}     \\

\noalign{\vskip 3mm}
\textit{Fishery parameters}                                          &                       &                        &             \\
Standard deviation of beta distribution for implementation error     & $\sigma_{h}$          & 0.1                    & \citep{pestes2008}            \\
Frequency of assessment (years)                                      & $f_{\mathrm{assess}}$ & 5                      &             \\
\bottomrule
\end{tabular}
\label{t:pars}
\end{table}
\clearpage

%\bibliographystyle{ecology3}
%\bibliography{jshort,ms}
\begin{thebibliography}{8}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
\expandafter\ifx\csname url\endcsname\relax
  \def\url#1{{\tt #1}}\fi
\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi

\bibitem[{Brett(1952)}]{brett1952}
Brett, J.~R.
\newblock 1952.
\newblock Temperature Tolerance in Young Pacific Salmon, Genus
  \textit{Oncorhynchus}.
\newblock J. Fish. Res. Bd. Can. 9:265--323.

\bibitem[{Cooper and Mangel(1999)}]{cooper1999}
Cooper, A.~B., and M.~Mangel.
\newblock 1999.
\newblock The dangers of ignoring metapopulation structure for the conservation
  of salmonids.
\newblock Fish. Bull. 97:213--226.

\bibitem[{Dorner et~al.(2008)Dorner, Peterman, and Haeseker}]{dorner2008}
Dorner, B., R.~M. Peterman, and S.~L. Haeseker.
\newblock 2008.
\newblock Historical trends in productivity of 120 {Pacific} pink, chum, and
  sockeye salmon stocks reconstructed by using a {Kalman} filter.
\newblock Can. J. Fish. Aquat. Sci. 65:1842--1866.

\bibitem[{Eliason et~al.(2011)Eliason, Clark, Hague, Hanson, Gallagher,
  Jeffries, Gale, Patterson, Hinch, and Farrell}]{eliason2011}
Eliason, E.~J., T.~D. Clark, M.~J. Hague, L.~M. Hanson, Z.~S. Gallagher, K.~M.
  Jeffries, M.~K. Gale, D.~A. Patterson, S.~G. Hinch, and A.~P. Farrell.
\newblock 2011.
\newblock Differences in thermal tolerance among sockeye salmon populations.
\newblock Science 332:109--112.

\bibitem[{Mantua et~al.(2010)Mantua, Tohver, and Hamlet}]{mantua2010}
Mantua, N., I.~Tohver, and A.~Hamlet.
\newblock 2010.
\newblock Climate change impacts on streamflow extremes and summertime stream
  temperature and their possible consequences for freshwater salmon habitat in
  {Washington State}.
\newblock Climatic Change 102:187--223.

\bibitem[{Pestes et~al.(2008)Pestes, Peterman, Bradford, and Wood}]{pestes2008}
Pestes, L.~R., R.~M. Peterman, M.~J. Bradford, and C.~C. Wood.
\newblock 2008.
\newblock Bayesian decision analysis for evaluating management options to
  promote recovery of a depleted salmon population.
\newblock Conserv. Biol. 22:351--361.

\bibitem[{Quinn(2005)}]{quinn2005}
Quinn, T.~P.
\newblock 2005.
\newblock The Behaviour and Ecology of Pacific Salmon and Trout.
\newblock American Fisheries Society, Bethesda, Maryland.

\bibitem[{Thorson et~al.(2014)Thorson, Jensen, and Zipkin}]{thorson2014a}
Thorson, J.~T., O.~P. Jensen, and E.~F. Zipkin.
\newblock 2014.
\newblock How variable is recruitment for exploited marine fishes? A
  hierarchical model for testing life history theory.
\newblock Can. J. Fish. Aquat. Sci. 71:973--983.

\end{thebibliography}

\end{document}

### 2014-07-10 inst/ms/appendices/Appendix-C.tex
Add separate appendix files

--- 2014-07-10 inst/ms/appendices/Appendix-B.tex
@@
 \documentclass[12pt]{article}
+\usepackage{graphicx}
 
-\usepackage[round]{natbib}
-\bibpunct{}{}{;}{a}{}{}
 \setlength\parindent{0in}
-\usepackage[originalparameters]{ragged2e}
-\usepackage{array}
-\usepackage{booktabs}
 
@@
 
-\textsc{Appendix B.} Simulation input parameters and default values.
+\textsc{Appendix C.} An example straying matrix.
 
@@
 
-\begin{table}[h!]
-\centering
-\footnotesize
-\caption{Input parameters to the salmon metapopulation simulation with default values.}
-\smallskip
-\begin{tabular}{>{\RaggedRight}p{7.8cm}p{1.1cm}p{2.5cm}>{\RaggedRight}p{3.3cm}}
-\toprule
-Description                                                          & Symbol                & Value                  & Reference  \\
-\midrule
-
-\textit{Population dynamics parameters}                              &                       &                        &             \\
-Stock-recruit residual standard deviation (on log scale)             & $\sigma_r$            & 0.7                    & \citep{thorson2014a}  \\
-AR(1) serial correlation of stock-recruit residuals                  & $\rho_w$              & 0.4                    & \citep{thorson2014a}  \\
-Fraction of fish that stray from natal streams                       & $f_{\mathrm{stray}}$  & 0.02                   & \citep{quinn2005} and references therein  \\
-Exponential rate of decay of straying with distance                  & $m$                   & 0.1                    & \citep{cooper1999}      \\
-Range of maximum productivities                                      & $a_i^{\mathrm{max}}$  & 2.2--2.9             &  \citep{dorner2008}   \\
-
-\noalign{\vskip 3mm}
-\textit{Environmental parameters}                                    &                       &                        &             \\
-Width parameter for thermal-tolerance curves for populations $i$ 1 to $n$  (values generate widths in line with listed references) & $W_i$                 & 0.08--0.04--0.08       &  \citep{brett1952, eliason2011}           \\
-Optimum environmental value for populations $i$ 1 to $n$             & $e_i^{\mathrm{opt}}$  & 13--19                 &  \citep{eliason2011}           \\
-Standard deviation of annual temperature fluctuations          & $\sigma_d$            & 2                      & \citep{eliason2011}      \\
-AR(1) autocorrelation of annual temperature fluctuations       & $\rho_e$              & 0.1                    &             \\
-Annual increase in stream temperature in degrees Celcius             & $\beta_e$             & 0.04                   & \citep{mantua2010}     \\
-
-\noalign{\vskip 3mm}
-\textit{Fishery parameters}                                          &                       &                        &             \\
-Standard deviation of beta distribution for implementation error     & $\sigma_{h}$          & 0.1                    & \citep{pestes2008}            \\
-Frequency of assessment (years)                                      & $f_{\mathrm{assess}}$ & 5                      &             \\
-\bottomrule
-\end{tabular}
-\label{t:pars}
-\end{table}
-\clearpage
-
-%\bibliographystyle{ecology3}
-%\bibliography{jshort,ms}
-\begin{thebibliography}{8}
-\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
-\expandafter\ifx\csname url\endcsname\relax
-  \def\url#1{{\tt #1}}\fi
-\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
-
-\bibitem[{Brett(1952)}]{brett1952}
-Brett, J.~R.
-\newblock 1952.
-\newblock Temperature Tolerance in Young Pacific Salmon, Genus
-  \textit{Oncorhynchus}.
-\newblock J. Fish. Res. Bd. Can. 9:265--323.
-
-\bibitem[{Cooper and Mangel(1999)}]{cooper1999}
-Cooper, A.~B., and M.~Mangel.
-\newblock 1999.
-\newblock The dangers of ignoring metapopulation structure for the conservation
-  of salmonids.
-\newblock Fish. Bull. 97:213--226.
-
-\bibitem[{Dorner et~al.(2008)Dorner, Peterman, and Haeseker}]{dorner2008}
-Dorner, B., R.~M. Peterman, and S.~L. Haeseker.
-\newblock 2008.
-\newblock Historical trends in productivity of 120 {Pacific} pink, chum, and
-  sockeye salmon stocks reconstructed by using a {Kalman} filter.
-\newblock Can. J. Fish. Aquat. Sci. 65:1842--1866.
-
-\bibitem[{Eliason et~al.(2011)Eliason, Clark, Hague, Hanson, Gallagher,
-  Jeffries, Gale, Patterson, Hinch, and Farrell}]{eliason2011}
-Eliason, E.~J., T.~D. Clark, M.~J. Hague, L.~M. Hanson, Z.~S. Gallagher, K.~M.
-  Jeffries, M.~K. Gale, D.~A. Patterson, S.~G. Hinch, and A.~P. Farrell.
-\newblock 2011.
-\newblock Differences in thermal tolerance among sockeye salmon populations.
-\newblock Science 332:109--112.
-
-\bibitem[{Mantua et~al.(2010)Mantua, Tohver, and Hamlet}]{mantua2010}
-Mantua, N., I.~Tohver, and A.~Hamlet.
-\newblock 2010.
-\newblock Climate change impacts on streamflow extremes and summertime stream
-  temperature and their possible consequences for freshwater salmon habitat in
-  {Washington State}.
-\newblock Climatic Change 102:187--223.
-
-\bibitem[{Pestes et~al.(2008)Pestes, Peterman, Bradford, and Wood}]{pestes2008}
-Pestes, L.~R., R.~M. Peterman, M.~J. Bradford, and C.~C. Wood.
-\newblock 2008.
-\newblock Bayesian decision analysis for evaluating management options to
-  promote recovery of a depleted salmon population.
-\newblock Conserv. Biol. 22:351--361.
-
-\bibitem[{Quinn(2005)}]{quinn2005}
-Quinn, T.~P.
-\newblock 2005.
-\newblock The Behaviour and Ecology of Pacific Salmon and Trout.
-\newblock American Fisheries Society, Bethesda, Maryland.
-
-\bibitem[{Thorson et~al.(2014)Thorson, Jensen, and Zipkin}]{thorson2014a}
-Thorson, J.~T., O.~P. Jensen, and E.~F. Zipkin.
-\newblock 2014.
-\newblock How variable is recruitment for exploited marine fishes? A
-  hierarchical model for testing life history theory.
-\newblock Can. J. Fish. Aquat. Sci. 71:973--983.
-
-\end{thebibliography}
+\includegraphics[width=3.5in]{Fig-C1}\\
+\textsc{Fig.~C1.} An example straying matrix.
+The rows and columns represent different populations (indicated by population number).
+Dark blue indicates a high rate of straying and light blue indicates a low rate of straying.
 

### 2014-07-10 inst/ms/appendices/Appendix-D.tex
Add separate appendix files

--- 2014-07-10 inst/ms/appendices/Appendix-C.tex
@@
 
-\textsc{Appendix C.} An example straying matrix.
+\textsc{Appendix D.} Sensitivity illustration with alternative parameter values.
 
@@
 
-\includegraphics[width=3.5in]{Fig-C1}\\
-\textsc{Fig.~C1.} An example straying matrix.
-The rows and columns represent different populations (indicated by population number).
-Dark blue indicates a high rate of straying and light blue indicates a low rate of straying.
+\includegraphics[width=3.5in]{Fig-D1}\\
+\textsc{Fig.~D1.} The impact of increasing or decreasing various parameter values on metapopulation return abundance.
+The different coloured lines represent three example salmon populations.
+The base case represents the base-case values for the short-term environmental fluctuation scenario.
 

### 2014-07-10 inst/ms/appendices/Appendix-E.tex
Add separate appendix files

\documentclass[12pt]{article}
\usepackage{graphicx}

\setlength\parindent{0in}

\begin{document}

\textbf{Sean C. Anderson, Jonathan W. Moore, Michelle M. McClure, Nicholas K. Dulvy, Andrew B. Cooper. 2014. Portfolio conservation of metapopulations under climate change. Ecological Applications. VOL:pp--pp.}

\bigskip

\textsc{Appendix E.} An illustration of the correlation between populations.

\bigskip

\includegraphics[width=4.5in]{Fig-E1}\\
\textsc{Fig.~E1.} A comparison of the log(returns) between populations. The subpopulation IDs are coloured from warm tolerant (warm colours) to cool tolerant (cool colours). Note how populations 1 and 10 have asynchronous returns whereas populations with more similar thermal-tolerance curves (say populations 9 and 10) have more synchronous dynamics. Populations with thermal tolerance curves in the middle (e.g. population 6) are less correlated with other populations. Their population dynamics end up primarily driven by demographic stochasticity and less so by temperature-induced systematic changes in productivity.

\end{document}

### 2014-07-10 inst/ms/appendices/Appendix-F.tex
Add separate appendix files

\documentclass[12pt]{article}
\usepackage{graphicx}

\setlength\parindent{0in}

\begin{document}

\textbf{Sean C. Anderson, Jonathan W. Moore, Michelle M. McClure, Nicholas K. Dulvy, Andrew B. Cooper. 2014. Portfolio conservation of metapopulations under climate change. Ecological Applications. VOL:pp--pp.}

\bigskip

\textsc{Appendix F.} Example simulated time series from alternative conservation scenarios.

\bigskip

\clearpage

\includegraphics[width=4.3in]{Fig-F1}\\
\textsc{Fig.~F1.} Conserving a \textbf{full range} of response diversity (spatial conservation strategy) with \textbf{short-term} environmental fluctuations. This is the same as Fig.~3 but in colour.
\clearpage

\includegraphics[width=4.3in]{Fig-F2}\\
\textsc{Fig.~F2.} Conserving \textbf{one half} of response diversity (spatial conservation strategy) with \textbf{short-term} environmental fluctuations.
\clearpage

\includegraphics[width=4.3in]{Fig-F3}\\
\textsc{Fig.~F3.} Conserving a \textbf{full range} of response diversity (spatial conservation strategy) with \textbf{long-term} environmental change.
\clearpage

\includegraphics[width=4.3in]{Fig-F4}\\
\textsc{Fig.~F4.} Conserving \textbf{one half} of response diversity (spatial conservation strategy) with \textbf{long-term} environmental change.
\clearpage

\includegraphics[width=4.3in]{Fig-F5}\\
\textsc{Fig.~F5.} \textbf{Two populations} conserved with random response diversity and \textbf{short-term} environmental fluctuations.
\clearpage

\includegraphics[width=4.3in]{Fig-F6}\\
\textsc{Fig.~F6.} \textbf{Sixteen populations} conserved with random response diversity and \textbf{short-term} environmental fluctuations.
\clearpage

\includegraphics[width=4.3in]{Fig-F7}\\
\textsc{Fig.~F7.} \textbf{Two populations} conserved with random response diversity and \textbf{long-term} environmental change.
\clearpage

\includegraphics[width=4.3in]{Fig-F8}\\
\textsc{Fig.~F8.} \textbf{Sixteen populations} conserved with random response diversity and \textbf{long-term} environmental change.
\clearpage

\includegraphics[width=4.3in]{Fig-F9}\\
\textsc{Fig.~F9.} \textbf{Two populations} conserved with random response diversity and \textbf{long-term declining stream flow}.
\clearpage

\includegraphics[width=4.3in]{Fig-F10}\\
\textsc{Fig.~F10.} \textbf{Sixteen populations} conserved with random response diversity and \textbf{long-term declining stream flow}.

\end{document}

### 2014-07-10 inst/ms/appendices/pdf2png.R

@@ -1,0 +1,8 @@
+files <- list.files(path = ".", pattern = "Fig[A-Z0-9\\-]*\\.pdf")
+files_png <- sub("\\.pdf", ".png", files)
+
+for(i in 1:length(files)) {
+  message(files_png[i])
+  system(paste("convert -density 120 -quality 100 -trim",
+      files[i], files_png[i]))
+}

### 2014-07-10 inst/ms/anderson-etal-metafolio.tex
Appendix B1 -> C1 (typo)
Remove response doc; input .bbl file for publication
Add missing sup fig references
Get main .tex file ready to submit

@@
 \newlabel{t:pars}{{B1}{2}}
-\newlabel{f:stray}{{B1}{3}}
+\newlabel{f:stray}{{C1}{3}}
 \newlabel{f:eg-sens}{{D1}{4}}
@@
 \newlabel{f:eg-n-linear-sixteen}{{F8}{13}}
+\newlabel{f:eg-n-squeeze-two}{{F9}{14}}
+\newlabel{f:eg-n-squeeze-twelve}{{F10}{15}}
 
@@
 
-\newcommand{\R}[1]{\label{#1}\linelabel{#1}}
-\newcommand{\lr}[1]{line~\lineref{#1}}
+%\newcommand{\R}[1]{\label{#1}\linelabel{#1}}
+%\newcommand{\lr}[1]{line~\lineref{#1}}
 
@@
 \setlength\parindent{0in}
-\include{response}
+%\include{response}
 \setlength\parskip{0in}
@@
 
-\bibliographystyle{ecology3}
-\bibliography{jshort,ms}
+%\bibliographystyle{ecology3}
+%\bibliography{jshort,ms}
 %\input{anderson-etal-metafolio.bbl}
+\begin{thebibliography}{64}
+\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
+\expandafter\ifx\csname url\endcsname\relax
+  \def\url#1{{\tt #1}}\fi
+\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
+
+\bibitem[{Allendorf et~al.(1997)Allendorf, Bayles, Bottom, Currens, Frissell,
+  Hankin, Lichatowich, Nehlsen, Trotter, and Williams}]{allendorf1997}
+Allendorf, F.~W., D.~Bayles, D.~L. Bottom, K.~P. Currens, C.~A. Frissell,
+  D.~Hankin, J.~A. Lichatowich, W.~Nehlsen, P.~C. Trotter, and T.~H. Williams.
+\newblock 1997.
+\newblock Prioritizing {Pacific} salmon stocks for conservation.
+\newblock Conserv. Biol. 11:140--152.
+
+\bibitem[{Anderson(2014)}]{metafoliopkg}
+Anderson, S.~C. 2014.
+\newblock metafolio: Salmon metapopulation simulations for conservation
+  planning through portfolio optimization. R package version 0.1.0.
+  http://cran.r-project.org/package=metafolio.
+
+\bibitem[{Anderson et~al.(2013)Anderson, Cooper, and Dulvy}]{anderson2013}
+Anderson, S.~C., A.~B. Cooper, and N.~K. Dulvy.
+\newblock 2013.
+\newblock Ecological prophets: Quantifying metapopulation portfolio effects.
+\newblock Methods Ecol. Evol. 4:971--981.
+
+\bibitem[{Ando and Mallory(2012)}]{ando2012}
+Ando, A.~W., and M.~L. Mallory.
+\newblock 2012.
+\newblock Optimal portfolio design to reduce climate-related conservation
+  uncertainty in the {Prairie Pothole Region}.
+\newblock Proc. Natl. Acad. Sci. U.S.A. 109:6484--6489.
+
+\bibitem[{Balvanera et~al.(2006)Balvanera, Pfisterer, Buchmann, He,
+  Nakashizuka, Raffaelli, and Schmid}]{balvanera2006}
+Balvanera, P., A.~B. Pfisterer, N.~Buchmann, J.-S. He, T.~Nakashizuka,
+  D.~Raffaelli, and B.~Schmid.
+\newblock 2006.
+\newblock Quantifying the evidence for biodiversity effects on ecosystem
+  functioning and services.
+\newblock Ecol. Lett. 9:1146--1156.
+
+\bibitem[{Battin et~al.(2007)Battin, Wiley, Ruckelshaus, Palmer, Korb, Bartz,
+  and Imaki}]{battin2007}
+Battin, J., M.~W. Wiley, M.~H. Ruckelshaus, R.~N. Palmer, E.~Korb, K.~K. Bartz,
+  and H.~Imaki.
+\newblock 2007.
+\newblock Projected impacts of climate change on salmon habitat restoration.
+\newblock Proc. Natl. Acad. Sci. U.S.A. 104:6720--6725.
+
+\bibitem[{Beechie et~al.(2013)Beechie, Imaki, Greene, Wade, Wu, Pess, Roni,
+  Kimball, Stanford, Kiffney, and Mantua}]{beechie2013}
+Beechie, T., H.~Imaki, J.~Greene, A.~Wade, H.~Wu, G.~Pess, P.~Roni, J.~Kimball,
+  J.~Stanford, P.~Kiffney, and N.~Mantua.
+\newblock 2013.
+\newblock Restoring salmon habitat for a changing climate.
+\newblock River Res. Applic. 29:939--960.
+
+\bibitem[{Carlson and Satterthwaite(2011)}]{carlson2011}
+Carlson, S.~M., and W.~H. Satterthwaite.
+\newblock 2011.
+\newblock Weakened portfolio effect in a collapsed salmon population complex.
+\newblock Can. J. Fish. Aquat. Sci. 68:1579--1589.
+
+\bibitem[{Cooper and Mangel(1999)}]{cooper1999}
+Cooper, A.~B., and M.~Mangel.
+\newblock 1999.
+\newblock The dangers of ignoring metapopulation structure for the conservation
+  of salmonids.
+\newblock Fish. Bull. 97:213--226.
+
+\bibitem[{Crozier et~al.(2008)Crozier, Hendry, Lawson, Quinn, Mantua, Battin,
+  Shaw, and Huey}]{crozier2008}
+Crozier, L.~G., A.~P. Hendry, P.~W. Lawson, T.~Quinn, N.~J. Mantua, J.~Battin,
+  R.~G. Shaw, and R.~B. Huey.
+\newblock 2008.
+\newblock Potential responses to climate change in organisms with complex life
+  histories: evolution and plasticity in {Pacific} salmon.
+\newblock Evol. Appl. 1:252--270.
+
+\bibitem[{de~Mazancourt et~al.(2013)de~Mazancourt, Isbell, Larocque, Berendse,
+  De~Luca, Grace, Haegeman, Wayne~Polley, Roscher, Schmid, Tilman, van Ruijven,
+  Weigelt, Wilsey, and Loreau}]{demazancourt2013}
+de~Mazancourt, C., F.~Isbell, A.~Larocque, F.~Berendse, E.~De~Luca, J.~B.
+  Grace, B.~Haegeman, H.~Wayne~Polley, C.~Roscher, B.~Schmid, D.~Tilman, J.~van
+  Ruijven, A.~Weigelt, B.~J. Wilsey, and M.~Loreau.
+\newblock 2013.
+\newblock Predicting ecosystem stability from community composition and
+  biodiversity.
+\newblock Ecol. Lett. 16:617--625.
+
+\bibitem[{DFO(2005)}]{dfo2005wsp}
+DFO. 2005.
+\newblock Canada's policy for conservation of wild {Pacific} salmon.
+\newblock Technical report, Fisheries and Oceans Canada, Vancouver, BC.
+
+\bibitem[{Dorner et~al.(2008)Dorner, Peterman, and Haeseker}]{dorner2008}
+Dorner, B., R.~M. Peterman, and S.~L. Haeseker.
+\newblock 2008.
+\newblock Historical trends in productivity of 120 {Pacific} pink, chum, and
+  sockeye salmon stocks reconstructed by using a {Kalman} filter.
+\newblock Can. J. Fish. Aquat. Sci. 65:1842--1866.
+
+\bibitem[{Eliason et~al.(2011)Eliason, Clark, Hague, Hanson, Gallagher,
+  Jeffries, Gale, Patterson, Hinch, and Farrell}]{eliason2011}
+Eliason, E.~J., T.~D. Clark, M.~J. Hague, L.~M. Hanson, Z.~S. Gallagher, K.~M.
+  Jeffries, M.~K. Gale, D.~A. Patterson, S.~G. Hinch, and A.~P. Farrell.
+\newblock 2011.
+\newblock Differences in thermal tolerance among sockeye salmon populations.
+\newblock Science 332:109--112.
+
+\bibitem[{Elmqvist et~al.(2003)Elmqvist, Folke, Nystr{\"o}m, Peterson,
+  Bengtsson, Walker, and Norberg}]{elmqvist2003}
+Elmqvist, T., C.~Folke, M.~Nystr{\"o}m, G.~Peterson, J.~Bengtsson, B.~Walker,
+  and J.~Norberg.
+\newblock 2003.
+\newblock Response diversity, ecosystem change, and resilience.
+\newblock Front. Ecol. Environ. 1:488--494.
+
+\bibitem[{Figge(2004)}]{figge2004}
+Figge, F.
+\newblock 2004.
+\newblock Bio-folio: applying portfolio theory to biodiversity.
+\newblock Biodivers. Conserv. 13:827--849.
+
+\bibitem[{Fraser et~al.(2011)Fraser, Weir, Bernatchez, Hansen, and
+  Taylor}]{fraser2011}
+Fraser, D.~J., L.~K. Weir, L.~Bernatchez, M.~M. Hansen, and E.~B. Taylor.
+\newblock 2011.
+\newblock Extent and scale of local adaptation in salmonid fishes: review and
+  meta-analysis.
+\newblock Heredity 106:404--420.
+
+\bibitem[{Gustafson et~al.(2007)Gustafson, Waples, Myers, Weitkamp, Bryant,
+  Johnson, and Hard}]{gustafson2007}
+Gustafson, R.~G., R.~S. Waples, J.~M. Myers, L.~A. Weitkamp, G.~J. Bryant,
+  O.~W. Johnson, and J.~J. Hard.
+\newblock 2007.
+\newblock Pacific salmon extinctions: quantifying lost and remaining diversity.
+\newblock Conserv. Biol. 21:1009--1020.
+
+\bibitem[{Haak and Williams(2012)}]{haak2012}
+Haak, A.~L., and J.~E. Williams.
+\newblock 2012.
+\newblock Spreading the risk: native trout management in a warmer and
+  less-certain future.
+\newblock N. Am. J. Fish. Manage. 32:387--401.
+
+\bibitem[{Hannah et~al.(2002)Hannah, Midgley, Lovejoy, Bond, Bush, Lovett,
+  Scott, and Woodward}]{hannah2002a}
+Hannah, L., G.~F. Midgley, T.~Lovejoy, W.~J. Bond, M.~Bush, J.~C. Lovett,
+  D.~Scott, and F.~I. Woodward.
+\newblock 2002.
+\newblock Conservation of biodiversity in a changing climate.
+\newblock Conserv. Biol. 16:264--268.
+
+\bibitem[{Hilborn et~al.(2003)Hilborn, Quinn, Schindler, and
+  Rogers}]{hilborn2003}
+Hilborn, R.~W., T.~Quinn, D.~Schindler, and D.~Rogers.
+\newblock 2003.
+\newblock Biocomplexity and fisheries sustainability.
+\newblock Proc. Natl. Acad. Sci. U.S.A. 100:6564--6568.
+
+\bibitem[{Hilborn and Walters(1992)}]{hilborn1992}
+Hilborn, R.~W., and C.~Walters.
+\newblock 1992.
+\newblock Quantitative Fisheries Stock Assessment: Choice, Dynamics, and
+  Uncertainty.
+\newblock Chapman and Hall, London.
+
+\bibitem[{Isaak and Rieman(2013)}]{isaak2013}
+Isaak, D.~J., and B.~E. Rieman.
+\newblock 2013.
+\newblock Stream isotherm shifts from climate change and implications for
+  distributions of ectothermic organisms.
+\newblock Global Change Biol. 19:742--751.
+
+\bibitem[{Isaak et~al.(2007)Isaak, Thurow, Rieman, and Dunham}]{isaak2007}
+Isaak, D.~J., R.~F. Thurow, B.~E. Rieman, and J.~B. Dunham.
+\newblock 2007.
+\newblock Chinook salmon use of spawning patches: relative roles of habitat
+  quality, size, and connectivity.
+\newblock Ecol. Appl. 17:352--364.
+
+\bibitem[{Isaak et~al.(2012)Isaak, Wollrab, Horan, and Chandler}]{isaak2012}
+Isaak, D.~J., S.~Wollrab, D.~Horan, and G.~Chandler.
+\newblock 2012.
+\newblock Climate change effects on stream and river temperatures across the
+  northwest U.S. from 1980--2009 and implications for salmonid fishes.
+\newblock Climatic Change 113:499--524.
+
+\bibitem[{Ives and Carpenter(2007)}]{ives2007}
+Ives, A.~R., and S.~R. Carpenter.
+\newblock 2007.
+\newblock Stability and diversity of ecosystems.
+\newblock Science 317:58--62.
+
+\bibitem[{Jentsch et~al.(2007)Jentsch, Kreyling, and
+  Beierkuhnlein}]{jentsch2007}
+Jentsch, A., J.~Kreyling, and C.~Beierkuhnlein.
+\newblock 2007.
+\newblock A new generation of climate-change experiments: events, not trends.
+\newblock Front. Ecol. Environ. 5:365--374.
+
+\bibitem[{Johnson and Schindler(2013)}]{johnson2013a}
+Johnson, S.~P., and D.~E. Schindler.
+\newblock 2013.
+\newblock Marine trophic diversity in an anadromous fish is linked to its
+  life-history variation in fresh water.
+\newblock Biol. Lett. 9:20120824.
+
+\bibitem[{Koellner and Schmitz(2006)}]{koellner2006}
+Koellner, T., and O.~J. Schmitz.
+\newblock 2006.
+\newblock Biodiversity, ecosystem function, and investment risk.
+\newblock BioScience 56:977--985.
+
+\bibitem[{Lackey(2003)}]{lackey2003}
+Lackey, R.~T.
+\newblock 2003.
+\newblock {Pacific Northwest} salmon: forecasting their status in 2100.
+\newblock Rev. Fish. Sci. 11:35--88.
+
+\bibitem[{Lee and Jetz(2008)}]{lee2008}
+Lee, T.~M., and W.~Jetz.
+\newblock 2008.
+\newblock Future battlegrounds for conservation under global change.
+\newblock Proc. R. Soc. Lond. B 275:1261--1270.
+
+\bibitem[{Luce et~al.(2013)Luce, Abatzoglou, and Holden}]{luce2013}
+Luce, C.~H., J.~T. Abatzoglou, and Z.~A. Holden.
+\newblock 2013.
+\newblock The missing mountain water: slower westerlies decrease orographic
+  enhancement in the {Pacific Northwest USA}.
+\newblock Science 342:1360--1364.
+
+\bibitem[{Luce and Holden(2009)}]{luce2009}
+Luce, C.~H., and Z.~A. Holden.
+\newblock 2009.
+\newblock Declining annual streamflow distributions in the {Pacific Northwest
+  United States}, 1948--2006.
+\newblock Geophys. Res. Lett. 36.
+
+\bibitem[{Mantua et~al.(2010)Mantua, Tohver, and Hamlet}]{mantua2010}
+Mantua, N., I.~Tohver, and A.~Hamlet.
+\newblock 2010.
+\newblock Climate change impacts on streamflow extremes and summertime stream
+  temperature and their possible consequences for freshwater salmon habitat in
+  {Washington State}.
+\newblock Climatic Change 102:187--223.
+
+\bibitem[{Margules and Pressey(2000)}]{margules2000}
+Margules, C.~R., and R.~L. Pressey.
+\newblock 2000.
+\newblock Systematic conservation planning.
+\newblock Nature 405:243--253.
+
+\bibitem[{Markowitz(1952)}]{markowitz1952}
+Markowitz, H.
+\newblock 1952.
+\newblock Portfolio Selection.
+\newblock J. Finance 7:77--91.
+
+\bibitem[{McClure et~al.(2008{\natexlab{{\em a\/}}})McClure, Carlson, Beechie,
+  Pess, Jorgensen, Sogard, Sultan, Holzer, Travis, Sanderson, Power, and
+  Carmichael}]{mcclure2008a}
+McClure, M.~M., S.~M. Carlson, T.~J. Beechie, G.~R. Pess, J.~C. Jorgensen,
+  S.~M. Sogard, S.~E. Sultan, D.~M. Holzer, J.~Travis, B.~L. Sanderson, M.~E.
+  Power, and R.~W. Carmichael.
+\newblock 2008{\natexlab{{\em a\/}}}.
+\newblock Evolutionary consequences of habitat loss for {Pacific} anadromous
+  salmonids.
+\newblock Evol. Appl. 1:300--318.
+
+\bibitem[{McClure et~al.(2008{\natexlab{{\em b\/}}})McClure, Utter, Baldwin,
+  Carmichael, Hassemer, Howell, Spruell, Cooney, Schaller, and
+  Petrosky}]{mcclure2008b}
+McClure, M.~M., F.~M. Utter, C.~Baldwin, R.~W. Carmichael, P.~F. Hassemer,
+  P.~J. Howell, P.~Spruell, T.~D. Cooney, H.~A. Schaller, and C.~E. Petrosky.
+\newblock 2008{\natexlab{{\em b\/}}}.
+\newblock Evolutionary effects of alternative artificial propagation programs:
+  implications for viability of endangered anadromous salmonids.
+\newblock Evol. Appl. 1:356--375.
+
+\bibitem[{McCullough(1999)}]{mccullough1999}
+McCullough, D.~A., 1999.
+\newblock A review and synthesis of effects of alterations to the water
+  temperature regime on freshwater life stages of salmonids, with special
+  reference to {Chinook} salmon.
+\newblock EPA 910-R-99-010, US Environmental Protection Agency.
+
+\bibitem[{McElhany et~al.(2000)McElhany, Ruckelshaus, Ford, Wainwright, and
+  Bjorkstedt}]{mcelhany2000}
+McElhany, P., M.~H. Ruckelshaus, M.~J. Ford, T.~C. Wainwright, and E.~P.
+  Bjorkstedt, 2000.
+\newblock {Viable Salmonid Populations} and the recovery of {Evolutionarily
+  Significant Units}.
+\newblock Technical Report NOAA Tech. Memo. NMFS-NWFSC-42, U.S. Deptartment of
+  Commerce.
+
+\bibitem[{Moore et~al.(2010)Moore, Mcclure, Rogers, and Schindler}]{moore2010}
+Moore, J.~W., M.~Mcclure, L.~A. Rogers, and D.~E. Schindler.
+\newblock 2010.
+\newblock Synchronization and portfolio performance of threatened salmon.
+\newblock Conserv. Lett. 3:340--348.
+
+\bibitem[{Mori et~al.(2013)Mori, Furukawa, and Sasaki}]{mori2013}
+Mori, A.~S., T.~Furukawa, and T.~Sasaki.
+\newblock 2013.
+\newblock Response diversity determines the resilience of ecosystems to
+  environmental change.
+\newblock Biol. Rev. 88:349--364.
+
+\bibitem[{Patterson et~al.(2007)Patterson, Skibo, Barnes, Hills, and
+  Macdonald}]{patterson2007}
+Patterson, D.~A., K.~M. Skibo, D.~P. Barnes, J.~A. Hills, and J.~S. Macdonald.
+\newblock 2007.
+\newblock The influence of water temperature on time to surface for adult
+  sockeye salmon carcasses and the limitations in estimating salmon carcasses
+  in the {Fraser River, British Columbia}.
+\newblock N. Am. J. Fish. Manage. 27:878--884.
+
+\bibitem[{Pestes et~al.(2008)Pestes, Peterman, Bradford, and Wood}]{pestes2008}
+Pestes, L.~R., R.~M. Peterman, M.~J. Bradford, and C.~C. Wood.
+\newblock 2008.
+\newblock Bayesian decision analysis for evaluating management options to
+  promote recovery of a depleted salmon population.
+\newblock Conserv. Biol. 22:351--361.
+
+\bibitem[{Policansky and Magnuson(1998)}]{policansky1998}
+Policansky, D., and J.~J. Magnuson.
+\newblock 1998.
+\newblock Genetics, metapopulations, and ecosystem management of fisheries.
+\newblock Ecol. Appl. 8:S119--S123.
+
+\bibitem[{Pressey et~al.(2007)Pressey, Cabeza, Watts, Cowling, and
+  Wilson}]{pressey2007}
+Pressey, R.~L., M.~Cabeza, M.~E. Watts, R.~M. Cowling, and K.~A. Wilson.
+\newblock 2007.
+\newblock Conservation planning in a changing world.
+\newblock Trends Ecol. Evolut. 22:583--592.
+
+\bibitem[{Quinn(2005)}]{quinn2005}
+Quinn, T.~P.
+\newblock 2005.
+\newblock The Behaviour and Ecology of Pacific Salmon and Trout.
+\newblock American Fisheries Society, Bethesda, Maryland.
+
+\bibitem[{{\textsf{R} Core Team}(2013)}]{r2013}
+{\textsf{R} Core Team}, 2013.
+\newblock \textsf{R}: {A} language and environment for statistical computing.
+\newblock \textsf{R} Foundation for Statistical Computing, Vienna, Austria.
+
+\bibitem[{Rogers and Schindler(2008)}]{rogers2008}
+Rogers, L.~A., and D.~E. Schindler.
+\newblock 2008.
+\newblock Asynchrony in population dynamics of sockeye salmon in southwest
+  {Alaska}.
+\newblock Oikos 117:1578--1586.
+
+\bibitem[{Ruckelshaus et~al.(2002)Ruckelshaus, Levin, Johnson, and
+  Kareiva}]{ruckelshaus2002}
+Ruckelshaus, M.~H., P.~Levin, J.~B. Johnson, and P.~M. Kareiva.
+\newblock 2002.
+\newblock The {Pacific} salmon wars: what science brings to the challenge of
+  recovering species.
+\newblock Annu. Rev. Ecol. Syst. 33:665--706.
+
+\bibitem[{Scheuerell et~al.(2006)Scheuerell, Hilborn, Ruckelshaus, Bartz,
+  Lagueux, Haas, and Rawson}]{scheuerell2006}
+Scheuerell, M.~D., R.~Hilborn, M.~H. Ruckelshaus, K.~K. Bartz, K.~M. Lagueux,
+  A.~D. Haas, and K.~Rawson.
+\newblock 2006.
+\newblock The {Shiraz} model: a tool for incorporating anthropogenic effects
+  and fish--habitat relationships in conservation planning.
+\newblock Can. J. Fish. Aquat. Sci. 63:1596--1607.
+
+\bibitem[{Schindler et~al.(2008)Schindler, Augerot, Fleishman, Mantua, Riddell,
+  Ruckelshaus, Seeb, and Webster}]{schindler2008}
+Schindler, D.~E., X.~Augerot, E.~Fleishman, N.~J. Mantua, B.~Riddell,
+  M.~Ruckelshaus, J.~Seeb, and M.~Webster.
+\newblock 2008.
+\newblock Climate change, ecosystem impacts, and management for {Pacific}
+  salmon.
+\newblock Fisheries 33:502--506.
+
+\bibitem[{Schindler et~al.(2010)Schindler, Hilborn, Chasco, Boatright, Quinn,
+  Rogers, and Webster}]{schindler2010}
+Schindler, D.~E., R.~Hilborn, B.~Chasco, C.~P. Boatright, T.~P. Quinn, L.~A.
+  Rogers, and M.~S. Webster.
+\newblock 2010.
+\newblock Population diversity and the portfolio effect in an exploited
+  species.
+\newblock Nature 465:609--612.
+
+\bibitem[{Schtickzelle and Quinn(2007)}]{schtickzelle2007}
+Schtickzelle, N., and T.~P. Quinn.
+\newblock 2007.
+\newblock A metapopulation perspective for salmon and other anadromous fish.
+\newblock Fish Fish. 8:297--314.
+
+\bibitem[{Stockwell et~al.(2003)Stockwell, Hendry, and
+  Kinnison}]{stockwell2003}
+Stockwell, C., A.~Hendry, and M.~Kinnison.
+\newblock 2003.
+\newblock Contemporary evolution meets conservation biology.
+\newblock Trends. Ecol. Evol. 18:94--101.
+
+\bibitem[{Thibaut et~al.(2012)Thibaut, Connolly, and Sweatman}]{thibaut2012}
+Thibaut, L., S.~R. Connolly, and H.~P.~A. Sweatman.
+\newblock 2012.
+\newblock Diversity and stability of herbivorous fishes on coral reefs.
+\newblock Ecology 93:891--901.
+
+\bibitem[{Thomas(2010)}]{thomas2010}
+Thomas, C.~D.
+\newblock 2010.
+\newblock Climate, climate change and range boundaries.
+\newblock Divers. Distrib. 16:488--495.
+
+\bibitem[{Thomas et~al.(2004)Thomas, Cameron, Green, Bakkenes, Beaumont,
+  Collingham, Erasmus, de~Siqueira, Grainger, Hannah, Hughes, Huntley, van
+  Jaarsveld, Midgley, Miles, Ortega-Huerta, Townsend~Peterson, Phillips, and
+  Williams}]{thomas2004}
+Thomas, C.~D., A.~Cameron, R.~E. Green, M.~Bakkenes, L.~J. Beaumont, Y.~C.
+  Collingham, B.~F.~N. Erasmus, M.~F. de~Siqueira, A.~Grainger, L.~Hannah,
+  L.~Hughes, B.~Huntley, A.~S. van Jaarsveld, G.~F. Midgley, L.~Miles, M.~A.
+  Ortega-Huerta, A.~Townsend~Peterson, O.~L. Phillips, and S.~E. Williams.
+\newblock 2004.
+\newblock Extinction risk from climate change.
+\newblock Nature 427:145--148.
+
+\bibitem[{Thorson et~al.(2014{\natexlab{{\em a\/}}})Thorson, Jensen, and
+  Zipkin}]{thorson2014a}
+Thorson, J.~T., O.~P. Jensen, and E.~F. Zipkin.
+\newblock 2014{\natexlab{{\em a\/}}}.
+\newblock How variable is recruitment for exploited marine fishes? A
+  hierarchical model for testing life history theory.
+\newblock Can. J. Fish. Aquat. Sci. 71:973--983.
+
+\bibitem[{Thorson et~al.(2014{\natexlab{{\em b\/}}})Thorson, Scheuerell, Buhle,
+  and Copeland}]{thorson2014}
+Thorson, J.~T., M.~D. Scheuerell, E.~R. Buhle, and T.~Copeland.
+\newblock 2014{\natexlab{{\em b\/}}}.
+\newblock Spatial variation buffers temporal fluctuations in early juvenile
+  survival for an endangered {Pacific} salmon.
+\newblock J. Anim. Ecol. 83:157--167.
+
+\bibitem[{Walther et~al.(2002)Walther, Post, Convey, Menzel, Parmesan, Beebee,
+  Fromentin, Hoegh-Guldberg, and Bairlein}]{walther2002}
+Walther, G.-R., E.~Post, P.~Convey, A.~Menzel, C.~Parmesan, T.~J.~C. Beebee,
+  J.-M. Fromentin, O.~Hoegh-Guldberg, and F.~Bairlein.
+\newblock 2002.
+\newblock Ecological responses to recent climate change.
+\newblock Nature 416:389--395.
+
+\bibitem[{Waples et~al.(2009)Waples, Beechie, and Pess}]{waples2009}
+Waples, R., T.~Beechie, and G.~R. Pess.
+\newblock 2009.
+\newblock Evolutionary history, habitat disturbance regimes, and anthropogenic
+  changes: what do these mean for resilience of {Pacific} salmon populations?
+\newblock Ecol. Soc. 14:3.
+
+\bibitem[{Winfree and Kremen(2009)}]{winfree2009}
+Winfree, R., and C.~Kremen.
+\newblock 2009.
+\newblock Are ecosystem services stabilized by differences among species? {A}
+  test using crop pollination.
+\newblock Proc. R. Soc. Lond. B 276:229--237.
+
+\bibitem[{Yeakel et~al.(2014)Yeakel, Moore, Guimar{\~a}es, and
+  de~Aguiar}]{yeakel2014}
+Yeakel, J.~D., J.~W. Moore, P.~R. Guimar{\~a}es, and M.~A.~M. de~Aguiar.
+\newblock 2014.
+\newblock Synchronisation and stability in river metapopulation networks.
+\newblock Ecol. Lett. 17:272--283.
+
+\end{thebibliography}
 
@@
 
- \setcounter{page}{1}
- \nolinenumbers
- \include{som}
+ %\setcounter{page}{1}
+ %\nolinenumbers
+ %\include{som}
 

### 2014-07-10 inst/ms/anderson-etal-metafolio.tex
Copy over .tex manuscript for submission

@@
 
-\input{ms}
+%\input{ms}
+\noindent
+\textsuperscript{1}Earth to Ocean Research Group, Department of Biological Sciences, Simon Fraser University, Burnaby BC, V5A 1S6, Canada
+
+\noindent
+\textsuperscript{2}School of Resource and Environmental Management, Simon Fraser University, Burnaby, BC, V5A 1S6, Canada
+
+\noindent
+\textsuperscript{3}Fishery Resource Analysis and Monitoring Division, Northwest Fisheries Science Center, National Marine Fisheries Service, National Oceanic and Atmospheric Administration, 2725 Montlake Blvd.
+E., Seattle, WA 98112, USA
+
+\noindent
+\textsuperscript{*}Corresponding author: Sean C.
+Anderson; Earth to Ocean Research Group, Department of Biological Sciences, Simon Fraser University, Burnaby BC, V5A 1S6; Phone: 1-778-782-3989; E-mail: sean\_anderson@sfu.ca
+
+\clearpage
+
+\section{Abstract}\label{abstract}
+
+Climate change will likely lead to increasing population variability and extinction risk.
+Theoretically, greater population diversity should buffer against rising climate variability, and this theory is often invoked as a reason for greater conservation.
+However, this has rarely been quantified.
+Here we show how a portfolio approach to managing population diversity can inform metapopulation conservation priorities in a changing world.
+We develop a salmon metapopulation model where productivity is driven by spatially-distributed thermal tolerance and patterns of short- and long-term climate change.
+We then implement spatial conservation scenarios that control population carrying capacities and evaluate the metapopulation portfolios as a financial manager might --- along axes of conservation risk and return.
+We show that preserving a diversity of thermal tolerances minimizes risk given environmental stochasticity and ensures persistence given long-term environmental change.
+When the thermal tolerances of populations are unknown, doubling the number of populations conserved may nearly halve metapopulation variability.
+However, this reduction in variability can come at the expense of long-term persistence if climate change increasingly restricts available habitat --- forcing ecological managers to balance society's desire for short-term stability and long-term viability.
+Our findings suggest the importance of conserving the processes that promote thermal-tolerance diversity, such as genetic diversity, habitat heterogeneity, and natural disturbance regimes, and demonstrate that diverse natural portfolios may be critical for metapopulation conservation in the face of increasing climate variability and change.
+
+\noindent
+\textit{Keywords}: biocomplexity, ecosystem based management, Pacific salmon, portfolio effect, prioritization, range contraction, response diversity, risk assessment, stability-diversity, stochastic simulation
+
+\noindent
+\textit{Running head}: Metapopulation portfolio conservation
+
+\section{Introduction}\label{introduction}
+
+Untangling the mechanisms that underpin the stability of ecological systems is a critical focus of ecology \citep[e.g.][]{ives2007, demazancourt2013}.
+Decades of research has focused on the role of species richness and functional diversity in driving stability; however, recent research has highlighted that the drivers of ecological stability are more complex and multidimensional than previously thought \citep[e.g.][]{balvanera2006, ives2007, demazancourt2013}.
+Two key drivers of population stability that have been comparatively understudied are response diversity \citep{winfree2009, mori2013} --- different responses to the environment by functionally similar species or populations \citep{elmqvist2003} --- and the role of metapopulations \citep{schtickzelle2007}.
+Here, we examine the role of response diversity conservation in stabilizing metapopulations given projected changes in climate.
+With unprecedented loss of biodiversity and levels of anthropogenic environmental change, it is more critical than ever to consider conservation approaches that maintain system stability in the face of environmental uncertainty \citep{lee2008, ando2012}.
+
+Typically, conservation actions to maintain system stability and thereby reduce risk are driven by an \emph{ad hoc} combination of scientific information, political influences, and feasibility \citep{margules2000}; the management of financial portfolios provides another way of considering risk \citep[e.g.][]{figge2004, koellner2006, ando2012, haak2012}.
+Economists work to minimize risk and maximize returns by building a portfolio of individual investments (called assets) with different attributes.
+For example, different financial sectors can be expected to perform uniquely in some economic conditions; when one rises in value another may fall.
+Modern Portfolio Theory proposes that out of all possible portfolios, there is a small subset of portfolios that maximizes expected return for a level of risk or minimizes risk for a level of return (called the efficient frontier), and that only by considering risk and return in tandem can an investor achieve maximum benefit from a portfolio \citep{markowitz1952}.
+
+Similarly, expected growth rate and variance of a metapopulation is a function of the variance, covariance, and size of the individual populations \citep{moore2010, carlson2011, anderson2013}.
+An ecological portfolio approach to managing risk for a metapopulation might therefore consider how conservation actions affect the weight of each population in a metapopulation portfolio.
+This investment weight could represent the conservation budget or the habitat conserved for each population.
+The population growth rate is then analogous to the financial rate of return and the variability of that growth rate a metric of risk.
+Environmental conditions could represent the financial market conditions.
+Given this interpretation, ecological managers could consider how various conservation strategies affect the expected risk and return of their ecological portfolio.
+These risk and return elements are central to ecological management and conservation --- management aims to ensure stability over environmental variability (risk), and increase population abundance (return).
+Different scenarios may suggest different desired trade-offs between the two.
+For example, a manager with a healthy population might prioritize short-term stability, while a manager with an endangered population might try to balance the two, or prioritize population growth initially.
+
+Managing Pacific salmon under the uncertainty of climate change is an ideal scenario to consider through the lens of portfolio theory for four reasons.
+(1) The migration of Pacific salmon biomass profoundly influences aquatic and terrestrial coastal ecosystems throughout the North Pacific ocean from Korea to California \citep{quinn2005}.
+(2) Pacific salmon form metapopulations \citep[e.g.][]{policansky1998, cooper1999, schtickzelle2007} and we can consider, for example, the metapopulation in a river-catchment as a portfolio and the stream populations as assets \citep{schindler2010, moore2010, carlson2011, anderson2013, yeakel2014}.
+Fisheries often integrate across multiple populations, acting as investors in the salmon portfolio \citep{hilborn2003}.
+Fisheries managers and conservation agencies can act as portfolio managers by choosing which salmon habitat to prioritize for protection or restoration.
+(3) Many Pacific salmon metapopulations are highly threatened \citep[e.g.][]{gustafson2007} and will likely become more at risk as threats such as overfishing, damming, logging, and particularly changing climate, intensify \citep[e.g.][]{lackey2003}.
+Indeed, recovery goals for Pacific salmon are often set at the metapopulation level \citep{mcelhany2000}, and knowing what minimizes risk to the metapopulation can help choose efficient conservation actions \citep{policansky1998, mcelhany2000}.
+(4) Given the scale and variety of the threats facing salmon, some prioritization will be required to recover these highly-valued, even iconic species \citep{allendorf1997, ruckelshaus2002}.
+
+Two key mechanisms can generate the asynchrony in metapopulation dynamics that is critical to a diversified portfolio.
+First, localized habitat features can filter larger-scale environments, generating unique conditions for populations \citep{schindler2008} (\emph{sensu} the Moran effect).
+Second, salmon populations may respond differently to environmental variability (i.e.~response diversity \citep{elmqvist2003} and biocomplexity \citep{hilborn2003}) arising from unique local adaptations and traits \citep{fraser2011, eliason2011, thorson2014}.
+In reality, these mechanisms can interact.
+For example, salmon response diversity in the marine environment can be driven by adaptation to localized freshwater environments \citep{johnson2013a}.
+
+In addition to posing perhaps the greatest threat to global biodiversity in general \citep{thomas2004}, climate warming poses a particular threat to riverine species whose ranges are largely confined to existing habitat \citep{thomas2010}.
+Among these species, salmon are strongly affected by climate warming \citep[e.g.][]{patterson2007}.
+Warmer water can lead to massive mortality of salmon populations \citep[e.g.][]{patterson2007} and indirectly impact salmon productivity through alterations to snow-melt timing and extreme hydrological events \citep{crozier2008}.
+Due to these effects, adverse stream temperatures are already impeding recovery of some Pacific salmon populations \citep{mccullough1999} and are expected to make recovery targets more difficult to achieve \citep{battin2007}.
+However, despite the evidence that warming impacts salmon, salmon also show evidence of response diversity and local adaptation to temperature.
+For example, thermal tolerance of sockeye salmon in the Fraser River, British Columbia, Canada, varies within streams according to historical environmental conditions \citep{eliason2011}.
+
+Here we ask how portfolio theory can inform spatial approaches to prioritizing metapopulation conservation in a changing world.
+To answer this, we develop a salmon metapopulation simulation in which spatially-distributed thermal tolerance and patterns of short- and long-term climatic change drive population-specific productivity.
+We then implement scenarios that prioritize alternative sets of populations and evaluate the salmon portfolios along risk-return axes, as a financial portfolio manager might.
+We show that conserving a diversity of thermal tolerances buffers metapopulation risk given short-term climate forcing and ensures metapopulation persistence given long-term climate warming.
+We then show that dividing conservation among more populations buffers risk regardless of thermal-tolerance diversity or climate trend, but possibly at the expense of long-term growth rate and persistence when available habitat declines over time.
+We conclude that considering metapopulations through portfolio theory provides a useful additional dimension through which we can evaluate conservation strategies.
+
+\section{Methods}\label{methods}
+
+We developed a 100-year salmon metapopulation simulation model that includes both population dynamics and harvesting along with process, observation, and implementation uncertainty (Fig.~\ref{f:flowchart}).
+We tested different conservation scenarios under two kinds of environmental regimes (short-term climate variability and long-term climate change) and in cases where habitat capacity remained constant or declined over time.
+We provide a package \texttt{metafolio} \citep{metafoliopkg} for the statistical software \textsf{R} \citep{r2013} as an appendix, to carry out the simulations and analyses described in this paper (\somR).
+
+\subsection{Defining the ecological portfolio}\label{defining-the-ecological-portfolio}
+
+In our ecological portfolios, we defined assets as stream-level populations and portfolios as salmon metapopulations.
+The specific configuration of our model refers to salmon that spend extended time rearing in freshwaters (e.g.~steelhead {[}\emph{Oncorhynchus mykiss}{]}, sockeye salmon {[}\emph{O.
+nerka}{]}, coho salmon {[}\emph{O.
+kisutch}{]}, and stream-type Chinook salmon {[}\emph{O.
+tshawytscha}{]}), which will likely be more impacted by changes to stream temperature and flow \citep{mantua2010}.
+We use the terms \emph{stream} and \emph{populations} interchangeably to represent the portfolio assets.
+We defined the portfolio investors as the stakeholders in the fishery and metapopulation performance.
+For example, the investors could be conservation agencies, First Nations groups, or civil society as a whole.
+The fisheries management agency then becomes the portfolio manager.
+We defined the asset value as the abundance of returning salmon in each stream and the value of the portfolio as the overall metapopulation abundance.
+
+In this scenario, the equivalent to financial rate of return is the generation-to-generation metapopulation growth rate, calculated as the first difference of the log salmon returns.
+We defined the financial asset investment weights as the capacity of the stream populations --- specifically the unfished equilibrium stock size --- since maintaining or restoring habitat requires money, time, and resources and habitat size itself is a strong predictor of the occupancy of salmon \citep{isaak2007}.
+Investment in a population therefore represents investing in salmon habitat conservation or restoration and the risk and return from investment strategies become emergent properties of our metapopulation model.
+
+\subsection{Salmon metapopulation dynamics}\label{salmon-metapopulation-dynamics}
+
+The salmon metapopulation dynamics in our simulation were governed by a spawner-return relationship with demographic stochasticity and straying between populations.
+We defined the spawner-return relationship with a Ricker model,
+
+\[R_{i(t+1)} = S_{i(t)}e^{a_{i(t)}(1-S_{i(t)}/b_i) + w_{i(t)}}\]
+
+\noindent
+where $i$ represents a population, $t$ a generation time, $R$ the number of returns, $S$ the number of spawners, $a$ the productivity parameter (which can vary with the environment), and $b$ the density-dependent term (which is used as the asset weights in the portfolios).
+The term $w_{i(t)}$ represents first-order autocorrelated error.
+Formally, $w_{i(t)} = w_{i(t-1)} \rho_w + r_{i(t)}$, where $r_{i(t)}$ represents independent and normally-distributed error with standard deviation of $\sigma_r$, mean of $-\sigma_r^2 / 2$ (bias corrected so the expected value after exponentiation is 1), and correlation between subsequent generation values of $\rho_w$.
+We set $\sigma_r = 0.7$ and $\rho_w = 0.4$ to match the mean values for salmonids in \citet{thorson2014a}.
+
+We manipulated the capacity and productivity parameters $b_i$ and $a_{i(t)}$ as part of the portfolio simulation.
+The capacity parameters $b_i$ were controlled by the investment weights in the populations.
+For example, a large investment in a stream was represented by a larger unfished equilibrium stock size $b$ for stream $i$.
+The productivity parameters $a_{i(t)}$ were controlled by the interaction between a temperature time series and the population thermal-tolerance performance curves.
+In a different context, investment could represent improving the productivity ($a_i$) parameters, say through culling, to offset mortality increases due to changing temperatures.
+However, such a scenario is unlikely in the case of an endangered species where population levels are often well below levels where culling would increase productivity.
+
+We generated the thermal-tolerance curves according to
+
+\[a_{i(t)} = \begin{cases} a_i^{\mathrm{max}} -
+W_i (e_t - e_i^{\mathrm{opt}})^2,
+& \text{if } a_{i(t)} > 0\\ 0, & \text{if } a_{i(t)} \leq
+0 \end{cases}\]
+
+\noindent
+where $W_i$ controls the width of the curve for population $i$, $e_t$ represents the environmental value at generation $t$, $e_i^{\mathrm{opt}}$ represents the optimal temperature for population $i$, and $a_i^{\mathrm{max}}$ represents the maximum possible $a$ value for population $i$.
+We set the $W_i$ parameters (evenly spaced values increasing and decreasing between 0.08 and 0.04) to generate widths approximately as shown in \citet{eliason2011}.
+We set the area under each curve to 30 units to create $a_i^{\mathrm{max}}$ values ranging roughly between 2.2 and 2.9 as in \citet{dorner2008}.
+These parameter values created some warm-tolerant populations, some cold-tolerant populations, and some populations with a wider range of thermal-tolerance but a lower maximum productivity (Fig.~\ref{f:curves}a).
+Although we refer to a thermal-tolerance curve because temperature is a dominant driver of salmon productivity \citep[e.g.][]{mccullough1999, patterson2007, eliason2011}, our model could apply to any environmental tolerance (e.g.~tolerance to stream flow volume or changes in snow melt timing; \citeauthor{crozier2008} \citeyear{crozier2008}).
+
+We implemented straying as in \citet{cooper1999}.
+We arranged the populations in a line and salmon were more likely to stray to streams near their natal stream (\somstray).
+Two parameters controlled the straying: the fraction of fish $f_{\mathrm{stray}}$ (0.02) that stray from their natal stream in any generation and the rate $m$ (0.1) at which this straying between streams decays with distance
+
+\[\mathrm{strays}_{ij(t)} = f_{\mathrm{stray}} R_{j(t)} \frac{e^{-m \lvert i-j
+\rvert }} {\displaystyle\sum\limits_{ \substack{k = 1 \\ k \neq j}}^{n} e^{-m
+\lvert k-j \rvert }}\]
+
+\noindent
+where $R_{j(t)}$ is the number of returning salmon at generation $t$ whose natal stream was stream $j$.
+The subscript $k$ represents a stream ID and $n$ the number of populations.
+The denominator is a normalizing constant to ensure the desired fraction of fish stray.
+Our simulation did not account for the homogenization of diversity due to straying.
+For example, all salmon in one population maintained the same thermal-tolerance curve regardless of how many salmon it received from another stream.
+
+\subsection{Fishing}\label{fishing}
+
+Our simulation used a simple set of rules to establish the exploitation rate of fisheries and the remainder left to spawn (escapement target).
+First, to establish a range of spawner-return values and to mimic the start of an open-access fishery, for the first 30 years we drew the fraction of fish harvested randomly from a uniform distribution between 0.1 and 0.9.
+We discarded these initial 30 years as a burn-in period.
+Then, every five years for the remaining 100 years of our simulation, we fitted a spawner-return function to the cumulative data for individual populations.
+The target escapement rate $E_{\mathrm{tar}}$ (a proportion per year) was set based on \citet{hilborn1992} as
+
+\[E_{\mathrm{tar}} = \frac{R}{b (0.5 - 0.07a)} \label{eq:esc}\]
+
+\noindent
+where $R$ represents the return abundance and $a$ and $b$ represent the Ricker model parameters.
+The target harvest rate is then a function of returns and the escapement target ($H_{\mathrm{tar}} = R - E_{\mathrm{tar}}$).
+We included implementation uncertainty in the actual harvest rate $H_{\mathrm{act}}$ as a function of the target harvest rate and a beta distribution with location parameter $\alpha_h$, shape parameter $\beta_h$, and standard deviation of $\sigma_h$ (set to 0.1 as observed for similar data in \citet{pestes2008}).
+
+\[\begin{aligned}
+\alpha_h &= H_{\mathrm{tar}}^2 \left( \frac{1 -
+H_{\mathrm{tar}}}{\sigma_h^2} - \frac{1}{H_{\mathrm{tar}}} \right)\\ \beta_h
+&= \alpha_h \left({\frac{1}{H_{\mathrm{tar}}} - 1}\right)\\
+H_{\mathrm{act}} &= \mathrm{beta}(\alpha_h, \beta_h).
+\end{aligned}\]
+
+\subsection{Environmental dynamics}\label{environmental-dynamics}
+
+Environmental dynamics typically have both short- and long-term fluctuations, such as annual variability and directional climatic warming.
+We evaluated portfolio performance under these two components separately in our initial scenarios and combined in our final scenario.
+We did not explicitly model a cyclical climate trend, such as the Pacific Decadal Oscillation, but the effect of such a trend would largely be a product of the short-term variability and long-term trend.
+We represented short-term dynamics $e_{\mathrm{short}(t)}$ as a stationary first-order autoregressive process, AR(1), with correlation $\rho_e$ (0.1)
+
+\[e_{\mathrm{short}(t)} = e_{t-1} \rho_e + d_t, d_t \sim \mathrm{N}(\mu_d, \sigma_d^2)\]
+
+\noindent
+where $d_t$ represents normally distributed deviations of some mean $\mu_d$ and standard deviation $\sigma_d$.
+We set $\mu_d$ to $16\,^{\circ}\mathrm{C}$ and $\sigma_d$ to $2\,^{\circ}\mathrm{C}$, to approximately match the stream temperature variation in \citet{eliason2011}.
+We represented long-term environmental dynamics $e_{\mathrm{long}(t)}$ as a linear shift in the temperature through time
+
+\[e_{\mathrm{long}(t)} = e_0 + \beta_e t\]
+
+\noindent
+where $e_0$ represents the starting temperature up until the burn-in period ends and $\beta_e$ represents the annual increase in temperature.
+We set $e_0 = 15\,^{\circ}\mathrm{C}$ and $\beta_e = 0.04\,^{\circ}\mathrm{C} / \mathrm{generation}$ to obtain an increase in stream temperature of $4\,^{\circ}\mathrm{C}$ over the next century (assuming one generation equals one year) ending at or above the optimum thermal optimum of all populations.
+This increase approximately matches predicted increases in stream temperature --- relative to the 1980s, stream temperatures in the Pacific Northwest have already increased by approximately $0.2\,^{\circ}\mathrm{C}$/decade \citep{isaak2012}, and are predicted to increase $2$ to $5\,^{\circ}\mathrm{C}$ by 2080 \citep{mantua2010}.
+
+We summarize the chosen parameter values in \somparam.
+Combining salmon population dynamics, fishing, and environmental dynamics, we illustrate the components of an example simulation in Fig.~\ref{f:ts} and the effect of varying population, fishing, and environmental parameters from their base values on metapopulation abundance in \somsens.
+
+\subsection{Conservation scenarios}\label{conservation-scenarios}
+
+\emph{Spatial conservation scenarios}: We evaluated four spatial conservation scenarios (Fig.~\ref{f:curves}b--e).
+We conserved four populations ($b_i = 1000$) and set the unfished equilibrium abundance of the six remaining populations to near elimination ($b_i = 5$) at the start of the simulation.
+These reduced populations could still receive straying salmon but were unlikely to rebuild on their own to a substantial abundance.
+The four spatial scenarios we considered were:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  Conserve a full range of thermal tolerances (conserve some cool-, some intermediate-, and some warm-tolerant populations; Fig.~\ref{f:curves}b).
+\item
+  Conserve the middle section of the metapopulation (conserve the most thermal-tolerant populations with the widest response curves; Fig.~\ref{f:curves}c).
+\item
+  Conserve the lower half of the metapopulation (conserve cool-tolerant populations; Fig.~\ref{f:curves}d).
+\item
+  Conserve the upper half of the metapopulation (conserve warm-tolerant populations; Fig.~\ref{f:curves}e).
+\end{enumerate}
+
+\emph{Unknown thermal tolerances}: In reality we rarely know precise levels of thermal response diversity.
+We therefore also considered cases where conservation was randomly assigned with respect to thermal tolerance but where conservation effort ($\sum\limits_{i=1}^n b_i = 2000$) could be distributed across different numbers of streams.
+We considered conserving from two to 16 streams with thermal tolerance distributed along the same range as in the spatial scenarios.
+As in the spatial strategies, we reduced the capacity of the remaining streams to the nominal level of $b_i = 5$.
+
+\emph{Declining habitat availability}: Habitat capacity in the Pacific Northwest is likely shrinking over time as salmon populations are squeezed between warming temperatures reducing habitat from below and declining stream flows reducing the habitat that remains from above.
+For example, temperature isotherms are shifting upstream at 1--25 km/decade in low gradient streams that Chinook use for spawning \citep{isaak2013}.
+At the same time, summer-fall stream flow volumes have been decreasing 10--30\% across the Pacific Northwest over the past 50 years \citep{luce2009} and are likely to continue declining \citep{luce2013}.
+We therefore considered a scenario where habitat capacity declined by a constant amount across all populations.
+We reduced the $b$ parameters by 0.85 units per generation so that some of the smaller populations would reach near extinction by the end of the simulation, as is likely for smaller isolated populations within this century \citep[e.g.][]{gustafson2007}.
+In this scenario, we considered cases where thermal tolerance was unknown but conservation effort could be distributed across between 16 and two streams.
+Climate followed a combination of the same long-term warming and short-term variability as before.
+For many Pacific salmon metapopulations, this scenario represents the most realistic scenario investigated.
+
+\section{Results}\label{results}
+
+\subsection{Spatial conservation scenarios}\label{spatial-conservation-scenarios}
+
+\emph{Given short-term environmental fluctuations} (strong interannual variation), conserving a wide range of thermal tolerances is the safest choice because it reduces overall risk to an ecological portfolio (Fig.~\ref{f:sp}a; \somts~Figs.~\ref{f:eg-sp-arma-full}, \ref{f:eg-sp-arma-half}).
+The average variance of metapopulation growth rate was 1.6 times lower given balanced thermal tolerance conservation (conserving a full range of thermal tolerances or the middle section vs.~the upper or lower half).
+Thermal tolerance diversity also led to more consistent stability --- there was less spread in variance across simulated metapopulations (width of quantiles from left to right in Fig.~\ref{f:sp}a).
+These increases in stability occurred despite the portfolios being comprised of warm- and cool-thriving populations that individually showed greater variation in response to environmental variability than populations with wide thermal tolerance curves.
+We can see the mechanism behind these portfolio properties by inspecting example population time series (Fig.~\ref{f:sp}c, d).
+If only the upper or lower half of thermal tolerances is conserved, the portfolio tends to alternate between performing well and poorly, depending on the environmental conditions, resulting in a riskier portfolio (Fig.~\ref{f:sp}e).
+This risk is buffered when a diversity of thermal tolerances is conserved (Fig.~\ref{f:sp}c) and the resulting asynchrony in population abundance (\somcor).
+
+\emph{Given long-term environmental change}, such as climate warming, an ecological manager is hedging his or her bets on the environmental trend and how the populations will respond by conserving a range of thermal tolerances.
+The choice of which populations to conserve affects the ``rate of return'' (metapopulation growth rate) properties of an ecological portfolio (Fig.~\ref{f:sp}b; \somts~Figs.~\ref{f:eg-sp-linear-full}, \ref{f:eg-sp-linear-half}).
+The typical metapopulation growth rate when thermal tolerances were balanced was near zero --- the metapopulation neither increased nor decreased in abundance in the long run.
+The example metapopulation abundance time series (Fig.~\ref{f:sp}d, f) illustrate the mechanism: by conserving a range of thermal tolerances, when one population is doing poorly, another is doing well and the metapopulation abundance remains stationary through time.
+If a manager had invested only in the populations that were doing well at the beginning they would have had the lowest metapopulation growth rate at the end (purple portfolios in Fig.~\ref{f:sp}f).
+
+\subsection{Unknown thermal tolerances}\label{unknown-thermal-tolerances}
+
+In a scenario where the distribution of population-level thermal tolerances are unknown, portfolio optimization informs us that investing in more populations buffers portfolio risk regardless of environmental trend (Fig.~\ref{f:n}).
+\emph{Given short-term environmental fluctuations}, conserving more populations buffers portfolio risk (Fig.~\ref{f:n}a, c, d; \somts~Figs.~\ref{f:eg-n-arma-two}, \ref{f:eg-n-arma-sixteen}).
+For example, a metapopulation with 16 conserved populations is on average 1.7 times less variable than a metapopulation with only eight.
+At the same time, the random conservation of thermal tolerances creates an increased spread of possible metapopulation risk given fewer populations conserved (increasing quantile width from left to right in Fig.~\ref{f:n}a).
+
+\emph{Given long-term environmental change}, conserving more populations also buffers portfolio risk (Fig.~\ref{f:n}b; \somts~Figs.~\ref{f:eg-n-linear-two}, \ref{f:eg-n-linear-sixteen}).
+Furthermore, in comparison to the short-term environmental noise scenario, the long-term environmental change creates a greater spread of possible metapopulation growth rates.
+For example, the height of the 75\% quantile of the mean metapopulation growth rate for the two-population systems (light grey polygons) is larger given long-term change than short-term change.
+
+\subsection{Declining habitat availability}\label{declining-habitat-availability}
+
+Given a reduction in stream flow over time along with climate change and climate variability, a manager encounters a risk-return trade-off when deciding how many populations to distribute conservation efforts across (Fig.~\ref{f:squeeze}; \somts~Figs.~\ref{f:eg-n-squeeze-two}, \ref{f:eg-n-squeeze-twelve}).
+Conserving more populations buffers portfolio risk, but at the expense of expected metapopulation growth rate.
+For example, the mean metapopulation variance was 2.7 times lower when 12 populations were conserved instead of four, but the expected metapopulation growth rate was 2.0 times lower when 16 populations were conserved instead of eight.
+The conservation scenarios represent an efficient frontier where a manager must choose whether to hedge his or her bets on a smaller number of populations and take on greater expected variability or conserve more populations and accept a lower expected metapopulation growth rate.
+
+\section{Discussion}\label{discussion}
+
+The importance of conserving populations with a diversity of responses to the environment is a key assumption of conservation ecology, but has rarely been tested quantitatively \citep{mori2013}.
+We show how maintaining populations with a variety of thermal tolerances reduces risk caused by short-term environmental stochasticity and optimizes chances for long-term persistence given climate change.
+Further, conserving more populations reduces metapopulation variability but possibly at the expense of long-term metapopulation growth rate if available habitat is squeezed by climate change.
+In this discussion, we begin by linking our model with real-world conservation issues for Pacific Northwest salmon.
+We then consider broader implications for metapopulation conservation of any species and ecological stability in general.
+
+\subsection{Implications for salmon conservation}\label{implications-for-salmon-conservation}
+
+Our results emphasize the importance of promoting ecological conditions that promote diversity of environmental response to the environment if stability is to be maintained in the face of environmental uncertainty.
+This suggests three clear conservation actions.
+First, since habitat heterogeneity can lead to local adaptation \citep[e.g.][]{fraser2011}, our results emphasize the need to maintain a diversity of salmon habitat \citep{rogers2008}.
+Second, if conservation actions must be prioritized, then our model suggests we should focus on populations that aren't spatially contiguous to maximize diversity of response to the environment.
+Third, our results demonstrate the advantages of avoiding structures that artificially remove diversity of environmental response.
+For salmon, dams are a prominent example \citep{mcclure2008a}.
+Dams can have a double impact whereby their introduction selectively eliminates a large swath of contiguous habitat, perhaps analogous to our upper- or lower-half scenarios in Fig.~\ref{f:sp}, and then mitigation approaches such as hatcheries can further reduce response diversity if not carefully managed \citep{mcclure2008b}.
+In fact, salmon habitat lost to dams in the western U.S.
+has been biased towards warmer, drier, higher habitats \citep{mcclure2008a} and our findings suggest the resulting loss of warm-tolerant species may compound the risk to current metapopulations in the face of global warming.
+
+The goals of existing salmon management structures in the western US and Canada support a portfolio conservation perspective.
+In the US, salmon populations are divided into Evolutionarily Significant Units (ESUs), groups of populations that are reproductively isolated and share a common evolutionary heritage, and finer-scale Viable Salmonid Populations (VSPs), populations that are demographically independent of other populations over a 100-year time frame \citep{mcelhany2000}.
+In Canada, the rough equivalent to the ESU is a Conservation Unit (CU), which consists of a group of salmon that are reproductively isolated and that if lost would be unlikely to recolonize in a reasonable time frame \citep{dfo2005wsp}.
+A salmon portfolio in our model could represent an ESU or CU and the lessons learned from our models are thus directly applicable to management guidelines in the Pacific Northwest.
+In fact, a number of VSP guidelines agree with our findings.
+For example, VSP guidelines suggest maintaining diversity in a variety of forms, focusing conservation efforts not just where salmon are currently abundant, and maintaining metapopulations with some populations near each other and others further apart \citep{mcelhany2000}.
+
+However, salmon populations in the Pacific Northwest are already heavily impacted \citep[e.g.][]{gustafson2007} and VSP and CU recovery goals have not yet been achieved for most populations.
+Since European-Americans arrived, 29\% of 1400 historical salmon populations in the Pacific Northwest and California have been lost \citep{gustafson2007}.
+Furthermore, 44\% of salmon habitat in the western US (in the lower 48 states) has been lost to dams and other freshwater blockages \citep{mcclure2008a}.
+Changes to habitat, combined with increasing climate variability, has led to disturbance regimes that differ substantially in the frequency, magnitude, and duration from historical patterns, and threaten the resilience of salmon populations \citep{waples2009}.
+Many remaining populations rely on hatcheries for long-term population viability --- creating substantial evolutionary risks such as outbreeding depression, genetic homogenization, reduced effective population size, and domestication of fish (adaption to artificial environments and reduced fitness in wild environments) \citep{mcclure2008b}.
+Reduction of long-term reliance on hatcheries, accompanied by habitat restoration through, for example, restoring connectivity of floodplains and stream flow regimes, remains a critical component of long-term salmon sustainability in the Pacific Northwest --- particularly given predicted patterns of climate change \citep{beechie2013}.
+
+Our model complements other simulation-based salmon-habitat prioritization models.
+While these other models tend to focus on detailed assessment of individual fish stocks, our model is the first to consider the role of response diversity in buffering risk for metapopulations as a whole.
+The Shiraz model is one complementary prioritization scheme \citep{scheuerell2006}.
+It focuses on detailed conditioning of the habitat-population-dynamics relationship at multiple life-history stages for a single salmon population.
+Whereas the Shiraz model can be applied to an entire watershed, it combines the populations together as a single unit thereby ignoring the role of population-level environmental response diversity.
+A second salmon prioritization model proposes combining population viability measures with an assessment of the genetic consequences of losing particular populations \citep{allendorf1997}.
+This model, however, also focuses on the assessment of individual stocks without considering their covariance and therefore the performance of the salmon portfolio as a whole.
+Our model does not replace these prioritization schemes.
+Rather, it proposes an additional focus on prioritization that optimizes metapopulation growth \emph{and} risk and that considers diversity of tolerance to environmental conditions.
+
+While our model captures many relevant aspects of salmon life history and environmental dynamics, it ignores others that could be investigated in future analyses and might improve our understanding of salmon portfolio conservation.
+First, some salmon populations, such as ocean-type Chinook, tend to spawn further downstream than stream-type salmon.
+Ocean-type Chinook may therefore be less affected by declining stream flow and be able to shift upstream to avoid shifting isotherms \citep{mantua2010}.
+A model could consider evolutionary adaptation by having populations adopt more ocean-type-like characteristics.
+Second, our model ignores lost thermal-tolerance diversity from populations that reach low population sizes and are reestablished by straying from nearby streams.
+An individual-based model might more accurately penalize for this lost diversity and emphasize the need to define lower limits on the investment weights in a salmon conservation portfolio.
+Third, our model ignores fine-scale within-stream spatial and temporal environmental fluctuations.
+Fine-scale extremes in temperature and stream flow may be particularly important to population dynamics \citep{mantua2010} and could be incorporated into a future analysis.
+Such a model might show an increased benefit of portfolio optimization if the impact of increased magnitude and frequency of local climate extremes is important in addition to the mean trend \citep{jentsch2007}.
+
+\subsection{Broad ecological implications and conservation priorities}\label{broad-ecological-implications-and-conservation-priorities}
+
+To promote the stabilizing effect of a diversified ecological portfolio, there are two key components to identify: (1) the environmental drivers to which a varied response might occur, and (2) the conservation actions that can increase or decrease the diversity of response.
+A third component, identifying the traits and behaviours that mediate population responses to the environment may provide further insight into the mechanisms.
+Environmental drivers of response can include, for example, changes to temperature, habitat availability, air quality, water chemistry, or extreme weather \citep{elmqvist2003}.
+Identifying conservation actions that promote environmental response diversity is critical to developing stable ecological systems \citep{mori2013}.
+However, merely measuring environmental response diversity in real ecological systems is challenging \citep[albeit possible;][]{thibaut2012}.
+Therefore, one realistic solution may be to create general guidelines from a small number of intensively-monitored systems in which we can associate changes in synchrony of populations with changes in conservation regimes \citep[e.g.][]{moore2010, carlson2011}.
+Another solution may be to monitor the diversity of environmental conditions themselves (e.g.~temperature, stream flow, and gravel size in the case of salmon) since we know that traits affecting response to environmental conditions are heritable and are likely to adapt to local conditions \citep{carlson2011} possibly producing diversity of response to subsequent disturbances.
+
+We suggest a number of specific extensions to our simulation model.
+First, the environment-thermal-tolerance mechanism could be expanded --- the distribution of environmental tolerance across a metapopulation does not necessarily follow a linear gradient, different forms of environmental tolerance could interact, and environmental conditions could affect populations through mechanisms other than productivity.
+Second, in addition to other taxa, our model could be extended to ecological communities or meta-communities after accounting for species interactions.
+Third, without any modifications, our model could consider the Moran or environmental-filter concept whereby populations experience increasingly different environmental forces at further distances \citep{schindler2008, rogers2008}.
+Fourth, a model could consider the contribution of contemporary evolution \citep{stockwell2003}.
+These rapid adaptations to changes in the environment could strongly influence portfolio performance and emphasize the importance of maintaining genetic diversity and a variety of local habitat.
+Finally, our model could be conditioned on a system of interest --- say a particular river basin in our example --- and the metapopulation portfolio could be optimized across conservation and restoration options as part of a formal decision analysis.
+
+Management decisions for exploited species often come with a trade-off between conservation and revenue generation.
+Our findings when habitat capacity declined over time illustrate another kind of trade-off more similar to the trade-off described by \citet{markowitz1952} in his seminal financial portfolio work.
+In this case, managers must navigate a trade-off between expected risk and return of the metapopulation/portfolio growth rate itself.
+No position along this trade-off is inherently better than another unless considered in the context of societal values.
+Does society value short-term stability or a greater assurance of long-term persistence?
+The optimal choice likely lies somewhere in the middle and parameterizing our model to a specific metapopulation could illustrate the nature of the trade-off and aid conservation decision making.
+However, if environmental tolerance could be targeted for conservation as in Fig.~\ref{f:sp}, a manager could likely achieve portfolios closer to the efficient frontier in Fig.~\ref{f:squeeze}.
+In other words, a manager could achieve a lower expected variance for the same expected growth rate or a higher expected growth rate for the same expected variance --- a better conservation outcome in either case.
+
+Conservation planning is inherently a spatial activity \citep{pressey2007} and our results can inform how we approach spatial conservation planning.
+First, our results suggest focusing on conserving the processes and mechanisms underlying stability, not just biodiversity itself \citep{pressey2007, beechie2013}.
+In particular, our results suggest that response diversity should be a mainstream element of conservation, not just species and functional diversity \citep{mori2013}.
+Our analysis also illustrates how conserving a portfolio of populations, ideally selected for a wide range of environmental tolerance, can help integrate across environmental uncertainty when spatial planning \citep{ando2012}.
+This is particularly important given the uncertainty surrounding the future ecological responses to climate change \citep{walther2002}.
+Finally, the increasing rapidness and variability of environmental change necessitates a dynamic approach in which spatial planning is reevaluated at regular intervals \citep{hannah2002a} --- perhaps testing for changes in population and species asynchrony in addition to changes in local productivity and variability.
+Combined, our results detail a pathway through which population diversity in environmental tolerance can underpin the stability of ecological systems.
+This pathway highlights that diverse natural portfolios may be critical for the conservation of metapopulations in the face of increasing climate variability and change.
+
+\section{Acknowledgements}\label{acknowledgements}
+
+We thank T.A.
+Branch, J.D.
+Yeakel, S.M.
+O'Regan, S.A.
+Pardo, L.N.K.
+Davidson, and C.C.
+Phillis for helpful discussions and comments on earlier drafts.
+We thank D.J.
+Isaak, an anonymous reviewer, and O.P.
+Jensen for suggestions that greatly improved the manuscript.
+We are particularly grateful to D.J.
+Isaak for suggesting and carefully outlining the declining stream flow scenario.
+Funding was provided by Simon Fraser University, NSERC (ABC, NKD, SCA), the Canada Research Chairs Program (NKD), the Liber Ero Chair of Coastal Science and Management (JWM), Fulbright Canada (SCA), and a Garfield Weston Foundation/B.C.
+Packers Ltd.
+Graduate Fellowship in Marine Sciences (SCA).
 
@@
 \begin{center}
-\includegraphics[height=5.5in]{../examples/simulation-diagram3-sansserif-bw}\\
+\includegraphics[height=5.5in]{Fig1}\\
 \textsc{Fig.} 1
 \clearpage
-\includegraphics[width=2.4in]{../examples/thermal-curve-scenarios}\\
+\includegraphics[width=2.4in]{Fig2}\\
 \textsc{Fig.} 2
 \clearpage
-\includegraphics[width=4.2in]{../examples/spatial-arma-sim-full}\\
+\includegraphics[width=4.2in]{Fig3}\\
 \textsc{Fig.} 3
 \clearpage
-\includegraphics[width=4.5in]{../examples/spatial-mv}\\
+\includegraphics[width=4.5in]{Fig4}\\
 \textsc{Fig.} 4
 \clearpage
-\includegraphics[width=4.5in]{../examples/cons-plans-n}\\
+\includegraphics[width=4.5in]{Fig5}\\
 \textsc{Fig.} 5
 \clearpage
-\includegraphics[width=3.0in]{../examples/cons-plans-squeeze}\\
+\includegraphics[width=3.0in]{Fig6}\\
 \textsc{Fig.} 6

### 2014-07-10 inst/ms/appendices/pdf2eps.R
Add pdf2eps script

@@ -1,0 +1,6 @@
+files <- list.files(path = ".", pattern = "Fig[A-Z0-9\\-]*\\.pdf")
+
+for(i in 1:length(files)) {
+  message(files[i])
+  system(paste("pdftops -eps", files[i]))
+}

### 2014-08-04 inst/examples/suppl-1.Rnw
Edits for ESA editing; see details
Move first appendix to supplement
Reorder appendices
Create new supplement describing code

\documentclass[10pt]{article}
\usepackage{geometry}
\usepackage{url}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}

\setlength{\parskip}{10pt}
\setlength{\parindent}{0cm}

\textheight 21.0cm

\begin{document}

<<setup, include=FALSE, cache=FALSE>>=
library(knitr)
# set global chunk options
opts_chunk$set(fig.path='figure/', cache=TRUE, autodep = TRUE,
  fig.align='center', fig.width=5, fig.height=5,
  messages=FALSE, results=FALSE)
opts_knit$set(out.format = "latex")
opts_chunk$set(warning=FALSE, tidy=FALSE, refresh=TRUE)
@

\title{Recreating the analyses with the \texttt{metafolio} \textsf{R} package}

\author{
Sean C. Anderson\textsuperscript{1*},
Jonathan W. Moore\textsuperscript{1,2},
Michelle M. McClure\textsuperscript{3},\\
Nicholas K. Dulvy\textsuperscript{1}
Andrew B. Cooper\textsuperscript{2}
}

\date{}
\maketitle

\textsuperscript{1}Earth to Ocean Research Group, Department of Biological Sciences, Simon Fraser University, Burnaby BC, V5A 1S6, Canada

\textsuperscript{2}School of Resource and Environmental Management, Simon Fraser University, Burnaby, BC, V5A 1S6, Canada

\textsuperscript{3}Fishery Resource Analysis and Monitoring Division, Northwest Fisheries Science Center, National Marine Fisheries Service, National Oceanic and Atmospheric Administration, 2725 Montlake Blvd. E., Seattle, WA 98112, USA

\noindent
\textsuperscript{*}Corresponding author

\bigskip

This supplement accompanies the paper \textit{Portfolio conservation of metapopulations under climate change} in Ecological Applications.

\section{Loading the package}

The \texttt{metafolio} \textsf{R} package is a tool to simulate metapopulations
and apply financial portfolio optimization concepts to those metapopulations.
The package is primarily intended for salmon simulations, but could be adapted
for other taxonomic groups. In this document, we describe how to install and
load the package, and run the analyses in our paper.

\texttt{metafolio} is available on CRAN:
\url{http://cran.r-project.org/package=metafolio} and can be installed with:

<<install, eval=FALSE>>=
install.packages("metafolio")
@

The analyses in this paper used \texttt{metafolio} version \texttt{0.1.0}.

Alternatively, you can view the code and install the package from\\ \texttt{http://github.com/seananderson/metafolio}. Installing the package from GitHub will require that you have a \texttt{C++} compiler installed.

\clearpage
Load the package with:

<<load>>=
library("metafolio")
@

The included vignette describes the package and illustrates some example simulations. You can view the vignette with:

<<vignette, eval=FALSE>>=
vignette("metafolio")
@

You can view the help for the package with:

<<help, eval=FALSE>>=
?metafolio
help(package = "metafolio")
@

The figures from this paper can be re-created by downloading the source code from CRAN or GitHub and sourcing the file \texttt{README.R} in the \texttt{inst/examples} folder:

<<source-readme, eval=FALSE>>=
setwd("metafolio/inst/examples")
source("README.R")
@

We have copied some examples from that code below. See the source code on GitHub for the full analysis. To save time compiling this document, we have run fewer iterations in these examples than in the paper.

\section{Prioritizing spatial response diversity}

Setup the scenarios:

<<setup-response-diversity>>=
set.seed(1)
w_plans <- list()
w_plans[[1]] <- c(5, 1000, 5, 1000, 5, 5, 1000, 5, 1000, 5)
w_plans[[2]] <- c(5, 5, 5, 1000, 1000, 1000, 1000, 5, 5, 5)
w_plans[[3]] <- c(rep(1000, 4), rep(5, 6))
w_plans[[4]] <- rev(w_plans[[3]])
plans_name_sp <- c("Full response range", "Most stable only",
  "Lower half", "Upper half")
n_trials <- 200 # number of trials at each conservation plan
num_pops <- c(10, 10, 10, 10)
n_plans <- length(num_pops) # number of plans
w <- list()
for(i in 1:n_plans) { # loop over plans
 w[[i]] <- list()
 for(j in 1:n_trials) { # loop over trials
   w[[i]][[j]] <- matrix(w_plans[[i]], nrow = 1)
 }
}
@

Here's one example of generating a time series plot:

<<ts-example, fig.width=5, fig.height=7, out.width="4in">>=
set.seed(123)
arma_env_params <- list(mean_value = 16, ar = 0.1, sigma_env = 3, ma = 0)
eg_arma <- meta_sim(b = w[[1]][[1]], n_pop = 10, env_params = arma_env_params,
  env_type = "arma", assess_freq = 5)
plot_sim_ts(eg_arma, years_to_show = 100, burn = 30)
@

Now, we'll run the simulations:

<<run-resp-diversity>>=
x_arma_sp <- run_cons_plans(w, env_type = "arma", env_params =
  arma_env_params, show_progress = FALSE)
x_arma_sp$plans_port <- NULL # save space

linear_env_params <- list(min_value = 15, max_value = 19, sigma_env = 0.001,
  start_t = 30)
x_linear_sp <- run_cons_plans(w, env_type = "linear", env_params =
  linear_env_params, max_a = thermal_integration(10),
  show_progress = FALSE)
x_linear_sp$plans_port <- NULL # save space
@

And plot the output:

<<plot-resp-div, fig.width=8>>=
cols <- RColorBrewer::brewer.pal(5, "Dark2")
xlim <- c(0.18, 0.80)
ylim <- c(-0.027, 0.027)
par(mfrow = c(1, 2))
par(las = 1, cex = 0.8, mar = c(0, 0, 0, 0), oma = c(4, 5.2, 1.5, .5),
  tck = -0.02, mgp = c(2, .6, 0))
plot_cons_plans(x_arma_sp$plans_mv, plans_name = plans_name_sp, cols = cols,
  add_all_efs = FALSE, xlim = xlim, ylim = ylim, add_legend = FALSE)
mtext("(a) Short-term environmental fluctuations", side = 3, line = 0.2, cex =
  0.8, adj = 0.05)
par(las = 0)
mtext("Mean of metapopulation growth rate", side = 2, line = 3, outer = FALSE,
  cex = 0.8)
par(las = 1)
plot_cons_plans(x_linear_sp$plans_mv, plans_name = plans_name_sp, cols = cols,
  add_all_efs = FALSE, xlim = xlim, ylim = ylim, y_axis = FALSE, add_legend =
  TRUE, legend_pos = "bottomright")
mtext("(b) Long-term environmental change", side = 3, line = 0.2, cex = 0.8,
  adj = 0.05)
mtext("Variance of metapopulation growth rate", side = 1, line = 2.25,
  outer = FALSE, cex = 0.8, adj = -1)
@

\section{Numbers of populations conserved}

Here, we'll show the effect of increasing the number of streams that conservation effort is distributed across. We'll conserve the populations without knowing how response diversity is distributed.

We'll set up the scenarios:

<<number, out.width="4in", fig.width=5, fig.height=7>>=
set.seed(1)
# In this version we start with a set amount of b and can split it up among many
# or invest it in a few
n_trials <- 200 # number of trials at each n conservation plan
num_pops <- c(2, 4, 8, 16) # n pops to conserve
b_conserve <- 2000 / num_pops
n_plans <- length(num_pops) # number of plans
w <- list()
for(i in 1:n_plans) { # loop over number conserved
 w[[i]] <- list()
 for(j in 1:n_trials) { # loop over trials
   w[[i]][[j]] <- matrix(rep(b_conserve[i], 16), nrow = 1)
   # conserve num_pops[i] populations; wipe out rest:
   w[[i]][[j]][-sample(1:16, num_pops[i])] <- 5
 }
}
plans_name_n <- paste(num_pops, "populations")
@

Now we'll run the simulations for the stationary climate variability and the non-stationary scenarios:

<<run-numbers>>=
x_arma_n <- run_cons_plans(w, env_type = "arma", env_params =
  arma_env_params, max_a = thermal_integration(16),
  show_progress = FALSE)
x_arma_n$plans_port <- NULL # to save space
@

<<linear>>=
linear_env_params <- list(min_value = 15, max_value = 19, sigma_env = 0.001,
  start_t = 30)
x_linear_n <- run_cons_plans(w, env_type = "linear",
  env_params = linear_env_params, max_a = thermal_integration(16),
  show_progress = FALSE)
x_linear_n$plans_port <- NULL # save space
@

And plot the output:

<<plot-number, fig.width=8>>=
cols <- RColorBrewer::brewer.pal(5, "Greys")[c(2:5)]
xlim <- c(0.008, 0.90)
ylim <- c(-0.034, 0.027)
par(mfrow = c(1, 2))
par(las = 1, cex = 0.8, mar = c(0, 0, 0, 0), oma = c(4, 5.2, 1.5, .5),
  tck = -0.02, mgp = c(2, .6, 0))
plot_cons_plans(x_arma_n$plans_mv, plans_name = plans_name_n, cols = cols,
  add_all_efs = FALSE, xlim = xlim, ylim = ylim, add_legend = FALSE)
mtext("(a) Short-term environmental fluctuations", side = 3, line = 0.2,
  cex = 0.8, adj = 0.05)
par(las = 0)
mtext("Mean of metapopulation growth rate", side = 2, line = 3,
  outer = FALSE, cex = 0.8)
par(las = 1)
plot_cons_plans(x_linear_n$plans_mv, plans_name = plans_name_n, cols = cols,
  add_all_efs = FALSE, xlim = xlim, ylim = ylim, y_axis = FALSE, add_legend = TRUE)
mtext("(b) Long-term environmental change", side = 3, line = 0.2, cex = 0.8,
  adj = 0.05)
par(xpd = NA)
mtext("Variance of metapopulation growth rate", side = 1, line = 2.25,
  outer = FALSE, cex = 0.8, adj = -1)
@

\section{Reduction of habitat over time}

Here we'll reduce the available habitat by a fixed quantity at every generation. The climate will follow a combination of short-term variability and long-term change.

Set up the scenarios:

<<squeeze-setup>>=
set.seed(1)
n_trials <- 200 # number of trials at each n conservation plan
num_pops <- c(2, 4, 8, 12, 16) # n pops to conserve
b_conserve <- 2000 / num_pops
n_plans <- length(num_pops)
w <- list()
for(i in 1:n_plans) { # loop over number conserved
 w[[i]] <- list()
 for(j in 1:n_trials) { # loop over trials
   w[[i]][[j]] <- matrix(rep(b_conserve[i], 16), nrow = 1)
   # conserve num_pops[i] populations; wipe out rest:
   w[[i]][[j]][-sample(1:16, num_pops[i])] <- 5
 }
}
plans_name_n <- paste(num_pops, "populations")
cols <- RColorBrewer::brewer.pal(6, "Greys")[c(2:6)]
@

Run the simulations:

<<sqeeze-run>>=
linear_arma_env_params <- list(min_value = 15, max_value = 19,
  start_t = 30, mean_value = 16, ar = 0.1, sigma_env = 2, ma = 0)
x_linear_arma_n <- run_cons_plans(w, env_type = "linear_arma", env_params =
  linear_arma_env_params, max_a = thermal_integration(16), decrease_b = 0.85,
  show_progress = FALSE)
x_linear_arma_n$plans_port <- NULL # save space
@

Plot the output:

<<squeeze-plot, out.width="3.75in">>=
xlim <- c(0.08, 0.9)
ylim <- c(-0.038, 0.028)
par(las = 1, cex = 0.8, mar = c(0, 0, 0, 0), oma = c(4, 5.2, 1.8, .5),
  tck = -0.02, mgp = c(2, .5, 0))
plot_cons_plans(x_linear_arma_n$plans_mv, plans_name = plans_name_n, cols = cols,
  add_all_efs = FALSE, xlim = xlim, ylim = ylim, add_legend = TRUE,
  add_poly = TRUE, legend_pos = "bottomright")
mtext("Reduction in stream flow", side = 3, line = .4,
  cex = 0.8, adj = 0.05)
mtext("Variance of metapopulation growth rate", side = 1, line = 2.25,
  outer = FALSE, cex = 0.8)
par(las = 0)
mtext("Mean of metapopulation growth rate", side = 2, line = 3,
  outer = FALSE, cex = 0.8)
@

\end{document}

### 2014-08-04 inst/examples/suppl-1.tex
Edits for ESA editing; see details
Move first appendix to supplement
Reorder appendices
Create new supplement describing code

\documentclass[10pt]{article}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}
\usepackage{geometry}
\usepackage{url}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}

\setlength{\parskip}{10pt}
\setlength{\parindent}{0cm}

\textheight 21.0cm
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

\title{Recreating the analyses with the \texttt{metafolio} \textsf{R} package}

\author{
Sean C. Anderson\textsuperscript{1*},
Jonathan W. Moore\textsuperscript{1,2},
Michelle M. McClure\textsuperscript{3},\\
Nicholas K. Dulvy\textsuperscript{1}
Andrew B. Cooper\textsuperscript{2}
}

\date{}
\maketitle

\textsuperscript{1}Earth to Ocean Research Group, Department of Biological Sciences, Simon Fraser University, Burnaby BC, V5A 1S6, Canada

\textsuperscript{2}School of Resource and Environmental Management, Simon Fraser University, Burnaby, BC, V5A 1S6, Canada

\textsuperscript{3}Fishery Resource Analysis and Monitoring Division, Northwest Fisheries Science Center, National Marine Fisheries Service, National Oceanic and Atmospheric Administration, 2725 Montlake Blvd. E., Seattle, WA 98112, USA

\noindent
\textsuperscript{*}Corresponding author

\bigskip

This supplement accompanies the paper \textit{Portfolio conservation of metapopulations under climate change} in Ecological Applications.

\section{Loading the package}

The \texttt{metafolio} \textsf{R} package is a tool to simulate metapopulations
and apply financial portfolio optimization concepts to those metapopulations.
The package is primarily intended for salmon simulations, but could be adapted
for other taxonomic groups. In this document, we describe how to install and
load the package, and run the analyses in our paper.

\texttt{metafolio} is available on CRAN:
\url{http://cran.r-project.org/package=metafolio} and can be installed with:

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{install.packages}\hlstd{(}\hlstr{"metafolio"}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}

The analyses in this paper used \texttt{metafolio} version \texttt{0.1.0}.

Alternatively, you can view the code and install the package from\\ \texttt{http://github.com/seananderson/metafolio}. Installing the package from GitHub will require that you have a \texttt{C++} compiler installed.

\clearpage
Load the package with:

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{library}\hlstd{(}\hlstr{"metafolio"}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}

The included vignette describes the package and illustrates some example simulations. You can view the vignette with:

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{vignette}\hlstd{(}\hlstr{"metafolio"}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}

You can view the help for the package with:

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlopt{?}\hlstd{metafolio}
\hlkwd{help}\hlstd{(}\hlkwc{package} \hlstd{=} \hlstr{"metafolio"}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}

The figures from this paper can be re-created by downloading the source code from CRAN or GitHub and sourcing the file \texttt{README.R} in the \texttt{inst/examples} folder:

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{setwd}\hlstd{(}\hlstr{"metafolio/inst/examples"}\hlstd{)}
\hlkwd{source}\hlstd{(}\hlstr{"README.R"}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}

We have copied some examples from that code below. See the source code on GitHub for the full analysis. To save time compiling this document, we have run fewer iterations in these examples than in the paper.

\section{Prioritizing spatial response diversity}

Setup the scenarios:

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{set.seed}\hlstd{(}\hlnum{1}\hlstd{)}
\hlstd{w_plans} \hlkwb{<-} \hlkwd{list}\hlstd{()}
\hlstd{w_plans[[}\hlnum{1}\hlstd{]]} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlnum{5}\hlstd{,} \hlnum{1000}\hlstd{,} \hlnum{5}\hlstd{,} \hlnum{1000}\hlstd{,} \hlnum{5}\hlstd{,} \hlnum{5}\hlstd{,} \hlnum{1000}\hlstd{,} \hlnum{5}\hlstd{,} \hlnum{1000}\hlstd{,} \hlnum{5}\hlstd{)}
\hlstd{w_plans[[}\hlnum{2}\hlstd{]]} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlnum{5}\hlstd{,} \hlnum{5}\hlstd{,} \hlnum{5}\hlstd{,} \hlnum{1000}\hlstd{,} \hlnum{1000}\hlstd{,} \hlnum{1000}\hlstd{,} \hlnum{1000}\hlstd{,} \hlnum{5}\hlstd{,} \hlnum{5}\hlstd{,} \hlnum{5}\hlstd{)}
\hlstd{w_plans[[}\hlnum{3}\hlstd{]]} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlkwd{rep}\hlstd{(}\hlnum{1000}\hlstd{,} \hlnum{4}\hlstd{),} \hlkwd{rep}\hlstd{(}\hlnum{5}\hlstd{,} \hlnum{6}\hlstd{))}
\hlstd{w_plans[[}\hlnum{4}\hlstd{]]} \hlkwb{<-} \hlkwd{rev}\hlstd{(w_plans[[}\hlnum{3}\hlstd{]])}
\hlstd{plans_name_sp} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlstr{"Full response range"}\hlstd{,} \hlstr{"Most stable only"}\hlstd{,}
  \hlstr{"Lower half"}\hlstd{,} \hlstr{"Upper half"}\hlstd{)}
\hlstd{n_trials} \hlkwb{<-} \hlnum{200} \hlcom{# number of trials at each conservation plan}
\hlstd{num_pops} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlnum{10}\hlstd{,} \hlnum{10}\hlstd{,} \hlnum{10}\hlstd{,} \hlnum{10}\hlstd{)}
\hlstd{n_plans} \hlkwb{<-} \hlkwd{length}\hlstd{(num_pops)} \hlcom{# number of plans}
\hlstd{w} \hlkwb{<-} \hlkwd{list}\hlstd{()}
\hlkwa{for}\hlstd{(i} \hlkwa{in} \hlnum{1}\hlopt{:}\hlstd{n_plans) \{} \hlcom{# loop over plans}
 \hlstd{w[[i]]} \hlkwb{<-} \hlkwd{list}\hlstd{()}
 \hlkwa{for}\hlstd{(j} \hlkwa{in} \hlnum{1}\hlopt{:}\hlstd{n_trials) \{} \hlcom{# loop over trials}
   \hlstd{w[[i]][[j]]} \hlkwb{<-} \hlkwd{matrix}\hlstd{(w_plans[[i]],} \hlkwc{nrow} \hlstd{=} \hlnum{1}\hlstd{)}
 \hlstd{\}}
\hlstd{\}}
\end{alltt}
\end{kframe}
\end{knitrout}

Here's one example of generating a time series plot:

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{set.seed}\hlstd{(}\hlnum{123}\hlstd{)}
\hlstd{arma_env_params} \hlkwb{<-} \hlkwd{list}\hlstd{(}\hlkwc{mean_value} \hlstd{=} \hlnum{16}\hlstd{,} \hlkwc{ar} \hlstd{=} \hlnum{0.1}\hlstd{,} \hlkwc{sigma_env} \hlstd{=} \hlnum{3}\hlstd{,} \hlkwc{ma} \hlstd{=} \hlnum{0}\hlstd{)}
\hlstd{eg_arma} \hlkwb{<-} \hlkwd{meta_sim}\hlstd{(}\hlkwc{b} \hlstd{= w[[}\hlnum{1}\hlstd{]][[}\hlnum{1}\hlstd{]],} \hlkwc{n_pop} \hlstd{=} \hlnum{10}\hlstd{,} \hlkwc{env_params} \hlstd{= arma_env_params,}
  \hlkwc{env_type} \hlstd{=} \hlstr{"arma"}\hlstd{,} \hlkwc{assess_freq} \hlstd{=} \hlnum{5}\hlstd{)}
\hlkwd{plot_sim_ts}\hlstd{(eg_arma,} \hlkwc{years_to_show} \hlstd{=} \hlnum{100}\hlstd{,} \hlkwc{burn} \hlstd{=} \hlnum{30}\hlstd{)}
\end{alltt}
\end{kframe}

{\centering \includegraphics[width=4in]{figure/ts-example}

}

\end{knitrout}

Now, we'll run the simulations:

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{x_arma_sp} \hlkwb{<-} \hlkwd{run_cons_plans}\hlstd{(w,} \hlkwc{env_type} \hlstd{=} \hlstr{"arma"}\hlstd{,} \hlkwc{env_params} \hlstd{=}
  \hlstd{arma_env_params,} \hlkwc{show_progress} \hlstd{=} \hlnum{FALSE}\hlstd{)}
\hlstd{x_arma_sp}\hlopt{$}\hlstd{plans_port} \hlkwb{<-} \hlkwa{NULL} \hlcom{# save space}

\hlstd{linear_env_params} \hlkwb{<-} \hlkwd{list}\hlstd{(}\hlkwc{min_value} \hlstd{=} \hlnum{15}\hlstd{,} \hlkwc{max_value} \hlstd{=} \hlnum{19}\hlstd{,} \hlkwc{sigma_env} \hlstd{=} \hlnum{0.001}\hlstd{,}
  \hlkwc{start_t} \hlstd{=} \hlnum{30}\hlstd{)}
\hlstd{x_linear_sp} \hlkwb{<-} \hlkwd{run_cons_plans}\hlstd{(w,} \hlkwc{env_type} \hlstd{=} \hlstr{"linear"}\hlstd{,} \hlkwc{env_params} \hlstd{=}
  \hlstd{linear_env_params,} \hlkwc{max_a} \hlstd{=} \hlkwd{thermal_integration}\hlstd{(}\hlnum{10}\hlstd{),}
  \hlkwc{show_progress} \hlstd{=} \hlnum{FALSE}\hlstd{)}
\hlstd{x_linear_sp}\hlopt{$}\hlstd{plans_port} \hlkwb{<-} \hlkwa{NULL} \hlcom{# save space}
\end{alltt}
\end{kframe}
\end{knitrout}

And plot the output:

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{cols} \hlkwb{<-} \hlstd{RColorBrewer::}\hlkwd{brewer.pal}\hlstd{(}\hlnum{5}\hlstd{,} \hlstr{"Dark2"}\hlstd{)}
\hlstd{xlim} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlnum{0.18}\hlstd{,} \hlnum{0.80}\hlstd{)}
\hlstd{ylim} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlopt{-}\hlnum{0.027}\hlstd{,} \hlnum{0.027}\hlstd{)}
\hlkwd{par}\hlstd{(}\hlkwc{mfrow} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{2}\hlstd{))}
\hlkwd{par}\hlstd{(}\hlkwc{las} \hlstd{=} \hlnum{1}\hlstd{,} \hlkwc{cex} \hlstd{=} \hlnum{0.8}\hlstd{,} \hlkwc{mar} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{0}\hlstd{,} \hlnum{0}\hlstd{,} \hlnum{0}\hlstd{,} \hlnum{0}\hlstd{),} \hlkwc{oma} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{4}\hlstd{,} \hlnum{5.2}\hlstd{,} \hlnum{1.5}\hlstd{,} \hlnum{.5}\hlstd{),}
  \hlkwc{tck} \hlstd{=} \hlopt{-}\hlnum{0.02}\hlstd{,} \hlkwc{mgp} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{2}\hlstd{,} \hlnum{.6}\hlstd{,} \hlnum{0}\hlstd{))}
\hlkwd{plot_cons_plans}\hlstd{(x_arma_sp}\hlopt{$}\hlstd{plans_mv,} \hlkwc{plans_name} \hlstd{= plans_name_sp,} \hlkwc{cols} \hlstd{= cols,}
  \hlkwc{add_all_efs} \hlstd{=} \hlnum{FALSE}\hlstd{,} \hlkwc{xlim} \hlstd{= xlim,} \hlkwc{ylim} \hlstd{= ylim,} \hlkwc{add_legend} \hlstd{=} \hlnum{FALSE}\hlstd{)}
\hlkwd{mtext}\hlstd{(}\hlstr{"(a) Short-term environmental fluctuations"}\hlstd{,} \hlkwc{side} \hlstd{=} \hlnum{3}\hlstd{,} \hlkwc{line} \hlstd{=} \hlnum{0.2}\hlstd{,} \hlkwc{cex} \hlstd{=}
  \hlnum{0.8}\hlstd{,} \hlkwc{adj} \hlstd{=} \hlnum{0.05}\hlstd{)}
\hlkwd{par}\hlstd{(}\hlkwc{las} \hlstd{=} \hlnum{0}\hlstd{)}
\hlkwd{mtext}\hlstd{(}\hlstr{"Mean of metapopulation growth rate"}\hlstd{,} \hlkwc{side} \hlstd{=} \hlnum{2}\hlstd{,} \hlkwc{line} \hlstd{=} \hlnum{3}\hlstd{,} \hlkwc{outer} \hlstd{=} \hlnum{FALSE}\hlstd{,}
  \hlkwc{cex} \hlstd{=} \hlnum{0.8}\hlstd{)}
\hlkwd{par}\hlstd{(}\hlkwc{las} \hlstd{=} \hlnum{1}\hlstd{)}
\hlkwd{plot_cons_plans}\hlstd{(x_linear_sp}\hlopt{$}\hlstd{plans_mv,} \hlkwc{plans_name} \hlstd{= plans_name_sp,} \hlkwc{cols} \hlstd{= cols,}
  \hlkwc{add_all_efs} \hlstd{=} \hlnum{FALSE}\hlstd{,} \hlkwc{xlim} \hlstd{= xlim,} \hlkwc{ylim} \hlstd{= ylim,} \hlkwc{y_axis} \hlstd{=} \hlnum{FALSE}\hlstd{,} \hlkwc{add_legend} \hlstd{=}
  \hlnum{TRUE}\hlstd{,} \hlkwc{legend_pos} \hlstd{=} \hlstr{"bottomright"}\hlstd{)}
\hlkwd{mtext}\hlstd{(}\hlstr{"(b) Long-term environmental change"}\hlstd{,} \hlkwc{side} \hlstd{=} \hlnum{3}\hlstd{,} \hlkwc{line} \hlstd{=} \hlnum{0.2}\hlstd{,} \hlkwc{cex} \hlstd{=} \hlnum{0.8}\hlstd{,}
  \hlkwc{adj} \hlstd{=} \hlnum{0.05}\hlstd{)}
\hlkwd{mtext}\hlstd{(}\hlstr{"Variance of metapopulation growth rate"}\hlstd{,} \hlkwc{side} \hlstd{=} \hlnum{1}\hlstd{,} \hlkwc{line} \hlstd{=} \hlnum{2.25}\hlstd{,}
  \hlkwc{outer} \hlstd{=} \hlnum{FALSE}\hlstd{,} \hlkwc{cex} \hlstd{=} \hlnum{0.8}\hlstd{,} \hlkwc{adj} \hlstd{=} \hlopt{-}\hlnum{1}\hlstd{)}
\end{alltt}
\end{kframe}

{\centering \includegraphics[width=\maxwidth]{figure/plot-resp-div}

}

\end{knitrout}

\section{Numbers of populations conserved}

Here, we'll show the effect of increasing the number of streams that conservation effort is distributed across. We'll conserve the populations without knowing how response diversity is distributed.

We'll set up the scenarios:

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{set.seed}\hlstd{(}\hlnum{1}\hlstd{)}
\hlcom{# In this version we start with a set amount of b and can split it up among many}
\hlcom{# or invest it in a few}
\hlstd{n_trials} \hlkwb{<-} \hlnum{200} \hlcom{# number of trials at each n conservation plan}
\hlstd{num_pops} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlnum{2}\hlstd{,} \hlnum{4}\hlstd{,} \hlnum{8}\hlstd{,} \hlnum{16}\hlstd{)} \hlcom{# n pops to conserve}
\hlstd{b_conserve} \hlkwb{<-} \hlnum{2000} \hlopt{/} \hlstd{num_pops}
\hlstd{n_plans} \hlkwb{<-} \hlkwd{length}\hlstd{(num_pops)} \hlcom{# number of plans}
\hlstd{w} \hlkwb{<-} \hlkwd{list}\hlstd{()}
\hlkwa{for}\hlstd{(i} \hlkwa{in} \hlnum{1}\hlopt{:}\hlstd{n_plans) \{} \hlcom{# loop over number conserved}
 \hlstd{w[[i]]} \hlkwb{<-} \hlkwd{list}\hlstd{()}
 \hlkwa{for}\hlstd{(j} \hlkwa{in} \hlnum{1}\hlopt{:}\hlstd{n_trials) \{} \hlcom{# loop over trials}
   \hlstd{w[[i]][[j]]} \hlkwb{<-} \hlkwd{matrix}\hlstd{(}\hlkwd{rep}\hlstd{(b_conserve[i],} \hlnum{16}\hlstd{),} \hlkwc{nrow} \hlstd{=} \hlnum{1}\hlstd{)}
   \hlcom{# conserve num_pops[i] populations; wipe out rest:}
   \hlstd{w[[i]][[j]][}\hlopt{-}\hlkwd{sample}\hlstd{(}\hlnum{1}\hlopt{:}\hlnum{16}\hlstd{, num_pops[i])]} \hlkwb{<-} \hlnum{5}
 \hlstd{\}}
\hlstd{\}}
\hlstd{plans_name_n} \hlkwb{<-} \hlkwd{paste}\hlstd{(num_pops,} \hlstr{"populations"}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}

Now we'll run the simulations for the stationary climate variability and the non-stationary scenarios:

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{x_arma_n} \hlkwb{<-} \hlkwd{run_cons_plans}\hlstd{(w,} \hlkwc{env_type} \hlstd{=} \hlstr{"arma"}\hlstd{,} \hlkwc{env_params} \hlstd{=}
  \hlstd{arma_env_params,} \hlkwc{max_a} \hlstd{=} \hlkwd{thermal_integration}\hlstd{(}\hlnum{16}\hlstd{),}
  \hlkwc{show_progress} \hlstd{=} \hlnum{FALSE}\hlstd{)}
\hlstd{x_arma_n}\hlopt{$}\hlstd{plans_port} \hlkwb{<-} \hlkwa{NULL} \hlcom{# to save space}
\end{alltt}
\end{kframe}
\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{linear_env_params} \hlkwb{<-} \hlkwd{list}\hlstd{(}\hlkwc{min_value} \hlstd{=} \hlnum{15}\hlstd{,} \hlkwc{max_value} \hlstd{=} \hlnum{19}\hlstd{,} \hlkwc{sigma_env} \hlstd{=} \hlnum{0.001}\hlstd{,}
  \hlkwc{start_t} \hlstd{=} \hlnum{30}\hlstd{)}
\hlstd{x_linear_n} \hlkwb{<-} \hlkwd{run_cons_plans}\hlstd{(w,} \hlkwc{env_type} \hlstd{=} \hlstr{"linear"}\hlstd{,}
  \hlkwc{env_params} \hlstd{= linear_env_params,} \hlkwc{max_a} \hlstd{=} \hlkwd{thermal_integration}\hlstd{(}\hlnum{16}\hlstd{),}
  \hlkwc{show_progress} \hlstd{=} \hlnum{FALSE}\hlstd{)}
\hlstd{x_linear_n}\hlopt{$}\hlstd{plans_port} \hlkwb{<-} \hlkwa{NULL} \hlcom{# save space}
\end{alltt}
\end{kframe}
\end{knitrout}

And plot the output:

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{cols} \hlkwb{<-} \hlstd{RColorBrewer::}\hlkwd{brewer.pal}\hlstd{(}\hlnum{5}\hlstd{,} \hlstr{"Greys"}\hlstd{)[}\hlkwd{c}\hlstd{(}\hlnum{2}\hlopt{:}\hlnum{5}\hlstd{)]}
\hlstd{xlim} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlnum{0.008}\hlstd{,} \hlnum{0.90}\hlstd{)}
\hlstd{ylim} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlopt{-}\hlnum{0.034}\hlstd{,} \hlnum{0.027}\hlstd{)}
\hlkwd{par}\hlstd{(}\hlkwc{mfrow} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{2}\hlstd{))}
\hlkwd{par}\hlstd{(}\hlkwc{las} \hlstd{=} \hlnum{1}\hlstd{,} \hlkwc{cex} \hlstd{=} \hlnum{0.8}\hlstd{,} \hlkwc{mar} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{0}\hlstd{,} \hlnum{0}\hlstd{,} \hlnum{0}\hlstd{,} \hlnum{0}\hlstd{),} \hlkwc{oma} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{4}\hlstd{,} \hlnum{5.2}\hlstd{,} \hlnum{1.5}\hlstd{,} \hlnum{.5}\hlstd{),}
  \hlkwc{tck} \hlstd{=} \hlopt{-}\hlnum{0.02}\hlstd{,} \hlkwc{mgp} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{2}\hlstd{,} \hlnum{.6}\hlstd{,} \hlnum{0}\hlstd{))}
\hlkwd{plot_cons_plans}\hlstd{(x_arma_n}\hlopt{$}\hlstd{plans_mv,} \hlkwc{plans_name} \hlstd{= plans_name_n,} \hlkwc{cols} \hlstd{= cols,}
  \hlkwc{add_all_efs} \hlstd{=} \hlnum{FALSE}\hlstd{,} \hlkwc{xlim} \hlstd{= xlim,} \hlkwc{ylim} \hlstd{= ylim,} \hlkwc{add_legend} \hlstd{=} \hlnum{FALSE}\hlstd{)}
\hlkwd{mtext}\hlstd{(}\hlstr{"(a) Short-term environmental fluctuations"}\hlstd{,} \hlkwc{side} \hlstd{=} \hlnum{3}\hlstd{,} \hlkwc{line} \hlstd{=} \hlnum{0.2}\hlstd{,}
  \hlkwc{cex} \hlstd{=} \hlnum{0.8}\hlstd{,} \hlkwc{adj} \hlstd{=} \hlnum{0.05}\hlstd{)}
\hlkwd{par}\hlstd{(}\hlkwc{las} \hlstd{=} \hlnum{0}\hlstd{)}
\hlkwd{mtext}\hlstd{(}\hlstr{"Mean of metapopulation growth rate"}\hlstd{,} \hlkwc{side} \hlstd{=} \hlnum{2}\hlstd{,} \hlkwc{line} \hlstd{=} \hlnum{3}\hlstd{,}
  \hlkwc{outer} \hlstd{=} \hlnum{FALSE}\hlstd{,} \hlkwc{cex} \hlstd{=} \hlnum{0.8}\hlstd{)}
\hlkwd{par}\hlstd{(}\hlkwc{las} \hlstd{=} \hlnum{1}\hlstd{)}
\hlkwd{plot_cons_plans}\hlstd{(x_linear_n}\hlopt{$}\hlstd{plans_mv,} \hlkwc{plans_name} \hlstd{= plans_name_n,} \hlkwc{cols} \hlstd{= cols,}
  \hlkwc{add_all_efs} \hlstd{=} \hlnum{FALSE}\hlstd{,} \hlkwc{xlim} \hlstd{= xlim,} \hlkwc{ylim} \hlstd{= ylim,} \hlkwc{y_axis} \hlstd{=} \hlnum{FALSE}\hlstd{,} \hlkwc{add_legend} \hlstd{=} \hlnum{TRUE}\hlstd{)}
\hlkwd{mtext}\hlstd{(}\hlstr{"(b) Long-term environmental change"}\hlstd{,} \hlkwc{side} \hlstd{=} \hlnum{3}\hlstd{,} \hlkwc{line} \hlstd{=} \hlnum{0.2}\hlstd{,} \hlkwc{cex} \hlstd{=} \hlnum{0.8}\hlstd{,}
  \hlkwc{adj} \hlstd{=} \hlnum{0.05}\hlstd{)}
\hlkwd{par}\hlstd{(}\hlkwc{xpd} \hlstd{=} \hlnum{NA}\hlstd{)}
\hlkwd{mtext}\hlstd{(}\hlstr{"Variance of metapopulation growth rate"}\hlstd{,} \hlkwc{side} \hlstd{=} \hlnum{1}\hlstd{,} \hlkwc{line} \hlstd{=} \hlnum{2.25}\hlstd{,}
  \hlkwc{outer} \hlstd{=} \hlnum{FALSE}\hlstd{,} \hlkwc{cex} \hlstd{=} \hlnum{0.8}\hlstd{,} \hlkwc{adj} \hlstd{=} \hlopt{-}\hlnum{1}\hlstd{)}
\end{alltt}
\end{kframe}

{\centering \includegraphics[width=\maxwidth]{figure/plot-number}

}

\end{knitrout}

\section{Reduction of habitat over time}

Here we'll reduce the available habitat by a fixed quantity at every generation. The climate will follow a combination of short-term variability and long-term change.

Set up the scenarios:

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{set.seed}\hlstd{(}\hlnum{1}\hlstd{)}
\hlstd{n_trials} \hlkwb{<-} \hlnum{200} \hlcom{# number of trials at each n conservation plan}
\hlstd{num_pops} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlnum{2}\hlstd{,} \hlnum{4}\hlstd{,} \hlnum{8}\hlstd{,} \hlnum{12}\hlstd{,} \hlnum{16}\hlstd{)} \hlcom{# n pops to conserve}
\hlstd{b_conserve} \hlkwb{<-} \hlnum{2000} \hlopt{/} \hlstd{num_pops}
\hlstd{n_plans} \hlkwb{<-} \hlkwd{length}\hlstd{(num_pops)}
\hlstd{w} \hlkwb{<-} \hlkwd{list}\hlstd{()}
\hlkwa{for}\hlstd{(i} \hlkwa{in} \hlnum{1}\hlopt{:}\hlstd{n_plans) \{} \hlcom{# loop over number conserved}
 \hlstd{w[[i]]} \hlkwb{<-} \hlkwd{list}\hlstd{()}
 \hlkwa{for}\hlstd{(j} \hlkwa{in} \hlnum{1}\hlopt{:}\hlstd{n_trials) \{} \hlcom{# loop over trials}
   \hlstd{w[[i]][[j]]} \hlkwb{<-} \hlkwd{matrix}\hlstd{(}\hlkwd{rep}\hlstd{(b_conserve[i],} \hlnum{16}\hlstd{),} \hlkwc{nrow} \hlstd{=} \hlnum{1}\hlstd{)}
   \hlcom{# conserve num_pops[i] populations; wipe out rest:}
   \hlstd{w[[i]][[j]][}\hlopt{-}\hlkwd{sample}\hlstd{(}\hlnum{1}\hlopt{:}\hlnum{16}\hlstd{, num_pops[i])]} \hlkwb{<-} \hlnum{5}
 \hlstd{\}}
\hlstd{\}}
\hlstd{plans_name_n} \hlkwb{<-} \hlkwd{paste}\hlstd{(num_pops,} \hlstr{"populations"}\hlstd{)}
\hlstd{cols} \hlkwb{<-} \hlstd{RColorBrewer::}\hlkwd{brewer.pal}\hlstd{(}\hlnum{6}\hlstd{,} \hlstr{"Greys"}\hlstd{)[}\hlkwd{c}\hlstd{(}\hlnum{2}\hlopt{:}\hlnum{6}\hlstd{)]}
\end{alltt}
\end{kframe}
\end{knitrout}

Run the simulations:

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{linear_arma_env_params} \hlkwb{<-} \hlkwd{list}\hlstd{(}\hlkwc{min_value} \hlstd{=} \hlnum{15}\hlstd{,} \hlkwc{max_value} \hlstd{=} \hlnum{19}\hlstd{,}
  \hlkwc{start_t} \hlstd{=} \hlnum{30}\hlstd{,} \hlkwc{mean_value} \hlstd{=} \hlnum{16}\hlstd{,} \hlkwc{ar} \hlstd{=} \hlnum{0.1}\hlstd{,} \hlkwc{sigma_env} \hlstd{=} \hlnum{2}\hlstd{,} \hlkwc{ma} \hlstd{=} \hlnum{0}\hlstd{)}
\hlstd{x_linear_arma_n} \hlkwb{<-} \hlkwd{run_cons_plans}\hlstd{(w,} \hlkwc{env_type} \hlstd{=} \hlstr{"linear_arma"}\hlstd{,} \hlkwc{env_params} \hlstd{=}
  \hlstd{linear_arma_env_params,} \hlkwc{max_a} \hlstd{=} \hlkwd{thermal_integration}\hlstd{(}\hlnum{16}\hlstd{),} \hlkwc{decrease_b} \hlstd{=} \hlnum{0.85}\hlstd{,}
  \hlkwc{show_progress} \hlstd{=} \hlnum{FALSE}\hlstd{)}
\hlstd{x_linear_arma_n}\hlopt{$}\hlstd{plans_port} \hlkwb{<-} \hlkwa{NULL} \hlcom{# save space}
\end{alltt}
\end{kframe}
\end{knitrout}

Plot the output:

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{xlim} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlnum{0.08}\hlstd{,} \hlnum{0.9}\hlstd{)}
\hlstd{ylim} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlopt{-}\hlnum{0.038}\hlstd{,} \hlnum{0.028}\hlstd{)}
\hlkwd{par}\hlstd{(}\hlkwc{las} \hlstd{=} \hlnum{1}\hlstd{,} \hlkwc{cex} \hlstd{=} \hlnum{0.8}\hlstd{,} \hlkwc{mar} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{0}\hlstd{,} \hlnum{0}\hlstd{,} \hlnum{0}\hlstd{,} \hlnum{0}\hlstd{),} \hlkwc{oma} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{4}\hlstd{,} \hlnum{5.2}\hlstd{,} \hlnum{1.8}\hlstd{,} \hlnum{.5}\hlstd{),}
  \hlkwc{tck} \hlstd{=} \hlopt{-}\hlnum{0.02}\hlstd{,} \hlkwc{mgp} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{2}\hlstd{,} \hlnum{.5}\hlstd{,} \hlnum{0}\hlstd{))}
\hlkwd{plot_cons_plans}\hlstd{(x_linear_arma_n}\hlopt{$}\hlstd{plans_mv,} \hlkwc{plans_name} \hlstd{= plans_name_n,} \hlkwc{cols} \hlstd{= cols,}
  \hlkwc{add_all_efs} \hlstd{=} \hlnum{FALSE}\hlstd{,} \hlkwc{xlim} \hlstd{= xlim,} \hlkwc{ylim} \hlstd{= ylim,} \hlkwc{add_legend} \hlstd{=} \hlnum{TRUE}\hlstd{,}
  \hlkwc{add_poly} \hlstd{=} \hlnum{TRUE}\hlstd{,} \hlkwc{legend_pos} \hlstd{=} \hlstr{"bottomright"}\hlstd{)}
\hlkwd{mtext}\hlstd{(}\hlstr{"Reduction in stream flow"}\hlstd{,} \hlkwc{side} \hlstd{=} \hlnum{3}\hlstd{,} \hlkwc{line} \hlstd{=} \hlnum{.4}\hlstd{,}
  \hlkwc{cex} \hlstd{=} \hlnum{0.8}\hlstd{,} \hlkwc{adj} \hlstd{=} \hlnum{0.05}\hlstd{)}
\hlkwd{mtext}\hlstd{(}\hlstr{"Variance of metapopulation growth rate"}\hlstd{,} \hlkwc{side} \hlstd{=} \hlnum{1}\hlstd{,} \hlkwc{line} \hlstd{=} \hlnum{2.25}\hlstd{,}
  \hlkwc{outer} \hlstd{=} \hlnum{FALSE}\hlstd{,} \hlkwc{cex} \hlstd{=} \hlnum{0.8}\hlstd{)}
\hlkwd{par}\hlstd{(}\hlkwc{las} \hlstd{=} \hlnum{0}\hlstd{)}
\hlkwd{mtext}\hlstd{(}\hlstr{"Mean of metapopulation growth rate"}\hlstd{,} \hlkwc{side} \hlstd{=} \hlnum{2}\hlstd{,} \hlkwc{line} \hlstd{=} \hlnum{3}\hlstd{,}
  \hlkwc{outer} \hlstd{=} \hlnum{FALSE}\hlstd{,} \hlkwc{cex} \hlstd{=} \hlnum{0.8}\hlstd{)}
\end{alltt}
\end{kframe}

{\centering \includegraphics[width=3.75in]{figure/squeeze-plot}

}

\end{knitrout}

\end{document}

### 2014-08-04 inst/ms/anderson-etal-metafolio.tex
Prep ms for final submission - clean up
Edits for ESA editing; see details
Move first appendix to supplement
Reorder appendices
Create new supplement describing code

@@
 \usepackage{lineno}
-\usepackage{xcolor}
-%\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
 \usepackage{booktabs}
@@
 \usepackage{array}
-%\usepackage{mathptmx}
 
@@
 
-\newcommand{\somR}{Appendix A}
+\newcommand{\somR}{Supplement}
+\newcommand{\somstray}{Appendix A}
 \newcommand{\somparam}{Appendix B}
-\newcommand{\somstray}{Appendix C}
-\newcommand{\somsens}{Appendix D}
+\newcommand{\somsens}{Appendix C}
+\newcommand{\somts}{Appendix D}
 \newcommand{\somcor}{Appendix E}
-\newcommand{\somts}{Appendix F}
 
 % SOM:
-\newlabel{t:pars}{{B1}{2}}
-\newlabel{f:stray}{{C1}{3}}
-\newlabel{f:eg-sens}{{D1}{4}}
+\newlabel{f:stray}{{A1}{2}}
+\newlabel{t:pars}{{B1}{3}}
+\newlabel{f:eg-sens}{{C1}{4}}
+\newlabel{f:eg-sp-arma-full}{{D1}{6}}
+\newlabel{f:eg-sp-arma-half}{{D2}{7}}
+\newlabel{f:eg-sp-linear-full}{{D3}{8}}
+\newlabel{f:eg-sp-linear-half}{{D4}{9}}
+\newlabel{f:eg-n-arma-two}{{D5}{10}}
+\newlabel{f:eg-n-arma-sixteen}{{D6}{11}}
+\newlabel{f:eg-n-linear-two}{{D7}{12}}
+\newlabel{f:eg-n-linear-sixteen}{{D8}{13}}
+\newlabel{f:eg-n-squeeze-two}{{D9}{14}}
+\newlabel{f:eg-n-squeeze-twelve}{{D10}{15}}
 \newlabel{f:ret-corr}{{E1}{5}}
-\newlabel{f:eg-sp-arma-full}{{F1}{6}}
-\newlabel{f:eg-sp-arma-half}{{F2}{7}}
-\newlabel{f:eg-sp-linear-full}{{F3}{8}}
-\newlabel{f:eg-sp-linear-half}{{F4}{9}}
-\newlabel{f:eg-n-arma-two}{{F5}{10}}
-\newlabel{f:eg-n-arma-sixteen}{{F6}{11}}
-\newlabel{f:eg-n-linear-two}{{F7}{12}}
-\newlabel{f:eg-n-linear-sixteen}{{F8}{13}}
-\newlabel{f:eg-n-squeeze-two}{{F9}{14}}
-\newlabel{f:eg-n-squeeze-twelve}{{F10}{15}}
 
@@
 \clubpenalty=10000
-
-%\usepackage{titlesec}
-%\titlespacing\section{0pt}{6pt plus 4pt minus 2pt}{-8pt plus 2pt minus 2pt}
-%\titlespacing\subsection{0pt}{6pt plus 4pt minus 2pt}{-8pt plus 2pt minus 2pt}
 
@@
 \hyphenation{meta-pop-ulation meta-pop-ulations sub-pop-ulations sub-pop-ulation e-con-o-mist en-vi-ron-men-tal pri-or-i-ti-za-tion}
-
-%\newcommand{\R}[1]{\label{#1}\linelabel{#1}}
-%\newcommand{\lr}[1]{line~\lineref{#1}}
 
@@
 
-%\modulolinenumbers[2]
 \begin{spacing}{1.9}
-%\begin{spacing}{1.2}
 \setlength{\parindent}{0.9cm}
@@
 We tested different conservation scenarios under two kinds of environmental regimes (short-term climate variability and long-term climate change) and in cases where habitat capacity remained constant or declined over time.
-We provide a package \texttt{metafolio} \citep{metafoliopkg} for the statistical software \textsf{R} \citep{r2013} as an appendix, to carry out the simulations and analyses described in this paper (\somR).
+We provide a package \texttt{metafolio} \citep{metafoliopkg} for the statistical software \textsf{R} \citep{r2013}, to carry out the simulations and analyses described in this paper (\somR).
 
@@
 \emph{Declining habitat availability}: Habitat capacity in the Pacific Northwest is likely shrinking over time as salmon populations are squeezed between warming temperatures reducing habitat from below and declining stream flows reducing the habitat that remains from above.
-For example, temperature isotherms are shifting upstream at 1--25 km/decade in low gradient streams that Chinook use for spawning \citep{isaak2013}.
+For example, temperature isotherms are shifting upstream at 1--10 km/decade in low gradient streams that Chinook use for spawning \citep{isaak2013}.
 At the same time, summer-fall stream flow volumes have been decreasing 10--30\% across the Pacific Northwest over the past 50 years \citep{luce2009} and are likely to continue declining \citep{luce2013}.
@@
 Anderson, S.~C. 2014.
-\newblock metafolio: Salmon metapopulation simulations for conservation
-  planning through portfolio optimization. R package version 0.1.0.
-  http://cran.r-project.org/package=metafolio.
+\newblock \texttt{metafolio}: Metapopulation simulations for conserving salmon through
+  portfolio optimization. \textsf{R} package version 0.1.0.
+  \texttt{http://cran.r-project.org/package=metafolio}.
 
@@
 \noindent
-\textsc{Appendix A.} The \texttt{metafolio} \textsf{R} package.
+\textsc{Supplement} Recreating the analyses with the \texttt{metafolio} \textsf{R} package.
+
+\noindent
+\textsc{Appendix A.} An example straying matrix.
 
@@
 \noindent
-\textsc{Appendix C.} An example straying matrix.
-
-\noindent
-\textsc{Appendix D.} Sensitivity illustration with alternative parameter values.
+\textsc{Appendix C.} Sensitivity illustration with alternative parameter values.
+
+\noindent
+\textsc{Appendix D.} Example simulated time series from alternative conservation scenarios.
 
@@
 
-\noindent
-\textsc{Appendix F.} Example simulated time series from alternative conservation scenarios.
-
 \clearpage
@@
 
-%\begin{center}
-%\end{center}
 \textsc{Fig. 1}. Flow chart of the salmon-metapopulation simulation.
@@
 Also shown are (b) example metapopulation growth rate and (c) abundance time series from the 2 and 16 population scenarios.
-Regression lines in (b) illustrate a decreasing growth rate through time.
 
@@
 
-\section{Figures}
-
-\begin{center}
-\includegraphics[height=5.5in]{Fig1}\\
-\textsc{Fig.} 1
-\clearpage
-\includegraphics[width=2.4in]{Fig2}\\
-\textsc{Fig.} 2
-\clearpage
-\includegraphics[width=4.2in]{Fig3}\\
-\textsc{Fig.} 3
-\clearpage
-\includegraphics[width=4.5in]{Fig4}\\
-\textsc{Fig.} 4
-\clearpage
-\includegraphics[width=4.5in]{Fig5}\\
-\textsc{Fig.} 5
-\clearpage
-\includegraphics[width=3.0in]{Fig6}\\
-\textsc{Fig.} 6
-\clearpage
-\end{center}
+%\section{Figures}
+
+% \begin{center}
+% \includegraphics[height=5.5in]{Fig1}\\
+% \textsc{Fig.} 1
+% \clearpage
+% \includegraphics[width=2.4in]{Fig2}\\
+% \textsc{Fig.} 2
+% \clearpage
+% \includegraphics[width=4.2in]{Fig3}\\
+% \textsc{Fig.} 3
+% \clearpage
+% \includegraphics[width=4.5in]{Fig4}\\
+% \textsc{Fig.} 4
+% \clearpage
+% \includegraphics[width=4.5in]{Fig5}\\
+% \textsc{Fig.} 5
+% \clearpage
+% \includegraphics[width=3.0in]{Fig6}\\
+% \textsc{Fig.} 6
+% \clearpage
+% \end{center}
 

### 2014-08-04 inst/ms/appendices/Appendix-C.tex
Edits for ESA editing; see details
Move first appendix to supplement
Reorder appendices
Create new supplement describing code

@@
 Portfolio conservation of metapopulations under climate change.
-Ecological Applications.
-VOL:pp--pp.}
+\textit{Ecological Applications} VOL:pp--pp.}
 
@@
 
-\textsc{Appendix C.} An example straying matrix.
+\textsc{Appendix C.} Sensitivity illustration with alternative parameter values.
 
@@
 
-\includegraphics[width=3.5in]{Fig-C1}\\
-\textsc{Fig.~C1.} An example straying matrix.
-The rows and columns represent different populations (indicated by population number).
-Dark blue indicates a high rate of straying and light blue indicates a low rate of straying.
+\includegraphics[width=3.2in]{Fig-C1}\\
+\textsc{Fig.~C1.} The impact of increasing or decreasing various parameter values on metapopulation return abundance.
+The different coloured lines represent three example salmon populations.
+The base case represents the base-case values for the short-term environmental fluctuation scenario.
 

### 2014-08-04 R/run_cons_plans.R

@@ -107,7 +107,7 @@ run_cons_plans <- function(w, env_type, env_params, show_progress =
     plans_port_n[[i]][[j]] <- temp$port_out
     }
     if(show_progress)
-      print(paste("Completed", i, "of", length(w), "conservation plans to evaluate"))
+      message(paste("Completed", i, "of", length(w), "conservation plans to evaluate"))
   }
 
   # here we have a nested list that is 4 elements deep

### 2014-08-04 inst/ms/appendices/Appendix-A.tex
Add back 2 appendices

@@
 \documentclass[12pt]{article}
+\usepackage{graphicx}
 
@@
 
-\textsc{Appendix A.} The \texttt{metafolio} \textsf{R} package.
+\textsc{Appendix A.} An example straying matrix.
 
@@
 
-The \texttt{metafolio} \textsf{R} package contains the functions and code to carry out the analyses in our paper. \texttt{metafolio}. The package can be installed from CRAN with:
-
-\begin{verbatim}
-install.packages("metafolio")
-\end{verbatim}
-
-%Alternatively, you can view the code and install the package from\\ \texttt{http://github.com/seananderson/metafolio}
-%\bigskip
-
-The included vignette describes the package and illustrates some example simulations. You can view the vignette with:
-
-\begin{verbatim}
-vignette("metafolio")
-\end{verbatim}
-
-You can view the help for the package with:
-
-\begin{verbatim}
-?metafolio
-help(package = "metafolio")
-\end{verbatim}
-
-The figures from this paper can be re-created by downloading the code from GitHub and sourcing the file \texttt{README.R} in the \texttt{inst/examples} folder:
-
-\begin{verbatim}
-setwd("metafolio/inst/examples")
-source("README.R")
-\end{verbatim}
+\includegraphics[width=3.5in]{Fig-A1}\\
+\textsc{Fig.~A1.} An example straying matrix.
+The rows and columns represent different populations (indicated by population number).
+Dark blue indicates a high rate of straying and light blue indicates a low rate of straying.
 

### 2014-08-04 inst/ms/appendices/Appendix-D.tex
Reduce fig size
Add back 2 appendices

--- 2014-07-10 inst/ms/appendices/Appendix-C.tex
@@
 
-\textsc{Appendix C.} An example straying matrix.
+\textsc{Appendix D.} Example simulated time series from alternative conservation scenarios.
 
@@
 
-\includegraphics[width=3.5in]{Fig-C1}\\
-\textsc{Fig.~C1.} An example straying matrix.
-The rows and columns represent different populations (indicated by population number).
-Dark blue indicates a high rate of straying and light blue indicates a low rate of straying.
+\clearpage
+
+\includegraphics[width=4.3in]{Fig-D1}\\
+\textsc{Fig.~D1.} Conserving a \textbf{full range} of response diversity (spatial conservation strategy) with \textbf{short-term} environmental fluctuations. This is the same as Fig.~3 but in colour.
+\clearpage
+
+\includegraphics[width=4.3in]{Fig-D2}\\
+\textsc{Fig.~D2.} Conserving \textbf{one half} of response diversity (spatial conservation strategy) with \textbf{short-term} environmental fluctuations.
+\clearpage
+
+\includegraphics[width=4.3in]{Fig-D3}\\
+\textsc{Fig.~D3.} Conserving a \textbf{full range} of response diversity (spatial conservation strategy) with \textbf{long-term} environmental change.
+\clearpage
+
+\includegraphics[width=4.3in]{Fig-D4}\\
+\textsc{Fig.~D4.} Conserving \textbf{one half} of response diversity (spatial conservation strategy) with \textbf{long-term} environmental change.
+\clearpage
+
+\includegraphics[width=4.3in]{Fig-D5}\\
+\textsc{Fig.~D5.} \textbf{Two populations} conserved with random response diversity and \textbf{short-term} environmental fluctuations.
+\clearpage
+
+\includegraphics[width=4.3in]{Fig-D6}\\
+\textsc{Fig.~D6.} \textbf{Sixteen populations} conserved with random response diversity and \textbf{short-term} environmental fluctuations.
+\clearpage
+
+\includegraphics[width=4.3in]{Fig-D7}\\
+\textsc{Fig.~D7.} \textbf{Two populations} conserved with random response diversity and \textbf{long-term} environmental change.
+\clearpage
+
+\includegraphics[width=4.3in]{Fig-D8}\\
+\textsc{Fig.~D8.} \textbf{Sixteen populations} conserved with random response diversity and \textbf{long-term} environmental change.
+\clearpage
+
+\includegraphics[width=4.3in]{Fig-D9}\\
+\textsc{Fig.~D9.} \textbf{Two populations} conserved with random response diversity and \textbf{long-term declining stream flow}.
+\clearpage
+
+\includegraphics[width=4.3in]{Fig-D10}\\
+\textsc{Fig.~D10.} \textbf{Sixteen populations} conserved with random response diversity and \textbf{long-term declining stream flow}.
 

### 2014-10-20 inst/ms/anderson-etal-metafolio.tex
Tweak ms for author preprint format

@@
 \usepackage{cite}
-\usepackage[originalparameters]{ragged2e}
+%\usepackage[originalparameters]{ragged2e}
 \usepackage{array}
@@
 \newcommand{\somcor}{Appendix E}
+
+% Linux Libertine:
+\usepackage{textcomp}
+\usepackage[sb]{libertine}
+\usepackage[varqu,varl]{inconsolata}% sans serif typewriter
+\usepackage[libertine,bigdelims,vvarbb]{newtxmath} % bb from STIX
+\usepackage[cal=boondoxo]{mathalfa} % mathcal
+\useosf % osf for text, not math
+\usepackage[supstfm=libertinesups,%
+  supscaled=1.2,%
+  raised=-.13em]{superiors}
 
@@
 
+\usepackage{xcolor}
+\usepackage{adjustbox}
+\usepackage{verbatim}
+\definecolor{shadecolor}{rgb}{.9, .9, .9}
+
+\newenvironment{code}%
+   {\par\noindent\adjustbox{margin=1ex,bgcolor=shadecolor,margin=0ex \medskipamount}\bgroup\varwidth\linewidth\verbatim}%
+   {\endverbatim\endvarwidth\egroup}
+
+\newenvironment{code2}%
+   {\par\noindent\adjustbox{margin=1ex,bgcolor=shadecolor,margin=0ex \medskipamount}\bgroup\minipage\linewidth\verbatim}%
+   {\endverbatim\endminipage\egroup}
+
 \begin{document}
-\raggedright
+%\raggedright
 
@@
 \setcounter{page}{1}
-\linenumbers
-
-\begin{spacing}{1.9}
+%\linenumbers
+
+\begin{spacing}{1.3}
 \setlength{\parindent}{0.9cm}
@@
 Anderson; Earth to Ocean Research Group, Department of Biological Sciences, Simon Fraser University, Burnaby BC, V5A 1S6; Phone: 1-778-782-3989; E-mail: sean\_anderson@sfu.ca
+
+\bigskip
+\bigskip
+\begin{code}
+This document is an author preprint of the following published paper:
+Anderson, S.C., J.W. Moore, M.M. McClure, N.K. Dulvy, A.B. Cooper. Portfolio
+conservation of metapopulations under climate change. Ecological Applications.
+http://doi.org/10.1890/14-0266.1
+\end{code}
 
@@
 \textit{Running head}: Metapopulation portfolio conservation
+
+\clearpage
 
@@
 
+\begin{center}
+\includegraphics[width=4in]{Fig1.pdf}
+\end{center}
+\bigskip
+\noindent
 \textsc{Fig. 1}. Flow chart of the salmon-metapopulation simulation.
 
-%\clearpage
-
-%\begin{center}
-%\includegraphics[width=2.9in]{../examples/thermal-curve-scenarios}
-%\end{center}
+\clearpage
+
+\begin{center}
+\includegraphics[width=2.9in]{Fig2.pdf}
+\end{center}
 \bigskip
@@
 
-%\clearpage
-
-%\begin{center}
-%\includegraphics[width=4.0in]{../examples/spatial-arma-sim-full}
-%\end{center}
+\clearpage
+
+\begin{center}
+\includegraphics[width=4.0in]{Fig3.pdf}
+\end{center}
 \bigskip
@@
 We show, from top to bottom, the temperature signal, the resulting productivity parameter (Ricker $a$), the salmon returns, fisheries catch, salmon escapement, salmon straying from their natal streams, salmon joining from other streams, spawner-return residuals on a log scale, and the estimated $a$ and $b$ parameters in the fitted Ricker curve.
-The shaded lines indicate populations that thrive at low (light grey) to high (dark grey) temperatures.
-
-%\clearpage
-
-%\begin{center}
-%\includegraphics[width=4.5in]{../examples/spatial-mv}
-%\end{center}
+The coloured lines indicate populations that thrive at low (blue) to high (red) temperatures.
+
+\clearpage
+
+\begin{center}
+\includegraphics[width=4.5in]{Fig4.pdf}
+\end{center}
 \bigskip
@@
 
-%\clearpage
-
-%\begin{center}
-%\includegraphics[width=4.5in]{../examples/cons-plans-n}
-%\end{center}
+\clearpage
+
+\begin{center}
+\includegraphics[width=4.5in]{Fig5.pdf}
+\end{center}
 \bigskip
@@
 
+\begin{center}
+\includegraphics[width=3in]{Fig6.pdf}
+\end{center}
 \bigskip
